### 2020-01-31 misc/meeting_notes.Rmd
Added meeting notes

---
output:
  pdf_document: default
  html_document: default
---

## Meeting notes
* Company helps appeal assessment values for residential and commercial 
* Both residential and commercial are important to the model, although the majority of company's profits come from corporate clients
* Tax classes (classified by government, with 01 residential, 05 light industrial, 06 commercial being the importnat classes)
* Some condos + commercial complexes have split classes
* Mill rate is used interchangeably with tax rate, and is unique to a  specific code (called mill rate because its over 1000)
* eg: “12.3* assessment /1000” mill rate is 12.3
* Three things he is interested in primarily:
     1. Predicting mill rate (published in April but assessments published in January) which changes every year, during the blank period from jan-april
     2. Predicting next years assessment values (once mill rates are known, for the period between May-Dec)
     3. Predicting final tax that will be owed
* Their service is used to help with tax budgeting
* Sometimes values are missing, but we still need to predict these things given what we do have available for a given property
* Municipal government has budget, and determines mill rates for each tax code in order to raise the budget amount (from other funds also, but estimated about 90% of fund raised by property taxes)
* Mentioned it would be interesting to quantify relationship between mill rates and assessment values, but accurate prediction is priority
* Roll number uniquely identifies properties across datasets (see handout for the form of this identifier)
* 2016-2020 data is available (2015?)
* Vancouver municipality is a special case when it comes to tax coding - this should be treated as such and is not comparable to other municipalities
* Sales data is available (they bought this from government); each row is a sales transaction (filter to market sales only, to leave name transfers out of it)
* Currently if property assessments go up 10%, they are assuming mill rate goes down 6-7% due to inflation and larger budgets and predicting like this (it's not working well)
* Assumes large chunk of variability in mill rate is due to assessment (and not just municipal budget amount)
* Table handout is for changes in assessment values (are they averages? he's not sure but we could check)
* Exemption exists but is maybe not that important
* Whatever information is used for assessment by the government is also available to us

### 2020-02-25 src/tax2016-2018.Rmd
filtered tax for municipalities in metro vancouver from 2016-2018 and computed percentage change of tax from 2017-2018.

---
title: "budget"
author: "Xuechun Lu"
date: "25/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
```

```{r}
schedule2016<-read_excel(here("data/tax", "schedule706_2016.xls"))
schedule2017<-read_excel(here("data/tax", "schedule706_2017.xls"))
schedule2018<-read_excel(here("data/tax", "schedule706_2018.xls"))

colnames(schedule2016) <- schedule2016[1,]
schedule2016<-schedule2016[-c(1),]
colnames(schedule2017) <- schedule2017[1,]
schedule2017<-schedule2017[-c(1),]
colnames(schedule2018) <- schedule2018[1,]
schedule2018<-schedule2018[-c(1),]

municipality.list = c(
  "Burnaby", 
  "Coquitlam", 
  "Delta", 
  "Langley", 
  #"Langley - Township",
  "Maple Ridge",
  #"Maple Ridge Rural", 
  "North Vancouver",
  #"North Vancouver - Dist",
  "Pitt Meadows", 
  "Port Coquitlam", 
  "Port Moody", 
  "Richmond", 
  "Surrey", 
  "Vancouver", 
  "White Rock", 
  "West Vancouver", 
  "Bowen Island", 
  "Anmore", 
  "Belcarra",
  "Lions Bay")

tax.2016<-schedule2016 %>% filter(Municipalities %in% municipality.list) %>% 
  select("Municipalities","Type","Current Year Taxes Collected")
tax.2017<-schedule2017 %>% filter(Municipalities %in% municipality.list) %>% 
  select("Municipalities","Type","Current Year Taxes Collected")
tax.2018<-schedule2018 %>% filter(Municipalities %in% municipality.list) %>% 
  select("Municipalities","Type","Current Year Taxes Collected")

# rename
tax.2016[7,1]<-"Langley - City"
tax.2016[8,1]<-"Langley - Township"
tax.2016[11,1]<-"North Vancouver - City"
tax.2016[12,1]<-"North Vancouver - Dist"

tax.2017[7,1]<-"Langley - City"
tax.2017[8,1]<-"Langley - Township"
tax.2017[11,1]<-"North Vancouver - City"
tax.2017[12,1]<-"North Vancouver - Dist"

tax.2018[7,1]<-"Langley - City"
tax.2018[8,1]<-"Langley - Township"
tax.2018[11,1]<-"North Vancouver - City"
tax.2018[12,1]<-"North Vancouver - Dist"

tax.2016<-tax.2016 %>% select("Municipalities","Current Year Taxes Collected")
tax.2017<-tax.2017 %>% select("Municipalities","Current Year Taxes Collected")
tax.2018<-tax.2018 %>% select("Municipalities","Current Year Taxes Collected")

names(tax.2016)[2]<-paste("2016")
names(tax.2017)[2]<-paste("2017")
names(tax.2018)[2]<-paste("2018")
tax.full<-tax.2016 %>% left_join(tax.2017, by = "Municipalities") %>% 
   left_join(tax.2018, by = "Municipalities")

# convert chr to numeric
tax.full[,2]<-as.numeric(tax.full$`2016`)
tax.full[,3]<-as.numeric(tax.full$`2017`)
tax.full[,4]<-as.numeric(tax.full$`2018`)

# compute percentage change

Municipalities<-tax.full$Municipalities
pct_2017<-(tax.full$`2017`-tax.full$`2016`)/tax.full$`2016`*100
pct_2018<-tax.pct[,3]<-(tax.full$`2018`-tax.full$`2017`)/tax.full$`2017`*100
tax.pct<-data.frame(Municipalities,pct_2017,pct_2018)
```

Note: 
- tax.full contains tax from 2016-2018 for 20 municipalities
- tax.pct contains percentage change of tax from 2017-2018.

### 2020-02-25 src/tax2016-2018.Rmd
Update tax2016-2018.Rmd
add output file

@@
 tax.pct<-data.frame(Municipalities,pct_2017,pct_2018)
+
+write.csv(tax.full, here("data","tax_final.csv"))
+write.csv(tax.pct, here("data","tax_pct.csv"))
 ```

### 2020-02-26 src/draft.Rmd
create draft

---
title: "draft"
author: "Xuechun Lu"
date: "25/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(tidyverse)
```

1. read the data and aggregate them
```{r}
tax_pct<-read_csv(here("data", "tax_pct.csv"))
assessment_pct<-read_csv(here("data", "assessment_pct.csv"))
# omit 2016
assessment_pct<-na.omit(assessment_pct)

```

### 2020-02-26 src/draft.Rmd
added draft template
aggregated assessment and tax by year and municipality

@@
 assessment_pct<-na.omit(assessment_pct)
+assessment_pct<-assessment_pct[,-c(1)]
+names(assessment_pct)[2]<-paste("Municipalities")
+tax_pct<-tax_pct[,-c(1)]
 
 ```
+Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge Rural and Maple Ridge with the same tax.
+
+```{r}
+assessment2017<-assessment_pct  %>%  filter(Year=="2017")
+tax2017<-tax_pct[,1:2]
+assessment2017<-assessment2017 %>% left_join(tax2017, by = c("Municipalities"))
+names(assessment2017)[9]<-paste("tax")
+
+assessment2018<-assessment_pct  %>%  filter(Year=="2018")
+tax2018<-tax_pct[,c(1,3)]
+assessment2018<-assessment2018 %>% left_join(tax2018, by = c("Municipalities"))
+names(assessment2018)[9]<-paste("tax")
+
+assessment2019<-assessment_pct  %>%  filter(Year=="2019")
+
+data_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
+write.csv(data_final, here("data","data_final.csv"))
+```
+
+2. plots

### 2020-02-26 src/draft.Rmd
Update draft.Rmd
aggregated assessment and tax
Update draft.Rmd
imputed tax 2019 by avg of 2016 and 2018 tax pct

--- 2020-02-26 src/draft.Rmd
@@
 library(tidyverse)
+library(dplyr)
+library(here)
+library(readxl)
 ```
@@
 assessment_pct<-na.omit(assessment_pct)
+assessment_pct<-assessment_pct[,-c(1)]
+names(assessment_pct)[2]<-paste("Municipalities")
+tax_pct<-tax_pct[,-c(1)]
 

### 2020-02-26 src/draft.Rmd
Add files via upload

--- 2020-02-26 src/draft.Rmd
@@
 
+2019 tax pct is missing, we imputed 2019 tax by the average of previous tax pct
+```{r}
+tax_pct[,4]<-(tax_pct$pct_2017+tax_pct$pct_2018)/2
+tax_pct
+names(tax_pct)[4]<-paste("pct_2019")
+tax_pct
+```
+
 ```{r}
@@
 assessment2019<-assessment_pct  %>%  filter(Year=="2019")
+tax2019<-tax_pct[,c(1,4)]
+assessment2019<-assessment2019 %>% left_join(tax2019, by = c("Municipalities"))
+names(assessment2019)[9]<-paste("tax")
 
@@
 write.csv(data_final, here("data","data_final.csv"))
+
+data_final
+
 ```
@@
 2. plots
+```{r}
+
+```

### 2020-02-27 src/draft.Rmd
Update draft.Rmd
EDA plots updated

@@
 
-2. plots
+2. pct_plots
+plot pct change of mill rates
 ```{r}
+dat_1<-data_final %>% subset(Municipalities == "Vancouver") %>% subset(TaxClassCode == "01")
+dat_1
+dat_1$rate_pct
+plot(dat_1$rate_pct)
+
+municipality.list = c(
+  "Burnaby", 
+  "Coquitlam", 
+  "Delta", 
+  "Langley - City", 
+  "Langley - Township",
+  "Maple Ridge",
+  "Maple Ridge Rural", 
+  "North Vancouver - City",
+  "North Vancouver - Dist",
+  "Pitt Meadows", 
+  "Port Coquitlam", 
+  "Port Moody", 
+  "Richmond", 
+  "Surrey", 
+  "Vancouver", 
+  "White Rock", 
+  "West Vancouver", 
+  "Bowen Island", 
+  "Anmore", 
+  "Belcarra",
+  "Lions Bay")
+tax.list<-c("01","05","06")
+tax.list[1]
+# overtall trend for tax class 1
+for (i in 1:3){
+  plot<-data_final %>% 
+  filter(Municipalities %in% municipality.list) %>%
+  filter(TaxClassCode == tax.list[i]) %>%
+  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
+    geom_line();
+  print(plot)
+}
+
+# for each municipality
+
+for (i in 1:21){
+  plot<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  filter(TaxClassCode %in% tax.list) %>% 
+  ggplot(aes(x=Year, y=rate_pct, group=TaxClassCode, color=TaxClassCode)) +
+    geom_line()
+  print(plot)
+}
 
 ```
+
+plot pct change of assessment
+```{r}
+# total trend of assessment 
+for (i in 1:21){
+  plot<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  filter(TaxClassCode %in% tax.list) %>% 
+  ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
+    geom_line()
+  print(plot)
+}
+```
+
+plot pct change of tax(budget)
+```{r}
+# total trend for budget 
+for (i in 1:21){
+  plot<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  filter(TaxClassCode %in% tax.list) %>% 
+  ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode)) +
+    geom_line()
+  print(plot)
+}
+```
+
+```{r}
+# plot rate and assessment for each municipality.
+
+for (i in 1:21){
+  test1<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  filter(TaxClassCode %in% tax.list) %>% 
+  ggplot(aes(x=Year, y=rate_pct, group=TaxClassCode, color=TaxClassCode)) +
+   geom_line()
+
+  test2<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  filter(TaxClassCode %in% tax.list) %>% 
+  ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
+   geom_line()
+
+  test3<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  filter(TaxClassCode %in% tax.list) %>% 
+  ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode))+
+   geom_line()
+
+  multiplot(test1,test2,test3)
+
+}
+```

### 2020-02-27 src/draft.Rmd
add tax 2019

@@
 ---
-title: "draft"
-author: "Xuechun Lu"
+title: "EDA"
+author: "Xuechun Lu, Yuting Wen"
 date: "25/02/2020"
@@
 
+```
+
+2019 tax is missing, we imputed 2019 tax using previous values
+```{r}
+tax_final <- read.csv("~/Desktop/tax_final.csv")
+head(tax_final)
+
+tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))
+head(tax_modified)
 ```

### 2020-02-27 src/draft.Rmd
add mill rate/ assessed val v.s year plots

@@
 library(readxl)
+```
+```{r}
+# all municipalities
+municipality.list = c(
+  "Burnaby", 
+  "Coquitlam", 
+  "Delta", 
+  "Langley - City", 
+  "Langley - Township",
+  "Maple Ridge",
+  "Maple Ridge Rural", 
+  "North Vancouver - City",
+  "North Vancouver - Dist",
+  "Pitt Meadows", 
+  "Port Coquitlam", 
+  "Port Moody", 
+  "Richmond", 
+  "Surrey", 
+  "Vancouver", 
+  "White Rock", 
+  "West Vancouver", 
+  "Bowen Island", 
+  "Anmore", 
+  "Belcarra",
+  "Lions Bay")
 ```
@@
 ```{r}
-tax_final <- read.csv("~/Desktop/tax_final.csv")
+tax_final <- read.csv(here("data","tax_final.csv"))
 head(tax_final)
@@
 
-2. pct_plots
+2. val_plots
+```{r}
+# mill rate vesus year for each municipality and tax class
+assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
+dat_rate <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, rate)
+head(dat_rate)
+
+dat_rate_1 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
+head(dat_rate_1)
+rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line()
+rate_class1
+
+dat_rate_5 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
+head(dat_rate_5)
+rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line()
+rate_class5
+
+dat_rate_6 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6") 
+head(dat_rate_6)
+rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line()
+rate_class6
+
+# plot assessed value vesus year for each municipality and tax class
+dat_assess <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, assessTotal)
+head(dat_assess)
+
+dat_assess_1 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
+head(dat_assess_1)
+assess_class1 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 1")
+assess_class1
+
+dat_assess_5 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
+head(dat_assess_5)
+assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 5")
+assess_class5
+
+dat_assess_6 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6") 
+head(dat_assess_6)
+assess_class6 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6")
+assess_class6
+```
+
+3. pct_plots
 plot pct change of mill rates
@@
 
-municipality.list = c(
-  "Burnaby", 
-  "Coquitlam", 
-  "Delta", 
-  "Langley - City", 
-  "Langley - Township",
-  "Maple Ridge",
-  "Maple Ridge Rural", 
-  "North Vancouver - City",
-  "North Vancouver - Dist",
-  "Pitt Meadows", 
-  "Port Coquitlam", 
-  "Port Moody", 
-  "Richmond", 
-  "Surrey", 
-  "Vancouver", 
-  "White Rock", 
-  "West Vancouver", 
-  "Bowen Island", 
-  "Anmore", 
-  "Belcarra",
-  "Lions Bay")
 tax.list<-c("01","05","06")

### 2020-02-27 src/draft.Rmd
add budget v.s year plot

@@
 library(tidyverse)
+library(ggplot2)
 library(dplyr)
@@
 head(dat_rate_1)
-rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line()
+rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 1")
 rate_class1
@@
 head(dat_rate_5)
-rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line()
+rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 5")
 rate_class5
@@
 head(dat_rate_6)
-rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line()
+rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6")
 rate_class6
@@
 assess_class6
+
+# plot budgets vesus year
+dat_tax_long <- gather(tax_modified, Year, TaxAmount, X2016:X2019, factor_key=TRUE)
+dat_tax_long
+budget_plot <- dat_tax_long %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_line() +ggtitle("Tax v.s Year")
+budget_plot
 ```

### 2020-02-27 src/draft.Rmd
add dummy boxplot

@@
 ## R Markdown
+libraries and global variables
 ```{r}
@@
 ```{r}
+
 # all municipalities
@@
 ```
+
+4. box plots of mill rates, grouped by region
+```{r}
+rate_boxes <- ggplot(dat_rate, aes(x=AddressAssessorMunicipalityDesc,y=rate,fill=TaxClassCode)) + geom_boxplot() + facet_wrap(~TaxClassCode)
+rate_boxes
+```

### 2020-02-27 src/draft.Rmd
added multiplot and aes

@@
 
+multiplot(rate_class1, rate_class5, rate_class6)
+
 # plot assessed value vesus year for each municipality and tax class
@@
 assess_class6
+
+multiplot(assess_class1, assess_class5, assess_class6)
 
@@
 rate_boxes <- ggplot(dat_rate, aes(x=AddressAssessorMunicipalityDesc,y=rate,fill=TaxClassCode)) + geom_boxplot() + facet_wrap(~TaxClassCode)
-rate_boxes
+rate_boxes + theme(axis.text.x = element_text(angle = 90, hjust = 1))
 ```

### 2020-02-27 src/draft.Rmd
added initial correlation procession

@@
 
+# tax_final is the budget
 tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))
+
+# tax_modified contains all the budget from 2016 to 2019
 head(tax_modified)
@@
 ```{r}
-rate_boxes <- ggplot(dat_rate, aes(x=AddressAssessorMunicipalityDesc,y=rate,fill=TaxClassCode)) + geom_boxplot() + facet_wrap(~TaxClassCode)
-rate_boxes + theme(axis.text.x = element_text(angle = 90, hjust = 1))
+rate_boxes <- ggplot(dat_rate, aes(x=AddressAssessorMunicipalityDesc,y=rate,fill=TaxClassCode)) + geom_boxplot() + facet_wrap(~TaxClassCode) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
+rate_boxes 
 ```
+5. 
+```{r}
+#tax_final
+#assessment_aggregate
+
+#colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
+#head(tax_modified)
+#dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
+
+#tax.all <- full_join(assessment_aggregate, tax_modified, by = 'Municipalities') 
+#tax.all
+#(tax.all <- tax.all[,-5])
+# correlation by Municipalities?
+#mill.budget.cor <- tax.all  %>% summarize(cor_2017 = cor(pct_2017, X2017), cor_2018 = cor(pct_2018, X2018), cor_2019 = cor(pct_2019, X2019))
+#mill.budget.cor
+
+years <- c(2017, 2018, 2019)
+city.cor <- data.frame(Municipality = municipality.list)
+corr.list <- vector(length = 21)
+corr.list
+for(city in municipality.list){
+  corr <- assessment_aggregate %>% filter(AddressAssessorMunicipalityDesc == city) %>% summarize(cor = cor(assessTotal, rate))
+  #print(corr)
+  corr.list <- cbind(corr.list, corr)
+}
+corr.list
+city.cor <- cbind(city.cor,corr.list)
+#city.cor
+
+```

### 2020-02-27 src/draft.Rmd
add acf, plus format modify

@@
 ```
-5. 
+
+5. Test correlations between past mill rates and assessed values and government budget.
 ```{r}
@@
 #city.cor
+```
 
+6. Compute autocorrelations of past mill rates to test the independence of mill rates. 
+```{r}
+# all white noise
+for (i in 1:21) {
+  print(municipality.list[i])
+  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "1") %>% select(rate)
+  if (municipality.list[i] != "Maple Ridge Rural") {
+    as.ts(dat_tmp)
+    acf(dat_tmp)
+  }
+}
+for (i in 1:21) {
+  print(municipality.list[i])
+  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "5") %>% select(rate)
+  if (municipality.list[i] != "Maple Ridge Rural" & municipality.list[i] != "White Rock" & municipality.list[i] != "Anmore" & municipality.list[i] != "Belcarra" & municipality.list[i] != "Lions Bay") {
+    as.ts(dat_tmp)
+    acf(dat_tmp)
+  }
+}
+for (i in 1:21) {
+  print(municipality.list[i])
+  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "6") %>% select(rate)
+  if (municipality.list[i] != "Maple Ridge Rural") {
+    as.ts(dat_tmp)
+    acf(dat_tmp)
+  }
+}
 ```

### 2020-02-27 src/draft.Rmd
clean up EDA

@@
 title: "EDA"
-author: "Xuechun Lu, Yuting Wen"
+author: "Xuechun Lu, Yuting Wen, Peter Han"
 date: "25/02/2020"
@@
 ## R Markdown
-libraries and global variables
+libraries
 ```{r}
@@
 library(readxl)
+
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+  numPlots = length(plots)
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+ if (numPlots==1) {
+    print(plots[[1]])
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
+
 ```
+
+1. data processing
 ```{r}
-
 # all municipalities
@@
   "Lions Bay")
-```
 
-1. read the data and aggregate them
-```{r}
+# budget and assessment data import and process
 tax_pct<-read_csv(here("data", "tax_pct.csv"))
@@
 tax_pct<-tax_pct[,-c(1)]
-
+head(tax_pct)
 ```
@@
 names(tax_pct)[4]<-paste("pct_2019")
-tax_pct
+head(tax_pct)
 ```
 
+aggregate tax and assessment
 ```{r}
@@
 
-data_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
-write.csv(data_final, here("data","data_final.csv"))
+pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
+write.csv(pct_final, here("data","data_final.csv"))
 
-data_final
+# all pct data aggregated
+head(pct_final)
 
@@
 
-# tax_modified contains all the budget from 2016 to 2019
+# tax_modified contains all the budget values from 2016 to 2019
 head(tax_modified)
+```
+
+global variables
+```{r}
+# all pct aggregated
+pct_final
+# all budget values
+tax_modified
 ```
@@
 head(dat_rate_1)
-rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 1")
-rate_class1
+rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 1") + theme(legend.position = "none")
+#rate_class1
 
@@
 head(dat_rate_5)
-rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 5")
-rate_class5
+rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 5") + theme(legend.position = "none")
+#rate_class5
 
@@
 rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6")
-rate_class6
+#rate_class6
 
@@
 head(dat_assess_1)
-assess_class1 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 1")
-assess_class1
+assess_class1 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 1") + theme(legend.position = "none")
+#assess_class1
 
@@
 head(dat_assess_5)
-assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 5")
-assess_class5
+assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 5") + theme(legend.position = "none")
+#assess_class5
 
@@
 assess_class6 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6")
-assess_class6
+#assess_class6
 
@@
 3. pct_plots
-plot pct change of mill rates
+pct change of mill rates for each tax class
 ```{r}
-dat_1<-data_final %>% subset(Municipalities == "Vancouver") %>% subset(TaxClassCode == "01")
-dat_1
-dat_1$rate_pct
-plot(dat_1$rate_pct)
-
 tax.list<-c("01","05","06")
@@
 for (i in 1:3){
-  plot<-data_final %>% 
+  plot<-pct_final %>% 
   filter(Municipalities %in% municipality.list) %>%
@@
   ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_line();
+    geom_line() + ggtitle(tax.list[i])
   print(plot)
 }
-
+```
+pct change of mill rate for each municipality
+```{r}
 # for each municipality
@@
 for (i in 1:21){
-  plot<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
   filter(TaxClassCode %in% tax.list) %>% 
   ggplot(aes(x=Year, y=rate_pct, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line()
+    geom_line() + ggtitle(municipality.list[i])
   print(plot)
 }
-
 ```
@@
 for (i in 1:21){
-  plot<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
   filter(TaxClassCode %in% tax.list) %>% 
   ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line()
+    geom_line() + ggtitle(municipality.list[i])
   print(plot)
@@
 for (i in 1:21){
-  plot<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
   filter(TaxClassCode %in% tax.list) %>% 
   ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line()
+    geom_line() + ggtitle(municipality.list[i])
   print(plot)
@@
 for (i in 1:21){
-  test1<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  test1<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
   filter(TaxClassCode %in% tax.list) %>% 
@@
 
-  test2<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  test2<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
   filter(TaxClassCode %in% tax.list) %>% 
@@
 
-  test3<-data_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  test3<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
   filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode))+
+  ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode))+
    geom_line()

### 2020-02-27 src/draft.Rmd
hide unnecessary display

@@
 tax_pct<-tax_pct[,-c(1)]
-head(tax_pct)
+#head(tax_pct)
 ```
@@
 tax_pct[,4]<-(tax_pct$pct_2017+tax_pct$pct_2018)/2
-tax_pct
+#tax_pct
 names(tax_pct)[4]<-paste("pct_2019")
-head(tax_pct)
+#head(tax_pct)
 ```
@@
 # all pct data aggregated
-head(pct_final)
+#head(pct_final)
 
@@
 tax_final <- read.csv(here("data","tax_final.csv"))
-head(tax_final)
+#head(tax_final)
 
@@
 # tax_modified contains all the budget values from 2016 to 2019
-head(tax_modified)
+#head(tax_modified)
 ```
@@
 dat_rate <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, rate)
-head(dat_rate)
+#head(dat_rate)
 
 dat_rate_1 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
-head(dat_rate_1)
+#head(dat_rate_1)
 rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 1") + theme(legend.position = "none")
@@
 dat_rate_5 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
-head(dat_rate_5)
+#head(dat_rate_5)
 rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 5") + theme(legend.position = "none")
@@
 dat_rate_6 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6") 
-head(dat_rate_6)
+#head(dat_rate_6)
 rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6")
@@
 dat_assess <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, assessTotal)
-head(dat_assess)
+#head(dat_assess)
 
 dat_assess_1 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
-head(dat_assess_1)
+#head(dat_assess_1)
 assess_class1 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 1") + theme(legend.position = "none")
@@
 dat_assess_5 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
-head(dat_assess_5)
+#head(dat_assess_5)
 assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 5") + theme(legend.position = "none")
@@
 dat_assess_6 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6") 
-head(dat_assess_6)
+#head(dat_assess_6)
 assess_class6 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6")
@@
 tax.list<-c("01","05","06")
-tax.list[1]
+#tax.list[1]
 # overtall trend for tax class 1
@@
 # for each municipality
-
 for (i in 1:21){
@@
 ```{r}
-# plot rate and assessment for each municipality.
+# plot rate, assessment, and budget for each municipality.
 

### 2020-02-27 src/draft.Rmd
add total pct comparison graph

@@
 #head(dat_rate_6)
-rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6")
+rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6") + theme(legend.position = "none")
 #rate_class6
 
-multiplot(rate_class1, rate_class5, rate_class6)
+#multiplot(rate_class1, rate_class5, rate_class6)
 
@@
 #head(dat_assess_6)
-assess_class6 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6")
+assess_class6 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6") + theme(legend.position = "none")
 #assess_class6
 
-multiplot(assess_class1, assess_class5, assess_class6)
+#multiplot(assess_class1, assess_class5, assess_class6)
 
@@
 dat_tax_long <- gather(tax_modified, Year, TaxAmount, X2016:X2019, factor_key=TRUE)
-dat_tax_long
+#dat_tax_long
 budget_plot <- dat_tax_long %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_line() +ggtitle("Tax v.s Year")
-budget_plot
+#budget_plot
+
+multiplot(rate_class1,assess_class1,budget_plot)
+multiplot(rate_class5,assess_class5,budget_plot)
+multiplot(rate_class6,assess_class6,budget_plot)
 ```
@@
 tax.list<-c("01","05","06")
-#tax.list[1]
-# overtall trend for tax class 1
-for (i in 1:3){
-  plot<-pct_final %>% 
+pct_rate1 <- pct_final %>% 
   filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == tax.list[i]) %>%
+  filter(TaxClassCode == "01") %>%
   ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle(tax.list[i])
-  print(plot)
-}
+    geom_line() + ggtitle("taxClass1 pct rate v.s year") + theme(legend.position = "none")
+
+pct_rate5 <- pct_final %>% 
+  filter(Municipalities %in% municipality.list) %>%
+  filter(TaxClassCode == "05") %>%
+  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
+    geom_line() + ggtitle("taxClass5 pct rate v.s year") + theme(legend.position = "none")
+
+pct_rate6 <- pct_final %>% 
+  filter(Municipalities %in% municipality.list) %>%
+  filter(TaxClassCode == "06") %>%
+  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
+    geom_line() + ggtitle("taxClass6 pct rate v.s year") + theme(legend.position = "none")
+
+pct_assess1 <- pct_final %>% 
+  filter(Municipalities %in% municipality.list) %>%
+  filter(TaxClassCode == "01") %>%
+  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
+    geom_line() + ggtitle("taxClass1 pct assessment v.s year") + theme(legend.position = "none")
+
+pct_assess5 <- pct_final %>% 
+  filter(Municipalities %in% municipality.list) %>%
+  filter(TaxClassCode == "05") %>%
+  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
+    geom_line() + ggtitle("taxClass5 pct assessment v.s year") + theme(legend.position = "none")
+
+pct_assess6 <- pct_final %>% 
+  filter(Municipalities %in% municipality.list) %>%
+  filter(TaxClassCode == "06") %>%
+  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
+    geom_line() + ggtitle("taxClass6 pct assessment v.s year") + theme(legend.position = "none")
+
+tax_156 <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  filter(TaxClassCode %in% tax.list) %>% 
+  ggplot(aes(x=Year, y=tax)) +
+   geom_line() + theme(legend.position = "none") + ggtitle("tax amount")
+
+multiplot(pct_rate1,pct_assess1,tax_156)
+multiplot(pct_rate5,pct_assess5,tax_156)
+multiplot(pct_rate6,pct_assess6,tax_156)
 ```
@@
   filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode))+
-   geom_line()
+  ggplot(aes(x=Year, y=tax)) +
+   geom_line() + theme(legend.position = "none")
 
@@
 ```
+
+model fitting
+```{r}
+linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
+summary(linear1)
+
+linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
+summary(linear2)
+```

### 2020-02-27 src/draft.Rmd
added plots

@@
 date: "25/02/2020"
-output: html_document
+output:
+  github_document
 ---
@@
 libraries
-```{r}
+```{r, message=F, warning=F}
 library(tidyverse)
@@
 library(readxl)
+```
 
+```{r}
 multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
@@
 1. data processing
-```{r}
+```{r, warning=F, message=F}
 # all municipalities
@@
 ```{r}
-tax.list<-c("01","05","06")
 pct_rate1 <- pct_final %>% 
@@
 
-tax_156 <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=tax)) +
-   geom_line() + theme(legend.position = "none") + ggtitle("tax amount")
-
-multiplot(pct_rate1,pct_assess1,tax_156)
-multiplot(pct_rate5,pct_assess5,tax_156)
-multiplot(pct_rate6,pct_assess6,tax_156)
+multiplot(pct_rate1,pct_assess1)
+multiplot(pct_rate5,pct_assess5)
+multiplot(pct_rate6,pct_assess6)
 ```
@@
 # for each municipality
+tax.list <- c("01","05","06")
 for (i in 1:21){

### 2020-02-28 src/draft.Rmd
corr added

@@
 plot pct change of tax(budget)
-```{r}
+```{r, warning=F}
 # total trend for budget 
@@
 ```{r}
-#tax_final
-#assessment_aggregate
+colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
+head(tax_modified)
+dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
 
-#colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
-#head(tax_modified)
-#dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
+tax.all <- full_join(assessment_aggregate, tax_modified, by = 'Municipalities')
 
-#tax.all <- full_join(assessment_aggregate, tax_modified, by = 'Municipalities') 
-#tax.all
-#(tax.all <- tax.all[,-5])
-# correlation by Municipalities?
-#mill.budget.cor <- tax.all  %>% summarize(cor_2017 = cor(pct_2017, X2017), cor_2018 = cor(pct_2018, X2018), cor_2019 = cor(pct_2019, X2019))
-#mill.budget.cor
+tax_modified1 <- gather(tax_modified, Year, tax, '2016':'2019', factor_key = TRUE)%>%
+  rename(AddressAssessorMunicipalityDesc = Municipalities)
+
+assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
+
+correlation <- assessment_aggregate %>%
+  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
+  summarise(cor(rate, tax))
+
+correlation2 <- assessment_aggregate %>%
+  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
+  summarise(cor(assessTotal, rate))
 

### 2020-02-28 src/draft.Rmd
model added

@@
 ```
+
+3. Fit linear model
+```{r}
+
+linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
+summary(linear1)
+
+linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
+summary(linear2)
+library(glmnet)
+
+x_matrix<-pct_final[,-c(8)]
+x_matrix[,1]<-factor(x_matrix$Year)
+x_matrix[,2]<-factor(x_matrix$Municipalities)
+x_matrix[,3]<-factor(x_matrix$TaxClassCode)
+x_matrix<-x_matrix[,-c(5,6,7)]
+x_matrix<-as.matrix(x_matrix)
+
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x_matrix, pct_final$rate_pct, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
+
+summary(ridge_reg)
+
+```

### 2020-02-28 src/draft.Rmd
data added

@@
 names(tax_pct)[4]<-paste("pct_2019")
+tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567))
 #head(tax_pct)

### 2020-02-28 src/draft.Rmd
1. corrected a mistake in imputation of "Maple Ridge Rural" in tax_pct. 2. Refit the model without "Langley - City" and "Langley - Township"

--- 2020-02-28 src/draft.Rmd
@@
 names(tax_pct)[4]<-paste("pct_2019")
+tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567,5.945346))
+
 #head(tax_pct)
@@
 
-tax.all <- full_join(assessment_aggregate, tax_modified, by = 'Municipalities')
+#tax.all <- full_join(assessment_aggregate, tax_modified, by = 'Municipalities')
 
@@
 ```
+r^2 is too small.
+we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400).
+We ignored the two municipalities and refit the linear model and observed that r^@ improved to 0.4882.
 
-3. Fit linear model
+```{r}
+test<-pct_final %>% filter(Municipalities!="Langley - City") %>% 
+  filter(Municipalities!="Langley - Township")
+linear_test<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
+summary(linear_test)
+```
+
+Lasso
 ```{r}
 
-linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
-summary(linear1)
-
-linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
-summary(linear2)
 library(glmnet)

### 2020-02-28 src/draft.Rmd
added model fitting

@@
 we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400).
-We ignored the two municipalities and refit the linear model and observed that r^@ improved to 0.4882.
+We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882.
 
@@
 
+Ridge Regression!
+reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r
+```{r}
+library(glmnet)
+library(dummies)
+dummy_year<-dummy(test$Year)
+dummy_municipal<-dummy(test$Municipalities)
+dummy_taxclass<-dummy(test$TaxClassCode)
+# build x matrix
+x<-cbind(dummy_year,dummy_municipal)
+x<-cbind(x,dummy_taxclass)
+x<-cbind(x,as.numeric(test$assessTotal_pct))
+x<-cbind(x,as.numeric(test$tax))
+y<-test$rate_pct
+lambdas <- 10^seq(2, -3, by = -.1)
+dim(x)
+colnames(x)[26]<-paste("assessment")
+colnames(x)[27]<-paste("tax")
+colnames(x)
+
+cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+optimal_lambda
+predictions_train <- predict(cv_ridge, s = optimal_lambda, newx = x)
+
+# Compute R^2 from true and predicted values
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
+
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquare = R_square
+)
+
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+eval_results(y, predictions_train, train)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+eval_results(y_test, predictions_test, test)
+
+```
+Ridge regression doesn't seem to work well. R square = 0.1967394.
+
 Lasso 
 ```{r}
+lambdas <- 10^seq(2, -3, by = -.1)
 
-library(glmnet)
+# Setting alpha = 1 implements lasso regression
+lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 5)
 
-x_matrix<-pct_final[,-c(8)]
-x_matrix[,1]<-factor(x_matrix$Year)
-x_matrix[,2]<-factor(x_matrix$Municipalities)
-x_matrix[,3]<-factor(x_matrix$TaxClassCode)
-x_matrix<-x_matrix[,-c(5,6,7)]
-x_matrix<-as.matrix(x_matrix)
+# Best 
+lambda_best <- lasso_reg$lambda.min;lambda_best
 
-lambdas <- 10^seq(2, -3, by = -.1)
-ridge_reg = glmnet(x_matrix, pct_final$rate_pct, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
+lasso_model <- glmnet(x, y, alpha = 1, lambda = lambda_best, standardize = TRUE)
 
-summary(ridge_reg)
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+eval_results(y, predictions_train, train)
+# R squared = 0.2000649	
+```
+Lasso doesn't seem to work well, but it is still better than ridge. R squared = 0.2000649.
 
+Elastic net
+```{r}
+library(caret)
+tibble::as_tibble(test)
+cv_5 = trainControl(method = "cv", number = 5)
+ridge = train(
+  rate_pct ~ Year + Municipalities + TaxClassCode + assessTotal_pct + tax, data = test,
+  method = "glmnet",
+  trControl = cv_5
+)
+ridge
 ```
+apha = 1 which implies lasso! refer to lasso.

### 2020-02-28 src/draft.Rmd
reference added

@@
 
-Elastic net
+Elastic net: 
+reference: https://daviddalpiaz.github.io/r4sl/elastic-net.html
 ```{r}

### 2020-02-28 src/draft.Rmd
eda improved

--- 2020-02-28 src/draft.Rmd
@@
 title: "EDA"
-author: "Xuechun Lu, Yuting Wen, Peter Han"
+author: "Xuechun Lu, Yuting Wen, Peter Han, Yuetong Liu"
 date: "25/02/2020"
@@
 names(tax_pct)[4]<-paste("pct_2019")
-tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567,5.945346))
-
+tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567))
 #head(tax_pct)
@@
 #head(dat_rate_1)
-rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 1") + theme(legend.position = "none")
+rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 1")
 #rate_class1
@@
 #head(dat_rate_5)
-rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 5") + theme(legend.position = "none")
+rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 5")
 #rate_class5
@@
 #head(dat_rate_6)
-rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6") + theme(legend.position = "none")
+rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6")
 #rate_class6
-
-#multiplot(rate_class1, rate_class5, rate_class6)
 
@@
 #head(dat_assess_1)
-assess_class1 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 1") + theme(legend.position = "none")
+assess_class1 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 1")
 #assess_class1
@@
 #head(dat_assess_5)
-assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 5") + theme(legend.position = "none")
+assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 5")
 #assess_class5
@@
 #head(dat_assess_6)
-assess_class6 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6") + theme(legend.position = "none")
+assess_class6 <- dat_assess_6 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6")
 #assess_class6
-
-#multiplot(assess_class1, assess_class5, assess_class6)
 
@@
 budget_plot <- dat_tax_long %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_line() +ggtitle("Tax v.s Year")
-#budget_plot
+budget_plot
 
-multiplot(rate_class1,assess_class1,budget_plot)
-multiplot(rate_class5,assess_class5,budget_plot)
-multiplot(rate_class6,assess_class6,budget_plot)
+```
+
+better EDA
+```{r}
+# Vancouver/West Vancouver are special cases for assessment, we omit them for taxClass 1 and 6
+rate_class1 <- dat_rate_1 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 1 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+assess_class1 <- dat_assess_1 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 1 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+multiplot(rate_class1 + theme(legend.position = "none"), assess_class1 + theme(legend.position = "none"))
+
+rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 5") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 5") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+multiplot(rate_class5 + theme(legend.position = "none"),assess_class5 + theme(legend.position = "none"))
+
+rate_class6 <- dat_rate_6 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 6 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+assess_class6 <- dat_assess_6 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 6 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+multiplot(rate_class6 + theme(legend.position = "none"),assess_class6 + theme(legend.position = "none"))
+
+budget_plot <- dat_tax_long %>% filter(Municipalities != "Vancouver" & Municipalities != "West Vancouver") %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_line() +ggtitle("Tax v.s Year")
+budget_plot
 ```
@@
   ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass1 pct rate v.s year") + theme(legend.position = "none")
+    geom_point() + ggtitle("taxClass1 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
@@
   ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass5 pct rate v.s year") + theme(legend.position = "none")
+    geom_point() + ggtitle("taxClass5 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
@@
   ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass6 pct rate v.s year") + theme(legend.position = "none")
+    geom_point() + ggtitle("taxClass6 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
@@
   ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass1 pct assessment v.s year") + theme(legend.position = "none")
+    geom_point() + ggtitle("taxClass1 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
@@
   ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass5 pct assessment v.s year") + theme(legend.position = "none")
+    geom_point() + ggtitle("taxClass5 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
@@
   ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass6 pct assessment v.s year") + theme(legend.position = "none")
+    geom_point() + ggtitle("taxClass6 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
@@
 
-#tax.all <- full_join(assessment_aggregate, tax_modified, by = 'Municipalities')
-
 tax_modified1 <- gather(tax_modified, Year, tax, '2016':'2019', factor_key = TRUE)%>%
@@
 ```
-r^2 is too small.
-we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400).
-We ignored the two municipalities and refit the linear model and observed that r^@ improved to 0.4882.
 
-```{r}
-test<-pct_final %>% filter(Municipalities!="Langley - City") %>% 
-  filter(Municipalities!="Langley - Township")
-linear_test<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
-summary(linear_test)
-```
-
-Lasso
+3. Fit linear model
 ```{r}
 
+linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
+summary(linear1)
+
+linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
+summary(linear2)
 library(glmnet)

### 2020-02-28 src/draft.Rmd

@@
 names(tax_pct)[4]<-paste("pct_2019")
-tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567))
+tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567,5.945346))
+
 #head(tax_pct)
@@
 ```
+r^2 is too small.
+we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400).
+We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882.
 
-3. Fit linear model
 ```{r}
+test<-pct_final %>% filter(Municipalities!="Langley - City") %>% 
+  filter(Municipalities!="Langley - Township")
+linear_test<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
+summary(linear_test)
+```
 
-linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
-summary(linear1)
+Ridge Regression!
+reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r
+```{r}
+library(glmnet)
+library(dummies)
+dummy_year<-dummy(test$Year)
+dummy_municipal<-dummy(test$Municipalities)
+dummy_taxclass<-dummy(test$TaxClassCode)
+# build x matrix
+x<-cbind(dummy_year,dummy_municipal)
+x<-cbind(x,dummy_taxclass)
+x<-cbind(x,as.numeric(test$assessTotal_pct))
+x<-cbind(x,as.numeric(test$tax))
+y<-test$rate_pct
+lambdas <- 10^seq(2, -3, by = -.1)
+dim(x)
+colnames(x)[26]<-paste("assessment")
+colnames(x)[27]<-paste("tax")
+colnames(x)
 
-linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
-summary(linear2)
-library(glmnet)
+cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas)
+optimal_lambda <- cv_ridge$lambda.min
+optimal_lambda
+predictions_train <- predict(cv_ridge, s = optimal_lambda, newx = x)
 
-x_matrix<-pct_final[,-c(8)]
-x_matrix[,1]<-factor(x_matrix$Year)
-x_matrix[,2]<-factor(x_matrix$Municipalities)
-x_matrix[,3]<-factor(x_matrix$TaxClassCode)
-x_matrix<-x_matrix[,-c(5,6,7)]
-x_matrix<-as.matrix(x_matrix)
+# Compute R^2 from true and predicted values
+eval_results <- function(true, predicted, df) {
+  SSE <- sum((predicted - true)^2)
+  SST <- sum((true - mean(true))^2)
+  R_square <- 1 - SSE / SST
+  RMSE = sqrt(SSE/nrow(df))
 
-lambdas <- 10^seq(2, -3, by = -.1)
-ridge_reg = glmnet(x_matrix, pct_final$rate_pct, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
+  # Model performance metrics
+data.frame(
+  RMSE = RMSE,
+  Rsquare = R_square
+)
 
-summary(ridge_reg)
+}
+
+# Prediction and evaluation on train data
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
+eval_results(y, predictions_train, train)
+
+# Prediction and evaluation on test data
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+eval_results(y_test, predictions_test, test)
 
 ```
+Ridge regression doesn't seem to work well. R square = 0.1967394.
+
+Lasso 
+```{r}
+lambdas <- 10^seq(2, -3, by = -.1)
+
+# Setting alpha = 1 implements lasso regression
+lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 5)
+
+# Best 
+lambda_best <- lasso_reg$lambda.min;lambda_best
+
+lasso_model <- glmnet(x, y, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
+eval_results(y, predictions_train, train)
+# R squared = 0.2000649	
+```
+Lasso doesn't seem to work well, but it is still better than ridge. R squared = 0.2000649.
+
+Elastic net: 
+reference: https://daviddalpiaz.github.io/r4sl/elastic-net.html
+```{r}
+library(caret)
+tibble::as_tibble(test)
+cv_5 = trainControl(method = "cv", number = 5)
+ridge = train(
+  rate_pct ~ Year + Municipalities + TaxClassCode + assessTotal_pct + tax, data = test,
+  method = "glmnet",
+  trControl = cv_5
+)
+ridge
+```
+apha = 1 which implies lasso! refer to lasso.

### 2020-02-28 src/EDA&mode_fitting.Rmd
cleanup and add graphs

--- 2020-02-28 src/draft.Rmd
@@
 #head(tax_pct)
+
+assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
 ```
@@
 
-2. val_plots
+check match between budget and assessTotal*mill_rate
+```{r}
+# get predicted tax income for all tax classes and each municipality
+assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/100
+names(assessment_aggregate)[10] <- paste("predTaxIncome")
+
+dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
+
+dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
+names(dat_pred_income)[3] <- paste("predTaxIncome")
+head(dat_pred_income)
+```
+
+2. plots of mill rate v.s assessments for each tax class
 ```{r}
 # mill rate vesus year for each municipality and tax class
-assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
 dat_rate <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, rate)
@@
 dat_rate_1 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
-#head(dat_rate_1)
-rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 1")
-#rate_class1
-
 dat_rate_5 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
-#head(dat_rate_5)
-rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 5")
-#rate_class5
-
 dat_rate_6 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6")
-#head(dat_rate_6)
-rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6")
-#rate_class6
 
@@
 dat_assess <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, assessTotal)
-#head(dat_assess)
 
 dat_assess_1 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
-#head(dat_assess_1)
-assess_class1 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 1")
-#assess_class1
-
 dat_assess_5 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
-#head(dat_assess_5)
-assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 5")
-#assess_class5
-
 dat_assess_6 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6")
-#head(dat_assess_6)
-assess_class6 <- dat_assess_6 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6")
-#assess_class6
 
@@
 dat_tax_long <- gather(tax_modified, Year, TaxAmount, X2016:X2019, factor_key=TRUE)
-#dat_tax_long
-budget_plot <- dat_tax_long %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_line() +ggtitle("Tax v.s Year")
-budget_plot
 
-```
-
-better EDA
-```{r}
 # Vancouver/West Vancouver are special cases for assessment, we omit them for taxClass 1 and 6
@@
 rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 5") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
 assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 5") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
@@
 
-budget_plot <- dat_tax_long %>% filter(Municipalities != "Vancouver" & Municipalities != "West Vancouver") %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_line() +ggtitle("Tax v.s Year")
+budget_plot <- dat_tax_long %>% filter(Municipalities != "Vancouver" & Municipalities != "West Vancouver") %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_point() + ggtitle("Tax v.s Year") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 budget_plot
@@
 
-3. pct_plots
-pct change of mill rates for each tax class
+3. plots of mill rate pct change v.s assessment pct change for each tax class
 ```{r}
@@
 ```
-pct change of mill rate for each municipality
+
+4. pct change of mill rate and assessment for each municipality
 ```{r}
@@
 for (i in 1:21){
-  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  if (municipality.list[i] != "Lions Bay") {
+    plot_rate_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
     filter(TaxClassCode %in% tax.list) %>% 
@@
       geom_line() + ggtitle(municipality.list[i])
-  print(plot)
+
+    plot_assess_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
+    filter(TaxClassCode %in% tax.list) %>% 
+    ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
+      geom_line() + ggtitle(municipality.list[i])
+
+    plot_tax_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
+  filter(TaxClassCode %in% tax.list) %>% 
+  ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode)) +
+    geom_line() + ggtitle(municipality.list[i])
+
+    multiplot(plot_rate_pct,plot_assess_pct,plot_tax_pct)
+  }
 }
@@
 
-plot pct change of assessment
-```{r}
-# total trend of assessment 
-for (i in 1:21){
-  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line() + ggtitle(municipality.list[i])
-  print(plot)
-}
-```
-
-plot pct change of tax(budget)
-```{r, warning=F}
-# total trend for budget 
-for (i in 1:21){
-  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line() + ggtitle(municipality.list[i])
-  print(plot)
-}
-```
-
-```{r}
-# plot rate, assessment, and budget for each municipality.
-
-for (i in 1:21){
-  test1<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=rate_pct, group=TaxClassCode, color=TaxClassCode)) +
-   geom_line()
-
-  test2<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
-   geom_line()
-
-  test3<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=tax)) +
-   geom_line() + theme(legend.position = "none")
-
-  multiplot(test1,test2,test3)
-
-}
-```
-
-4. box plots of mill rates, grouped by region
+5. box plots of mill rates, grouped by region
 ```{r}
@@
 
-5. Test correlations between past mill rates and assessed values and government budget.
+6. Test correlations between past mill rates and assessed values and government budget.
 ```{r}
+# aggregate tax and mill rate & assessment
 colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
-head(tax_modified)
 dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
@@
 assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
+assessment_aggregate <- assessment_aggregate[,-c(3,11)]
+names(assessment_aggregate)[10] <- paste("tax")
 
-correlation <- assessment_aggregate %>%
+# correlation between mill rate and budget for each municipality
+rate_tax_correlation <- assessment_aggregate %>%
   group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
   summarise(cor(rate, tax))
-
-correlation2 <- assessment_aggregate %>%
+rate_tax_correlation
+# correlation between mill rate and assessment
+rate_assess_correlation2 <- assessment_aggregate %>%
   group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
   summarise(cor(assessTotal, rate))
-
-years <- c(2017, 2018, 2019)
-city.cor <- data.frame(Municipality = municipality.list)
-corr.list <- vector(length = 21)
-corr.list
-for(city in municipality.list){
-  corr <- assessment_aggregate %>% filter(AddressAssessorMunicipalityDesc == city) %>% summarize(cor = cor(assessTotal, rate))
-  #print(corr)
-  corr.list <- cbind(corr.list, corr)
-}
-corr.list
-city.cor <- cbind(city.cor,corr.list)
-#city.cor
+rate_assess_correlation2
 ```

### 2020-02-28 src/draft.Rmd
fixed RIDGE/LASSO/ELASTIC NET regression.

--- 2020-02-28 src/draft.Rmd
@@
 
-cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas)
+lambdas <- 10^seq(2, -3, by = -.1)
+ridge_reg = glmnet(x, y, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
+set.seed(450)
+cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas, nfolds=10)
 optimal_lambda <- cv_ridge$lambda.min
 optimal_lambda
-predictions_train <- predict(cv_ridge, s = optimal_lambda, newx = x)
+predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
 
 # Compute R^2 from true and predicted values
-eval_results <- function(true, predicted, df) {
+eval_results <- function(true, predicted) {
   SSE <- sum((predicted - true)^2)
@@
   R_square <- 1 - SSE / SST
-  RMSE = sqrt(SSE/nrow(df))
-
+  RMSE = sqrt(SSE/nrow(predicted))
 # Model performance metrics
@@
 
-# Prediction and evaluation on train data
 predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
-eval_results(y, predictions_train, train)
-
-# Prediction and evaluation on test data
-predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
-eval_results(y_test, predictions_test, test)
+eval_results(y, predictions_train)
 
 ```
-Ridge regression doesn't seem to work well. R square = 0.1967394.
+RMSE=7.159649
+Rsquare=0.2136659
+Ridge regression doesn't seem to work well.
 
@@
 # Setting alpha = 1 implements lasso regression
-lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 5)
+set.seed(450)
+lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
 
@@
 predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
-eval_results(y, predictions_train, train)
-# R squared = 0.2000649	
+eval_results(y, predictions_train)
 ```
-Lasso doesn't seem to work well, but it is still better than ridge. R squared = 0.2000649.
+RMSE = 7.221299, Rsquare=0.2000649
 
@@
 tibble::as_tibble(test)
-cv_5 = trainControl(method = "cv", number = 5)
+cv_10 = trainControl(method = "cv", number = 10)
 ridge = train(
@@
   method = "glmnet",
-  trControl = cv_5
+  trControl = cv_10
 )

### 2020-02-29 src/draft.Rmd
goodness of fit (with all data) updated. RMSE computed for each model.

@@
 model fitting
+(i) linear model
+Not too much difference between linear 1 and linear 2 (with and without year).
+This probably implies adding year doesn't make a difference.
 ```{r}
@@
 summary(linear2)
+
+# goodness of fit (all data)
+library(broom)
+linear1_fit<-augment(linear1)
+sqrt(sum((linear1_fit$.resid)^2)/nrow(pct_final)) # 5.614843
+linear2_fit<-augment(linear2)
+sqrt(sum((linear2_fit$.resid)^2)/nrow(pct_final)) # 5.614843
+
 ```
-r^2 is too small.
+linear 1: Multiple R-squared:  0.4796,	Adjusted R-squared:  0.155 
+linear 2: Multiple R-squared:  0.4796,	Adjusted R-squared:  0.155
+
 we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400).
-We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882.
+We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882 (Multiple R-squared:  0.4882,	Adjusted R-squared:  0.1605).
+Still not too much difference bewteen the model with and without variable year.
 
@@
 summary(linear_test)
+
+# goodness of fit
+fit_linear_test<-augment(linear_test)
+sqrt(sum((fit_linear_test$.resid)^2)/nrow(test)) # 5.775885
+
+linear_test_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
+summary(linear_test_2)
+
+# goodness of fit
+fit_linear_test_2<-augment(linear_test_2)
+sqrt(sum((fit_linear_test_2$.resid)^2)/nrow(test)) # 5.775885
+
 ```
+Question: How to judge goodness of fit? Multiple R square? Adjusted R square? or mean prediction error?
+**TODO: create training/test sets to test the prediction power!
 
-Ridge Regression!
+(ii) Ridge, Lasso and elastic net.
+
+Ridge Regression
 reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r

### 2020-02-29 src/EDA&mode_fitting.Rmd
added graphs and basic nn model

@@
 library(readxl)
+require(neuralnet)
 ```
@@
 assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
+assessment_final <- read.csv(here("data","assessment_final.csv"))
+assessment_final <- assessment_final[,-1]
+assessment_final <- assessment_final[complete.cases(assessment_final),]
 ```
@@
 # get predicted tax income for all tax classes and each municipality
-assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/100
+assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
 names(assessment_aggregate)[10] <- paste("predTaxIncome")
@@
 dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
-
 dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
 names(dat_pred_income)[3] <- paste("predTaxIncome")
-head(dat_pred_income)
+
+tax_refactored <- tax_modified
+names(tax_refactored)[3] <- paste("2016")
+names(tax_refactored)[4] <- paste("2017")
+names(tax_refactored)[5] <- paste("2018")
+names(tax_refactored)[6] <- paste("2019")
+dat_true_income <- gather(tax_refactored, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
+dat_true_income <- dat_true_income[,-1]
+
+income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
+income_budget_compare <- gather(income_budget_compare, TaxType, Amount, TaxAmount:predTaxIncome, factor_key=TRUE)
+
+compare_full_plot <- income_budget_compare %>% filter(Municipalities!="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=Municipalities)) + geom_line()
+
+compare_plot <- income_budget_compare %>% filter(Municipalities=="Vancouver" | Municipalities=="Surrey" | Municipalities=="Burnaby" | Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome")
+compare_plot
 ```
@@
 apha = 1 which implies lasso! refer to lasso.
+
+## neural network
+```{r}
+nn=neuralnet(TaxClassTaxRate~Year+TaxClassCode+AssessorAreaCode+AddressAssessorMunicipalityCode+AssessedValueAmt+CurrentYearTotal+TaxOwingAmountTotalCalculated,data=assessment_final, hidden=20, act.fct = "logistic", linear.output = FALSE)
+plot(nn)
+```

### 2020-02-29 src/draft.Rmd

@@
 pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
+pct_final$tax<-as.numeric(pct_final$tax)
 write.csv(pct_final, here("data","data_final.csv"))
@@
 linear1_fit<-augment(linear1)
-sqrt(sum((linear1_fit$.resid)^2)/nrow(pct_final)) # 5.614843
+sqrt(sum((linear1_fit$.resid)^2)/nrow(pct_final)) # 6.672786
 linear2_fit<-augment(linear2)
-sqrt(sum((linear2_fit$.resid)^2)/nrow(pct_final)) # 5.614843
+sqrt(sum((linear2_fit$.resid)^2)/nrow(pct_final)) # 7.248359
 
 ```
-linear 1: Multiple R-squared:  0.4796,	Adjusted R-squared:  0.155 
-linear 2: Multiple R-squared:  0.4796,	Adjusted R-squared:  0.155
+linear 1: Residual standard error: 7.296 on 138 degrees of freedom
+Multiple R-squared:  0.265,	Adjusted R-squared:  0.1265 
+PMSE = 6.672786
 
-we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400).
-We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882 (Multiple R-squared:  0.4882,	Adjusted R-squared:  0.1605).
-Still not too much difference bewteen the model with and without variable year.
+linear 2: Residual standard error: 7.869 on 140 degrees of freedom
+Multiple R-squared:  0.1327,	Adjusted R-squared:  -0.01593 
+F-statistic: 0.8929 on 24 and 140 DF,  p-value: 0.6112
+PMSE = 7.248359
+
+we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400). We ignored the two municipalities and refit the linear model
 
@@
 fit_linear_test<-augment(linear_test)
-sqrt(sum((fit_linear_test$.resid)^2)/nrow(test)) # 5.775885
+sqrt(sum((fit_linear_test$.resid)^2)/nrow(test)) # 6.767493
 
@@
 fit_linear_test_2<-augment(linear_test_2)
-sqrt(sum((fit_linear_test_2$.resid)^2)/nrow(test)) # 5.775885
+sqrt(sum((fit_linear_test_2$.resid)^2)/nrow(test)) #  7.50047
 
 ```
+After filtering Langely
+with variable year:
+Residual standard error: 7.429 on 122 degrees of freedom
+Multiple R-squared:  0.2974,	Adjusted R-squared:  0.1592 
+F-statistic: 2.152 on 24 and 122 DF,  p-value: 0.003567
+PMSE: 6.767493
+
+without variable year
+Residual standard error: 8.167 on 124 degrees of freedom
+Multiple R-squared:  0.137,	Adjusted R-squared:  -0.01609 
+F-statistic: 0.8949 on 22 and 124 DF,  p-value: 0.6014
+PMSE: 7.50047
+
+```{r}
+linear3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=pct_final)
+summary(linear3)
+fit_linear3<-augment(linear3)
+sqrt(sum((fit_linear3$.resid)^2)/nrow(test)) #  7.50047
+
+```
+
 Question: How to judge goodness of fit? Multiple R square? Adjusted R square? or mean prediction error?
@@
 apha = 1 which implies lasso! refer to lasso.
+
+Prediction power
+train/test
+```{r}
+set.seed(450)
+train_ind<-sample(165,120)
+train1<-pct_final[train_ind,]
+test1<-pct_final[-train_ind,]
+```
+
+linear model
+```{r}
+# full model with year
+newx<-test1[,-c(5,6,7,8)]
+y<-test1[,c(8)]
+linear_1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
+resid<-predict(linear_1,newdata = newx) - y
+sqrt(sum(resid^2)/nrow(test1))
+# pmse = 6.5688
+
+# full model without year
+linear_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
+resid<-predict(linear_2,newdata = newx) - y
+sqrt(sum(resid^2)/nrow(test1))
+# 6.855429
+
+# reduced variable
+linear_3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=train1)
+resid<-predict(linear_3,newdata = newx) - y
+sqrt(sum(resid^2)/nrow(test1))
+# 6.570848
+```
+Lasso
+```{r}
+# create the whole matrix
+dummy_year<-dummy(pct_final$Year)
+dummy_municipal<-dummy(pct_final$Municipalities)
+dummy_taxclass<-dummy(pct_final$TaxClassCode)
+# build x matrix
+x<-cbind(dummy_year,dummy_municipal)
+x<-cbind(x,dummy_taxclass)
+x<-cbind(x,as.numeric(pct_final$assessTotal_pct))
+x<-cbind(x,as.numeric(pct_final$tax))
+y<-as.matrix(pct_final$rate_pct)
+dim(x) # 165  29
+dim(y)
+# creat x_train matrix and y_train
+x_train<-x[train_ind,]
+y_train<-y[train_ind,]
+# create x_test matrix
+x_test<-x[-train_ind,]
+y_test<-y[-train_ind,]
+
+lambdas <- 10^seq(2, -3, by = -.1)
+# Setting alpha = 1 implements lasso regression
+set.seed(450)
+lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+# Best 
+lambda_best <- lasso_reg$lambda.min;lambda_best
+
+lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+eval_results(y_test, predictions_test)
+# 7.012128	
+```
+
+Ridge
+```{r}
+ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
+set.seed(450)
+cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
+optimal_lambda <- cv_ridge$lambda.min
+optimal_lambda
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+eval_results(y_test, predictions_test)
+# 6.885151	
+```

### 2020-03-01 src/EDA&mode_fitting.Rmd
reformat everything

@@
 summary(linear2)
+
+# goodness of fit (all data)
+library(broom)
+linear1_fit<-augment(linear1)
+sqrt(sum((linear1_fit$.resid)^2)/nrow(pct_final)) # 6.672786
+linear2_fit<-augment(linear2)
+sqrt(sum((linear2_fit$.resid)^2)/nrow(pct_final)) # 7.248359
+
 ```
-r^2 is too small.
-we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400).
-We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882.
+linear 1: Residual standard error: 7.296 on 138 degrees of freedom
+Multiple R-squared:  0.265,	Adjusted R-squared:  0.1265 
+PMSE = 6.672786
+
+linear 2: Residual standard error: 7.869 on 140 degrees of freedom
+Multiple R-squared:  0.1327,	Adjusted R-squared:  -0.01593 
+F-statistic: 0.8929 on 24 and 140 DF,  p-value: 0.6112
+PMSE = 7.248359
+
+we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400). We ignored the two municipalities and refit the linear model
 
@@
 summary(linear_test)
+
+# goodness of fit
+fit_linear_test<-augment(linear_test)
+sqrt(sum((fit_linear_test$.resid)^2)/nrow(test)) # 6.767493
+
+linear_test_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
+summary(linear_test_2)
+
+# goodness of fit
+fit_linear_test_2<-augment(linear_test_2)
+sqrt(sum((fit_linear_test_2$.resid)^2)/nrow(test)) #  7.50047
+
+```
+After filtering Langely
+with variable year:
+Residual standard error: 7.429 on 122 degrees of freedom
+Multiple R-squared:  0.2974,	Adjusted R-squared:  0.1592 
+F-statistic: 2.152 on 24 and 122 DF,  p-value: 0.003567
+PMSE: 6.767493
+
+without variable year
+Residual standard error: 8.167 on 124 degrees of freedom
+Multiple R-squared:  0.137,	Adjusted R-squared:  -0.01609 
+F-statistic: 0.8949 on 22 and 124 DF,  p-value: 0.6014
+PMSE: 7.50047
+
+```{r}
+linear3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=pct_final)
+summary(linear3)
+fit_linear3<-augment(linear3)
+sqrt(sum((fit_linear3$.resid)^2)/nrow(test)) #  7.50047
+
 ```
 
-Ridge Regression!
+Question: How to judge goodness of fit? Multiple R square? Adjusted R square? or mean prediction error?
+**TODO: create training/test sets to test the prediction power!
+
+(ii) Ridge, Lasso and elastic net.
+
+Ridge Regression
 reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r
@@
 apha = 1 which implies lasso! refer to lasso.
+
+Prediction power
+train/test
+```{r}
+set.seed(450)
+train_ind<-sample(165,120)
+train1<-pct_final[train_ind,]
+test1<-pct_final[-train_ind,]
+```
+
+linear model
+```{r}
+# full model with year
+newx<-test1[,-c(5,6,7,8)]
+y<-test1[,c(8)]
+linear_1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
+resid<-predict(linear_1,newdata = newx) - y
+sqrt(sum(resid^2)/nrow(test1))
+# pmse = 6.5688
+
+# full model without year
+linear_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
+resid<-predict(linear_2,newdata = newx) - y
+sqrt(sum(resid^2)/nrow(test1))
+# 6.855429
+
+# reduced variable
+linear_3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=train1)
+resid<-predict(linear_3,newdata = newx) - y
+sqrt(sum(resid^2)/nrow(test1))
+# 6.570848
+```
+Lasso
+```{r}
+# create the whole matrix
+dummy_year<-dummy(pct_final$Year)
+dummy_municipal<-dummy(pct_final$Municipalities)
+dummy_taxclass<-dummy(pct_final$TaxClassCode)
+# build x matrix
+x<-cbind(dummy_year,dummy_municipal)
+x<-cbind(x,dummy_taxclass)
+x<-cbind(x,as.numeric(pct_final$assessTotal_pct))
+x<-cbind(x,as.numeric(pct_final$tax))
+y<-as.matrix(pct_final$rate_pct)
+dim(x) # 165  29
+dim(y)
+# creat x_train matrix and y_train
+x_train<-x[train_ind,]
+y_train<-y[train_ind,]
+# create x_test matrix
+x_test<-x[-train_ind,]
+y_test<-y[-train_ind,]
+
+lambdas <- 10^seq(2, -3, by = -.1)
+# Setting alpha = 1 implements lasso regression
+set.seed(450)
+lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+
+# Best 
+lambda_best <- lasso_reg$lambda.min;lambda_best
+
+lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+
+predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
+eval_results(y_test, predictions_test)
+# 7.012128	
+```
+
+Ridge
+```{r}
+ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
+set.seed(450)
+cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
+optimal_lambda <- cv_ridge$lambda.min
+optimal_lambda
+predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
+eval_results(y_test, predictions_test)
+# 6.885151	
+```
 

### 2020-03-01 src/EDA&mode_fitting.Rmd
Add Correlation Plots

@@
 rate_assess_correlation2
+
+## Correlation for continuous variables
+rate_class <- pct_final %>% ggplot(aes(x=assessTotal_pct,y=rate_pct,group=Municipalities,color=Municipalities)) + geom_point() + ggtitle("mill rate and tax") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+assess_class <- pct_final %>% ggplot(aes(x=tax,y=rate_pct,group=Municipalities,color=Municipalities)) + geom_point() + ggtitle("assessTotal for taxClass 1 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+rate_class
+
+assess_class
+
+## Correlation for categorical variables
+TaxClassCode <- aov(rate_pct ~ TaxClassCode, data = pct_final)
+summary(TaxClassCode)
+
+Year <- aov(rate_pct ~ Year, data = pct_final)
+summary(Year)
+
+Municipalities <- aov(rate_pct ~ Municipalities, data = pct_final)
+summary(Municipalities)
+
+TaxClassCode <- aov(rate_pct ~ TaxClassCode, data = pct_final)
+summary(TaxClassCode)
 ```
 
-6. Compute autocorrelations of past mill rates to test the independence of mill rates. 
+7. Compute autocorrelations of past mill rates to test the independence of mill rates. 
 ```{r}

### 2020-03-01 src/EDA&mode_fitting.Rmd
Cor Plot

@@
 ## Correlation for continuous variables
-rate_class <- pct_final %>% ggplot(aes(x=assessTotal_pct,y=rate_pct,group=Municipalities,color=Municipalities)) + geom_point() + ggtitle("mill rate and tax") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and assess - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
-assess_class <- pct_final %>% ggplot(aes(x=tax,y=rate_pct,group=Municipalities,color=Municipalities)) + geom_point() + ggtitle("assessTotal for taxClass 1 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-rate_class
+tax_class <- assessment_aggregate %>% ggplot(aes(x=predTaxIncome,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
@@
 
+tax_class
+
 ## Correlation for categorical variables
-TaxClassCode <- aov(rate_pct ~ TaxClassCode, data = pct_final)
+TaxClassCode <- aov(rate ~ TaxClassCode, data = assessment_aggregate)
 summary(TaxClassCode)
@@
 
-Municipalities <- aov(rate_pct ~ Municipalities, data = pct_final)
+Municipalities <- aov(rate ~ AddressAssessorMunicipalityDesc, data = assessment_aggregate)
 summary(Municipalities)
-
-TaxClassCode <- aov(rate_pct ~ TaxClassCode, data = pct_final)
-summary(TaxClassCode)
 ```
@@
 summary(linear1)
+
+# linear1<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+predTaxIncome,data=assessment_aggregate)
+# summary(linear1)
+
+linear<-lm(rate~.,data=assessment_aggregate)
+summary(linear)
 

### 2020-03-01 src/EDA&mode_fitting.Rmd
added budget graphs

--- 2020-03-01 src/EDA&mode_fitting.Rmd
@@
 assessment_final <- assessment_final[complete.cases(assessment_final),]
+assessment_final <- assessment_final[,-4]
 ```
@@
 
-compare_plot <- income_budget_compare %>% filter(Municipalities=="Vancouver" | Municipalities=="Surrey" | Municipalities=="Burnaby" | Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome")
-compare_plot
+compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome for Vancouver")
+compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Surrey") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome for Surrey")
+compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome for Burnaby")
+compare_plot4 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome for Richmond")
+multiplot(compare_plot1,compare_plot2,compare_plot3,compare_plot4,cols=2)
 ```
@@
 ```
+
+```{r}
+# in progress
+#library(keras)
+#nn_dat <- to_categorical(assessment_final, 10)
+#train_data <- scale(assessment_final)
+
+```

### 2020-03-01 src/EDA&mode_fitting.Rmd

@@
 rate_assess_correlation2
+
+## Correlation for continuous variables
+assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and assess - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+tax_class <- assessment_aggregate %>% ggplot(aes(x=predTaxIncome,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+assess_class
+
+tax_class
+
+## Correlation for categorical variables
+TaxClassCode <- aov(rate ~ TaxClassCode, data = assessment_aggregate)
+summary(TaxClassCode)
+
+Year <- aov(rate_pct ~ Year, data = pct_final)
+summary(Year)
+
+Municipalities <- aov(rate ~ AddressAssessorMunicipalityDesc, data = assessment_aggregate)
+summary(Municipalities)
 ```
 
-6. Compute autocorrelations of past mill rates to test the independence of mill rates. 
+7. Compute autocorrelations of past mill rates to test the independence of mill rates. 
 ```{r}

### 2020-03-01 src/EDA&mode_fitting.Rmd
models using orginal value are updated now.

@@
 (i) linear model
-Not too much difference between linear 1 and linear 2 (with and without year).
-This probably implies adding year doesn't make a difference.
 ```{r}
@@
 
-# linear1<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+predTaxIncome,data=assessment_aggregate)
-# summary(linear1)
+linear_full<-lm(rate~.,data=assessment_aggregate)
+summary(linear_full)
 
-linear<-lm(rate~.,data=assessment_aggregate)
-summary(linear)
-
-linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
-summary(linear2)
-
-# goodness of fit (all data)
 library(broom)
-linear1_fit<-augment(linear1)
-sqrt(sum((linear1_fit$.resid)^2)/nrow(pct_final)) # 6.672786
-linear2_fit<-augment(linear2)
-sqrt(sum((linear2_fit$.resid)^2)/nrow(pct_final)) # 7.248359
+linear_full_fit<-augment(linear_full)
+sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) #  2.772333
 
 ```
-linear 1: Residual standard error: 7.296 on 138 degrees of freedom
-Multiple R-squared:  0.265,	Adjusted R-squared:  0.1265 
-PMSE = 6.672786
+Multiple R-squared:  0.7803,	Adjusted R-squared:  0.7503 
+MSPE = 2.772333
 
-linear 2: Residual standard error: 7.869 on 140 degrees of freedom
-Multiple R-squared:  0.1327,	Adjusted R-squared:  -0.01593 
-F-statistic: 0.8929 on 24 and 140 DF,  p-value: 0.6112
-PMSE = 7.248359
+Next: variable selection
+with only significant variables
+```{r}
+reduced<-lm(rate~Year+TaxClassCode+AddressAssessorMunicipalityDesc, data=assessment_aggregate)
+summary(reduced)
 
-we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400). We ignored the two municipalities and refit the linear model
-
-```{r}
-test<-pct_final %>% filter(Municipalities!="Langley - City") %>% 
-  filter(Municipalities!="Langley - Township")
-linear_test<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
-summary(linear_test)
-
-# goodness of fit
-fit_linear_test<-augment(linear_test)
-sqrt(sum((fit_linear_test$.resid)^2)/nrow(test)) # 6.767493
-
-linear_test_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
-summary(linear_test_2)
-
-# goodness of fit
-fit_linear_test_2<-augment(linear_test_2)
-sqrt(sum((fit_linear_test_2$.resid)^2)/nrow(test)) #  7.50047
+reduced_fit<-augment(reduced)
+sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate)) #  2.861619
 
 ```
-After filtering Langely
-with variable year:
-Residual standard error: 7.429 on 122 degrees of freedom
-Multiple R-squared:  0.2974,	Adjusted R-squared:  0.1592 
-F-statistic: 2.152 on 24 and 122 DF,  p-value: 0.003567
-PMSE: 6.767493
+Multiple R-squared:  0.7659,	Adjusted R-squared:  0.7408 
+MSPE = 2.861619
+With fewer variables, adjusted R-squared declines.
 
-without variable year
-Residual standard error: 8.167 on 124 degrees of freedom
-Multiple R-squared:  0.137,	Adjusted R-squared:  -0.01609 
-F-statistic: 0.8949 on 22 and 124 DF,  p-value: 0.6014
-PMSE: 7.50047
-
-```{r}
-linear3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=pct_final)
-summary(linear3)
-fit_linear3<-augment(linear3)
-sqrt(sum((fit_linear3$.resid)^2)/nrow(test)) #  7.50047
-
-```
-
-Question: How to judge goodness of fit? Multiple R square? Adjusted R square? or mean prediction error?
-**TODO: create training/test sets to test the prediction power!
+By comparing the two models, the fit of full model is better.
 
@@
 library(dummies)
-dummy_year<-dummy(test$Year)
-dummy_municipal<-dummy(test$Municipalities)
-dummy_taxclass<-dummy(test$TaxClassCode)
+dummy_year<-dummy(assessment_aggregate$Year)
+dummy_municipal<-dummy(assessment_aggregate$AddressAssessorMunicipalityDesc)
+dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
 # build x matrix
-x<-cbind(dummy_year,dummy_municipal)
-x<-cbind(x,dummy_taxclass)
-x<-cbind(x,as.numeric(test$assessTotal_pct))
-x<-cbind(x,as.numeric(test$tax))
-y<-test$rate_pct
+x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$predTaxIncome,assessment_aggregate$tax)
+
+y<-assessment_aggregate$rate
 lambdas <- 10^seq(2, -3, by = -.1)
 dim(x)
-colnames(x)[26]<-paste("assessment")
-colnames(x)[27]<-paste("tax")
-colnames(x)
 
@@
   R_square <- 1 - SSE / SST
-  RMSE = sqrt(SSE/nrow(predicted))
+  MSPE = sqrt(SSE/nrow(predicted))
 # Model performance metrics
 data.frame(
-  RMSE = RMSE,
+  MSPE = MSPE,
   Rsquare = R_square
@@
 ```
-RMSE=7.159649
-Rsquare=0.2136659
-Ridge regression doesn't seem to work well.
+MSPE = 1.992244
+Rsquare=0.8865213
 
@@
 ```{r}
-lambdas <- 10^seq(2, -3, by = -.1)
-
 # Setting alpha = 1 implements lasso regression
@@
 ```
-RMSE = 7.221299, Rsquare=0.2000649
+MSPE = 1.985639
+Rsquare=0.8872724
 
@@
 library(caret)
-tibble::as_tibble(test)
+tibble::as_tibble(assessment_aggregate)
 cv_10 = trainControl(method = "cv", number = 10)
-ridge = train(
-  rate_pct ~ Year + Municipalities + TaxClassCode + assessTotal_pct + tax, data = test,
+elastic_net = train(
+  rate~., data = assessment_aggregate,
   method = "glmnet",
@@
 )
-ridge
+elastic_net
 ```
-apha = 1 which implies lasso! refer to lasso.
+The final values used for the model were alpha = 0.1 and lambda = 0.09334029.
+MSPE = 3.066083  
+R squared = 0.7459584
+
+Among these all the models, lasso seems to have the best goodness of fit.
 
@@
 set.seed(450)
-train_ind<-sample(165,120)
-train1<-pct_final[train_ind,]
-test1<-pct_final[-train_ind,]
+train_ind<-sample(218,218-50)
+train<-assessment_aggregate[train_ind,]
+test<-assessment_aggregate[-train_ind,]
 ```
 
-linear model
+Full linear model
 ```{r}
-# full model with year
-newx<-test1[,-c(5,6,7,8)]
-y<-test1[,c(8)]
-linear_1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
+
+newx<-test[,-c(8)]
+y<-test[,c(8)]
+linear_1<-lm(rate ~.,data=train)
 resid<-predict(linear_1,newdata = newx) - y
-sqrt(sum(resid^2)/nrow(test1))
-# pmse = 6.5688
+sqrt(sum(resid^2)/nrow(test))
+```
+MSPE = 3.52065
 
-# full model without year
-linear_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
+Reduced model
+```{r}
+# reduced variable
+linear_2<-lm(rate~Year+TaxClassCode+AddressAssessorMunicipalityDesc,data=train)
 resid<-predict(linear_2,newdata = newx) - y
-sqrt(sum(resid^2)/nrow(test1))
-# 6.855429
+sqrt(sum(resid^2)/nrow(test))
+# 3.516749
+```
+MSPE = 3.516749
 
-# reduced variable
-linear_3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=train1)
-resid<-predict(linear_3,newdata = newx) - y
-sqrt(sum(resid^2)/nrow(test1))
-# 6.570848
-```
 Lasso
@@
 # create the whole matrix
-dummy_year<-dummy(pct_final$Year)
-dummy_municipal<-dummy(pct_final$Municipalities)
-dummy_taxclass<-dummy(pct_final$TaxClassCode)
-# build x matrix
-x<-cbind(dummy_year,dummy_municipal)
-x<-cbind(x,dummy_taxclass)
-x<-cbind(x,as.numeric(pct_final$assessTotal_pct))
-x<-cbind(x,as.numeric(pct_final$tax))
-y<-as.matrix(pct_final$rate_pct)
+y<-as.matrix(assessment_aggregate$rate)
 dim(x) # 165  29
@@
 
-lambdas <- 10^seq(2, -3, by = -.1)
 # Setting alpha = 1 implements lasso regression
@@
 eval_results(y_test, predictions_test)
-# 7.012128	
 ```
+MSPE = 2.530686
 
@@
 eval_results(y_test, predictions_test)
-# 6.885151	
 ```
+MSPE = 2.568047
+
+Elastic Net
+```{r}
+tibble::as_tibble(assessment_aggregate[train_ind,])
+cv_10 = trainControl(method = "cv", number = 10)
+elastic_net = train(
+  rate~., data = assessment_aggregate[train_ind,],
+  method = "glmnet",
+  trControl = cv_10
+)
+elastic_net
+# RMSE was used to select the optimal model using the smallest value.
+# The final values used for the model were alpha = 0.55 and lambda = 0.009140377.
+
+ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0.55, family = 'gaussian', lambda = 0.009140377)
+predictions_test <- predict(ridge_reg, newx = x_test)
+eval_results(y_test, predictions_test)
+```
+MSPE = 2.566096
 

### 2020-03-01 src/EDA&mode_fitting.Rmd
typo corrected.

@@
 
-ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0.55, family = 'gaussian', lambda = 0.009140377)
-predictions_test <- predict(ridge_reg, newx = x_test)
+elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0.55, family = 'gaussian', lambda = 0.009140377)
+predictions_test <- predict(elastic_reg, newx = x_test)
 eval_results(y_test, predictions_test)

### 2020-03-01 src/EDA&mode_fitting.Rmd
a few errors in lm model fixed

--- 2020-03-01 src/EDA&mode_fitting.Rmd
@@
 
-linear_full<-lm(rate~.,data=assessment_aggregate)
+colnames(assessment_aggregate)
+linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+predTaxIncome,data=assessment_aggregate)
 summary(linear_full)
@@
 linear_full_fit<-augment(linear_full)
-sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) #  2.772333
-
+sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
 ```
-Multiple R-squared:  0.7803,	Adjusted R-squared:  0.7503 
-MSPE = 2.772333
+Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8708 
+MSPE = 1.984109
 
@@
 ```{r}
-reduced<-lm(rate~Year+TaxClassCode+AddressAssessorMunicipalityDesc, data=assessment_aggregate)
+reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc), data=assessment_aggregate)
 summary(reduced)
@@
 reduced_fit<-augment(reduced)
-sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate)) #  2.861619
-
+sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate))
 ```
-Multiple R-squared:  0.7659,	Adjusted R-squared:  0.7408 
-MSPE = 2.861619
+Multiple R-squared:  0.8826,	Adjusted R-squared:  0.8679 
+MSPE = 2.02679
 With fewer variables, adjusted R-squared declines.
 
-By comparing the two models, the fit of full model is better.
+By comparing the two models, the fit of full model is the better of the two.
 
@@
 elastic_net = train(
-  rate~., data = assessment_aggregate,
+  rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+predTaxIncome, data = assessment_aggregate,
   method = "glmnet",
@@
 ```
-The final values used for the model were alpha = 0.1 and lambda = 0.09334029.
-MSPE = 3.066083  
-R squared = 0.7459584
+The final values used for the model were alpha = 0.1 and lambda = 0.007102922.
+ alpha  lambda       RMSE      Rsquared   MAE     
+  0.10   0.007102922  2.238720  0.8623328  1.711568
 
@@
 y<-test[,c(8)]
-linear_1<-lm(rate ~.,data=train)
+linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+predTaxIncome,data=train)
 resid<-predict(linear_1,newdata = newx) - y
@@
 ```
-MSPE = 3.52065
+MSPE =  2.592136
 
@@
 # reduced variable
-linear_2<-lm(rate~Year+TaxClassCode+AddressAssessorMunicipalityDesc,data=train)
+linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc),data=train)
 resid<-predict(linear_2,newdata = newx) - y
 sqrt(sum(resid^2)/nrow(test))
-# 3.516749
 ```
-MSPE = 3.516749
+MSPE = 2.537555
 
@@
 elastic_net = train(
-  rate~., data = assessment_aggregate[train_ind,],
+ rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+predTaxIncome, 
+ data = assessment_aggregate[train_ind,],
   method = "glmnet",
@@
 elastic_net
+
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were alpha = 0.55 and lambda = 0.009140377.
+# The final values used for the model were alpha = 1 and lambda = 0.006549203.
 
-ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0.55, family = 'gaussian', lambda = 0.009140377)
-predictions_test <- predict(ridge_reg, newx = x_test)
+elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.006549203)
+predictions_test <- predict(elastic_reg, newx = x_test)
 eval_results(y_test, predictions_test)
 ```
-MSPE = 2.566096
+MSPE = 2.559923
 

### 2020-03-01 src/EDA&mode_fitting.Rmd
delete highly correlated variable predTaxIncome in all models

--- 2020-03-01 src/EDA&mode_fitting.Rmd
@@
 
-linear_full<-lm(rate~.,data=assessment_aggregate)
+colnames(assessment_aggregate)
+linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
 summary(linear_full)
@@
 linear_full_fit<-augment(linear_full)
-sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) #  2.772333
-
+sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
 ```
-Multiple R-squared:  0.7803,	Adjusted R-squared:  0.7503 
-MSPE = 2.772333
+Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8707 
+MSPE =  1.984285
 
@@
 ```{r}
-reduced<-lm(rate~Year+TaxClassCode+AddressAssessorMunicipalityDesc, data=assessment_aggregate)
+reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal, data=assessment_aggregate)
 summary(reduced)
@@
 reduced_fit<-augment(reduced)
-sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate)) #  2.861619
-
+sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate))
 ```
-Multiple R-squared:  0.7659,	Adjusted R-squared:  0.7408 
-MSPE = 2.861619
-With fewer variables, adjusted R-squared declines.
-
-By comparing the two models, the fit of full model is better.
+Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8721 
+MSPE = 1.984541
 
@@
 # build x matrix
-x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$predTaxIncome,assessment_aggregate$tax)
+x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)
 
@@
 ```
-MSPE = 1.992244
-Rsquare=0.8865213
+MSPE = 1.989583
+Rsquare=	0.8868242
 
@@
 ```
-MSPE = 1.985639
-Rsquare=0.8872724
+MSPE = 1.985463	
+Rsquare=0.8872924
 
@@
 elastic_net = train(
-  rate~., data = assessment_aggregate,
+  rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
   method = "glmnet",
@@
 ```
-The final values used for the model were alpha = 0.1 and lambda = 0.09334029.
-MSPE = 3.066083  
-R squared = 0.7459584
-
-Among these all the models, lasso seems to have the best goodness of fit.
+The final values used for the model were alpha = 0.1 and lambda = 0.007102922.
+   alpha  lambda       RMSE      Rsquared   MAE     
+  0.10   0.007102922  2.236598  0.8625398  1.708885
 
@@
 
-newx<-test[,-c(8)]
+newx<-test[,-c(8,9)]
 y<-test[,c(8)]
-linear_1<-lm(rate ~.,data=train)
+linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
 resid<-predict(linear_1,newdata = newx) - y
@@
 ```
-MSPE = 3.52065
+MSPE =  2.590233
 
@@
 # reduced variable
-linear_2<-lm(rate~Year+TaxClassCode+AddressAssessorMunicipalityDesc,data=train)
+linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal,data=train)
 resid<-predict(linear_2,newdata = newx) - y
 sqrt(sum(resid^2)/nrow(test))
-# 3.516749
 ```
-MSPE = 3.516749
+MSPE = 2.523658
 
@@
 ```
-MSPE = 2.530686
+MSPE = 2.528047
 
@@
 ```
-MSPE = 2.568047
+MSPE = 2.567499
 
@@
 elastic_net = train(
-  rate~., data = assessment_aggregate[train_ind,],
+ rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
+ data = assessment_aggregate[train_ind,],
   method = "glmnet",
@@
 elastic_net
+
 # RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were alpha = 0.55 and lambda = 0.009140377.
+# The final values used for the model were alpha = 1 and lambda = 0.06549203.
 
-ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0.55, family = 'gaussian', lambda = 0.009140377)
-predictions_test <- predict(ridge_reg, newx = x_test)
+elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.06549203)
+predictions_test <- predict(elastic_reg, newx = x_test)
 eval_results(y_test, predictions_test)
 ```
-MSPE = 2.566096
+MSPE = 2.54861
 

### 2020-03-02 src/EDA&mode_fitting.Rmd
add a few plots and update the categorical variable in aov (need to use factor function for year)

@@
 
-tax_class <- assessment_aggregate %>% ggplot(aes(x=predTaxIncome,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and landTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and improvementTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and landTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and propertyCount - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
 assess_class
-
 tax_class
+land_class
+improvementTotal_class
+propertyCount_class
 
 ## Correlation for categorical variables
-TaxClassCode <- aov(rate ~ TaxClassCode, data = assessment_aggregate)
+TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
 summary(TaxClassCode)
 
-Year <- aov(rate_pct ~ Year, data = pct_final)
+Year <- aov(rate_pct ~ factor(Year), data = pct_final)
 summary(Year)
 
-Municipalities <- aov(rate ~ AddressAssessorMunicipalityDesc, data = assessment_aggregate)
+Municipalities <- aov(rate ~ factor(AddressAssessorMunicipalityDesc), data = assessment_aggregate)
 summary(Municipalities)

### 2020-03-03 doc/RealEstate_Xuechun_Lu.Rmd
Add files via upload

---
title: "Individual Report"
author: "Real Estate - Xuechun Lu"
date: "01/03/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Summary
A real estate analyst in Vancouver was interested in predicting future property tax for his company by predicting future mill rates. His expectation was to obtain a good prediction model to predict mill rates in each municipality in Metro Vancouver given the past data. In this study, data cleaning was done to prepare for the data analysis and raw data was transformed into summary statistics for each municipality. Exploratory data analysis (EDA) shows that mill rate is strongly correlated with year, municipality, tax class, assessment total, improvement, land and  property count, and weakly correlated with municipal budget. Linear regression models (ordinary linear regression, Lasso, Ridge and Elastic Net regression) were built to explore the relationship between mill rate and potential features. Among all models, ordinary linear regression with reduced features gives the slightly better goodness of fit and prediction power.

# Introduction
The goal of this study is to predict the 2019 mill rate for the following 21 municipalities in Metro Vancouver:

- Burnaby
- Coquitlam
- Delta
- Langley - City
- Langley - Township
- Maple Ridge
- Maple Ridge Rural 
- North Vancouver - City
- North Vancouver - Dist
- Pitt Meadows
- Port Coquitlam
- Port Moody
- Richmond
- Surrey
- Vancouver
- White Rock
- West Vancouver
- Bowen Island
- Anmore
- Belcarra
- Lions Bay

In this study, EDA was first done to visualize the trend of mill rats, and relationships between mill rate and all potential features. Based on results of EDA, we built ordinary linear regression, Lasso, Ridge and Elastic Net models and evaluated their goodness of fit (how well data fit models) using multiple R squared and adjusted R squared respectively. To further test model performance, prediction power (how well models can predict future values) for each model was examined by using mean squared prediction error (MSPE).

# Data Description

The real estate analyst provided past property assessment data with 2.4 million data entries from 2016 to 2019. We filtered information about commercial and residential properties (tax class 1, 5 and 6) in all 21 municipalities in Metro Vancouver.

Variables relevant to property tax prediction are selected as below:

Variable name | Interpretation | Type
--------------|--------|------------------------
rate | Mill Rates (2016-2019) | Quantitative
assessTotal| Assessment (2016-2019) | Quantitative
landTotal | Land Total (2016 - 2019) | Quantitative
ImprovementTotal|Improvement Total (2016 - 2019) | Quantitative
propertyCount|Number of Properties (2016-2019) | Quantitative
tax | Municpal Budget | Quantitatiive
TaxClassCode|Tax Class Code (01, 05, 06) | Categorical
AddressAssessorMunicipalityDesc | Area Code (Municipality) | Categorical
Year |Year (2016, 2017, 2018, 2019)| Categorical

External Data:

The analyst client also proposed that municipal budget could be related to mill rate, however, this data was not available neither in the data set he provided or online. Therefore, we collected external data from the Government of British Columbia to approximate the municipal budget:
Municipal Budget of Cities in Metro Vancouver (Taxes Imposed & Collected, Schedule 706)

The data listed were all cleaned to reduce their dimensions before data analysis and were all aggregated to  “assessment_aggregate.csv”.

# Method

## 1. Exploratory Data Analysis (EDA)

Before model construction and prediction of mill rates, EDA needs to be done to visualize the main characteristics of our dataset. In this study, EDA mainly focuses on exploring the relationships between mill rate and potential features. We first plotted correlations between mill rate and quantitative features (assessTotal, improvementTotal, propertyCount and landTotal). As for categorical features (TaxClassCode, Year and AddressAssessorMunicipalityDesc), instead of plotting correlations for all combination of them (3\*4\*21), ANOVA analysis was used to test if the correlation between mill rate and each categorical feature is significant.

## 2. Measure of goodness of fit and prediction power

In this study, we first built various models. The evaluation of model performances is based on goodness of fit (how well the model explains the data) and prediction power (how well it can predict future values).

Goodness of fit is defined as the extent to which the sample data are consistent with the model. Multiple R squared and adjusted R squared are the most well known measures of goodness of fit and higher multiple R squared and adjusted R squared generally indicate better goodness of fit. Moreover, for ordinary linear regression (OLR), reduction of number of features decreases multiple R squared. Therefore, when comparing OLR with different number of features, adjusted R squared will be looked at, instead of R squared.

Prediction power measures how well models can predict the future values. We used mean squared prediction error (MSPE) to examine prediction performances across all the models.  MSPE is the mean squared prediction error calculated from the testing set. To obtain it, we divided the data set into training data and testing data, which would be used as model construction and prediction respectively. Generally, smaller MSPE indicates better prediction power.

## 3. Ordinary Linear Model

The features we are interested in are AddressAssessorMunicipalityDesc (municipality), Year, TaxClassCode, assessTotal, landTotal, improvementTotal, propertyCount, predTaxIncome and tax (municipal budget).

We first built the ordinary linear model using all features available (OLR full) and computed its R squared, adjusted R squared and MSPE on the whole data set as measures of goodness of fit. Next, based on its R output, we selected the significant variables and re-built a OLR model using only these variables (OLR reduced), and we computed goodness of fit by repeating the previous step.

## 4. Ridge, Lasso and Elastic Net

Except for ordinary linear regression, we are also interested in the performance of advanced linear regressions like Ridge, Lasso and Elastic Net as they are also in the linear regression family, but have a different objection function to optimize. Among the three models, weights are assigned to features, and we are interested in minimizing the sum of squares of errors plus a penalty term. Ridge regression takes penalty in sum of absolute values of weights, whereas Lasso takes penalty in sum of squared absolute values of weights. Elastic net is a combination of Ridge and Lasso.

The reason for considering advanced linear regressions is that assigning weights to features might improve the goodness of fit and prediction power as each feature might influence the mill rate to a different extent.

Similar to the model construction for OLR, we first built the three models using all data and evaluated their goodness of fit. Then we split the data into training and testing sets and computed MSPE to examine their prediction power.

```{r, echo=FALSE, message=F, warning=F}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
library(readxl)
require(neuralnet)
library(knitr)
library(broom)
library(glmnet)
library(dummies)
library(caret)
library(neuralnet)
```

```{r, echo=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {
    print(plots[[1]])
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

```{r, echo=FALSE, warning=F, message=F}
# all municipalities
municipality.list = c(
  "Burnaby", 
  "Coquitlam", 
  "Delta", 
  "Langley - City", 
  "Langley - Township",
  "Maple Ridge",
  "Maple Ridge Rural", 
  "North Vancouver - City",
  "North Vancouver - Dist",
  "Pitt Meadows", 
  "Port Coquitlam", 
  "Port Moody", 
  "Richmond", 
  "Surrey", 
  "Vancouver", 
  "White Rock", 
  "West Vancouver", 
  "Bowen Island", 
  "Anmore", 
  "Belcarra",
  "Lions Bay")

assessment_aggregate <- read.csv(here("RealEstate/data","assessment_aggregate.csv"))

tax_final <- read.csv(here("RealEstate/data","tax_final.csv"))
#head(tax_final)

# tax_final is the budget
tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))

# tax_modified contains all the budget values from 2016 to 2019
#head(tax_modified)
```

# Results

## 1. EDA

```{r someVar, echo=FALSE}
# get predicted tax income for all tax classes and each municipality
assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
names(assessment_aggregate)[10] <- paste("predTaxIncome")

dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
names(dat_pred_income)[3] <- paste("predTaxIncome")

tax_refactored <- tax_modified
names(tax_refactored)[3] <- paste("2016")
names(tax_refactored)[4] <- paste("2017")
names(tax_refactored)[5] <- paste("2018")
names(tax_refactored)[6] <- paste("2019")
dat_true_income <- gather(tax_refactored, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
dat_true_income <- dat_true_income[,-1]

income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
income_budget_compare <- gather(income_budget_compare, TaxType, Amount, TaxAmount:predTaxIncome, factor_key=TRUE)

compare_full_plot <- income_budget_compare %>% filter(Municipalities!="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=Municipalities)) + geom_line()

compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Vancouver")
compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Surrey") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Surrey")
compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax or Burnaby")
compare_plot4 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Richmond")
multiplot(compare_plot1,compare_plot2,compare_plot3,compare_plot4,cols=2)
```

Plot 1. budget and tax to pay from 2016-2019 in 4 major municipalities

We first plot municipal budget (taxAmount) and predicted property tax to pay (predTaxIncome) against time from 2016 to 2019 in 4 major municipalities (Vancouver, Burnaby, Surrey and Richmond). In the plot above,  the predicted tax to pay is less than the municipal budget. Also, budget and predicted tax generally have an increasing trend and move together in the consecutive four years, which indicates a potential correlation between the two variables.

```{r, echo=F}
# aggregate tax and mill rate & assessment
colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)

tax_modified1 <- gather(tax_modified, Year, tax, '2016':'2019', factor_key = TRUE)%>%
  rename(AddressAssessorMunicipalityDesc = Municipalities)

assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
assessment_aggregate <- assessment_aggregate[,-c(3,11)]
names(assessment_aggregate)[10] <- paste("tax")

## Correlation for continuous variables
assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and assess - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and landTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and improvementTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and propertyCount - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

multiplot(assess_class,
tax_class,
land_class,improvementTotal_class,
          propertyCount_class, cols=2)

```

Plot 2. scatterplot with regression line for mill rate against assessTotal, tax, landTotal, improvementTotal and propertyCount.

Plot 2 shows the correlation between mill rate and five quantitative variables that likely affect the mill rate. Mill rate is negatively correlated with assessTotal, improvementTotal, propertyCount and landTotal considering all the 21 municipalities in Metro Vancouver, as the black linear regression lines all have a negative slope. In contrast, mill rate and budget is weakly correlated since the linear regression line is almost flat.

```{r, echo = F}
## Correlation for categorical variables
TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
#summary(TaxClassCode)

Year <- aov(rate ~ factor(Year), data = assessment_aggregate)
#summary(Year)

Municipalities <- aov(rate ~ factor(AddressAssessorMunicipalityDesc), data = assessment_aggregate)
#summary(Municipalities)

table<-
rbind(
c("factor(TaxClassCode)",2,5796,2897.9,340.6,"<2e-16 ***"),
c("factor(Year)",3,405,135.13,4.006,"0.00841 **"),
c("factor(AddressAssessorMunicipalityDesc)",19,979,51.53,"1.535", "0.0768 .")

)

colnames(table)<-paste(c("variable", "df", "sum sq", "mean sq", "F value", "Pr(>F)"))
#table
kable(table, caption = "ANOVA analysis of mill rate and three categorical variables")
print("Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1")
```

Table 2 lists the result of ANOVA analysis of mill rate and three categorical variables - TaxClassCode, Year and Municipality (AddressAssessorMunicipalityDesc). A variable with p-value (Pr(>F)) less than 5% is considered statistically significant at 5% significance level. From the table, TaxClassCode and Year are statistically significantly correlated with mill rate at 5% level.

# 2. Goodness of Fit
```{r, echo=F, results="hide"}
# full linear model
linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
summary(linear_full)
linear_full_fit<-augment(linear_full)
sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
# Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8707 
# MSPE = 1.984285
row1<-c("OLR full",0.8874, 0.8707,round(1.984285,4))

# reduced model
reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal, data=assessment_aggregate)
summary(reduced)
reduced_fit<-augment(reduced)
# Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8721 
sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate))
row2<-c("OLR reduced",0.8874,0.8721,round(1.984541,4))
```
```{r, echo=F}
# create a list of significant variables from the full linear model

colsname<-c("Variable", "Pr(>|t|)", "Significance level") 
table<-rbind(c("Intercept",                               "0.000111", "***"),
  c("Coquitlam", "0.001594", "**"),
  c("Delta",  "0.002579", "**"),
  c("Langley - City",  "0.020365", "*"),
  c("Langley - Township",   "0.023030", "*" ),
  c("Maple Ridge",      "4.18e-05", "***"),
  c("Pitt Meadows",   "2.41e-07", "***"),
  c("Coquitlam",    "1.60e-05", "***"),
  c("Port Moody",     "1.16e-05", "***"),
  c("West Vancouver",  "0.001003", "**"),
  c("White Rock",  "0.034041", "*"), 
  c("Year 2017",    "0.000655", "***"),
  c("Year 2018",   "1.30e-08", "***"),
  c("Year 2019",   "4.73e-13", "***"),
  c("TaxClassCode 5",          "< 2e-16",  "***"),
  c("TaxClassCode 6",    "< 2e-16", "***"),
  c("assessTotal",    "0.013752", "*"),  
  c("landTotal",   "0.013863", "*" ))

colnames(table)<-colsname
rownames(table)<-NULL
kable(table, caption="List of significant variables in OLR full")

print("Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1")

```

```{r, echo=F, results="hide"}
# Ridge
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$AddressAssessorMunicipalityDesc)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)
y<-assessment_aggregate$rate
lambdas <- 10^seq(2, -3, by = -.1)

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)

# Compute R^2 from true and predicted values
eval_results <- function(true, predicted) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  MSPE = sqrt(SSE/nrow(predicted))
# Model performance metrics
data.frame(
  MSPE = MSPE,
  Rsquare = R_square
)

}

predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
eval_results(y, predictions_train)

row3<-c("Ridge",round(0.8868242,4),"N/A",round(1.989583,4))

# Lasso
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
lambda_best <- lasso_reg$lambda.min;lambda_best
lasso_model <- glmnet(x, y, alpha = 1, lambda = lambda_best, standardize = TRUE)
predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
eval_results(y, predictions_train)

row4<-c("Lasso",round(0.8872924,4)	,"N/A",round(1.985463,4))

# Elastic Net
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
elastic_net
row5<-c("Elastic Net",round(0.8625398,4),"N/A",round(2.236598,4))

```

```{r, echo=F}
colsname<-c("Model", "Multiple R-squared", "Adjusted R-squared", "MSPE")
table<-rbind(row1,row2,row3,row4,row5)
colnames(table)<-colsname
rownames(table)<-NULL
kable(table, caption="Goodness of fit for all models")

```

We performed ordinary linear regression (OLR) with all variables first and computed multiple R-squared, adjust R-squared and mean square prediction error (MSPE). Since the summary statistics in Table 3 show that only Municipality, Year, TaxClassCode, assessTotal and landTotal are significant, we therefore constructed a reduced OLR model with only the significant variables.

Table 5 shows the comparison of the full model (OLR full) and the reduced model (OLR reduced). Multiple R-squared doesn’t seem to improve, but adjusted R-squared improves by a small amount. This indicates that with fewer variables, the goodness of fit improves slightly. In addition, with the three advanced linear models (Ridge, Lasso and Elastic Net), the family of OLR have slightly better fit and OLR reduced has the best fit as its adjusted R-squared is slighter higher than OLR full.

## 3. Prediction Power

```{r, echo=F, results="hide"}
set.seed(450)
train_ind<-sample(218,218-50)
train<-assessment_aggregate[train_ind,]
test<-assessment_aggregate[-train_ind,]
```

```{r, echo=F, results="hide", warning=F}
# full linear model
newx<-test[,-c(8,9)]
y<-test[,c(8)]
linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
resid<-predict(linear_1,newdata = newx) - y
row1<-c("OLR full", round(sqrt(sum(resid^2)/nrow(test)),4))

# reduced model
linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal,data=train)
resid<-predict(linear_2,newdata = newx) - y
row2<-c("OLR reduced",round(sqrt(sum(resid^2)/nrow(test)),4))
```

```{r, echo=F, results="hide", warning=F}
# lasso
# create the whole matrix
y<-as.matrix(assessment_aggregate$rate)
dim(x) # 165  29
dim(y)
# creat x_train matrix and y_train
x_train<-x[train_ind,]
y_train<-y[train_ind,]
# create x_test matrix
x_test<-x[-train_ind,]
y_test<-y[-train_ind,]

# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
lambda_best <- lasso_reg$lambda.min;lambda_best
lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
eval_results(y_test, predictions_test)
row4<-c("Lasso",round(2.528047,4))

# ridge
ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
eval_results(y_test, predictions_test)
row3<-c("Ridge",round(2.567499,4))

# elastic net
tibble::as_tibble(assessment_aggregate[train_ind,])
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
 rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
 data = assessment_aggregate[train_ind,],
  method = "glmnet",
  trControl = cv_10
)
elastic_net

# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were alpha = 1 and lambda = 0.06549203.

elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.06549203)
predictions_test <- predict(elastic_reg, newx = x_test)
eval_results(y_test, predictions_test)
row5<-c("Elastic Net",round(2.54861,4))

```

```{r, echo=F}
colsname<-c("Model", "MSPE")
table<-rbind(row1,row2,row3,row4,row5)
colnames(table)<-colsname
rownames(table)<-NULL
kable(table, caption="Prediction power for all models")
```

Table 5 shows that among the five models, OLR reduced has the smallest MSPE and thus the best prediction power. However, the difference between OLR reduced and other models is not big.

# Conclusion

The OLR shows significant variables related to mill rates are Year, Tax Class, Municipality, Assessment and Land Total

OLR reduced has slightly better goodness of fit of the model and prediction power among all models. However, this good fit is likely to be a result of over-fitness since for each municipality and tax class, we only have 4 consecutive data points from 2016 to 2019.

# References

Links to source of data:

- Schedule 706 (https://www2.gov.bc.ca/gov/content/governments/local-governments/facts-framework/statistics/statistics)

Code repository:

- Data Cleaning (https://github.com/STAT450-550/RealEstate/blob/450/src/Data_Cleaning.Rmd)

- Exploratory Data Analysis and Model Fitting (https://github.com/STAT450-550/RealEstate/blob/450/src/EDA%26mode_fitting.Rmd)

# Appendix

## Missing Value:

There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values
Based on client information, all properties in the same region, classcode, and year should have a unique class rate

- For entries with mill rate, we aggregated them into groups by region + classcode + year.

- For entries without mill rate, we found the group they belong to and assign them mill rate in that group.

Here is some exceptions found:

Some groups' mill rate is not unique:

- In Delta, properties in different neighbourhoods have slightly different mill rates. Since the variance is not significant, we take the mean as the overall mill rate in groups.

- In Vancouver, 2019, Class 01, one property's mill rate is different from others. It is regarded as an outlier.

- In Burnaby, 2019, Class 06, six properties' mill rate are different from others. They are regarded as outliers.

- In Langley, 2019, Class 06, mill rate is different between assessment type. After talking to the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.

- In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:

Year  | Region | Class |Number of Properties
------|--------|-------|----------
2016 | Belcarra | 06 | 9 
2016 | Lions Bay | 01 | 40 
2016 | Lions Bay | 06 | 25 
2016 | Maple Ridge Rural | 05 | 36 
2017 | Belcarra | 06 | 9 
2017 | Lions Bay | 01 | 39 
2017 | Lions Bay | 06 | 24 
2017 | Maple Ridge Rural | 05 | 36 
2018 | Maple Ridge Rural | 05 | 36 
2019 | Maple Ridge Rural | 05 | 38

### 2020-03-03 doc/RealEstate_Yuetong_Liu.Rmd
YT report

---
title: 'STAT 450 Project: Real Estate'
author: "Yuetong Liu, 43838168"
date: "3/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

Our client is a real estate data analyst in Metro Vancouver who analyzes the market trend to make a market prediction for his company and future client. He is interested in accurately predicting future mill rates in Metro Vancouver. There are three major sections of the project: Data cleaning, exploratory data analysis, and linear modelling. Data cleaning aggregated our data into summary statistics and also reduced the number of data entries. Exploratory data analysis analyzed correlation between mill rate and othee features in assessment data frame. Linear modelling gave the best goodness of fit and prediction power among reduced linear regression models, Ridge, Lasso and Elastic net.

## Introduction

Our goal is to predict the property tax in 2020, for each of the following 21 sub-regions in the Greater Vancouver area. A full linear model, a reduced linear model, LASSO, Ridge, and elastics net Regressions were used to train our predictive model. Correlations between mil rate, assessment value, and municipal budget were performed to see the significance of these factors in our predictive model.

Once we have finished the above tasks, we will use the mean square test error to evaluate the performance of our model.

## Data Description

Since we are interested in predicting mill rate, only the assessment data set provide by the client was used.There are approximately 2.4 million data entries across these 5 years.

The dataset was filter based on below conditions:

TaxClassCode: 01, 05, 06

MunicipalityDesc:

- Burnaby
- Coquitlam
- Delta
- Langley - City
- Langley - Township
- Maple Ridge
- Maple Ridge Rural
- North Vancouver - City 
- North Vancouver - Dist
- Pitt Meadows 
- Port Coquitlam
- Port Moody 
- Richmond
- Surrey
- Vancouver
- White Rock
- West Vancouver
- Bowen Island
- Anmore
- Belcarra
- Lions Bay

Year: 2016, 2017, 2018, 2019

Mill Rate and 4 features relevent to Mill Rate are selected:

- Mill Rates 
- Assessment Value
- Tax Class Code 
- Area Code 
- Year

We also collected external data from the Government of British Columbia which could be associated with the tax rate:

- Municipal Budget of Cities in Metro Vancouver (Taxes Imposed & Collected, Schedule 706)

Based on client information, all properties in the same region, classcode, and year should have a unique class rate. Therefore, all data were grouped by **region + classcode + year**. The assessement value and count of each properties within a group was summed up. 
Moreover, we decided to calculate the growth rate of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, Budget and Tax Rate for each municipality and their corresponding tax class.

## Methods

### Exploratory Analysis

Before we made any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to visualize the main characteristics of our dataset and found relationships between mill rate, assessment and municipal budget.

- It's assumed that the government aims to match its budget and its income by adjusting the mill rates; therefore, we plotted a graph of tax income for three different tax classes (01, 05, and 06) vs. budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct.

- Correlation analysis between mill rate and other features. We used scatter plot for numerical variables and ANOVA table for categorical variables.

### Measure of goodness of fit and prediction power

In this study, we first built various models and we evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.

**Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data. R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.

**Prediction power** measures how well models can predict the future values. In this study, we use mean squared prediction power to compare the prediction performance across all the models.  To do that, we will divide the data set into training data, used as building models, and testing data, used as computing mean squared prediction power.

### Linear model

The explanatory variables we are interested in are Year, TaxClassCode, Municipalities, assessTotal_pct and tax (budget). We first explore the full linear model using all the available features. Since there are potential outliers in the data set, it might be helpful to refit the model without the outliers to check if the fit of the model is improved.

### Selection of variables

A list of significant variables are included in the result of the full model. In the next step, we will refit the model with only the significant variables. Goodness of fit will decline with fewer features, however, the performance of prediction power might be improved.

### Ridge, Lasso and Elastic Net

Except for ordinary linear regression, we are also interested in the performance of Ridge, Lasso, and Elastic Net models as they are also in the linear regression family, but more advanced than ordinary linear regression. In these two models, weights are assigned to features. Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively. Elastic net is a combination of Ridge and Lasso.

We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.

### Neural Network

In the current report, we are mainly focusing on predicting the mill rate using Linear Model. In the future, we will discuss with our supervisor and try to fit a neural network model if possible (R package: neuralnet or keras). Neural Network has the ability to learn and model non-linear and complex relationships. It is flexible and can be used for both classification and prediction. Since we have a large number of data, it makes Neural Network a very good candidate for our prediction model.

Report on the performance of the Neural Network will be included in the next report.
\pagebreak

### Results: Exploratory Analysis

```{r budget, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
library(readxl)

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {
    print(plots[[1]])
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

# all municipalities
municipality.list = c(
  "Burnaby", 
  "Coquitlam", 
  "Delta", 
  "Langley - City", 
  "Langley - Township",
  "Maple Ridge",
  "Maple Ridge Rural", 
  "North Vancouver - City",
  "North Vancouver - Dist",
  "Pitt Meadows", 
  "Port Coquitlam", 
  "Port Moody", 
  "Richmond", 
  "Surrey", 
  "Vancouver", 
  "White Rock", 
  "West Vancouver", 
  "Bowen Island", 
  "Anmore", 
  "Belcarra",
  "Lions Bay")

# budget and assessment data import and process
tax_pct<-read_csv(here("data","tax_pct.csv"))
assessment_pct<-read_csv(here("data","assessment_pct.csv"))
# omit 2016
assessment_pct<-na.omit(assessment_pct)
assessment_pct<-assessment_pct[,-c(1)]
names(assessment_pct)[2]<-paste("Municipalities")
tax_pct<-tax_pct[,-c(1)]
#head(tax_pct)

assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
assessment_final <- read.csv(here("data","assessment_final.csv"))
assessment_final <- assessment_final[,-1]
assessment_final <- assessment_final[complete.cases(assessment_final),]
assessment_final <- assessment_final[,-4]

# Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge Rural and Maple Ridge with the same tax.

# 2019 tax pct is missing, we imputed 2019 tax by the average of previous tax pct
tax_pct[,4]<-(tax_pct$pct_2017+tax_pct$pct_2018)/2
#tax_pct
names(tax_pct)[4]<-paste("pct_2019")
tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567,5.945346))

# aggregate tax and assessment
assessment2017<-assessment_pct  %>%  filter(Year=="2017")
tax2017<-tax_pct[,1:2]
assessment2017<-assessment2017 %>% left_join(tax2017, by = c("Municipalities"))
names(assessment2017)[9]<-paste("tax")

assessment2018<-assessment_pct  %>%  filter(Year=="2018")
tax2018<-tax_pct[,c(1,3)]
assessment2018<-assessment2018 %>% left_join(tax2018, by = c("Municipalities"))
names(assessment2018)[9]<-paste("tax")

assessment2019<-assessment_pct  %>%  filter(Year=="2019")
tax2019<-tax_pct[,c(1,4)]
assessment2019<-assessment2019 %>% left_join(tax2019, by = c("Municipalities"))
names(assessment2019)[9]<-paste("tax")

pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
pct_final$tax<-as.numeric(pct_final$tax)
write.csv(pct_final, here("data","data_final.csv"))

# 2019 tax is missing, we imputed 2019 tax using previous values
tax_final <- read.csv(here("data","tax_final.csv"))
#head(tax_final)

# tax_final is the budget
tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))

# check match between budget and assessTotal*mill_rate
# get predicted tax income for all tax classes and each municipality
assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
names(assessment_aggregate)[10] <- paste("predTaxIncome")

dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
names(dat_pred_income)[3] <- paste("TaxIncome")

tax_refactored <- tax_modified
names(tax_refactored)[3] <- paste("2016")
names(tax_refactored)[4] <- paste("2017")
names(tax_refactored)[5] <- paste("2018")
names(tax_refactored)[6] <- paste("2019")
dat_true_income <- gather(tax_refactored, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
dat_true_income <- dat_true_income[,-1]
names(dat_true_income)[3] <- paste("Budget")

income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
income_budget_compare <- gather(income_budget_compare, TaxType, Amount, Budget:TaxIncome, factor_key=TRUE)

compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 1: Budget and Tax Income v.s. Year of Vancouver") 
compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 2: Budget and Tax Income v.s Year of Burnaby")
compare_plot1
compare_plot3
```

For each year, we predicted the tax income of three class codes (01, 05, 06) in each municipality by multiplying access total and mill rate. Then we plotted the predicted tax income vs municipal budget into line-chart to visualize their relationship. We chose Vancoouver and Burnaby because they were two largest municipalities, and the plots showed that the predicted tax amount was positively related to municipal budget.

```{r cor, echo=FALSE, warning=FALSE, message=FALSE}
# aggregate tax and mill rate & assessment
colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)

tax_modified1 <- gather(tax_modified, Year, tax, '2016':'2019', factor_key = TRUE)%>%
  rename(AddressAssessorMunicipalityDesc = Municipalities)

assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
assessment_aggregate <- assessment_aggregate[,-c(3,11)]
names(assessment_aggregate)[10] <- paste("tax")

# correlation between mill rate and budget for each municipality
rate_tax_correlation <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(rate, tax))
#rate_tax_correlation
# correlation between mill rate and assessment
rate_assess_correlation2 <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(assessTotal, rate))
#rate_assess_correlation2
names(assessment_aggregate)[1] <- paste("Municipality")
## Correlation for continuous variables
assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 3: Correlations between Mill Rates and Total Assessment Values of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 4: Correlations between Mill rates and Budgets of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 5: Correlations between Mill Rates and Total Assessment Values of Land of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 6: Correlations between Mill Rates and Total Assessment Values of Improvement of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 7: Correlations between Mill Rates and Total Number of Properties of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
```

```{r assess, echo=FALSE}
assess_class
tax_class
land_class
improvementTotal_class
propertyCount_class
```

These 5 scatter plots visualized the correlation between mill rate and 5 numerical features (assessment total, municipal budget, land assessment total, improvement assessment total, and the number of properties). In every scatter plot, Vancouver (the dark pink points) are significantly different from other points and should be regarded as an outlier/special case. There are some other municipalities (including Richmond, Surrey and Burnaby) that are also distinct in the scatter plots. We believe these municiplaties also have impacts on the correlation between mill rate and its features.

```{r anova, echo=FALSE, warning=FALSE, message=FALSE}
## Correlation for categorical variables
TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
#summary(TaxClassCode)

Year <- aov(rate_pct ~ factor(Year), data = pct_final)
#summary(Year)

Municipalities <- aov(rate ~ factor(Municipality), data = assessment_aggregate)
#summary(Municipalities)
```

Feature Name  | Pr(>F) 
------|--------
TaxClassCode | <2e-16 
Year | 4.2e-06 
Municipality | 0.0768

We used three 3 ANOVA tables to analyze the correlation between mill rate and each of the 3 categorical features (TaxClassCode, Year, Municipalities). TaxClassCode and Year showed a strong correlation with mill rate while the correlation between Municipality and mill rate is not significant at 5% level.

\pagebreak

## Results: Linear Model

```{r full, echo=FALSE, warning=FALSE, message=FALSE}
linear_full<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
# summary(linear_full)

library(broom)

linear_full_fit<-augment(linear_full)
mse_full <- sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
```
```{r reduced, echo=FALSE, warning=FALSE, message=FALSE}
reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
#summary(reduced)

reduced_fit<-augment(reduced)
mse_reduced <- sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate))
```

```{r Ridge, echo=FALSE, warning=FALSE, message=FALSE}
library(glmnet)
library(dummies)
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$Municipality)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)

y<-assessment_aggregate$rate
lambdas <- 10^seq(2, -3, by = -.1)
#dim(x)

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
#optimal_lambda
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)

# Compute R^2 from true and predicted values
eval_results <- function(true, predicted) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  MSPE = sqrt(SSE/nrow(predicted))
# Model performance metrics
data.frame(
  MSPE = MSPE,
  Rsquare = R_square
)

}

predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
ridge_mse <- eval_results(y, predictions_train)

```

```{r LASSO, echo=FALSE, warning=FALSE, message=FALSE}
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

# Best 
lambda_best <- lasso_reg$lambda.min;#lambda_best

lasso_model <- glmnet(x, y, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
lasso_mse <- eval_results(y, predictions_train)
```

```{r elastic_net, echo=FALSE, warning=FALSE, message=FALSE}
library(caret)
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
```

```{r Prediction power, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(450)
train_ind<-sample(218,218-50)
train<-assessment_aggregate[train_ind,]
test<-assessment_aggregate[-train_ind,]

# Full linear model
newx<-test[,-c(8,9)]
y<-test[,c(8)]
linear_1<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
resid<-predict(linear_1,newdata = newx) - y
full_mspe <- sqrt(sum(resid^2)/nrow(test))

# Reduced model
linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal,data=train)
resid<-predict(linear_2,newdata = newx) - y
reduced_mspe <- sqrt(sum(resid^2)/nrow(test))

# Lasso
# create the whole matrix
y<-as.matrix(assessment_aggregate$rate)
#dim(x) # 165  29
#dim(y)
# creat x_train matrix and y_train
x_train<-x[train_ind,]
y_train<-y[train_ind,]
# create x_test matrix
x_test<-x[-train_ind,]
y_test<-y[-train_ind,]

# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

# Best 
lambda_best <- lasso_reg$lambda.min;#lambda_best

lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
lasso_mspe <- eval_results(y_test, predictions_test)

# Ridge
ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
#optimal_lambda
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
ridge_mspe <- eval_results(y_test, predictions_test)

# Elastic Net
#tibble::as_tibble(assessment_aggregate[train_ind,])
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
 rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
 data = assessment_aggregate[train_ind,],
  method = "glmnet",
  trControl = cv_10
)

elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.06549203)
predictions_test <- predict(elastic_reg, newx = x_test)
elastic_net_mspe <- eval_results(y_test, predictions_test)
```

We fitted 5 models and compared the goodness of fit of each models.

Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE 
------|--------|--------|--------|--------
      OLR full | 0.8874| 0.8707| 1.9843| 2.5902
      OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237
      Ridge|  0.8868| NA| 1.9896| 2.5675
      LASSO| 0.8873| NA| 1.9855| 2.5280
      Elastic Net| 0.8625| NA| 2.2366| 2.5486

Here are the specifications of each indicator:

- **Multiple R squared**:  A measure of Rsquared for models that have multiple predictor variables. It will increase when adding predictors to your model.

- **Adjusted Rsquared**:  It controls against the increase of Multiple R squared and adds penalties for the number of predictors in the model. Therefore it shows a balance between the most parsimonious model, and the best fitting model.

- **MSE**: An estimator that measures the average of the squares of the errors: the average squared difference between the estimated values and the actual value.

- **PMSE**: It is the expected value of the squared difference between the fitted values implied by the predictive function and the values of the (unobservable) function (In our case, we randomly select 50 samples from the data frame as testing set to estimate the PMSE).

Except for Elastic Net model which have higher MSE and PMSE, there is no significant differences between these models.

## Conclusions
Among all five models, the full linear regression model is the most ideal prediction model since it make accurate prediction without losing information. There are some aspects that we could improve/implemented in the future:

- **Cross Validation**: The fitted reduced linear model was able to make the most accurate prediction across all our selected linear models, but there is still a very high chance that our model is overfitted. For our next report, we are going to use cross-validation to reduce the effect of overfitting.

- **Data Tranformation**: The client suggested that we can transform numerical factors such as assessment total into percentage change. This data transformation does not perform as well as we expected, which only yields an R^2 of around 0.2 across all of our fitted models. We believe the method that the client has suggested might leave out some important information about the housing market in each municipality. However, there could be other data transformation applied to our data, and the relationship between mill rate and other features might not be linear. GAM (Genalized addictive model) could be used to fit non-linear data.

- **Outlier**: In most correlation scatter plots, Vancouver always is an outlier. Since it makes up large portion of the market, we would like to further investigate Vancouver as a special case. We are also interested in the effect of different tax classes in predicting mill rate, since the P-value of ANOVA between tax class and mill rate was extremely small.

## References

Links to source of data:

- Schedule 706 (https://www2.gov.bc.ca/gov/content/governments/local-governments/facts-framework/statistics/statistics)

Code repository:

- Data Cleaning (https://github.com/STAT450-550/RealEstate/blob/450/src/Data_Cleaning.Rmd)

- Exploratory Data Analysis and Model Fitting (https://github.com/STAT450-550/RealEstate/blob/450/src/EDA%26mode_fitting.Rmd)

## Appendix

### Missing Value:

There are 1801 missing values in mill rate (TaxClassTaxRate). Since this is the predictor, we decided to impute these missing values
Based on client information, all properties in the same region, classcode, and year should have a unique class rate

- For entries with mill rate, aggregate them into groups by region + classcode + year.

- For entries without mill rate, find the group they belong to and assign them mill rate in that group.

Here is some exceptions found:

#### Some groups' mill rate is not unique:

- In Delta, properties in different neighbourhoods have slightly different mill rates. Since the variance is not significant, we take the mean as the overall mill rate in groups.

- In Vancouver, 2019, Class 01, one property's mill rate is different from others. It is regarded as an outlier.

- In Burnaby, 2019, Class 06, six properties' mill rate are different from others. They are regarded as outliers.

- In Langley, 2019, Class 06, mill rate is different between assessment type. After talking with the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.

#### In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:

Year  | Region | Class |Number of Properties
------|--------|-------|----------
2016 | Belcarra | 06 | 9 
2016 | Lions Bay | 01 | 40 
2016 | Lions Bay | 06 | 25 
2016 | Maple Ridge Rural | 05 | 36 
2017 | Belcarra | 06 | 9 
2017 | Lions Bay | 01 | 39 
2017 | Lions Bay | 06 | 24 
2017 | Maple Ridge Rural | 05 | 36 
2018 | Maple Ridge Rural | 05 | 36 
2019 | Maple Ridge Rural | 05 | 38

### 2020-03-03 doc/RealEstate_Xuechun_Lu.Rmd
Add files via upload

@@
 
-```{r someVar, echo=FALSE}
+```{r someVar, echo=T}
 # get predicted tax income for all tax classes and each municipality
@@
 
-```{r, echo=F}
+```{r, echo=T}
 # aggregate tax and mill rate & assessment
@@
 # 2. Goodness of Fit
-```{r, echo=F, results="hide"}
+```{r, echo=T, results="hide"}
 # full linear model
@@
 
-```{r, echo=F, results="hide"}
+```{r, echo=T, results="hide"}
 # Ridge
@@
 
-```{r, echo=F, results="hide"}
+```{r, echo=T, results="hide"}
 set.seed(450)
@@
 
-```{r, echo=F, results="hide", warning=F}
+```{r, echo=T, results="hide", warning=F}
 # full linear model
@@
 
-```{r, echo=F, results="hide", warning=F}
+```{r, echo=T, results="hide", warning=F}
 # lasso

### 2020-03-03 doc/RealEstate_Yuetong_Liu.Rmd
Individual Report

@@
 
-Our client is a real estate data analyst in Metro Vancouver who analyzes the market trend to make a market prediction for his company and future client.
-He is interested in accurately predicting future mill rates in Metro Vancouver.
+Our client is a real estate data analyst who is interested in accurately predicting future mill rates in Metro Vancouver.
 There are three major sections of the project: Data cleaning, exploratory data analysis, and linear modelling.
 Data cleaning aggregated our data into summary statistics and also reduced the number of data entries.
-Exploratory data analysis analyzed correlation between mill rate and othee features in assessment data frame.
+Exploratory data analysis analyzed correlation between mill rate and other features in assessment data frame.
 Linear modelling gave the best goodness of fit and prediction power among reduced linear regression models, Ridge, Lasso and Elastic net.
@@
 
-The dataset was filter based on below conditions:
+The dataset was filtered based on below conditions:
 
@@
 
-Mill Rate and 4 features relevent to Mill Rate are selected:
+Mill Rate and 4 features relevent to Mill Rate were selected:
 
@@
 Therefore, all data were grouped by **region + classcode + year**.
-The assessement value and count of each properties within a group was summed up.
+The assessement value and count of each properties within a group were summed up.
 Moreover, we decided to calculate the growth rate of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, Budget and Tax Rate for each municipality and their corresponding tax class.
@@
 
-Before we made any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to visualize the main characteristics of our dataset and found relationships between mill rate, assessment and municipal budget.
+Before making any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to visualize the main characteristics of our dataset and help with model selection.
+Here are two of our major approach:
 
@@
 
-**Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data.
-R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.
-
-**Prediction power** measures how well models can predict the future values.
-In this study, we use mean squared prediction power to compare the prediction performance across all the models.
-To do that, we will divide the data set into training data, used as building models, and testing data, used as computing mean squared prediction power.
-
 ### Linear model
@@
 Since there are potential outliers in the data set, it might be helpful to refit the model without the outliers to check if the fit of the model is improved.
-
-### Selection of variables
-
-A list of significant variables are included in the result of the full model.
-In the next step, we will refit the model with only the significant variables.
-Goodness of fit will decline with fewer features, however, the performance of prediction power might be improved.
 
@@
 
-We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.
+### Goodness of fit
 
-### Neural Network
+It is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data.
+R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.
 
-In the current report, we are mainly focusing on predicting the mill rate using Linear Model.
-In the future, we will discuss with our supervisor and try to fit a neural network model if possible (R package: neuralnet or keras).
-Neural Network has the ability to learn and model non-linear and complex relationships.
-It is flexible and can be used for both classification and prediction.
-Since we have a large number of data, it makes Neural Network a very good candidate for our prediction model.
+### Prediction power
 
-Report on the performance of the Neural Network will be included in the next report.
+It measures how well models can predict the future values.
+In this study, we use mean squared prediction power to compare the prediction performance across all the models.
+To do that, we will divide the data set into training data, used as building models, and testing data, used as computing mean squared prediction power.
+
 \pagebreak
@@
 multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+
   library(grid)
@@
  if (numPlots==1) {
+
     print(plots[[1]])
+
   } else {
+
     # Set up the page
@@
 
-# Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge Rural and Maple Ridge with the same tax.
+# Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge 
+# Rural and Maple Ridge with the same tax.
 
@@
 # aggregate tax and assessment
-assessment2017<-assessment_pct  %>%  filter(Year=="2017")
+assessment2017<-assessment_pct  %>%  
+  filter(Year=="2017")
 tax2017<-tax_pct[,1:2]
-assessment2017<-assessment2017 %>% left_join(tax2017, by = c("Municipalities"))
+assessment2017<-assessment2017 %>% 
+  left_join(tax2017, by = c("Municipalities"))
 names(assessment2017)[9]<-paste("tax")
 
-assessment2018<-assessment_pct  %>%  filter(Year=="2018")
+assessment2018<-assessment_pct  %>%  
+  filter(Year=="2018")
 tax2018<-tax_pct[,c(1,3)]
-assessment2018<-assessment2018 %>% left_join(tax2018, by = c("Municipalities"))
+assessment2018<-assessment2018 %>% 
+  left_join(tax2018, by = c("Municipalities"))
 names(assessment2018)[9]<-paste("tax")
 
-assessment2019<-assessment_pct  %>%  filter(Year=="2019")
+assessment2019<-assessment_pct  %>%  
+  filter(Year=="2019")
 tax2019<-tax_pct[,c(1,4)]
-assessment2019<-assessment2019 %>% left_join(tax2019, by = c("Municipalities"))
+assessment2019<-assessment2019 %>% 
+  left_join(tax2019, by = c("Municipalities"))
 names(assessment2019)[9]<-paste("tax")
 
-pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
+pct_final<-assessment2017 %>% 
+  full_join(assessment2018) %>% 
+  full_join(assessment2019)
 pct_final$tax<-as.numeric(pct_final$tax)
@@
 # tax_final is the budget
-tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))
+tax_modified <- tax_final  %>% 
+  mutate(X2019=tax_final$X2018*
+           (1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 +
+                  (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))
 
 # check match between budget and assessTotal*mill_rate
+
+```
+
+```{r}
 # get predicted tax income for all tax classes and each municipality
-assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
+assessment_aggregate[,10] <- 
+  assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
+```
+
+```{r graphr, echo= FALSE, warning=FALSE, message=FALSE}
 names(assessment_aggregate)[10] <- paste("predTaxIncome")
 
-dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
-dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
+dat_pred_income <- assessment_aggregate %>%
+  select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
+dat_pred_income <- 
+  aggregate(dat_pred_income$predTaxIncome,
+            by=list(Year=dat_pred_income$Year,
+                    Municipalities=
+                      dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
 names(dat_pred_income)[3] <- paste("TaxIncome")
@@
 names(tax_refactored)[6] <- paste("2019")
-dat_true_income <- gather(tax_refactored, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
+dat_true_income <- 
+  gather(tax_refactored, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
 dat_true_income <- dat_true_income[,-1]
@@
 
-income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
-income_budget_compare <- gather(income_budget_compare, TaxType, Amount, Budget:TaxIncome, factor_key=TRUE)
+income_budget_compare <- merge(dat_true_income,dat_pred_income,
+                               by=c("Municipalities","Year"))
+income_budget_compare <- gather(income_budget_compare, TaxType, Amount, 
+                                Budget:TaxIncome, factor_key=TRUE)
 
-compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 1: Budget and Tax Income v.s.
-Year of Vancouver")
-compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 2: Budget and Tax Income v.s Year of Burnaby")
+compare_plot1 <- income_budget_compare %>% 
+  filter(Municipalities=="Vancouver") %>% 
+  ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
+  geom_line() + 
+  ggtitle("Figure 1: Budget and Tax Income v.s. Year of Vancouver")
+
+compare_plot3 <- income_budget_compare %>% 
+  filter(Municipalities=="Burnaby") %>%
+  ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
+  geom_line() + 
+  ggtitle("Figure 2: Budget and Tax Income v.s Year of Burnaby")
 compare_plot1
@@
 
-assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
+assessment_aggregate <- merge(assessment_aggregate, tax_modified1, 
+                              by = c("AddressAssessorMunicipalityDesc","Year"))
 assessment_aggregate <- assessment_aggregate[,-c(3,11)]
@@
 
-# correlation between mill rate and budget for each municipality
-rate_tax_correlation <- assessment_aggregate %>%
-  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
-  summarise(cor(rate, tax))
-#rate_tax_correlation
-# correlation between mill rate and assessment
-rate_assess_correlation2 <- assessment_aggregate %>%
-  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
-  summarise(cor(assessTotal, rate))
 #rate_assess_correlation2
+
 names(assessment_aggregate)[1] <- paste("Municipality")
 ## Correlation for continuous variables
-assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 3: Correlations between Mill Rates and Total Assessment Values of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+assess_class <- assessment_aggregate %>% 
+  ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + 
+  geom_point() + 
+  ggtitle("Figure 3: Correlations between Mill Rates and Total Assessment Values of all municipalities") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
+  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 4: Correlations between Mill rates and Budgets of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+tax_class <- assessment_aggregate %>% 
+  ggplot(aes(x=tax,y=rate,group=Municipality,color=Municipality)) + 
+  geom_point() + 
+  ggtitle("Figure 4: Correlations between Mill rates and Budgets of all municipalities") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
+  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 5: Correlations between Mill Rates and Total Assessment Values of Land of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+land_class<-assessment_aggregate %>% 
+  ggplot(aes(x=landTotal,y=rate,group=Municipality,color=Municipality)) + 
+  geom_point() + 
+  ggtitle("Figure 5: Correlations between Mill Rates and Total Assessment Values of Land of all municipalities") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
+  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 6: Correlations between Mill Rates and Total Assessment Values of Improvement of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+improvementTotal_class<-assessment_aggregate %>% 
+  ggplot(aes(x=improvementTotal,y=rate,group=Municipality,color=Municipality)) + 
+  geom_point() + 
+  ggtitle("Figure 6: Correlations between Mill Rates and Total Assessment Values of Improvement of all municipalities") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
+  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 7: Correlations between Mill Rates and Total Number of Properties of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+propertyCount_class<-assessment_aggregate %>% 
+  ggplot(aes(x=propertyCount,y=rate,group=Municipality,color=Municipality)) + 
+  geom_point() + 
+  ggtitle("Figure 7: Correlations between Mill Rates and Total Number of Properties of all municipalities") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
+  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 ```
@@
 
-```{r anova, echo=FALSE, warning=FALSE, message=FALSE}
+```{r anova, echo=TRUE, warning=FALSE, message=FALSE}
 ## Correlation for categorical variables
@@
 
-```{r full, echo=FALSE, warning=FALSE, message=FALSE}
-linear_full<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
+```{r full, echo=TRUE, warning=FALSE, message=FALSE}
+## Full Linear model
+linear_full<-lm(rate~factor(Municipality)+factor(Year)+
+                  factor(TaxClassCode)+assessTotal+landTotal+
+                  improvementTotal+propertyCount+tax,data=assessment_aggregate)
 # summary(linear_full)
@@
 ```
-```{r reduced, echo=FALSE, warning=FALSE, message=FALSE}
-reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
+
+```{r reduced, echo=TRUE, warning=FALSE, message=FALSE}
+## Reduced Linear model
+reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+
+              factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
 #summary(reduced)
@@
 
-```{r Ridge, echo=FALSE, warning=FALSE, message=FALSE}
+```{r Ridge, echo=TRUE, warning=FALSE, message=FALSE}
+## Ridge
 library(glmnet)
@@
 # build x matrix
-x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)
+x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,
+         assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,
+         assessment_aggregate$propertyCount,assessment_aggregate$tax)
 
@@
 
-```{r LASSO, echo=FALSE, warning=FALSE, message=FALSE}
+```{r LASSO, echo=TRUE, warning=FALSE, message=FALSE}
+# LASSO
 # Setting alpha = 1 implements lasso regression
@@
 
-```{r elastic_net, echo=FALSE, warning=FALSE, message=FALSE}
+```{r elastic_net, echo=TRUE, warning=FALSE, message=FALSE}
+# Elastic Net
 library(caret)
@@
 elastic_net = train(
-  rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
+  rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+
+    landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
   method = "glmnet",
@@
 
-```{r Prediction power, echo=FALSE, warning=FALSE, message=FALSE}
+```{r Prediction power, echo=TRUE, warning=FALSE, message=FALSE}
+# Prediction power
 set.seed(450)
@@
 y<-test[,c(8)]
-linear_1<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
+linear_1<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+
+               assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
 resid<-predict(linear_1,newdata = newx) - y
@@
 # Reduced model
-linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal,data=train)
+linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+
+               assessTotal+landTotal,data=train)
 resid<-predict(linear_2,newdata = newx) - y
@@
 set.seed(450)
-lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
+lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, 
+                       standardize = TRUE, nfolds = 10)
 
@@
 
-lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
+lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, 
+                      standardize = TRUE)
 
@@
 # Ridge
-ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
+ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, 
+                   family = 'gaussian', lambda = lambdas)
 set.seed(450)
@@
 elastic_net = train(
- rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
+ rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+
+   landTotal+improvementTotal+propertyCount+tax, 
  data = assessment_aggregate[train_ind,],
@@
 
-elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.06549203)
+elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, 
+                     family = 'gaussian', lambda =  0.06549203)
 predictions_test <- predict(elastic_reg, newx = x_test)
@@
 ## Conclusions
-Among all five models, the full linear regression model is the most ideal prediction model since it make accurate prediction without losing information.
+
+Among all five models, the full linear regression model is the most ideal prediction model since it make accurate prediction without losing much information.
 There are some aspects that we could improve/implemented in the future:
@@
 Since it makes up large portion of the market, we would like to further investigate Vancouver as a special case.
-We are also interested in the effect of different tax classes in predicting mill rate, since the P-value of ANOVA between tax class and mill rate was extremely small.
+Moreover, we are also interested in the effect of different tax classes in predicting mill rate, since the P-value of ANOVA between tax class and mill rate was extremely small.
 
@@
 
-### Missing Value:
+### Missing Value and Imputation:
 
-There are 1801 missing values in mill rate (TaxClassTaxRate). Since this is the predictor, we decided to impute these missing values
-Based on client information, all properties in the same region, classcode, and year should have a unique class rate
+Of 700,000 data entries, there are 1801 missing values in mill rate. Since this is the predictor, we decided to impute these missing values.
+
+Based on client information, all properties in the same region, classcode, and year should have a unique class rate. Therefore, we followed the below procedure to impute mill rate.
 
@@
 
-#### In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:
+#### No mill rate in whole group:
+
+All entries' mill rates in some groups are missing; therefore all entries in these groups were removed. Here is the list of the groups:
 

### 2020-03-03 doc/RealEstate_Xuechun_Lu.Rmd
Add files via upload

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+# knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
 ```
@@
 
-dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
-dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
+dat_pred_income <- assessment_aggregate %>% select(Year,
+                                                   AddressAssessorMunicipalityDesc,
+                                                   TaxClassCode,
+                                                   predTaxIncome)
+dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,
+                             by=list(Year=dat_pred_income$Year,
+                                     Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),
+                             FUN=sum)
 names(dat_pred_income)[3] <- paste("predTaxIncome")
@@
 income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
-income_budget_compare <- gather(income_budget_compare, TaxType, Amount, TaxAmount:predTaxIncome, factor_key=TRUE)
+income_budget_compare <- gather(income_budget_compare, 
+                                TaxType, 
+                                Amount, 
+                                TaxAmount:predTaxIncome, 
+                                factor_key=TRUE)
 
-compare_full_plot <- income_budget_compare %>% filter(Municipalities!="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=Municipalities)) + geom_line()
+compare_full_plot <- income_budget_compare %>% filter(Municipalities!="Vancouver") %>% 
+  ggplot(
+  aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),
+      color=Municipalities)) + 
+  geom_line()
 
-compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Vancouver")
-compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Surrey") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Surrey")
-compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax or Burnaby")
-compare_plot4 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Richmond")
+compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>%
+  ggplot(
+    aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),
+        color=TaxType)) + 
+  geom_line() + 
+  ggtitle("budget v.s tax for Vancouver")
+
+compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Surrey") %>% 
+  ggplot(
+  aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),
+      color=TaxType)) + 
+  geom_line() + 
+  ggtitle("budget v.s tax for Surrey")
+
+compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% 
+  ggplot(
+  aes(x=Year,y=Amount,
+      group=interaction(Municipalities,TaxType),
+      color=TaxType)) +
+  geom_line() + 
+  ggtitle("budget v.s tax or Burnaby")
+
+compare_plot4 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% 
+  ggplot(
+  aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),
+      color=TaxType)) + 
+  geom_line() + 
+  ggtitle("budget v.s tax for Richmond")
+
 multiplot(compare_plot1,compare_plot2,compare_plot3,compare_plot4,cols=2)
@@
 
-assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
+assessment_aggregate <- merge(assessment_aggregate, 
+                              tax_modified1, 
+                              by = c("AddressAssessorMunicipalityDesc","Year"))
 assessment_aggregate <- assessment_aggregate[,-c(3,11)]
@@
 ## Correlation for continuous variables
-assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and assess - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")
+assess_class <- assessment_aggregate %>% 
+  ggplot(
+  aes(x=assessTotal,
+      y=rate,
+      group=AddressAssessorMunicipalityDesc,
+      color=AddressAssessorMunicipalityDesc)) + 
+  geom_point() + 
+  ggtitle("mill rate and assessTotal") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
+  theme(legend.position = "none")
 
-tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")
+tax_class <- assessment_aggregate %>% 
+  ggplot(
+  aes(x=tax,
+      y=rate,
+      group=AddressAssessorMunicipalityDesc,
+      color=AddressAssessorMunicipalityDesc)) + 
+  geom_point() + 
+  ggtitle("mill rate and budget") +
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
+  theme(legend.position = "none")
 
-land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and landTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")
+land_class<-assessment_aggregate %>% ggplot(
+  aes(x=landTotal,
+      y=rate,
+      group=AddressAssessorMunicipalityDesc,
+      color=AddressAssessorMunicipalityDesc)) + 
+  geom_point() + 
+  ggtitle("mill rate and landTotal") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
+  theme(legend.position = "none")
 
-improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and improvementTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")
+improvementTotal_class<-assessment_aggregate %>% 
+  ggplot(
+  aes(x=improvementTotal,
+      y=rate,
+      group=AddressAssessorMunicipalityDesc,
+      color=AddressAssessorMunicipalityDesc)) + 
+  geom_point() + 
+  ggtitle("mill rate and improvementTotal") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
+  theme(legend.position = "none")
 
-propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and propertyCount - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")
+propertyCount_class<-assessment_aggregate %>% 
+  ggplot(
+  aes(x=propertyCount,
+      y=rate,
+      group=AddressAssessorMunicipalityDesc,
+      color=AddressAssessorMunicipalityDesc)) + 
+  geom_point() + 
+  ggtitle("mill rate and propertyCount") + 
+  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
+  theme(legend.position = "none")
 
-multiplot(assess_class,
+multiplot(
+  assess_class,
   tax_class,
-land_class,improvementTotal_class,
-          propertyCount_class, cols=2)
+  land_class,
+  improvementTotal_class,
+  propertyCount_class, 
+  cols=2)
 
@@
 
-Plot 2. scatterplot with regression line for mill rate against assessTotal, tax, landTotal, improvementTotal and propertyCount.
+Plot 2. scatterplot with regression line for mill rate against assessTotal, municial budget, landTotal, improvementTotal and propertyCount.
 
@@
 # full linear model
-linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
+linear_full<-lm(rate~
+                  factor(AddressAssessorMunicipalityDesc)+
+                  factor(Year)+
+                  factor(TaxClassCode)+
+                  assessTotal+
+                  landTotal+
+                  improvementTotal+
+                  propertyCount+
+                  tax,
+                data=assessment_aggregate)
+
 summary(linear_full)
@@
 # reduced model
-reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal, data=assessment_aggregate)
+reduced<-lm(rate~
+              factor(Year)+
+              factor(TaxClassCode)+
+              factor(AddressAssessorMunicipalityDesc)+
+              assessTotal+
+              landTotal, 
+            data=assessment_aggregate)
 summary(reduced)
@@
 ```
+
 ```{r, echo=F}
@@
 # build x matrix
-x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)
+x<-cbind(dummy_municipal,
+         dummy_year,
+         dummy_taxclass,
+         assessment_aggregate$assessTotal,
+         assessment_aggregate$landTotal,
+         assessment_aggregate$improvementTotal,
+         assessment_aggregate$propertyCount,
+         assessment_aggregate$tax)
+
 y<-assessment_aggregate$rate
@@
 elastic_net = train(
-  rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
+  rate~
+    factor(AddressAssessorMunicipalityDesc)+
+    factor(Year)+
+    factor(TaxClassCode)+
+    assessTotal+
+    landTotal+
+    improvementTotal+
+    propertyCount, 
+  data = assessment_aggregate,
   method = "glmnet",
@@
 y<-test[,c(8)]
-linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
+linear_1<-lm(rate~
+               factor(AddressAssessorMunicipalityDesc)+
+               factor(Year)+
+               factor(TaxClassCode)+
+               assessTotal+
+               landTotal+
+               improvementTotal+
+               propertyCount+
+               tax,
+             data=train)
 resid<-predict(linear_1,newdata = newx) - y
-row1<-c("OLR full", round(sqrt(sum(resid^2)/nrow(test)),4))
 
 # reduced model
-linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal,data=train)
+linear_2<-lm(rate~
+               factor(Year)+
+               factor(TaxClassCode)+
+               factor(AddressAssessorMunicipalityDesc)+
+               assessTotal+
+               landTotal,
+             data=train)
+
 resid<-predict(linear_2,newdata = newx) - y
-row2<-c("OLR reduced",round(sqrt(sum(resid^2)/nrow(test)),4))
 ```
@@
 eval_results(y_test, predictions_test)
-row4<-c("Lasso",round(2.528047,4))
 
@@
 eval_results(y_test, predictions_test)
-row3<-c("Ridge",round(2.567499,4))
 
@@
 elastic_net = train(
- rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
+ rate~
+   factor(AddressAssessorMunicipalityDesc)+
+   factor(Year)+
+   factor(TaxClassCode)+
+   assessTotal+landTotal+
+   improvementTotal+
+   propertyCount+
+   tax, 
  data = assessment_aggregate[train_ind,],
@@
 eval_results(y_test, predictions_test)
-row5<-c("Elastic Net",round(2.54861,4))
 
@@
 colsname<-c("Model", "MSPE")
+row1<-c("OLR full", round(sqrt(sum(resid^2)/nrow(test)),4))
+row2<-c("OLR reduced",round(sqrt(sum(resid^2)/nrow(test)),4))
+row3<-c("Ridge",round(2.567499,4))
+row4<-c("Lasso",round(2.528047,4))
+row5<-c("Elastic Net",round(2.54861,4))
 table<-rbind(row1,row2,row3,row4,row5)
@@
 
-The OLR shows significant variables related to mill rates are Year, Tax Class, Municipality, Assessment and Land Total
+The OLR shows significant variables related to mill rates are Year, Tax Class, Municipality, Assessment and Land Total.
 
@@
 However, this good fit is likely to be a result of over-fitness since for each municipality and tax class, we only have 4 consecutive data points from 2016 to 2019.
+
+In the current design, prediction power is measured by MSPE on one trial of testing set, which might cause a large variance of MSPE and thus an inaccurate prediction power.
+In the next stage, in order to get a more accurate prediction power, we will perform k-fold cross validation to decrease the variance of MSPE.
 
@@
 
-There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values
-Based on client information, all properties in the same region, classcode, and year should have a unique class rate
+1. Mill Rate
 
-- For entries with mill rate, we aggregated them into groups by region + classcode + year.
+There are 1801 missing values in mill rate(TaxClassTaxRate).
+Since properties belonging to the same group (same year, region and tax class) should share a common mill rate, missing values were imputed by the common mill rate of properties in the same group.
 
-- For entries without mill rate, we found the group they belong to and assign them mill rate in that group.
-
-Here is some exceptions found:
-
-Some groups' mill rate is not unique:
+However, some groups' mill rates are not unique or missing entirely. The treatments are as below:
 
@@
 2019 | Maple Ridge Rural | 05 | 38
+
+2. Municipal Budget
+
+We used total tax collected in each area to approximate municipal budget.
+However, the governement website only has data available up to 2018; total tax collected in 2019 is missing.
+In this study, tax collected in 2019 is imputed by average value of tax collected in the previous three years.

### 2020-03-03 doc/RealEstate_Yuting_Wen.Rmd
added report

---
title: 'STAT 450 Project: Real Estate'
author: "Yuting Wen, 44625168"
date: "3/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r process, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
library(readxl)

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {
    print(plots[[1]])
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

# all municipalities
municipality.list = c(
  "Burnaby", 
  "Coquitlam", 
  "Delta", 
  "Langley - City", 
  "Langley - Township",
  "Maple Ridge",
  "Maple Ridge Rural", 
  "North Vancouver - City",
  "North Vancouver - Dist",
  "Pitt Meadows", 
  "Port Coquitlam", 
  "Port Moody", 
  "Richmond", 
  "Surrey", 
  "Vancouver", 
  "White Rock", 
  "West Vancouver", 
  "Bowen Island", 
  "Anmore", 
  "Belcarra",
  "Lions Bay")

# budget and assessment data import and process
tax_pct<-read_csv(here("data","tax_pct.csv"))
assessment_pct<-read_csv(here("data","assessment_pct.csv"))
# omit 2016
assessment_pct<-na.omit(assessment_pct)
assessment_pct<-assessment_pct[,-c(1)]
names(assessment_pct)[2]<-paste("Municipalities")
tax_pct<-tax_pct[,-c(1)]
#head(tax_pct)

assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
assessment_final <- read.csv(here("data","assessment_final.csv"))
assessment_final <- assessment_final[,-1]
assessment_final <- assessment_final[complete.cases(assessment_final),]
assessment_final <- assessment_final[,-4]

# Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge Rural and Maple Ridge with the same tax.

# 2019 tax pct is missing, we imputed 2019 tax by the average of previous tax pct
tax_pct[,4]<-(tax_pct$pct_2017+tax_pct$pct_2018)/2
#tax_pct
names(tax_pct)[4]<-paste("pct_2019")
tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567,5.945346))

# aggregate tax and assessment
assessment2017<-assessment_pct  %>%  filter(Year=="2017")
tax2017<-tax_pct[,1:2]
assessment2017<-assessment2017 %>% left_join(tax2017, by = c("Municipalities"))
names(assessment2017)[9]<-paste("tax")

assessment2018<-assessment_pct  %>%  filter(Year=="2018")
tax2018<-tax_pct[,c(1,3)]
assessment2018<-assessment2018 %>% left_join(tax2018, by = c("Municipalities"))
names(assessment2018)[9]<-paste("tax")

assessment2019<-assessment_pct  %>%  filter(Year=="2019")
tax2019<-tax_pct[,c(1,4)]
assessment2019<-assessment2019 %>% left_join(tax2019, by = c("Municipalities"))
names(assessment2019)[9]<-paste("tax")

pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
pct_final$tax<-as.numeric(pct_final$tax)
write.csv(pct_final, here("data","data_final.csv"))

# 2019 tax is missing, we imputed 2019 tax using previous values
tax_final <- read.csv(here("data","tax_final.csv"))
#head(tax_final)

# tax_final is the budget
tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))

# check match between budget and assessTotal*mill_rate
# get predicted tax income for all tax classes and each municipality
assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
names(assessment_aggregate)[10] <- paste("predTaxIncome")

dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
names(dat_pred_income)[3] <- paste("TaxIncome")

tax_refactored <- tax_modified
names(tax_refactored)[3] <- paste("2016")
names(tax_refactored)[4] <- paste("2017")
names(tax_refactored)[5] <- paste("2018")
names(tax_refactored)[6] <- paste("2019")
dat_true_income <- gather(tax_refactored, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
dat_true_income <- dat_true_income[,-1]
names(dat_true_income)[3] <- paste("Budget")

income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
income_budget_compare <- gather(income_budget_compare, TaxType, Amount, Budget:TaxIncome, factor_key=TRUE)
```

## Summary

Our client is a real estate data analyst in Metro Vancouver who analyzes the market trend to make a market prediction for his company and future client. He is interested in a good prediction model that can accurately predict future mill rates, given the past data like the past mill rates and the past assessment values, in Metro Vancouver. Data cleaning, exploratory data analysis, and linear models with different kinds of regulazations were used to analyze the relationship between mill rate and other factors and predict future mill rates. Data cleaning is performed to aggregate our data into summary statistics and also reduce the number of data entries. Exploratory analysis has shown that there is a strong correlation between mill rate and municipal budget and strong correlation between mill rate and assessment in some municipalities. The results of exploratory analysis have also shown that there is no clear difference among full linear Regression model with all features (Year, TaxClassCode, Municipalities, assessTotal_pct, tax (budget), landTotal, improvementTotal, numProperties) and reduced linear regression models (Ridge, Lasso and Elastic net).

## Introduction

Our goal is to accurately predict the mill rate in 2020 based on past data from 2016 to 2019, for each of the following 21 sub-regions in the Greater Vancouver area:

- Burnaby 
- Coquitlam
- Delta
- Langley City, 
- Langley Township 
- Maple Ridge
- Maple Ridge Rural
- North Vancouver City
- North Vancouver Dist
- Pitt Meadows
- Port Coquitlam
- Port Moody
- Richmond
- Surrey
- Vancouver
- White Rock
- West Vancouver
- Bowen Island
- Anmore
- Belcarra
- Lions Bay

Correlations between mill rates, assessment values, municipal budgets, land assessment values, improvement assessment values, number of properties, tax class codes, municipalities, and year are performed to see the significance of these factors in our predictive model.

A full linear model with all features, a few reduced linear models with regularzations like the LASSO, Ridge Regression are used to train our predictive model. Once we have finished fitting those models, we will use the mean squared training error to evaluate the performance of our model.

## Data Description

To address these problems, we used past assessment data, which is provided by our client. We obtained information about commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019. There are approximately 2.4 million data entries across these 4 years. We selected variables that we believe were relevant to the property tax prediction. These selected variables are:

- Mill Rates (2016-2019)
- Assessment Total (2016-2019)
- Land Assessment Total (2016-2019)
- Improvement Assessment Total (2016-2019)
- Number of Properties (2016-2019)
- Tax Class Code (01, 05, 06)
- Area Code (21 Municipalities)
- Year (2016-2019)

We also collected external data from the Government of British Columbia which could be associated with the tax rate:

- Municipal Budget of Cities in Metro Vancouver (Taxes Imposed & Collected, Schedule 706)

The listing data was cleaned to reduce its dimension. We decided to calculate the sum of total assessment value, total land assessment value, total improvement assessment value, budget, and number of properties for each combination of municipality and their corresponding tax class. The mill rate is unique for each combination of municipality and tax class, we have merged this information into our data for each combination of municipality and tax class.

## Methods

### Exploratory Analysis

We did some exploratory analysis to visualize the main characteristics of our dataset and found relationships between mill rate and other factors, before we made any prediction on the future mill rate of Metro Vancouver’s real estate market.

We assume that the government aims to match its budget and its income by adjusting the mill rates, so we have plotted a graph of tax income for three different tax classes (01, 05, and 06) vs. budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct. Vancouver, Richmond, and Burnaby were chosen to illustrate such relationship as they are the most representing city in metro vancouver.

We also made scatterplot of mill rate vs. assessment, mil rate vs. municipal budget, mill rate vs. landTotal (total land assessment value), mill rate vs. improvementTotal (total improvement assessment value), and mill rate vs. numOfProperties (number of properties) to see the correlation between each of the two factors. ANOVA analysis was also performed to see the correlations between mill rate vs. tax class, mill rate vs. year and mill rate vs. municipality.

These exploratory analysis and plots helped us to find trends in data within each of its own categories.

### Measure of goodness of fit and prediction power

In this study, we first built various models then evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.

Goodness of fit is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data. R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.

Prediction power measures how well models can predict the future values. In this study, we use mean squared prediction error to compare the prediction performance across all the models. To do that, we will divide the data set into training data, used as building models, and test data, used as computing mean squared prediction error.

### Ordinary Linear Model

The explanatory variables we are interested in are Year, TaxClassCode, Municipalities, assessTotal, tax (budget), landTotal, improvementTotal, numOfProperties. We first explored the full linear model using all the available features; then the linear models with regularzations were performed since they potentially have better prediction power.

There are potential outliers in the data set, it might be helpful to refit the model without the outliers to check if the fit of the model is improved. Futher prediction models will be built here.

### Reduced Ordinary Linear Model (Variable Selection)

A list of significant variables are included in the result of the full model. In the next step, we will refit the model with only the significant variables. Goodness of fit will decline with fewer features, however, the performance of prediction power might be improved.

### Ridge, Lasso and Elastic Net

Except for ordinary linear regression, we are also interested in the performance of Ridge, Lasso, and Elastic Net models as they are also in the linear regression family, but more advanced than ordinary linear regression. In these three models, weights are assigned to features. Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively. Elastic net is a combination of Ridge and Lasso.

We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.

### Neural Network

In the current report, we are mainly focusing on predicting the mill rate using Linear Model. In the future, we will discuss with our supervisor and try to fit a neural network model if possible (R package: neuralnet or keras). Neural Network has the ability to learn and model non-linear and complex relationships. It is flexible and can be used for both classification and prediction. Since we have a large number of data, it makes Neural Network a very good candidate for our prediction model.

Report on the performance of the Neural Network will be included in the next report.

\pagebreak

### Results: Exploratory Analysis

We assume that the government aims to match its budget and its income by adjusting the mill rates, so we plotted the predicted tax income vs municipal budget through time into line-chart to visualize their relationship, to make sure our assumption is correct.

For each year from 2016 to 2019, we predicted the tax income of three class codes (01, 05, 06) in each municipality by multiplying access total and mill rate. Vancoouver, Richmond, and Burnaby were chosen because they were the three most representing municipalities in Metro Vancouver, and the plots showed that the predicted tax amount was positively related to municipal budget.

```{r budget, echo=FALSE, warning=FALSE, message=FALSE}
income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
  geom_line() + 
  ggtitle("Figure 1: Budget and Tax Income v.s. Year of Vancouver")
income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
  geom_line() + 
  ggtitle("Figure 2: Budget and Tax Income v.s Year of Richmond")
income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
  geom_line() + 
  ggtitle("Figure 3: Budget and Tax Income v.s Year of Burnaby")
```

### Correlation Analysis

We made 5 scatter plots to visualize the correlation between mill rate and 5 numerical features (assessment total, municipal budget, land assessment total, improvement assessment total, and the number of properties).

```{r cor, echo=FALSE, warning=FALSE, message=FALSE}
# aggregate tax and mill rate & assessment
colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)

tax_modified1 <- gather(tax_modified, Year, tax, '2016':'2019', factor_key = TRUE)%>%
  rename(AddressAssessorMunicipalityDesc = Municipalities)

assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
assessment_aggregate <- assessment_aggregate[,-c(3,11)]
names(assessment_aggregate)[10] <- paste("tax")

# correlation between mill rate and budget for each municipality
rate_tax_correlation <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(rate, tax))
#rate_tax_correlation
# correlation between mill rate and assessment
rate_assess_correlation2 <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(assessTotal, rate))
#rate_assess_correlation2
names(assessment_aggregate)[1] <- paste("Municipality")

## Correlation for continuous variables
assessment_aggregate %>% 
  ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 4: Correlation: Mill Rates and Total Assessment Values") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_aggregate %>% 
  ggplot(aes(x=tax,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + ggtitle("Figure 5: Correlation: Mill rates and Budgets") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_aggregate %>% 
  ggplot(aes(x=landTotal,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 6: Correlation: Mill Rates and Total Land Assessment Values") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_aggregate %>% 
  ggplot(aes(x=improvementTotal,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 7: Correlation: Mill Rates and Total Improvement Assessment Values") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_aggregate %>% 
  ggplot(aes(x=propertyCount,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 8: Correlation: Mill Rates and Total Number of Properties") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
```

As we can see:

Figure 4, 6, 7, and 8 show that mill rate is negatively correlated with total assessment values, total land assessment values, total improvement assessment values, total number of properties. Budget is not significantly correlated with mill rate as shown in Figure 5.

Outlier detection:

The plots above also show that Vancouver (the dark pink points) are significantly different from other points and should be regarded as an outlier (futher analysis needed here). Some other municipalities (including Richmond, Surrey and Burnaby) that are also distinct in the scatter plots.

### ANOVA Analysis
```{r anova, echo=FALSE, warning=FALSE, message=FALSE}
## Correlation for categorical variables
TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
#summary(TaxClassCode)

Year <- aov(rate_pct ~ factor(Year), data = pct_final)
#summary(Year)

Municipalities <- aov(rate ~ factor(Municipality), data = assessment_aggregate)
#summary(Municipalities)
```

                    Table 1: ANOVA analysis of mill rate and three categorical variables

Feature Name  | Pr(>F) | significance levels
------|--------|------
Tax Class Code | <2e-16 | ***(very significant)
Year | 4.2e-06 | **(significant)
Municipality | 0.0768 | .(not significant)

We used three 3 ANOVA tables to analyze the correlation between mill rate and each of the 3 categorical features (TaxClassCode, Year, Municipalities). The table showed that Tax Class Code and Year have a strong correlation with mill rate while the correlation between Municipality and mill rate is not significant at 5% level.

\pagebreak

## Results

### Variable Selection

                  Table 2: Significance level of all variables of full linear model

 Variable |                                 Pr(>t) | Significance Level 
 ------|------|------
(Intercept)                         |        0.000111 | ***
factor(Municipality)Belcarra        |        0.299387 |   
factor(Municipality)Bowen Island    |        0.303157 |   
factor(Municipality)Burnaby         |        0.604683 |   
factor(Municipality)Coquitlam       |        0.001594 | ** 
factor(Municipality)Delta           |        0.002579 | ** 
factor(Municipality)Langley - City  |        0.020365 | *  
factor(Municipality)Langley - Township  |    0.023030 | *  
factor(Municipality)Lions Bay           |    0.091427 | .  
factor(Municipality)Maple Ridge         |    4.18e-05 | ***
factor(Municipality)North Vancouver - City |  0.129059 |    
factor(Municipality)North Vancouver - Dist |  0.077124 | .  
factor(Municipality)Pitt Meadows        |    2.41e-07 | ***
factor(Municipality)Port Coquitlam      |    1.60e-05 | ***
factor(Municipality)Port Moody          |    1.16e-05 | ***
factor(Municipality)Richmond            |    0.894181 |   
factor(Municipality)Surrey              |    0.984723 |   
factor(Municipality)Vancouver           |    0.885521 |   
factor(Municipality)West Vancouver      |    0.001003 | ** 
factor(Municipality)White Rock          |    0.034041 | *  
factor(Year)2017                        |    0.000655 | ***
factor(Year)2018                        |    1.30e-08 | ***
factor(Year)2019                        |    4.73e-13 | ***
factor(TaxClassCode)5                   |    < 2e-16 | ***
factor(TaxClassCode)6                   |    < 2e-16 | ***
assessTotal                             |    0.013752 | *  
landTotal                               |    0.013863 | *  
improvementTotal                        |          NA |   
propertyCount                           |    0.825699 |   
tax                                     |    0.958323 |

After fitting the ordinary linear regression model with all features, the summary statistics in Table 2 showed that municipality, year, tax class code, total assessment value, and total land assess are significant, which indicated that total improvement assessment value, number of properties, and budget(tax) might not have strong impact for the linear model. Therefore we built an ordinary linear regression reduced model with only significant variables.

### Goodness of Fit

                          Table 3: Goodness of fit of all models

Model | Mutiple R-Squared | Adjusted R_Squared | MSE
------|--------|--------|--------|--------
      Ordinary Linear Regression full model | 0.8874| 0.8707| 1.9843
      Ordinary Linear Regression reduced model | 0.8874| 0.8721| 1.9845
      Ridge Regression model |  0.8868| NA| 1.9896
      LASSO Regression model | 0.8873| NA| 1.9855
      Elastic Net Regression model | 0.8625| NA| 2.2366

The table above is the goodness of fit of five models we fitted. All models have relatively high R squared and adjusted R squared values, which indicates that the above models are able to accurately fit the data. Also there is no significant differences among these models in terms of mean squared error, except for Elastic Net model which have higher MSE. We have not used cross validation to reduce model overfitting. Although the results are relatively similar and have high R_squared across all models, we believe using cross validation there will be a clearer difference between the goodness of fit of each model.

### Prediction Power

                        Table 4: Prediction Power of all models
Model | PMSE 
------|--------
      Ordinary Linear Regression full model | 2.5902
      Ordinary Linear Regression reduced model | 2.5237
      Ridge Regression model | 2.5675
      LASSO Regression model | 2.5280
      Elastic Net Regression model | 2.5486

The table above is the prediction power of five models we fitted. All models have relatively low error (PMSE), while there is also no clear difference among them. Futher analysis is needed here.

## Conclusions

From our exploratory data analysis, We have found that when comparing the correlation between mill rate against features in our models, mill rate is highly correlated with total assessment values, total land assessment values, total improvement assessment values, total number of properties, comparing with budget. Also Vancouver is an outlier in our plot. We would like to further investigate Vancouver as a special case.

We can also see that total assessment value, total land assessment value, tax class code, year and municipality are all significant in our linear regression model. We are also interested in the effect of different tax classes in predicting mill rate. Further analysis and prediction will be performed based on our hypothesis.

The linear regression models were able to accurately predict mill rate with relatively low error, but there is still a very high chance that our model is overfitted. For our next report, we are going to use cross-validation to reduce the effect of overfitting.

## References

Links to source of data:

- Schedule 706 (https://www2.gov.bc.ca/gov/content/governments/local-governments/facts-framework/statistics/statistics)

Code repository:

- Data Cleaning (https://github.com/STAT450-550/RealEstate/blob/450/src/Data_Cleaning.Rmd)

- Exploratory Data Analysis and Model Fitting (https://github.com/STAT450-550/RealEstate/blob/450/src/EDA%26mode_fitting.Rmd)

## Appendix

### Missing Value:

There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values based on client information; all properties in the same region, classcode, and year should have a unique class rate

- For entries with mill rate, we aggregated them into groups by region + classcode + year.

- For entries without mill rate, we found the group that they belong to and assign them mill rate in that group.

### Here is some exceptions found:

Some groups' mill rate is not unique:

- In Delta, properties in different neighbourhoods have slightly different mill rates. Since the variance is not significant, we take the mean as the overall mill rate in groups.

- In Vancouver, 2019, Class 01, one property's mill rate is different from others. It is regarded as an outlier.

- In Burnaby, 2019, Class 06, six properties' mill rate are different from others. They are regarded as outliers.

- In Langley, 2019, Class 06, mill rate is different between assessment type. After talking with the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.

- In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:

                          Table 4: Groups that miss mill rate

Year  | Region | Class |Number of Properties
------|--------|-------|----------
2016 | Belcarra | 06 | 9 
2016 | Lions Bay | 01 | 40 
2016 | Lions Bay | 06 | 25 
2016 | Maple Ridge Rural | 05 | 36 
2017 | Belcarra | 06 | 9 
2017 | Lions Bay | 01 | 39 
2017 | Lions Bay | 06 | 24 
2017 | Maple Ridge Rural | 05 | 36 
2018 | Maple Ridge Rural | 05 | 36 
2019 | Maple Ridge Rural | 05 | 38

### Model Fitting

#### Ordinary Linear Regression Full Model
```{r full, echo=TRUE, warning=FALSE, message=FALSE}
linear_full<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
#summary(linear_full)

library(broom)
linear_full_fit<-augment(linear_full)
mse_full <- sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
```

#### Ordinary Linear Regression Reduced Model
```{r reduced, echo=TRUE, warning=FALSE, message=FALSE}
reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
#summary(reduced)

reduced_fit<-augment(reduced)
mse_reduced <- sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate))
```

#### Ridge
```{r Ridge, echo=TRUE, warning=FALSE, message=FALSE}
library(glmnet)
library(dummies)
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$Municipality)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)

y<-assessment_aggregate$rate
lambdas <- 10^seq(2, -3, by = -.1)

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min

predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)

# Compute R^2 from true and predicted values
eval_results <- function(true, predicted) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  MSPE = sqrt(SSE/nrow(predicted))
# Model performance metrics
data.frame(
  MSPE = MSPE,
  Rsquare = R_square
)

}

predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
ridge_mse <- eval_results(y, predictions_train)
```

#### LASSO
```{r LASSO, echo=TRUE, warning=FALSE, message=FALSE}
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

# Best 
lambda_best <- lasso_reg$lambda.min;#lambda_best

lasso_model <- glmnet(x, y, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
lasso_mse <- eval_results(y, predictions_train)
```

#### Elastic Net
```{r elastic_net, echo=TRUE, warning=FALSE, message=FALSE}
library(caret)
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
```

#### Prediction Power
```{r Prediction power, echo=TRUE, warning=FALSE, message=FALSE}
set.seed(450)
train_ind<-sample(218,218-50)
train<-assessment_aggregate[train_ind,]
test<-assessment_aggregate[-train_ind,]

# Full linear model
newx<-test[,-c(8,9)]
y<-test[,c(8)]
linear_1<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
resid<-predict(linear_1,newdata = newx) - y
full_mspe <- sqrt(sum(resid^2)/nrow(test))

# Reduced model
linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal,data=train)
resid<-predict(linear_2,newdata = newx) - y
reduced_mspe <- sqrt(sum(resid^2)/nrow(test))

# Lasso
# create the whole matrix
y<-as.matrix(assessment_aggregate$rate)
#dim(x) # 165  29
#dim(y)
# creat x_train matrix and y_train
x_train<-x[train_ind,]
y_train<-y[train_ind,]
# create x_test matrix
x_test<-x[-train_ind,]
y_test<-y[-train_ind,]

# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

# Best 
lambda_best <- lasso_reg$lambda.min;#lambda_best

lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
lasso_mspe <- eval_results(y_test, predictions_test)

# Ridge
ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
#optimal_lambda
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
ridge_mspe <- eval_results(y_test, predictions_test)

# Elastic Net
#tibble::as_tibble(assessment_aggregate[train_ind,])
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
 rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
 data = assessment_aggregate[train_ind,],
  method = "glmnet",
  trControl = cv_10
)

elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.06549203)
predictions_test <- predict(elastic_reg, newx = x_test)
elastic_net_mspe <- eval_results(y_test, predictions_test)
```

### 2020-03-05 src/draft.Rmd
450 local changes

--- 2020-02-28 src/draft.Rmd
@@
 libraries
-```{r, message=F, warning=F}
+```{r}
 library(tidyverse)
@@
 library(readxl)
-```
 
-```{r}
 multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
@@
 names(tax_pct)[4]<-paste("pct_2019")
-tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567,5.945346))
-
 #head(tax_pct)
@@
 plot pct change of tax(budget)
-```{r, warning=F}
+```{r}
 # total trend for budget 
@@
 ```{r}
-colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
-head(tax_modified)
-dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
+#tax_final
+#assessment_aggregate
+
+#colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
+#head(tax_modified)
+#dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
 
 #tax.all <- full_join(assessment_aggregate, tax_modified, by = 'Municipalities') 
-
-tax_modified1 <- gather(tax_modified, Year, tax, '2016':'2019', factor_key = TRUE)%>%
-  rename(AddressAssessorMunicipalityDesc = Municipalities)
-
-assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
-
-correlation <- assessment_aggregate %>%
-  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
-  summarise(cor(rate, tax))
-
-correlation2 <- assessment_aggregate %>%
-  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
-  summarise(cor(assessTotal, rate))
+#tax.all
+#(tax.all <- tax.all[,-5])
+# correlation by Municipalities?
+#mill.budget.cor <- tax.all  %>% summarize(cor_2017 = cor(pct_2017, X2017), cor_2018 = cor(pct_2018, X2018), cor_2019 = cor(pct_2019, X2019))
+#mill.budget.cor
 
@@
 ```
-r^2 is too small.
-we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400).
-We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882.
-
-```{r}
-test<-pct_final %>% filter(Municipalities!="Langley - City") %>% 
-  filter(Municipalities!="Langley - Township")
-linear_test<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
-summary(linear_test)
-```
-
-Ridge Regression!
-reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r
-```{r}
-library(glmnet)
-library(dummies)
-dummy_year<-dummy(test$Year)
-dummy_municipal<-dummy(test$Municipalities)
-dummy_taxclass<-dummy(test$TaxClassCode)
-# build x matrix
-x<-cbind(dummy_year,dummy_municipal)
-x<-cbind(x,dummy_taxclass)
-x<-cbind(x,as.numeric(test$assessTotal_pct))
-x<-cbind(x,as.numeric(test$tax))
-y<-test$rate_pct
-lambdas <- 10^seq(2, -3, by = -.1)
-dim(x)
-colnames(x)[26]<-paste("assessment")
-colnames(x)[27]<-paste("tax")
-colnames(x)
-
-lambdas <- 10^seq(2, -3, by = -.1)
-ridge_reg = glmnet(x, y, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
-set.seed(450)
-cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas, nfolds=10)
-optimal_lambda <- cv_ridge$lambda.min
-optimal_lambda
-predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
-
-# Compute R^2 from true and predicted values
-eval_results <- function(true, predicted) {
-  SSE <- sum((predicted - true)^2)
-  SST <- sum((true - mean(true))^2)
-  R_square <- 1 - SSE / SST
-  RMSE = sqrt(SSE/nrow(predicted))
-# Model performance metrics
-data.frame(
-  RMSE = RMSE,
-  Rsquare = R_square
-)
-
-}
-
-predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
-eval_results(y, predictions_train)
-
-```
-RMSE=7.159649
-Rsquare=0.2136659
-Ridge regression doesn't seem to work well.
-
-Lasso 
-```{r}
-lambdas <- 10^seq(2, -3, by = -.1)
-
-# Setting alpha = 1 implements lasso regression
-set.seed(450)
-lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
-
-# Best 
-lambda_best <- lasso_reg$lambda.min;lambda_best
-
-lasso_model <- glmnet(x, y, alpha = 1, lambda = lambda_best, standardize = TRUE)
-
-predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
-eval_results(y, predictions_train)
-```
-RMSE = 7.221299, Rsquare=0.2000649
-
-Elastic net: 
-reference: https://daviddalpiaz.github.io/r4sl/elastic-net.html
-```{r}
-library(caret)
-tibble::as_tibble(test)
-cv_10 = trainControl(method = "cv", number = 10)
-ridge = train(
-  rate_pct ~ Year + Municipalities + TaxClassCode + assessTotal_pct + tax, data = test,
-  method = "glmnet",
-  trControl = cv_10
-)
-ridge
-```
-apha = 1 which implies lasso! refer to lasso.

### 2020-03-05 doc/RealEstate_Peter_Han.Rmd

--- 2020-03-03 doc/RealEstate_Yuting_Wen.Rmd
@@
 title: 'STAT 450 Project: Real Estate'
-author: "Yuting Wen, 44625168"
-date: "3/2/2020"
-output: pdf_document
+author: "Peter Han, 14912166"
+date: "3/3/2020"
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-```{r process, echo=FALSE, warning=FALSE, message=FALSE}
+## Summary
+
+Our client is a real estate analyst who analyzes the market trend and make market prediction for his company and future client.
+Data cleaning, exploratory data analysis were used in this project to analyze the relationship between mill rate and other factors.
+Data cleaning is performed to aggregate our data into summary statistics and also reduce the number of data entries.
+Exploratory analysis has shown there is a linear trend in mill rates over the years; it has also shown there is a strong correlation between mill rate and municipal budget and strong correlation between mill rate and assessment in some municipalities.
+
+## Introduction
+
+Our client is only interested in accurate predictive models.The goal is to predict the property tax in 2020, for each of the following 21 sub-regions in the Metro Vancouver area:
+
+- Burnaby
+- Coquitlam
+- Delta
+- Langley - City
+- Langley - Township
+- Maple Ridge
+- Maple Ridge Rural 
+- North Vancouver - City
+- North Vancouver - Dist
+- Pitt Meadows
+- Port Coquitlam
+- Port Moody
+- Richmond
+- Surrey
+- Vancouver
+- White Rock
+- West Vancouver
+- Bowen Island
+- Anmore
+- Belcarra
+- Lions Bay.
+
+Correlations between mil rate, assessment value (Land and Imporvement) and municipal budget were computed to see the significance of these factors in our predictive model.
+An ANOVA analysis had also been carried out to see the significance of categorical factor such as Year, Municipality and Tax Class Code.
+A Simple linear model, a linear model with the interaction term, LASSO, Ridge Regression and Elastic Net were used to train our predictive model.
+Once we have finished the above tasks, we will use the mean square test error to evaluate the performance of our model.
+
+## Data Description
+
+For exploratory data analysis and building predictive model, we used past property assessment data, which is provided by our client.
+It contains information about some commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019.
+We selected variables that we believe were relevant to the property tax prediction.
+These selected variables are:
+
+- Mill Rates (2016-2019)
+- Assessment (2016-2019)
+- Tax Class Code (01, 05, 06)
+- Area Code (Municipality)
+- Year
+- Land Total (2016 - 2019)
+- Improvement Total (2016 - 2019)
+- Number of properties
+
+As the client suggested, we have also collected external data from the Government of British Columbia which could be associated with the tax rate:
+
+- Municipal Budget of Cities in Metro Vancouver (Taxes Imposed & Collected, Schedule 706)
+
+We decided to calculate the sum of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, and Budget for each combination of municipality and the tax class to reduce the dimension of our data.
+The mill rate is unique for each combination of municipality and tax class, so we have added this information into our data for each municipality and its tax classes.
+
+## Methods
+
+### Exploratory Analysis
+
+Before we made any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs.
+Assessment, Mill Rate vs.
+Municipal Budget, Mill Rate vs.
+Land Total, Mill Rate vs.
+Imrpovement Total and Mill Rate vs.
+Number of Properties.
+
+The municipal government wants their tax income to match their annual spending in order to balance their bedget.
+Given this piece of information, we want to find and model the relationship between municipal tax income and municipal budget.
+We plotted a graph of tax income for three different tax classes (01, 05, and 06) vs.
+budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct.
+We chose to illustrate such relationship using scatter plot below for Vancouver, Richmond and Burnaby as they are the most representing city in Metro Vancouver.
+
+Mill rate is mainly affected by assessment and municipal budget, so we have plotted scatter plot of mill rate vs.
+assessment and mil rate vs.
+municipal budget to see the correlation between each pair of the two factors.
+ANOVA analysis is also performed to see the correlation between mill rate vs.
+tax class, mill rate vs.
+year and mill rate vs.
+municipality.
+
+### Measure of goodness of fit and prediction power
+
+In this study, we built a few linear models and evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.
+The definition of goodness of fit and prediction power is given below.
+
+ - **Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data.
+R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.
+
+ - **Prediction power** measures how well models can predict the future values.
+We use mean squared prediction power to compare the prediction performance across all of our fitted linear predictive models.
+To do that, we will divide the data set into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.
+
+### Linear model
+
+For our linear model, Year, TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties and tax (budget) are considered.
+We first explore the full linear model using all the available features.
+From our exploratory data analysis, we have found Vancouver, Richmond, Surrey and Burnaby are outliers.
+Further prediction model will be built based on these findings.
+We are also intesterd in how different tax class can affect our predictive model, so models based on different tax classes will also be considered in the future.
+For our reduced model, we build our linear predictor using only the significant variables from the full model (See table 4).
+The reduced model's performence in terms of goodness of fit will decrease with fewer features, however, the performance of prediction power might be improved.
+
+### Ridge, Lasso and Elastic Net
+
+Other than simple linear regression, we are also interested in the performance of Ridge Regression, Lasso, and Elastic Net.
+They are more advanced than simple linear regression as they reduce the variance of the model.
+In these three models, weights are assigned to features.
+Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively.
+Elastic net is a combination of Ridge and Lasso.
+
+We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.
+
+### Neural Network
+
+In the current report, we are mainly focusing on predicting the mill rate using Linear Model.
+In the future, we will discuss with our supervisor and try to fit a neural network model if possible (R package: neuralnet or keras).
+Neural Network has the ability to learn and model non-linear and complex relationships.
+It is flexible and can be used for both classification and prediction.
+Since we have a large number of data, it makes Neural Network a very good candidate for our predictive model.
+
+Report on the performance of the Neural Network will be included in the next report.
+\pagebreak
+
+## Results
+
+###Exploratory Analysis
+
+We assume that the government aims to match its budget and its income by adjusting the mill rates
+
+```{r budget, echo=FALSE, warning=FALSE, message=FALSE}
 library(tidyverse)
@@
 income_budget_compare <- gather(income_budget_compare, TaxType, Amount, Budget:TaxIncome, factor_key=TRUE)
+
+compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 1: Budget and Tax Income v.s.
+Year of Vancouver")
+compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 2: Budget and Tax Income v.s.
+Year of Richmond")
+compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 3: Budget and Tax Income v.s Year of Burnaby")
+compare_plot1
+compare_plot2
+compare_plot3
 ```
 
-## Summary
-
-Our client is a real estate data analyst in Metro Vancouver who analyzes the market trend to make a market prediction for his company and future client.
-He is interested in a good prediction model that can accurately predict future mill rates, given the past data like the past mill rates and the past assessment values, in Metro Vancouver.
-Data cleaning, exploratory data analysis, and linear models with different kinds of regulazations were used to analyze the relationship between mill rate and other factors and predict future mill rates.
-Data cleaning is performed to aggregate our data into summary statistics and also reduce the number of data entries.
-Exploratory analysis has shown that there is a strong correlation between mill rate and municipal budget and strong correlation between mill rate and assessment in some municipalities.
-The results of exploratory analysis have also shown that there is no clear difference among full linear Regression model with all features (Year, TaxClassCode, Municipalities, assessTotal_pct, tax (budget), landTotal, improvementTotal, numProperties) and reduced linear regression models (Ridge, Lasso and Elastic net).
-
-## Introduction
-
-Our goal is to accurately predict the mill rate in 2020 based on past data from 2016 to 2019, for each of the following 21 sub-regions in the Greater Vancouver area:
-
-- Burnaby 
-- Coquitlam
-- Delta
-- Langley City, 
-- Langley Township 
-- Maple Ridge
-- Maple Ridge Rural
-- North Vancouver City
-- North Vancouver Dist
-- Pitt Meadows
-- Port Coquitlam
-- Port Moody
-- Richmond
-- Surrey
-- Vancouver
-- White Rock
-- West Vancouver
-- Bowen Island
-- Anmore
-- Belcarra
-- Lions Bay
-
-Correlations between mill rates, assessment values, municipal budgets, land assessment values, improvement assessment values, number of properties, tax class codes, municipalities, and year are performed to see the significance of these factors in our predictive model.
-
-A full linear model with all features, a few reduced linear models with regularzations like the LASSO, Ridge Regression are used to train our predictive model.
-Once we have finished fitting those models, we will use the mean squared training error to evaluate the performance of our model.
-
-## Data Description
-
-To address these problems, we used past assessment data, which is provided by our client.
-We obtained information about commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019.
-There are approximately 2.4 million data entries across these 4 years.
-We selected variables that we believe were relevant to the property tax prediction.
-These selected variables are:
-
-- Mill Rates (2016-2019)
-- Assessment Total (2016-2019)
-- Land Assessment Total (2016-2019)
-- Improvement Assessment Total (2016-2019)
-- Number of Properties (2016-2019)
-- Tax Class Code (01, 05, 06)
-- Area Code (21 Municipalities)
-- Year (2016-2019)
-
-We also collected external data from the Government of British Columbia which could be associated with the tax rate:
-
-- Municipal Budget of Cities in Metro Vancouver (Taxes Imposed & Collected, Schedule 706)
-
-The listing data was cleaned to reduce its dimension.
-We decided to calculate the sum of total assessment value, total land assessment value, total improvement assessment value, budget, and number of properties for each combination of municipality and their corresponding tax class.
-The mill rate is unique for each combination of municipality and tax class, we have merged this information into our data for each combination of municipality and tax class.
-
-## Methods
-
-### Exploratory Analysis
-
-We did some exploratory analysis to visualize the main characteristics of our dataset and found relationships between mill rate and other factors, before we made any prediction on the future mill rate of Metro Vancouver’s real estate market.
-
-We assume that the government aims to match its budget and its income by adjusting the mill rates, so we have plotted a graph of tax income for three different tax classes (01, 05, and 06) vs.
-budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct.
-Vancouver, Richmond, and Burnaby were chosen to illustrate such relationship as they are the most representing city in metro vancouver.
-
-We also made scatterplot of mill rate vs.
-assessment, mil rate vs.
-municipal budget, mill rate vs.
-landTotal (total land assessment value), mill rate vs.
-improvementTotal (total improvement assessment value), and mill rate vs.
-numOfProperties (number of properties) to see the correlation between each of the two factors.
-ANOVA analysis was also performed to see the correlations between mill rate vs.
-tax class, mill rate vs.
-year and mill rate vs.
-municipality.
-
-These exploratory analysis and plots helped us to find trends in data within each of its own categories.
-
-### Measure of goodness of fit and prediction power
-
-In this study, we first built various models then evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.
-
-Goodness of fit is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data.
-R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.
-
-Prediction power measures how well models can predict the future values.
-In this study, we use mean squared prediction error to compare the prediction performance across all the models.
-To do that, we will divide the data set into training data, used as building models, and test data, used as computing mean squared prediction error.
-
-### Ordinary Linear Model
-
-The explanatory variables we are interested in are Year, TaxClassCode, Municipalities, assessTotal, tax (budget), landTotal, improvementTotal, numOfProperties.
-We first explored the full linear model using all the available features; then the linear models with regularzations were performed since they potentially have better prediction power.
-
-There are potential outliers in the data set, it might be helpful to refit the model without the outliers to check if the fit of the model is improved. Futher prediction models will be built here.
-
-### Reduced Ordinary Linear Model (Variable Selection)
-
-A list of significant variables are included in the result of the full model.
-In the next step, we will refit the model with only the significant variables.
-Goodness of fit will decline with fewer features, however, the performance of prediction power might be improved.
-
-### Ridge, Lasso and Elastic Net
-
-Except for ordinary linear regression, we are also interested in the performance of Ridge, Lasso, and Elastic Net models as they are also in the linear regression family, but more advanced than ordinary linear regression.
-In these three models, weights are assigned to features.
-Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively.
-Elastic net is a combination of Ridge and Lasso.
-
-We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.
-
-### Neural Network
-
-In the current report, we are mainly focusing on predicting the mill rate using Linear Model.
-In the future, we will discuss with our supervisor and try to fit a neural network model if possible (R package: neuralnet or keras).
-Neural Network has the ability to learn and model non-linear and complex relationships.
-It is flexible and can be used for both classification and prediction.
-Since we have a large number of data, it makes Neural Network a very good candidate for our prediction model.
-
-Report on the performance of the Neural Network will be included in the next report.
-
-\pagebreak
-
-### Results: Exploratory Analysis
-
-We assume that the government aims to match its budget and its income by adjusting the mill rates, so we plotted the predicted tax income vs municipal budget through time into line-chart to visualize their relationship, to make sure our assumption is correct.
-
-For each year from 2016 to 2019, we predicted the tax income of three class codes (01, 05, 06) in each municipality by multiplying access total and mill rate.
-Vancoouver, Richmond, and Burnaby were chosen because they were the three most representing municipalities in Metro Vancouver, and the plots showed that the predicted tax amount was positively related to municipal budget.
-
-```{r budget, echo=FALSE, warning=FALSE, message=FALSE}
-income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
-  geom_line() + 
-  ggtitle("Figure 1: Budget and Tax Income v.s. Year of Vancouver")
-income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
-  geom_line() + 
-  ggtitle("Figure 2: Budget and Tax Income v.s Year of Richmond")
-income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
-  geom_line() + 
-  ggtitle("Figure 3: Budget and Tax Income v.s Year of Burnaby")
-```
-
-### Correlation Analysis
-
-We made 5 scatter plots to visualize the correlation between mill rate and 5 numerical features (assessment total, municipal budget, land assessment total, improvement assessment total, and the number of properties).
+We calculated the total tax income of each municipality by summing the product of assessment total and mill rate of all tax class.
+Then we plotted the computed tax income vs municipal budget into line-chart to visualize their relationship.
+We chose Vancoouver, Richmond and Burnaby as representative because they are the three largest municipalities in Metro Vancouver.
+The plots showed that the computed tax amount was positively and approximatly linearly related to municipal budget through time.
 
@@
 names(assessment_aggregate)[1] <- paste("Municipality")
+## Correlation for continuous variables
+assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 4: Correlations: Mill Rates vs.
+Total Assessment Values") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-## Correlation for continuous variables
-assessment_aggregate %>% 
-  ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + 
-  geom_point() + 
-  ggtitle("Figure 4: Correlation: Mill Rates and Total Assessment Values") + 
-  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
-  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 5: Correlations: Mill rates vs.
+Budgets") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-assessment_aggregate %>% 
-  ggplot(aes(x=tax,y=rate,group=Municipality,color=Municipality)) + 
-  geom_point() + ggtitle("Figure 5: Correlation: Mill rates and Budgets") + 
-  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
-  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 6: Correlations: Mill Rates vs.
+Total Assessment Values of Land") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-assessment_aggregate %>% 
-  ggplot(aes(x=landTotal,y=rate,group=Municipality,color=Municipality)) + 
-  geom_point() + 
-  ggtitle("Figure 6: Correlation: Mill Rates and Total Land Assessment Values") + 
-  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
-  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 7: Correlations: Mill Rates vs.
+Total Assessment Values of Improvement") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-assessment_aggregate %>% 
-  ggplot(aes(x=improvementTotal,y=rate,group=Municipality,color=Municipality)) + 
-  geom_point() + 
-  ggtitle("Figure 7: Correlation: Mill Rates and Total Improvement Assessment Values") + 
-  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
-  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
-
-assessment_aggregate %>% 
-  ggplot(aes(x=propertyCount,y=rate,group=Municipality,color=Municipality)) + 
-  geom_point() + 
-  ggtitle("Figure 8: Correlation: Mill Rates and Total Number of Properties") + 
-  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
-  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 8: Correlations: Mill Rates vs.
+Total Number of Properties") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 ```
 
-As we can see:
+### Correlation Analysis
+```{r assess, echo=FALSE}
+assess_class
+tax_class
+land_class
+improvementTotal_class
+propertyCount_class
+```
 
-Figure 4, 6, 7, and 8 show that mill rate is negatively correlated with total assessment values, total land assessment values, total improvement assessment values, total number of properties.
-Budget is not significantly correlated with mill rate as shown in Figure 5.
-
-Outlier detection:
-
-The plots above also show that Vancouver (the dark pink points) are significantly different from other points and should be regarded as an outlier (futher analysis needed here).
-Some other municipalities (including Richmond, Surrey and Burnaby) that are also distinct in the scatter plots.
+The above scatter plots showed the correlation between each pair of mill rate vs.
+assessment total, municipal budget, land assessment total, improvement assessment total, and the number of properties.
+From all the scatter plots above we have noticed that Vancouver (the dark pink points) is significantly different from other points and should be treated as an outlier/special case.
+Richmond, Surrey and Burnaby are also distinct in the scatter plots and should be treated as outliers.
+We believe these municiplaties also have impacts on the correlation between mill rate and its features.
+Further analysis and prediction model will be used to explore and explain such phenomenon.
 
 ### ANOVA Analysis
+
 ```{r anova, echo=FALSE, warning=FALSE, message=FALSE}
@@
 
-                    Table 1: ANOVA analysis of mill rate and three categorical variables
+**Table 1: Result from ANOVA Analysis**
 
-Feature Name  | Pr(>F) | significance levels
-------|--------|------
-Tax Class Code | <2e-16 | ***(very significant)
-Year | 4.2e-06 | **(significant)
-Municipality | 0.0768 | .(not significant)
+Feature Name  | Pr(>F) 
+------|--------
+TaxClassCode | <2e-16 
+Year | 4.2e-06 
+Municipality | 0.0768
 
-We used three 3 ANOVA tables to analyze the correlation between mill rate and each of the 3 categorical features (TaxClassCode, Year, Municipalities).
-The table showed that Tax Class Code and Year have a strong correlation with mill rate while the correlation between Municipality and mill rate is not significant at 5% level.
+We used three 3 ANOVA analysis to analyze the correlation between mill rate and each of the 3 categorical features - TaxClassCode, Year, Municipalities.
+TaxClassCode and Year showed a strong correlation with mill rate at 5% significant lete, yet the correlation between Municipality and mill rate is not significant at 5% level.
 
@@
 
-## Results
+###Linear Model
 
-### Variable Selection
+**Table 2: Result from Model Used**
 
-                  Table 2: Significance level of all variables of full linear model
+Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE 
+------|--------|--------|--------|--------
+      OLR full | 0.8874| 0.8707| 1.9843| 2.5902
+      OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237
+      Ridge|  0.8868| NA| 1.9896| 2.5675
+      LASSO| 0.8873| NA| 1.9855| 2.5280
+      Elastic Net| 0.8625| NA| 2.2366| 2.5486
 
- Variable |                                 Pr(>t) | Significance Level 
- ------|------|------
-(Intercept)                         |        0.000111 | ***
-factor(Municipality)Belcarra        |        0.299387 |   
-factor(Municipality)Bowen Island    |        0.303157 |   
-factor(Municipality)Burnaby         |        0.604683 |   
-factor(Municipality)Coquitlam       |        0.001594 | ** 
-factor(Municipality)Delta           |        0.002579 | ** 
-factor(Municipality)Langley - City  |        0.020365 | *  
-factor(Municipality)Langley - Township  |    0.023030 | *  
-factor(Municipality)Lions Bay           |    0.091427 | .  
-factor(Municipality)Maple Ridge         |    4.18e-05 | ***
-factor(Municipality)North Vancouver - City |  0.129059 |    
-factor(Municipality)North Vancouver - Dist |  0.077124 | .  
-factor(Municipality)Pitt Meadows        |    2.41e-07 | ***
-factor(Municipality)Port Coquitlam      |    1.60e-05 | ***
-factor(Municipality)Port Moody          |    1.16e-05 | ***
-factor(Municipality)Richmond            |    0.894181 |   
-factor(Municipality)Surrey              |    0.984723 |   
-factor(Municipality)Vancouver           |    0.885521 |   
-factor(Municipality)West Vancouver      |    0.001003 | ** 
-factor(Municipality)White Rock          |    0.034041 | *  
-factor(Year)2017                        |    0.000655 | ***
-factor(Year)2018                        |    1.30e-08 | ***
-factor(Year)2019                        |    4.73e-13 | ***
-factor(TaxClassCode)5                   |    < 2e-16 | ***
-factor(TaxClassCode)6                   |    < 2e-16 | ***
-assessTotal                             |    0.013752 | *  
-landTotal                               |    0.013863 | *  
-improvementTotal                        |          NA |   
-propertyCount                           |    0.825699 |   
-tax                                     |    0.958323 |
+Above is the goodness of fit of five models.
+Except for the Elastic Net model which has higher MSE and PMSE, there are no significant differences between these models.
+We have not used cross-validation to reduce model overfitting.
+Although the results are relatively similar and have high R_squared across all models, we believe after using cross-validation there will be a clearer difference between the goodness of fit of each model.
 
-After fitting the ordinary linear regression model with all features, the summary statistics in Table 2 showed that municipality, year, tax class code, total assessment value, and total land assess are significant, which indicated that total improvement assessment value, number of properties, and budget(tax) might not have strong impact for the linear model.
-Therefore we built an ordinary linear regression reduced model with only significant variables.
+For a dtailed look of how the models were fitted and evaluated, please refer to the appendix.
 
-### Goodness of Fit
+## Conclusion
 
-                          Table 3: Goodness of fit of all models
-
-Model | Mutiple R-Squared | Adjusted R_Squared | MSE
-------|--------|--------|--------|--------
-      Ordinary Linear Regression full model | 0.8874| 0.8707| 1.9843
-      Ordinary Linear Regression reduced model | 0.8874| 0.8721| 1.9845
-      Ridge Regression model |  0.8868| NA| 1.9896
-      LASSO Regression model | 0.8873| NA| 1.9855
-      Elastic Net Regression model | 0.8625| NA| 2.2366
-
-The table above is the goodness of fit of five models we fitted.
-All models have relatively high R squared and adjusted R squared values, which indicates that the above models are able to accurately fit the data.
-Also there is no significant differences among these models in terms of mean squared error, except for Elastic Net model which have higher MSE.
-We have not used cross validation to reduce model overfitting.
-Although the results are relatively similar and have high R_squared across all models, we believe using cross validation there will be a clearer difference between the goodness of fit of each model.
-
-### Prediction Power
-
-                        Table 4: Prediction Power of all models
-Model | PMSE 
-------|--------
-      Ordinary Linear Regression full model | 2.5902
-      Ordinary Linear Regression reduced model | 2.5237
-      Ridge Regression model | 2.5675
-      LASSO Regression model | 2.5280
-      Elastic Net Regression model | 2.5486
-
-The table above is the prediction power of five models we fitted. All models have relatively low error (PMSE), while there is also no clear difference among them. Futher analysis is needed here.
-
-## Conclusions
-
-From our exploratory data analysis, We have found that when comparing the correlation between mill rate against features in our models, mill rate is highly correlated with total assessment values, total land assessment values, total improvement assessment values, total number of properties, comparing with budget.
-Also Vancouver is an outlier in our plot.
+From our exploratory data analysis, we can see that assessment total, land total, tax code, year and municipalities are all significant in our model.
+The client suggested that we can transform numerical factors such as assessment total into percentage change.
+This data transformation does not perform as well as we expected, which only yields an R^2 of around 0.2 across all of our fitted models.
+We believe the method that the client has suggested might leave out some important information about the housing market in each municipality.
+We have also found that when comparing the correlation between mill rate against features in our models, Vancouver is an outlier.
 We would like to further investigate Vancouver as a special case.
-
-We can also see that total assessment value, total land assessment value, tax class code, year and municipality are all significant in our linear regression model.
 We are also interested in the effect of different tax classes in predicting mill rate.
@@
 
-The linear regression models were able to accurately predict mill rate with relatively low error, but there is still a very high chance that our model is overfitted.
+All fitted linear model were able to make accurate prediction based on means squared error, but there is still a very high chance that our model is overfitted.
 For our next report, we are going to use cross-validation to reduce the effect of overfitting.
@@
 
-### Missing Value:
-
-There are 1801 missing values in mill rate(TaxClassTaxRate).
-We decided to impute these missing values based on client information; all properties in the same region, classcode, and year should have a unique class rate
-
-- For entries with mill rate, we aggregated them into groups by region + classcode + year.
-
-- For entries without mill rate, we found the group that they belong to and assign them mill rate in that group.
-
-### Here is some exceptions found:
-
-Some groups' mill rate is not unique:
-
-- In Delta, properties in different neighbourhoods have slightly different mill rates. Since the variance is not significant, we take the mean as the overall mill rate in groups.
-
-- In Vancouver, 2019, Class 01, one property's mill rate is different from others. It is regarded as an outlier.
-
-- In Burnaby, 2019, Class 06, six properties' mill rate are different from others. They are regarded as outliers.
-
-- In Langley, 2019, Class 06, mill rate is different between assessment type.
-After talking with the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.
-
-- In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:
-
-                          Table 4: Groups that miss mill rate
-
-Year  | Region | Class |Number of Properties
-------|--------|-------|----------
-2016 | Belcarra | 06 | 9 
-2016 | Lions Bay | 01 | 40 
-2016 | Lions Bay | 06 | 25 
-2016 | Maple Ridge Rural | 05 | 36 
-2017 | Belcarra | 06 | 9 
-2017 | Lions Bay | 01 | 39 
-2017 | Lions Bay | 06 | 24 
-2017 | Maple Ridge Rural | 05 | 36 
-2018 | Maple Ridge Rural | 05 | 36 
-2019 | Maple Ridge Rural | 05 | 38
-
-### Model Fitting
-
-#### Ordinary Linear Regression Full Model
-```{r full, echo=TRUE, warning=FALSE, message=FALSE}
+### Code of Linear Regressoin Model
+```{r full, warning=FALSE, message=FALSE}
 linear_full<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
-#summary(linear_full)
+# summary(linear_full)
 
 library(broom)
+
 linear_full_fit<-augment(linear_full)
@@
 ```
-
-#### Ordinary Linear Regression Reduced Model
-```{r reduced, echo=TRUE, warning=FALSE, message=FALSE}
+```{r reduced, echo=FALSE, warning=FALSE, message=FALSE}
 reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
@@
 
-#### Ridge
-```{r Ridge, echo=TRUE, warning=FALSE, message=FALSE}
+```{r Ridges, warning=FALSE, message=FALSE}
 library(glmnet)
@@
 lambdas <- 10^seq(2, -3, by = -.1)
+#dim(x)
 
@@
 optimal_lambda <- cv_ridge$lambda.min
-
+#optimal_lambda
 predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
@@
 ridge_mse <- eval_results(y, predictions_train)
+
 ```
 
-#### LASSO
-```{r LASSO, echo=TRUE, warning=FALSE, message=FALSE}
+```{r LASSO, warning=FALSE, message=FALSE}
 # Setting alpha = 1 implements lasso regression
@@
 
-#### Elastic Net
-```{r elastic_net, echo=TRUE, warning=FALSE, message=FALSE}
+```{r elastic_net, echo=FALSE, warning=FALSE, message=FALSE}
 library(caret)
@@
 
-#### Prediction Power
-```{r Prediction power, echo=TRUE, warning=FALSE, message=FALSE}
+```{r Prediction power, warning=FALSE, message=FALSE}
 set.seed(450)
@@
 ```
+
+### Missing Value:
+
+There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values
+Based on client information, all properties in the same region, classcode, and year should have a unique class rate
+
+- For entries with mill rate, we aggregated them into groups by region + classcode + year.
+
+- For entries without mill rate, we found the group they belong to and assign them mill rate in that group.
+
+###Here is some exceptions found:
+
+Some groups' mill rate is not unique:
+
+- In Delta, properties in different neighbourhoods have slightly different mill rates. Since the variance is not significant, we take the mean as the overall mill rate in groups.
+
+- In Vancouver, 2019, Class 01, one property's mill rate is different from others. It is regarded as an outlier.
+
+- In Burnaby, 2019, Class 06, six properties' mill rate are different from others. They are regarded as outliers.
+
+- In Langley, 2019, Class 06, mill rate is different between assessment type. After talking to the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.
+
+- In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:
+
+**Table 3: Data with missing mill rate**
+
+Year  | Region | Class |Number of Properties
+------|--------|-------|----------
+2016 | Belcarra | 06 | 9 
+2016 | Lions Bay | 01 | 40 
+2016 | Lions Bay | 06 | 25 
+2016 | Maple Ridge Rural | 05 | 36 
+2017 | Belcarra | 06 | 9 
+2017 | Lions Bay | 01 | 39 
+2017 | Lions Bay | 06 | 24 
+2017 | Maple Ridge Rural | 05 | 36 
+2018 | Maple Ridge Rural | 05 | 36 
+2019 | Maple Ridge Rural | 05 | 38
+
+###Here is the summary statistics from the full linear regression model
+
+**Table 4: Summary statistics of the full linear regression model**
+
+Variable |                                 Pr(>t) | Significance Level 
+ ------|------|------
+(Intercept)                         |        0.000111 | ***
+factor(Municipality)Belcarra        |        0.299387 |   
+factor(Municipality)Bowen Island    |        0.303157 |   
+factor(Municipality)Burnaby         |        0.604683 |   
+factor(Municipality)Coquitlam       |        0.001594 | ** 
+factor(Municipality)Delta           |        0.002579 | ** 
+factor(Municipality)Langley - City  |        0.020365 | *  
+factor(Municipality)Langley - Township  |    0.023030 | *  
+factor(Municipality)Lions Bay           |    0.091427 | .  
+factor(Municipality)Maple Ridge         |    4.18e-05 | ***
+factor(Municipality)North Vancouver - City |  0.129059 |    
+factor(Municipality)North Vancouver - Dist |  0.077124 | .  
+factor(Municipality)Pitt Meadows        |    2.41e-07 | ***
+factor(Municipality)Port Coquitlam      |    1.60e-05 | ***
+factor(Municipality)Port Moody          |    1.16e-05 | ***
+factor(Municipality)Richmond            |    0.894181 |   
+factor(Municipality)Surrey              |    0.984723 |   
+factor(Municipality)Vancouver           |    0.885521 |   
+factor(Municipality)West Vancouver      |    0.001003 | ** 
+factor(Municipality)White Rock          |    0.034041 | *  
+factor(Year)2017                        |    0.000655 | ***
+factor(Year)2018                        |    1.30e-08 | ***
+factor(Year)2019                        |    4.73e-13 | ***
+factor(TaxClassCode)5                   |    < 2e-16 | ***
+factor(TaxClassCode)6                   |    < 2e-16 | ***
+assessTotal                             |    0.013752 | *  
+landTotal                               |    0.013863 | *  
+improvementTotal                        |          NA |   
+propertyCount                           |    0.825699 |   
+tax                                     |    0.958323 |

### 2020-03-05 doc/RealEstate_Yuting_Wen.Rmd

@@
 He is interested in a good prediction model that can accurately predict future mill rates, given the past data like the past mill rates and the past assessment values, in Metro Vancouver.
-Data cleaning, exploratory data analysis, and linear models with different kinds of regulazations were used to analyze the relationship between mill rate and other factors and predict future mill rates.
+Data cleaning, exploratory data analysis, and linear models with different kinds of regularzations were used to analyze the relationship between mill rate and other factors and predict future mill rates.
 Data cleaning is performed to aggregate our data into summary statistics and also reduce the number of data entries.
-Exploratory analysis has shown that there is a strong correlation between mill rate and municipal budget and strong correlation between mill rate and assessment in some municipalities.
-The results of exploratory analysis have also shown that there is no clear difference among full linear Regression model with all features (Year, TaxClassCode, Municipalities, assessTotal_pct, tax (budget), landTotal, improvementTotal, numProperties) and reduced linear regression models (Ridge, Lasso and Elastic net).
+Exploratory analysis has shown that there is a strong correlation between mill rate and municipal budget and between mill rate and assessment in some municipalities.
+The results of exploratory analysis have also shown that there is no clear difference among full linear Regression model with all features (year, tax class code, municipality, total Assessment value, tax (budget), total land assessment value, total improvement assessment value, number of properties) and reduced linear regression models (Ridge, Lasso and Elastic net), while they still have relatively low error.
 
@@
 A full linear model with all features, a few reduced linear models with regularzations like the LASSO, Ridge Regression are used to train our predictive model.
-Once we have finished fitting those models, we will use the mean squared training error to evaluate the performance of our model.
+Once we have finished fitting those models, we will use the mean squared test error to evaluate the performance of our model.
 
@@
 municipal budget, mill rate vs.
-landTotal (total land assessment value), mill rate vs.
-improvementTotal (total improvement assessment value), and mill rate vs.
-numOfProperties (number of properties) to see the correlation between each of the two factors.
+total land assessment value, mill rate vs.
+total improvement assessment value, and mill rate vs.
+number of properties to see the correlation between each of the two factors.
 ANOVA analysis was also performed to see the correlations between mill rate vs.
@@
 These exploratory analysis and plots helped us to find trends in data within each of its own categories.
+
+We used R package called "ggplot" to visualize our results.
 

### 2020-03-16 src/2nd_draft.Rmd
2ndDraft

--- 2020-03-01 src/EDA&mode_fitting.Rmd
@@
 ---
-title: "EDA"
+title: "2nd_draft"
 author: "Xuechun Lu, Yuting Wen, Peter Han, Yuetong Liu"
-date: "25/02/2020"
+date: "15/03/2020"
 output:
@@
 require(neuralnet)
-```
-
-```{r}
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-  numPlots = length(plots)
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
- if (numPlots==1) {
-    print(plots[[1]])
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-
+library(plotly)
 ```
@@
 ```{r, warning=F, message=F}
-# all municipalities
-municipality.list = c(
-  "Burnaby", 
-  "Coquitlam", 
-  "Delta", 
-  "Langley - City", 
-  "Langley - Township",
-  "Maple Ridge",
-  "Maple Ridge Rural", 
-  "North Vancouver - City",
-  "North Vancouver - Dist",
-  "Pitt Meadows", 
-  "Port Coquitlam", 
-  "Port Moody", 
-  "Richmond", 
-  "Surrey", 
-  "Vancouver", 
-  "White Rock", 
-  "West Vancouver", 
-  "Bowen Island", 
-  "Anmore", 
-  "Belcarra",
-  "Lions Bay")
-
-# budget and assessment data import and process
-tax_pct<-read_csv(here("data", "tax_pct.csv"))
-assessment_pct<-read_csv(here("data", "assessment_pct.csv"))
-# omit 2016
-assessment_pct<-na.omit(assessment_pct)
-assessment_pct<-assessment_pct[,-c(1)]
-names(assessment_pct)[2]<-paste("Municipalities")
-tax_pct<-tax_pct[,-c(1)]
-#head(tax_pct)
-
-assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
-assessment_final <- read.csv(here("data","assessment_final.csv"))
-assessment_final <- assessment_final[,-1]
-assessment_final <- assessment_final[complete.cases(assessment_final),]
-assessment_final <- assessment_final[,-4]
-```
-Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge Rural and Maple Ridge with the same tax.
-
-2019 tax pct is missing, we imputed 2019 tax by the average of previous tax pct
-```{r}
-tax_pct[,4]<-(tax_pct$pct_2017+tax_pct$pct_2018)/2
-#tax_pct
-names(tax_pct)[4]<-paste("pct_2019")
-tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567,5.945346))
-
-#head(tax_pct)
+## Aggregated Data
+assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))[,-1]
 ```
 
-aggregate tax and assessment
+2. Test correlations between past mill rates and other features.
 ```{r}
-assessment2017<-assessment_pct  %>%  filter(Year=="2017")
-tax2017<-tax_pct[,1:2]
-assessment2017<-assessment2017 %>% left_join(tax2017, by = c("Municipalities"))
-names(assessment2017)[9]<-paste("tax")
+# continuous variables
+assessment_transform <- assessment_aggregate
+## TODO: mean, log transform, etc.
 
-assessment2018<-assessment_pct  %>%  filter(Year=="2018")
-tax2018<-tax_pct[,c(1,3)]
-assessment2018<-assessment2018 %>% left_join(tax2018, by = c("Municipalities"))
-names(assessment2018)[9]<-paste("tax")
+for(i in colnames(assessment_transform)[4:7]){
+  print(sprintf("correlation between mill rate and %s: %s.",
+                i, cor(assessment_aggregate$rate, eval(parse(text=paste0("assessment_aggregate$", i))))))
+  print(assessment_aggregate %>%
+                   ggplot(aes(x=!!as.name(i),
+                              y=rate,group=AddressAssessorMunicipalityDesc,
+                              color=AddressAssessorMunicipalityDesc)) + 
+                   geom_point() + 
+                   ggtitle(sprintf("mill rate vs %s ", i)) + 
+                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black"))
 
-assessment2019<-assessment_pct  %>%  filter(Year=="2019")
-tax2019<-tax_pct[,c(1,4)]
-assessment2019<-assessment2019 %>% left_join(tax2019, by = c("Municipalities"))
-names(assessment2019)[9]<-paste("tax")
+}
 
-pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
-pct_final$tax<-as.numeric(pct_final$tax)
-write.csv(pct_final, here("data","data_final.csv"))
-
-# all pct data aggregated
-#head(pct_final)
-
-```
-
-2019 tax is missing, we imputed 2019 tax using previous values
-```{r}
-tax_final <- read.csv(here("data","tax_final.csv"))
-#head(tax_final)
-
-# tax_final is the budget
-tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))
-
-# tax_modified contains all the budget values from 2016 to 2019
-#head(tax_modified)
-```
-
-global variables
-```{r}
-# all pct aggregated
-pct_final
-# all budget values
-tax_modified
-```
-
-check match between budget and assessTotal*mill_rate
-```{r}
-# get predicted tax income for all tax classes and each municipality
-assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
-names(assessment_aggregate)[10] <- paste("predTaxIncome")
-
-dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
-dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
-names(dat_pred_income)[3] <- paste("predTaxIncome")
-
-tax_refactored <- tax_modified
-names(tax_refactored)[3] <- paste("2016")
-names(tax_refactored)[4] <- paste("2017")
-names(tax_refactored)[5] <- paste("2018")
-names(tax_refactored)[6] <- paste("2019")
-dat_true_income <- gather(tax_refactored, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
-dat_true_income <- dat_true_income[,-1]
-
-income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
-income_budget_compare <- gather(income_budget_compare, TaxType, Amount, TaxAmount:predTaxIncome, factor_key=TRUE)
-
-compare_full_plot <- income_budget_compare %>% filter(Municipalities!="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=Municipalities)) + geom_line()
-
-compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome for Vancouver")
-compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Surrey") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome for Surrey")
-compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome for Burnaby")
-compare_plot4 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s taxIncome for Richmond")
-multiplot(compare_plot1,compare_plot2,compare_plot3,compare_plot4,cols=2)
-```
-
-2. plots of mill rate v.s assessments for each tax class
-```{r}
-# mill rate vesus year for each municipality and tax class
-dat_rate <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, rate)
-#head(dat_rate)
-
-dat_rate_1 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
-dat_rate_5 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
-dat_rate_6 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6")
-
-# plot assessed value vesus year for each municipality and tax class
-dat_assess <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, assessTotal)
-
-dat_assess_1 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
-dat_assess_5 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
-dat_assess_6 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6")
-
-# plot budgets vesus year
-dat_tax_long <- gather(tax_modified, Year, TaxAmount, X2016:X2019, factor_key=TRUE)
-
-# Vancouver/West Vancouver are special cases for assessment, we omit them for taxClass 1 and 6
-rate_class1 <- dat_rate_1 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 1 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-assess_class1 <- dat_assess_1 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 1 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-multiplot(rate_class1 + theme(legend.position = "none"), assess_class1 + theme(legend.position = "none"))
-
-rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 5") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 5") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-multiplot(rate_class5 + theme(legend.position = "none"),assess_class5 + theme(legend.position = "none"))
-
-rate_class6 <- dat_rate_6 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 6 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-assess_class6 <- dat_assess_6 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 6 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-multiplot(rate_class6 + theme(legend.position = "none"),assess_class6 + theme(legend.position = "none"))
-
-budget_plot <- dat_tax_long %>% filter(Municipalities != "Vancouver" & Municipalities != "West Vancouver") %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_point() + ggtitle("Tax v.s Year") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-budget_plot
-```
-
-3. plots of mill rate pct change v.s assessment pct change for each tax class
-```{r}
-pct_rate1 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "01") %>%
-  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_point() + ggtitle("taxClass1 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-pct_rate5 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "05") %>%
-  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_point() + ggtitle("taxClass5 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-pct_rate6 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "06") %>%
-  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_point() + ggtitle("taxClass6 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-pct_assess1 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "01") %>%
-  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_point() + ggtitle("taxClass1 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-pct_assess5 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "05") %>%
-  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_point() + ggtitle("taxClass5 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-pct_assess6 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "06") %>%
-  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_point() + ggtitle("taxClass6 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-multiplot(pct_rate1,pct_assess1)
-multiplot(pct_rate5,pct_assess5)
-multiplot(pct_rate6,pct_assess6)
-```
-
-4. pct change of mill rate and assessment for each municipality
-```{r}
-# for each municipality
-tax.list <- c("01","05","06")
-for (i in 1:21){
-  if (municipality.list[i] != "Lions Bay") {
-    plot_rate_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-    filter(TaxClassCode %in% tax.list) %>% 
-    ggplot(aes(x=Year, y=rate_pct, group=TaxClassCode, color=TaxClassCode)) +
-      geom_line() + ggtitle(municipality.list[i])
-
-    plot_assess_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-    filter(TaxClassCode %in% tax.list) %>% 
-    ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
-      geom_line() + ggtitle(municipality.list[i])
-
-    plot_tax_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line() + ggtitle(municipality.list[i])
-
-    multiplot(plot_rate_pct,plot_assess_pct,plot_tax_pct)
-  }
+# categorical variables
+for(i in colnames(assessment_transform)[1:3]){
+  print(ggplot(assessment_aggregate, aes(x=as.factor(!!as.name(i)),y=rate, fill=as.factor(!!as.name(i)))) + 
+  geom_boxplot()+ 
+  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
 }
@@
 
-5. box plots of mill rates, grouped by region
-```{r}
-rate_boxes <- ggplot(dat_rate, aes(x=AddressAssessorMunicipalityDesc,y=rate,fill=TaxClassCode)) + geom_boxplot() + facet_wrap(~TaxClassCode) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
-rate_boxes 
-```
-
-6. Test correlations between past mill rates and assessed values and government budget.
-```{r}
-# aggregate tax and mill rate & assessment
-colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
-dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
-
-tax_modified1 <- gather(tax_modified, Year, tax, '2016':'2019', factor_key = TRUE)%>%
-  rename(AddressAssessorMunicipalityDesc = Municipalities)
-
-assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
-assessment_aggregate <- assessment_aggregate[,-c(3,11)]
-names(assessment_aggregate)[10] <- paste("tax")
-
-# correlation between mill rate and budget for each municipality
-rate_tax_correlation <- assessment_aggregate %>%
-  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
-  summarise(cor(rate, tax))
-rate_tax_correlation
-# correlation between mill rate and assessment
-rate_assess_correlation2 <- assessment_aggregate %>%
-  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
-  summarise(cor(assessTotal, rate))
-rate_assess_correlation2
-
-## Correlation for continuous variables
-assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and assess - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-tax_class <- assessment_aggregate %>% ggplot(aes(x=predTaxIncome,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-assess_class
-
-tax_class
-
-## Correlation for categorical variables
-TaxClassCode <- aov(rate ~ TaxClassCode, data = assessment_aggregate)
-summary(TaxClassCode)
-
-Year <- aov(rate_pct ~ Year, data = pct_final)
-summary(Year)
-
-Municipalities <- aov(rate ~ AddressAssessorMunicipalityDesc, data = assessment_aggregate)
-summary(Municipalities)
-```
-
-7. Compute autocorrelations of past mill rates to test the independence of mill rates. 
-```{r}
-# all white noise
-for (i in 1:21) {
-  print(municipality.list[i])
-  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "1") %>% select(rate)
-  if (municipality.list[i] != "Maple Ridge Rural") {
-    as.ts(dat_tmp)
-    acf(dat_tmp)
-  }
-}
-for (i in 1:21) {
-  print(municipality.list[i])
-  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "5") %>% select(rate)
-  if (municipality.list[i] != "Maple Ridge Rural" & municipality.list[i] != "White Rock" & municipality.list[i] != "Anmore" & municipality.list[i] != "Belcarra" & municipality.list[i] != "Lions Bay") {
-    as.ts(dat_tmp)
-    acf(dat_tmp)
-  }
-}
-for (i in 1:21) {
-  print(municipality.list[i])
-  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "6") %>% select(rate)
-  if (municipality.list[i] != "Maple Ridge Rural") {
-    as.ts(dat_tmp)
-    acf(dat_tmp)
-  }
-}
-```
-
-model fitting
+3.model fitting
 (i) linear model
 ```{r}
-linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
-summary(linear1)
-
-colnames(assessment_aggregate)
-linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
+linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount,data=assessment_aggregate)
 summary(linear_full)
@@
 library(broom)
+
 linear_full_fit<-augment(linear_full)
@@
 predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
-eval_results(y, predictions_train)
+ridge_r <- eval_results(y, predictions_train)
 
@@
 predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
-eval_results(y, predictions_train)
+lasso_r <- eval_results(y, predictions_train)
 ```
@@
 )
+
 elastic_net
+
 ```
@@
 y<-test[,c(8)]
-linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
+linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount,data=train)
 resid<-predict(linear_1,newdata = newx) - y
@@
 elastic_net = train(
- rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
+ rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, 
  data = assessment_aggregate[train_ind,],
@@
 
-## neural network
+Summary on GOF
 ```{r}
-nn=neuralnet(TaxClassTaxRate~Year+TaxClassCode+AssessorAreaCode+AddressAssessorMunicipalityCode+AssessedValueAmt+CurrentYearTotal+TaxOwingAmountTotalCalculated,data=assessment_final, hidden=20, act.fct = "logistic", linear.output = FALSE)
-plot(nn)
-```
 
-```{r}
-# in progress
-#library(keras)
-#nn_dat <- to_categorical(assessment_final, 10)
-#train_data <- scale(assessment_final)
+GOF_table <- rbind(
+      c("OLR full", 0.8874, 0.8707, 1.9843, 2.5902),
+      c("OLR reduced", 0.8874, 0.8721, 1.9845, 2.5237),
+      c("Ridge", 0.8868, 0.8707, 1.9896, 2.5675),
+      c("LASSO", 0.8873, 0.8707, 1.9855, 2.5280),
+      c("Elastic Net", 0.8625, 0.8707, 2.2366, 2.5486))
+colnames(GOF_table) <- c("Model", "Mutiple R-Squared","Adjusted R_Squared", "MSE", "PMSE")
 

### 2020-03-16 src/2nd_draft.Rmd
Update 2nd_draft.Rmd
cv function added

@@
 ```
+
+```{r}
+dummy_year<-dummy(assessment_aggregate$Year)
+dummy_municipal<-dummy(assessment_aggregate$AddressAssessorMunicipalityDesc)
+dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
+# build x matrix
+x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)
+
+y<-assessment_aggregate$rate
+x<-cbind(x,y)
+# correct col name
+#assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax
+colnames(x)[28:32]<-c("assessTotal","landTotal","improvementTotal","propertyCount","tax")
+xm<-x[,1:(ncol(x)-1)]
+x.data.frame<-assessment_aggregate
+xm.data.frame<-assessment_aggregate[,-8]
+y.data.frame<-assessment_aggregate$rate
+n <- nrow(xm)
+k <- 10
+ii <- (1:n)%%k + 1
+set.seed(123)
+N <- 50
+mspe.el<-mspe.la <- mspe.f <- mspe.ri <- mspe.reduced <- rep(0, N)
+
+for (i in 1:N) {
+  ii <- sample(ii)
+  pr.el<-pr.la <- pr.f <- pr.ri <- pr.reduced <- rep(0, n)
+  for (j in 1:k) {
+    tmp.ri <- cv.glmnet(x = xm[ii != j, ], y = y[ii != j], lambda = lambdas, 
+                        nfolds = 10, alpha = 0, family = "gaussian")
+    tmp.la <- cv.glmnet(x = xm[ii != j, ], y = y[ii != j], lambda = lambdas, 
+                        nfolds = 10, alpha = 1, family = "gaussian")
+    tmp.reduced <- lm(rate ~ factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal, 
+                  data = x.data.frame[ii != j, ])
+    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
+               data = x.data.frame[ii != j, ])
+
+    tmp.elastic<-train(
+      rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
+      data = x.data.frame[ii != j, ],
+      method = "glmnet",
+      trControl = cv_10
+    )
+
+    pr.ri[ii == j] <- predict(tmp.ri, s = tmp.ri$lambda.min, newx = xm[ii == 
+                                                                    j, ])
+    pr.la[ii == j] <- predict(tmp.la, s = tmp.la$lambda.min, newx = xm[ii == 
+                                                                    j, ])
+    pr.reduced[ii == j] <- predict(tmp.reduced, newdata = x.data.frame[ii == j, ])
+    pr.f[ii == j] <- predict(tmp.full, newdata = x.data.frame[ii == j, ])
+
+    index<-which.min(tmp.elastic$results$RMSE)
+    pr.el[ii == j] <- predict(tmp.elastic, 
+                              alpha = tmp.elastic$results[index,1], 
+                              lambda = tmp.elastic$results[index,2], 
+                              newdata = xm.data.frame[ii == j, ])
+  }
+  mspe.ri[i]<-mean((y - pr.ri)^2)
+  mspe.la[i]<-mean((y - pr.la)^2)
+  mspe.reduced[i]<-mean((y - pr.reduced)^2)
+  mspe.f[i]<-mean((y - pr.f)^2)
+  mspe.el[i]<-mean((y - pr.el)^2)
+
+}
+boxplot(mspe.la, mspe.ri,mspe.el, mspe.reduced, mspe.f, 
+        names = c("LASSO", "Ridge", "Elastic Net", "Reduced", "Full"), 
+        col = c("yellow","steelblue", "gray80", "tomato", "springgreen"), 
+        cex.axis = 1, 
+        cex.lab = 1, 
+        cex.main = 2)
+mtext(expression(hat(MSPE)), side = 2, line = 2.5)
+
+```

### 2020-03-16 src/2nd_draft.Rmd
EDA fulfilled, full model changed

@@
 library(readxl)
-require(neuralnet)
-library(plotly)
+library(broom)
+library(glmnet)
+library(dummies)
 ```
@@
 ```{r}
-# continuous variables
+# ______________________________________________________________________________________________
+# CONTINUS variables
 assessment_transform <- assessment_aggregate
-## TODO: mean, log transform, etc.
 
-for(i in colnames(assessment_transform)[4:7]){
-  print(sprintf("correlation between mill rate and %s: %s.",
-                i, cor(assessment_aggregate$rate, eval(parse(text=paste0("assessment_aggregate$", i))))))
-  print(assessment_aggregate %>%
-                   ggplot(aes(x=!!as.name(i),
+# log transformation and take avg
+assessment_transform[,9] <- assessment_transform$assessTotal/assessment_transform$propertyCount
+names(assessment_transform)[9] <- paste("AvgTotalAssessmentValue")
+
+cor(assessment_transform$rate,assessment_transform$AvgTotalAssessmentValue)
+
+assessment_transform %>%
+                   ggplot(aes(x=AvgTotalAssessmentValue,
                               y=rate,group=AddressAssessorMunicipalityDesc,
@@
                    geom_point() + 
-                   ggtitle(sprintf("mill rate vs %s ", i)) + 
-                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black"))
+                   ggtitle(sprintf("mill rate vs average total assessment value")) + 
+                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
 
-}
-
-# categorical variables
-for(i in colnames(assessment_transform)[1:3]){
+# ______________________________________________________________________________________________
+# CATEGORICAL variables
+for(i in colnames(assessment_transform)[2:3]){
   print(ggplot(assessment_aggregate, aes(x=as.factor(!!as.name(i)),y=rate, fill=as.factor(!!as.name(i)))) + 
@@
 }
+
+# kruskal.test
+kruskal.test(assessment_transform$rate~assessment_transform$TaxClassCode,data=assessment_transform)
+kruskal.test(assessment_transform$rate~assessment_transform$AddressAssessorMunicipalityDesc,data=assessment_transform)
 ```
@@
 ```{r}
-linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount,data=assessment_aggregate)
+# based on EDA, we chose to fit model using municipality, tax class code, avg assess value
+linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,data=assessment_transform)
 summary(linear_full)
-
-library(broom)
 
@@
 ```{r}
-library(glmnet)
-library(dummies)
 dummy_year<-dummy(assessment_aggregate$Year)

### 2020-03-16 src/2nd_draft.Rmd
cross validation updated.

@@
 ```{r}
-dummy_year<-dummy(assessment_aggregate$Year)
-dummy_municipal<-dummy(assessment_aggregate$AddressAssessorMunicipalityDesc)
-dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
+dummy_year<-dummy(assessment_transform$Year)
+dummy_municipal<-dummy(assessment_transform$AddressAssessorMunicipalityDesc)
+dummy_taxclass<-dummy(assessment_transform$TaxClassCode)
 # build x matrix
-x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)
-
-y<-assessment_aggregate$rate
+x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_transform$AvgTotalAssessmentValue)
+y<-assessment_transform$rate
 x<-cbind(x,y)
@@
 #assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax
-colnames(x)[28:32]<-c("assessTotal","landTotal","improvementTotal","propertyCount","tax")
+colnames(x)[30]<-c("AvgTotalAssessmentValue")
 xm<-x[,1:(ncol(x)-1)]
-x.data.frame<-assessment_aggregate
-xm.data.frame<-assessment_aggregate[,-8]
-y.data.frame<-assessment_aggregate$rate
+x.data.frame<-assessment_transform
+xm.data.frame<-assessment_transform[,-8]
+y.data.frame<-assessment_transform$rate
 n <- nrow(xm)
@@
 N <- 50
-mspe.el<-mspe.la <- mspe.f <- mspe.ri <- mspe.reduced <- rep(0, N)
+mspe.n<-mspe.el<-mspe.la <- mspe.f <- mspe.ri <- mspe.reduced <- rep(0, N)
 
@@
   ii <- sample(ii)
-  pr.el<-pr.la <- pr.f <- pr.ri <- pr.reduced <- rep(0, n)
+  pr.n<-pr.el<-pr.la <- pr.f <- pr.ri <- pr.reduced <- rep(0, n)
   for (j in 1:k) {
@@
                         nfolds = 10, alpha = 1, family = "gaussian")
-    tmp.reduced <- lm(rate ~ factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal, 
+    tmp.reduced <- lm(rate ~ factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+AvgTotalAssessmentValue, 
                       data = x.data.frame[ii != j, ])
-    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
+    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+AvgTotalAssessmentValue, 
                    data = x.data.frame[ii != j, ])
+    # tmp.n<-lm(rate ~ 1, data = x.data.frame[ii != j, ])
 
-    tmp.elastic<-train(
-      rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
-      data = x.data.frame[ii != j, ],
-      method = "glmnet",
-      trControl = cv_10
-    )
+    # tmp.elastic<-train(
+    #   rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue, 
+    #   data = x.data.frame[ii != j, ],
+    #   method = "glmnet",
+    #   trControl = cv_10)
 
@@
     pr.f[ii == j] <- predict(tmp.full, newdata = x.data.frame[ii == j, ])
+    # pr.n[ii == j] <- predict(tmp.n, newdata = x.data.frame[ii == j, ])
 
@@
   mspe.f[i]<-mean((y - pr.f)^2)
+  # mspe.n[i]<-mean((y - pr.n)^2)
   mspe.el[i]<-mean((y - pr.el)^2)
@@
 }
-boxplot(mspe.la, mspe.ri,mspe.el, mspe.reduced, mspe.f, 
+
+boxplot(mspe.la, 
+        mspe.ri,
+        mspe.el, 
+        # mspe.n, 
+        mspe.reduced, 
+        mspe.f, 
         names = c("LASSO", "Ridge", "Elastic Net", "Reduced", "Full"), 
-        col = c("yellow","steelblue", "gray80", "tomato", "springgreen"), 
+        col = c("steelblue","gray80","pink", "tomato", "springgreen"), 
         cex.axis = 1, 
@@
         cex.main = 2)
+
 mtext(expression(hat(MSPE)), side = 2, line = 2.5)

### 2020-03-17 src/2nd_draft.Rmd
year deleted for all models

@@
 ```{r}
-dummy_year<-dummy(assessment_transform$Year)
 dummy_municipal<-dummy(assessment_transform$AddressAssessorMunicipalityDesc)
@@
 # build x matrix
-x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_transform$AvgTotalAssessmentValue)
+x<-cbind(dummy_municipal,dummy_taxclass,assessment_transform$AvgTotalAssessmentValue)
 y<-assessment_transform$rate
@@
 #assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax
-colnames(x)[30]<-c("AvgTotalAssessmentValue")
+colnames(x)[26]<-c("AvgTotalAssessmentValue")
+colnames(x)[27]<-c("rate")
 xm<-x[,1:(ncol(x)-1)]
 x.data.frame<-assessment_transform
-xm.data.frame<-assessment_transform[,-8]
+xm.data.frame<-assessment_transform[,-c(1,8)]
 y.data.frame<-assessment_transform$rate
@@
 mspe.n<-mspe.el<-mspe.la <- mspe.f <- mspe.ri <- mspe.reduced <- rep(0, N)
-
 for (i in 1:N) {
@@
                       data = x.data.frame[ii != j, ])
-    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+AvgTotalAssessmentValue, 
+    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+AvgTotalAssessmentValue, 
                    data = x.data.frame[ii != j, ])

### 2020-03-17 src/2nd_draft.Rmd
goodness of fit added

@@
 
-```{r}
+```{r, warning=FALSE}
 dummy_municipal<-dummy(assessment_transform$AddressAssessorMunicipalityDesc)
@@
 N <- 50
+mspe1.n<-mspe1.el<-mspe1.la <- mspe1.f <- mspe1.ri <- mspe1.reduced <- rep(0, N)
 mspe.n<-mspe.el<-mspe.la <- mspe.f <- mspe.ri <- mspe.reduced <- rep(0, N)
@@
   ii <- sample(ii)
+  pr1.n<-pr1.el<-pr1.la <- pr1.f <- pr1.ri <- pr1.reduced <- rep(0, n)
   pr.n<-pr.el<-pr.la <- pr.f <- pr.ri <- pr.reduced <- rep(0, n)
   for (j in 1:k) {
+    # ~ models trained!
     tmp.ri <- cv.glmnet(x = xm[ii != j, ], y = y[ii != j], lambda = lambdas, 
@@
 
+    # ~ prediction vector on train set
+    pr1.ri[ii != j] <- predict(tmp.ri, s = tmp.ri$lambda.min, newx = xm[ii !=
+                                                                         j, ])
+    pr1.la[ii != j] <- predict(tmp.la, s = tmp.la$lambda.min, newx = xm[ii !=
+                                                                         j, ])
+    pr1.reduced[ii != j] <- predict(tmp.reduced, newdata = x.data.frame[ii != j, ])
+    pr1.f[ii != j] <- predict(tmp.full, newdata = x.data.frame[ii != j, ])
+    # pr1.n[ii != j] <- predict(tmp.n, newdata = x.data.frame[ii != j, ])
+
+    # index<-which.min(tmp.elastic$results$RMSE)
+    # pr1.el[ii != j] <- predict(tmp.elastic,
+    #                           alpha = tmp.elastic$results[index,1],
+    #                           lambda = tmp.elastic$results[index,2],
+    #                           newdata = xm.data.frame[ii == j, ])
+
+    # ~ prediction vector on test set
     pr.ri[ii == j] <- predict(tmp.ri, s = tmp.ri$lambda.min, newx = xm[ii == 
@@
     # pr.n[ii == j] <- predict(tmp.n, newdata = x.data.frame[ii == j, ])
+    # index<-which.min(tmp.elastic$results$RMSE)
+    # pr.el[ii == j] <- predict(tmp.elastic,
+    #                           alpha = tmp.elastic$results[index,1],
+    #                           lambda = tmp.elastic$results[index,2],
+    #                           newdata = xm.data.frame[ii == j, ])
+  }
 
-    index<-which.min(tmp.elastic$results$RMSE)
-    pr.el[ii == j] <- predict(tmp.elastic,
-                              alpha = tmp.elastic$results[index,1],
-                              lambda = tmp.elastic$results[index,2],
-                              newdata = xm.data.frame[ii == j, ])
-  }
+  # goodness of fit
+  mspe1.ri[i]<-mean((y - pr1.ri)^2)
+  mspe1.la[i]<-mean((y - pr1.la)^2)
+  mspe1.reduced[i]<-mean((y - pr1.reduced)^2)
+  mspe1.f[i]<-mean((y - pr1.f)^2)
+  # mspe1.n[i]<-mean((y - pr1.n)^2)
+  # mspe1.el[i]<-mean((y - pr1.el)^2)
+
+  # prediction power
   mspe.ri[i]<-mean((y - pr.ri)^2)
@@
   # mspe.n[i]<-mean((y - pr.n)^2)
-  mspe.el[i]<-mean((y - pr.el)^2)
+  # mspe.el[i]<-mean((y - pr.el)^2)
 
@@
 
+boxplot(mspe1.la, 
+        mspe1.ri,
+        # mspe1.el, 
+        # mspe1.n, 
+        mspe1.reduced, 
+        mspe1.f, 
+        names = c("LASSO", 
+                  "Ridge", 
+                  # "Elastic Net", 
+                  "Reduced", 
+                  "Full"), 
+        col = c("steelblue","gray80",
+                # "pink", 
+                "tomato", "springgreen"), 
+        cex.axis = 1, 
+        cex.lab = 1, 
+        cex.main = 2,
+        main = "Goodness of Fit",
+        ylab = "MSE")
+
 boxplot(mspe.la, 
         mspe.ri,
-        mspe.el, 
+        # mspe.el, 
         # mspe.n, 
@@
         mspe.f, 
-        names = c("LASSO", "Ridge", "Elastic Net", "Reduced", "Full"), 
-        col = c("steelblue","gray80","pink", "tomato", "springgreen"), 
+        names = c("LASSO", "Ridge", 
+                  # "Elastic Net", 
+                  "Reduced", "Full"), 
+        col = c("steelblue","gray80",
+                # "pink", 
+                "tomato", "springgreen"), 
         cex.axis = 1, 
         cex.lab = 1, 
-        cex.main = 2)
+        cex.main = 2,
+        main = "Prediction Power",
+        ylab = "MSPE")
 

### 2020-03-17 src/2nd_draft.Rmd
y lab fixed

@@
         main = "Goodness of Fit",
-        ylab = "MSE")
+        ylab = expression(hat(MSE)))
 
@@
         main = "Prediction Power",
-        ylab = "MSPE")
-
-mtext(expression(hat(MSPE)), side = 2, line = 2.5)
+        ylab = expression(hat(MSPE)))
 

### 2020-03-17 src/2nd_draft.Rmd
Fit 2020

--- 2020-03-17 src/2nd_draft.Rmd
@@
 assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))[,-1]
+assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
+assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
 ```
@@
 ```
+
+```{r fit}
+## choose full model
+linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount,data=assessment_aggregate)
+
+## Fit and get PMSE
+assessment_2020.agg <- assessment_2020%>%
+  mutate(predict_rate = predict(linear_1,newdata = .))
+resid<-assessment_2020.agg$predict_rate - assessment_2020.agg$rate
+sqrt(sum(resid^2)/nrow(assessment_2020.agg))
+
+## Final result
+municipality.list = c(
+  "Burnaby", 
+  "Coquitlam", 
+  "Delta", 
+  "Langley - City", 
+  "Langley - Township",
+  "Maple Ridge",
+  "Maple Ridge Rural", 
+  "North Vancouver - City",
+  "North Vancouver - Dist",
+  "Pitt Meadows", 
+  "Port Coquitlam", 
+  "Port Moody", 
+  "Richmond", 
+  "Surrey", 
+  "Vancouver", 
+  "White Rock", 
+  "West Vancouver", 
+  "Bowen Island", 
+  "Anmore", 
+  "Belcarra",
+  "Lions Bay",
+  "New Westminster")
+
+## Raw Data
+a <- assessment_2020.full%>%
+  filter(TaxClassCode %in% c("01","05","06"),
+         AddressAssessorMunicipalityDesc %in% municipality.list)
+
+## Aggregate Data
+b <- assessment_2020.agg%>%
+  select(-c(Year, assessTotal : rate))%>%
+  mutate(TaxClassCode = as.factor(paste0("0",TaxClassCode)))
+
+assessment_2020.final<- merge(a, b, by =c("AddressAssessorMunicipalityDesc", "TaxClassCode"), all.x = TRUE)
+
+write.csv(assessment_2020.final, "../Data/Assessment/assessment_2020.final")
+write.csv(b, "../Data/assessment_2020.csv")
+```

### 2020-03-19 src/2nd_draft.Rmd
model modification

@@
 assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))[,-1]
-assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
-assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
+#assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
+#assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
 ```
@@
 names(assessment_transform)[9] <- paste("AvgTotalAssessmentValue")
+assessment_transform[,10] <- assessment_transform$landTotal/assessment_transform$propertyCount
+names(assessment_transform)[10] <- paste("AvgTotalLandValue")
+assessment_transform[,11] <- assessment_transform$improvementTotal/assessment_transform$propertyCount
+names(assessment_transform)[11] <- paste("AvgTotalImprovementValue")
 
 cor(assessment_transform$rate,assessment_transform$AvgTotalAssessmentValue)
+cor(assessment_transform$rate,assessment_transform$AvgTotalLandValue)
+cor(assessment_transform$rate,assessment_transform$AvgTotalImprovementValue)
 
@@
                    ggtitle(sprintf("mill rate vs average total assessment value")) + 
+                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+assessment_transform %>%
+                   ggplot(aes(x=AvgTotalLandValue,
+                              y=rate,group=AddressAssessorMunicipalityDesc,
+                              color=AddressAssessorMunicipalityDesc)) + 
+                   geom_point() + 
+                   ggtitle(sprintf("mill rate vs average total land value")) + 
+                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+assessment_transform %>%
+                   ggplot(aes(x=AvgTotalImprovementValue,
+                              y=rate,group=AddressAssessorMunicipalityDesc,
+                              color=AddressAssessorMunicipalityDesc)) + 
+                   geom_point() + 
+                   ggtitle(sprintf("mill rate vs average total improvement value")) + 
                    geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
@@
 3.model fitting
-(i) linear model
-```{r}
-# based on EDA, we chose to fit model using municipality, tax class code, avg assess value
-linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,data=assessment_transform)
-summary(linear_full)
-
-linear_full_fit<-augment(linear_full)
-sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
-```
-Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8707 
-MSPE =  1.984285
-
-Next: variable selection
-with only significant variables
-```{r}
-reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal, data=assessment_aggregate)
-summary(reduced)
-
-reduced_fit<-augment(reduced)
-sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate))
-```
-Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8721 
-MSPE = 1.984541
-
-(ii) Ridge, Lasso and elastic net.
-
-Ridge Regression
-reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r
-```{r}
-dummy_year<-dummy(assessment_aggregate$Year)
-dummy_municipal<-dummy(assessment_aggregate$AddressAssessorMunicipalityDesc)
-dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
-# build x matrix
-x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)
-
-y<-assessment_aggregate$rate
-lambdas <- 10^seq(2, -3, by = -.1)
-dim(x)
-
-lambdas <- 10^seq(2, -3, by = -.1)
-ridge_reg = glmnet(x, y, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
-set.seed(450)
-cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas, nfolds=10)
-optimal_lambda <- cv_ridge$lambda.min
-optimal_lambda
-predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
-
-# Compute R^2 from true and predicted values
-eval_results <- function(true, predicted) {
-  SSE <- sum((predicted - true)^2)
-  SST <- sum((true - mean(true))^2)
-  R_square <- 1 - SSE / SST
-  MSPE = sqrt(SSE/nrow(predicted))
-# Model performance metrics
-data.frame(
-  MSPE = MSPE,
-  Rsquare = R_square
-)
-
-}
-
-predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
-ridge_r <- eval_results(y, predictions_train)
-
-```
-MSPE = 1.989583
-Rsquare=	0.8868242
-
-Lasso 
-```{r}
-# Setting alpha = 1 implements lasso regression
-set.seed(450)
-lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
-
-# Best 
-lambda_best <- lasso_reg$lambda.min;lambda_best
-
-lasso_model <- glmnet(x, y, alpha = 1, lambda = lambda_best, standardize = TRUE)
-
-predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
-lasso_r <- eval_results(y, predictions_train)
-```
-MSPE = 1.985463	
-Rsquare=0.8872924
-
-Elastic net: 
-reference: https://daviddalpiaz.github.io/r4sl/elastic-net.html
-```{r}
-library(caret)
-tibble::as_tibble(assessment_aggregate)
-cv_10 = trainControl(method = "cv", number = 10)
-elastic_net = train(
-  rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
-  method = "glmnet",
-  trControl = cv_10
-)
-
-elastic_net
-
-```
-The final values used for the model were alpha = 0.1 and lambda = 0.007102922.
-   alpha  lambda       RMSE      Rsquared   MAE     
-  0.10   0.007102922  2.236598  0.8625398  1.708885
-
-Prediction power
-train/test
-```{r}
-set.seed(450)
-train_ind<-sample(218,218-50)
-train<-assessment_aggregate[train_ind,]
-test<-assessment_aggregate[-train_ind,]
-```
-
-Full linear model
-```{r}
-
-newx<-test[,-c(8,9)]
-y<-test[,c(8)]
-linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount,data=train)
-resid<-predict(linear_1,newdata = newx) - y
-sqrt(sum(resid^2)/nrow(test))
-```
-MSPE =  2.590233
-
-Reduced model
-```{r}
-# reduced variable
-linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal,data=train)
-resid<-predict(linear_2,newdata = newx) - y
-sqrt(sum(resid^2)/nrow(test))
-```
-MSPE = 2.523658
-
-Lasso
-```{r}
-# create the whole matrix
-y<-as.matrix(assessment_aggregate$rate)
-dim(x) # 165  29
-dim(y)
-# creat x_train matrix and y_train
-x_train<-x[train_ind,]
-y_train<-y[train_ind,]
-# create x_test matrix
-x_test<-x[-train_ind,]
-y_test<-y[-train_ind,]
-
-# Setting alpha = 1 implements lasso regression
-set.seed(450)
-lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
-
-# Best 
-lambda_best <- lasso_reg$lambda.min;lambda_best
-
-lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
-
-predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
-eval_results(y_test, predictions_test)
-```
-MSPE = 2.528047
-
-Ridge
-```{r}
-ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
-set.seed(450)
-cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
-optimal_lambda <- cv_ridge$lambda.min
-optimal_lambda
-predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
-eval_results(y_test, predictions_test)
-```
-MSPE = 2.567499
-
-Elastic Net
-```{r}
-tibble::as_tibble(assessment_aggregate[train_ind,])
-cv_10 = trainControl(method = "cv", number = 10)
-elastic_net = train(
- rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, 
- data = assessment_aggregate[train_ind,],
-  method = "glmnet",
-  trControl = cv_10
-)
-elastic_net
-
-# RMSE was used to select the optimal model using the smallest value.
-# The final values used for the model were alpha = 1 and lambda = 0.06549203.
-
-elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.06549203)
-predictions_test <- predict(elastic_reg, newx = x_test)
-eval_results(y_test, predictions_test)
-```
-MSPE = 2.54861
-
-Summary on GOF
-```{r}
-
-GOF_table <- rbind(
-      c("OLR full", 0.8874, 0.8707, 1.9843, 2.5902),
-      c("OLR reduced", 0.8874, 0.8721, 1.9845, 2.5237),
-      c("Ridge", 0.8868, 0.8707, 1.9896, 2.5675),
-      c("LASSO", 0.8873, 0.8707, 1.9855, 2.5280),
-      c("Elastic Net", 0.8625, 0.8707, 2.2366, 2.5486))
-colnames(GOF_table) <- c("Model", "Mutiple R-Squared","Adjusted R_Squared", "MSE", "PMSE")
-
-```
 
@@
 x.data.frame<-assessment_transform
-xm.data.frame<-assessment_transform[,-c(1,8)]
+xm.data.frame<-assessment_transform[,-c(1,8,9,10,11)]
 y.data.frame<-assessment_transform$rate
@@
                       data = x.data.frame[ii != j, ])
-    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+AvgTotalAssessmentValue, 
+    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, 
                    data = x.data.frame[ii != j, ])
@@
 ## choose full model
-linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount,data=assessment_aggregate)
+linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,data=assessment_aggregate)
 

### 2020-03-19 src/2nd_draft.Rmd
added text from first draft, strat revising..

@@
 
+## Summary
+
+Our client is a real estate property tax analyst who analyzes the market trend and makes market predictions for his company and future clients.
+The main objective of our project is to accurately predict future mill rate (property tax) in metro Vancouver for the following 3 property tax classes: Tax class 1: Residential, Tax class 5: Light industry and Tax class 6: Business and other.
+Data cleaning, exploratory data analysis are used in this project to analyze the relationship between mill rate and other factors.
+Data cleaning is performed to aggregate our data into summary statistics.
+Exploratory analysis has shown there is a linear trend in mill rates over the years; it has also shown there is a fairly strong correlation between mill rate and average assessment per property in different municipalities.
+Ordinary Linear Model and a variety of its variants are used to predict the mill rate.
+A full assessment of the performance of these models is shown below.
+
+## Introduction
+
+Our goal is to predict mill rates in 2020.
+We also seek to identify which explanatory variables are the most important in determining mill rates.
+Every year, the assessment value of each property is released at the beginning of the year, however, the mill rate is still unknown until Spring.
+Prediction of mill rate is a focus of interest because it gives an approximate property tax to pay for property owners.
+It is also important because it might affect future buyers’ purchasing incentives.
+The property tax rate has a fairly small margin to change.
+Mill rate is adjusted based on the total assessment in each city so the municipal government can use tax earning (total assessment * mill rate) to match their annual expense in order to balance the city’s budget.
+
+Correlation between the mill rate and each explanatory variable will be asses to pick the essential variables in our model.
+Then, a variety of linear models will be fitted using our selected variable.
+The best model is selected based on its prediction power and goodness of fit.
+
+## Data Description
+
+For exploratory data analysis and building predictive model, we used past property assessment data, which is provided by our client.
+It contains information about some commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019.
+We selected variables that we believe were relevant to the property tax prediction.
+These selected variables are:
+
+- Mill Rates (2016-2019)
+- Assessment (2016-2019)
+- Tax Class Code (01, 05, 06)
+- Area Code (Municipality)
+- Year
+- Land Total (2016 - 2019)
+- Improvement Total (2016 - 2019)
+- Number of properties
+
+As the client suggested, we have also collected external data from the Government of British Columbia which could be associated with the tax rate:
+
+- Municipal Budget of Cities in Metro Vancouver (Taxes Imposed & Collected, Schedule 706)
+
+We decided to calculate the sum of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, and Budget for each combination of municipality and the tax class to reduce the dimension of our data.
+The mill rate is unique for each combination of municipality and tax class, so we have added this information into our data for each municipality and its tax classes.
+
+## Methods
+
+### Exploratory Analysis
+
+Before we made any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs.
+Assessment, Mill Rate vs.
+Municipal Budget, Mill Rate vs.
+Land Total, Mill Rate vs.
+Imrpovement Total and Mill Rate vs.
+Number of Properties.
+
+The municipal government wants their tax income to match their annual spending in order to balance their bedget.
+Given this piece of information, we want to find and model the relationship between municipal tax income and municipal budget.
+We plotted a graph of tax income for three different tax classes (01, 05, and 06) vs.
+budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct.
+We chose to illustrate such relationship using scatter plot below for Vancouver, Richmond and Burnaby as they are the most representing city in Metro Vancouver.
+
+Mill rate is mainly affected by assessment and municipal budget, so we have plotted scatter plot of mill rate vs.
+assessment and mil rate vs.
+municipal budget to see the correlation between each pair of the two factors.
+ANOVA analysis is also performed to see the correlation between mill rate vs.
+tax class, mill rate vs.
+year and mill rate vs.
+municipality.
+
+### Measure of goodness of fit and prediction power
+
+In this study, we built a few linear models and evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.
+The definition of goodness of fit and prediction power is given below.
+
+ - **Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data.
+R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.
+
+ - **Prediction power** measures how well models can predict the future values.
+We use mean squared prediction power to compare the prediction performance across all of our fitted linear predictive models.
+To do that, we will divide the data set into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.
+
+### Linear model
+
+For our linear model, Year, TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties and tax (budget) are considered.
+We first explore the full linear model using all the available features.
+From our exploratory data analysis, we have found Vancouver, Richmond, Surrey and Burnaby are outliers.
+Further prediction model will be built based on these findings.
+We are also intesterd in how different tax class can affect our predictive model, so models based on different tax classes will also be considered in the future.
+For our reduced model, we build our linear predictor using only the significant variables from the full model (See table 4).
+The reduced model's performence in terms of goodness of fit will decrease with fewer features, however, the performance of prediction power might be improved.
+
+### Ridge, Lasso and Elastic Net
+
+Other than simple linear regression, we are also interested in the performance of Ridge Regression, Lasso, and Elastic Net.
+They are more advanced than simple linear regression as they reduce the variance of the model.
+In these three models, weights are assigned to features.
+Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively.
+Elastic net is a combination of Ridge and Lasso.
+
+We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.
+
+\pagebreak
+
+## Results
+
+###Exploratory Analysis
+
+We assume that the government aims to match its budget and its income by adjusting the mill rates
 1. data processing
@@
 ```
+
+We calculated the total tax income of each municipality by summing the product of assessment total and mill rate of all tax class.
+Then we plotted the computed tax income vs municipal budget into line-chart to visualize their relationship.
+We chose Vancoouver, Richmond and Burnaby as representative because they are the three largest municipalities in Metro Vancouver.
+The plots showed that the computed tax amount was positively and approximatly linearly related to municipal budget through time.
 
@@
 ```
+
+###Linear Model
+
+**Table 2: Result from Model Used**
+
+Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE 
+------|--------|--------|--------|--------
+      OLR full | 0.8874| 0.8707| 1.9843| 2.5902
+      OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237
+      Ridge|  0.8868| NA| 1.9896| 2.5675
+      LASSO| 0.8873| NA| 1.9855| 2.5280
+      Elastic Net| 0.8625| NA| 2.2366| 2.5486
+
+Above is the goodness of fit of five models.
+Except for the Elastic Net model which has higher MSE and PMSE, there are no significant differences between these models.
+We have not used cross-validation to reduce model overfitting.
+Although the results are relatively similar and have high R_squared across all models, we believe after using cross-validation there will be a clearer difference between the goodness of fit of each model.
+
+For a dtailed look of how the models were fitted and evaluated, please refer to the appendix.
+
+## Conclusion
+
+From our exploratory data analysis, we can see that assessment total, land total, tax code, year and municipalities are all significant in our model.
+The client suggested that we can transform numerical factors such as assessment total into percentage change.
+This data transformation does not perform as well as we expected, which only yields an R^2 of around 0.2 across all of our fitted models.
+We believe the method that the client has suggested might leave out some important information about the housing market in each municipality.
+We have also found that when comparing the correlation between mill rate against features in our models, Vancouver is an outlier.
+We would like to further investigate Vancouver as a special case.
+We are also interested in the effect of different tax classes in predicting mill rate.
+Further analysis and prediction will be performed based on our hypothesis.
+
+All fitted linear model were able to make accurate prediction based on means squared error, but there is still a very high chance that our model is overfitted.
+For our next report, we are going to use cross-validation to reduce the effect of overfitting.
+
+## References
+
+Links to source of data:
+
+- Schedule 706 (https://www2.gov.bc.ca/gov/content/governments/local-governments/facts-framework/statistics/statistics)
+
+Code repository:
+
+- Data Cleaning (https://github.com/STAT450-550/RealEstate/blob/450/src/Data_Cleaning.Rmd)
+
+- Exploratory Data Analysis and Model Fitting (https://github.com/STAT450-550/RealEstate/blob/450/src/EDA%26mode_fitting.Rmd)
+
+## Appendix
+
+### Missing Value:
+
+There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values
+Based on client information, all properties in the same region, classcode, and year should have a unique class rate
+
+- For entries with mill rate, we aggregated them into groups by region + classcode + year.
+
+- For entries without mill rate, we found the group they belong to and assign them mill rate in that group.
+
+###Here is some exceptions found:
+
+Some groups' mill rate is not unique:
+
+- In Delta, properties in different neighbourhoods have slightly different mill rates. Since the variance is not significant, we take the mean as the overall mill rate in groups.
+
+- In Vancouver, 2019, Class 01, one property's mill rate is different from others. It is regarded as an outlier.
+
+- In Burnaby, 2019, Class 06, six properties' mill rate are different from others. They are regarded as outliers.
+
+- In Langley, 2019, Class 06, mill rate is different between assessment type. After talking to the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.
+
+- In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:
+
+**Table 3: Data with missing mill rate**
+
+Year  | Region | Class |Number of Properties
+------|--------|-------|----------
+2016 | Belcarra | 06 | 9 
+2016 | Lions Bay | 01 | 40 
+2016 | Lions Bay | 06 | 25 
+2016 | Maple Ridge Rural | 05 | 36 
+2017 | Belcarra | 06 | 9 
+2017 | Lions Bay | 01 | 39 
+2017 | Lions Bay | 06 | 24 
+2017 | Maple Ridge Rural | 05 | 36 
+2018 | Maple Ridge Rural | 05 | 36 
+2019 | Maple Ridge Rural | 05 | 38

### 2020-03-19 src/2nd_draft.Rmd
format

@@
 1. data processing
-```{r, warning=F, message=F}
+```{r, warning=F, message=F, echo=FALSE}
 ## Aggregated Data
@@
 2. Test correlations between past mill rates and other features.
-```{r}
+```{r, echo=FALSE}
 # ______________________________________________________________________________________________
@@
 
-```{r, warning=FALSE}
+```{r, warning=FALSE, echo=FALSE}
+lambdas <- 10^seq(2, -3, by = -.1)
 dummy_municipal<-dummy(assessment_transform$AddressAssessorMunicipalityDesc)
@@
 
-```{r fit}
+```{r fit, echo=FALSE}
 ## choose full model
-linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,data=assessment_aggregate)
+linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,data=assessment_transform)
 

### 2020-03-19 src/2nd_draft.Rmd
cleanup

@@
 names(assessment_transform)[9] <- paste("AvgTotalAssessmentValue")
-assessment_transform[,10] <- assessment_transform$landTotal/assessment_transform$propertyCount
-names(assessment_transform)[10] <- paste("AvgTotalLandValue")
-assessment_transform[,11] <- assessment_transform$improvementTotal/assessment_transform$propertyCount
-names(assessment_transform)[11] <- paste("AvgTotalImprovementValue")
 
 cor(assessment_transform$rate,assessment_transform$AvgTotalAssessmentValue)
-cor(assessment_transform$rate,assessment_transform$AvgTotalLandValue)
-cor(assessment_transform$rate,assessment_transform$AvgTotalImprovementValue)
 
@@
                    ggtitle(sprintf("mill rate vs average total assessment value")) + 
-                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-assessment_transform %>%
-                   ggplot(aes(x=AvgTotalLandValue,
-                              y=rate,group=AddressAssessorMunicipalityDesc,
-                              color=AddressAssessorMunicipalityDesc)) + 
-                   geom_point() + 
-                   ggtitle(sprintf("mill rate vs average total land value")) + 
-                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-
-assessment_transform %>%
-                   ggplot(aes(x=AvgTotalImprovementValue,
-                              y=rate,group=AddressAssessorMunicipalityDesc,
-                              color=AddressAssessorMunicipalityDesc)) + 
-                   geom_point() + 
-                   ggtitle(sprintf("mill rate vs average total improvement value")) + 
                    geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
@@
 # correct col name
-#assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax
 colnames(x)[26]<-c("AvgTotalAssessmentValue")
@@
                    data = x.data.frame[ii != j, ])
-    # tmp.n<-lm(rate ~ 1, data = x.data.frame[ii != j, ])
-
-    # tmp.elastic<-train(
-    #   rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,
-    #   data = x.data.frame[ii != j, ],
-    #   method = "glmnet",
-    #   trControl = cv_10)
 
@@
     pr1.f[ii != j] <- predict(tmp.full, newdata = x.data.frame[ii != j, ])
-    # pr1.n[ii != j] <- predict(tmp.n, newdata = x.data.frame[ii != j, ])
-
-    # index<-which.min(tmp.elastic$results$RMSE)
-    # pr1.el[ii != j] <- predict(tmp.elastic,
-    #                           alpha = tmp.elastic$results[index,1],
-    #                           lambda = tmp.elastic$results[index,2],
-    #                           newdata = xm.data.frame[ii == j, ])
 
@@
     pr.f[ii == j] <- predict(tmp.full, newdata = x.data.frame[ii == j, ])
-    # pr.n[ii == j] <- predict(tmp.n, newdata = x.data.frame[ii == j, ])
-    # index<-which.min(tmp.elastic$results$RMSE)
-    # pr.el[ii == j] <- predict(tmp.elastic,
-    #                           alpha = tmp.elastic$results[index,1],
-    #                           lambda = tmp.elastic$results[index,2],
-    #                           newdata = xm.data.frame[ii == j, ])
   }
@@
   mspe1.f[i]<-mean((y - pr1.f)^2)
-  # mspe1.n[i]<-mean((y - pr1.n)^2)
-  # mspe1.el[i]<-mean((y - pr1.el)^2)
 
@@
   mspe.f[i]<-mean((y - pr.f)^2)
-  # mspe.n[i]<-mean((y - pr.n)^2)
-  # mspe.el[i]<-mean((y - pr.el)^2)
-
 }
@@
         mspe1.ri, 
-        # mspe1.el, 
-        # mspe1.n, 
         mspe1.reduced, 
@@
                   "Ridge",
-                  # "Elastic Net", 
                   "Reduced", 
@@
         mspe.ri,
-        # mspe.el, 
-        # mspe.n, 
         mspe.reduced, 
@@
         names = c("LASSO", "Ridge", 
-                  # "Elastic Net", 
                   "Reduced", "Full"), 
         col = c("steelblue","gray80", 
-                # "pink", 
                 "tomato", "springgreen"), 
@@
         ylab = expression(hat(MSPE)))
-
 ```
@@
 ```{r fit, echo=FALSE}
-## choose full model
+## choose transformed model
 linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+AvgTotalAssessmentValue,data=assessment_transform)

### 2020-03-19 src/2nd_draft.Rmd
more formatting

@@
 
-We assume that the government aims to match its budget and its income by adjusting the mill rates
-1. data processing
 ```{r, warning=F, message=F, echo=FALSE}
+## data setup
 ## Aggregated Data
@@
 
-We calculated the total tax income of each municipality by summing the product of assessment total and mill rate of all tax class.
-Then we plotted the computed tax income vs municipal budget into line-chart to visualize their relationship.
-We chose Vancoouver, Richmond and Burnaby as representative because they are the three largest municipalities in Metro Vancouver.
-The plots showed that the computed tax amount was positively and approximatly linearly related to municipal budget through time.
-
-2. Test correlations between past mill rates and other features.
 ```{r, echo=FALSE}
+## Test correlations between past mill rates and other features.
 # ______________________________________________________________________________________________
@@
 
-3.model fitting
-
 ```{r, warning=FALSE, echo=FALSE}
+## model fitting
+set.seed(123)
 lambdas <- 10^seq(2, -3, by = -.1)
@@
 dummy_taxclass<-dummy(assessment_transform$TaxClassCode)
+
 # build x matrix
@@
 x<-cbind(x,y)
+
 # correct col name
@@
 ii <- (1:n)%%k + 1
-set.seed(123)
 N <- 50
@@
 mspe.n<-mspe.el<-mspe.la <- mspe.f <- mspe.ri <- mspe.reduced <- rep(0, N)
+
 for (i in 1:N) {

### 2020-03-19 src/2nd_draft.Rmd
revised Data Description

--- 2020-03-19 src/2nd_draft.Rmd
@@
 
-Our client is a real estate property tax analyst who analyzes the market trend and makes market predictions for his company and future clients.
 The main objective of our project is to accurately predict future mill rate (property tax) in metro Vancouver for the following 3 property tax classes: Tax class 1: Residential, Tax class 5: Light industry and Tax class 6: Business and other.
@@
 Data cleaning is performed to aggregate our data into summary statistics.
-Exploratory analysis has shown there is a linear trend in mill rates over the years; it has also shown there is a fairly strong correlation between mill rate and average assessment per property in different municipalities.
-Ordinary Linear Model and a variety of its variants are used to predict the mill rate.
-A full assessment of the performance of these models is shown below.
+Exploratory analysis has shown there are strong relationships between mill rate and tax class and mill rate and municipalities; it has also shown there is a fairly strong correlation between mill rate and average assessment per property in different municipalities.
+Ordinary linear model, reduced ordinary linear model, Ridge Regression and LASSO are used to predict the mill rate.
+A full assessment of the performance - prediction power and goodness of fit of these models, is shown below.
 
@@
 
-Our goal is to predict mill rates in 2020.
+Our goal is to predict mill rates in 2020 in the following 22 municipalities:
+
+- Burnaby
+- Coquitlam
+- Delta
+- Langley - City
+- Langley - Township
+- Maple Ridge
+- Maple Ridge Rural
+- New Westminster
+- North Vancouver - City
+- North Vancouver - Dist
+- Pitt Meadows
+- Port Coquitlam
+- Port Moody
+- Richmond
+- Surrey
+- Vancouver
+- White Rock
+- West Vancouver
+- Bowen Island
+- Anmore
+- Belcarra
+- Lions Bay.
+
 We also seek to identify which explanatory variables are the most important in determining mill rates.
@@
 The property tax rate has a fairly small margin to change.
-Mill rate is adjusted based on the total assessment in each city so the municipal government can use tax earning (total assessment * mill rate) to match their annual expense in order to balance the city’s budget.
+Mill rate is adjusted based on the total assessment in each city so the municipal government can use tax earning (total assessment * mill rate) to match their annual expense to balance the city’s budget.
 
-Correlation between the mill rate and each explanatory variable will be asses to pick the essential variables in our model.
+Correlation between the mill rate and each explanatory variable will be used to pick the essential variables in our model.
 Then, a variety of linear models will be fitted using our selected variable.
@@
 
-For exploratory data analysis and building predictive model, we used past property assessment data, which is provided by our client.
-It contains information about some commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019.
+For exploratory data analysis and building predictive models, we used past property assessment data, which was provided by our client.
+It contained information about some commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019.
 We selected variables that we believe were relevant to the property tax prediction.
@@
 
-- Mill Rates (2016-2019)
-- Assessment (2016-2019)
-- Tax Class Code (01, 05, 06)
-- Area Code (Municipality)
+- Mill Rates ("rate": 2016-2019)
+- Assessment ("assessTotal": 2016-2019)
+- Tax Class Code ("TaxClassCode": 01, 05, 06)
+- Municipality ("AddressAssessorMunicipalityDesc"")
 - Year
-- Land Total (2016 - 2019)
-- Improvement Total (2016 - 2019)
-- Number of properties
+- Land Total ("landTotal": 2016 - 2019)
+- Improvement Total ("improvementTotal": 2016 - 2019)
+- Number of properties ("propertyCount")
 
-As the client suggested, we have also collected external data from the Government of British Columbia which could be associated with the tax rate:
-
-- Municipal Budget of Cities in Metro Vancouver (Taxes Imposed & Collected, Schedule 706)
-
-We decided to calculate the sum of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, and Budget for each combination of municipality and the tax class to reduce the dimension of our data.
+We decided to calculate the sum of TotalAssessedValue, AverageTotalAssessedValue, LandAssessedValue, and ImprovementAssessedValue for each combination of municipality and the tax class to reduce the dimension of our data.
 The mill rate is unique for each combination of municipality and tax class, so we have added this information into our data for each municipality and its tax classes.

### 2020-03-19 src/2nd_draft.Rmd
extension

@@
 
-Before we made any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs.
+Before any prediction on the future mill rate of Metro Vancouver’s real estate market were made, some exploratory analysis to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs.
 Assessment, Mill Rate vs.
-Municipal Budget, Mill Rate vs.
 Land Total, Mill Rate vs.
 Imrpovement Total and Mill Rate vs.
-Number of Properties.
+Number of Properties were performed.
 
-The municipal government wants their tax income to match their annual spending in order to balance their bedget.
-Given this piece of information, we want to find and model the relationship between municipal tax income and municipal budget.
-We plotted a graph of tax income for three different tax classes (01, 05, and 06) vs.
-budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct.
-We chose to illustrate such relationship using scatter plot below for Vancouver, Richmond and Burnaby as they are the most representing city in Metro Vancouver.
-
-Mill rate is mainly affected by assessment and municipal budget, so we have plotted scatter plot of mill rate vs.
-assessment and mil rate vs.
-municipal budget to see the correlation between each pair of the two factors.
-ANOVA analysis is also performed to see the correlation between mill rate vs.
-tax class, mill rate vs.
-year and mill rate vs.
+Mill rate is mainly affected by assessment, so some scatter plots of mill rate vs.
+various assessment were created to see the correlation between each pair of the two factors.
+KW analysis is also performed to see the correlation between mill rate vs.
+tax class and mill rate vs.
 municipality.

### 2020-03-19 doc/RealEstate_Peter_Han.Rmd
revisied Methods - EDA

@@
 - Maple Ridge Rural
+- New Westminster
 - North Vancouver - City

### 2020-03-19 src/2nd_draft.Rmd
revisied Methods - EDA

@@
 
-Before any prediction on the future mill rate of Metro Vancouver’s real estate market were made, some exploratory analysis to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs.
+Before any prediction on the future mill rate of Metro Vancouver’s real estate market was made, exploratory data analysis was performed to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs.
 Assessment, Mill Rate vs.
 Land Total, Mill Rate vs.
-Imrpovement Total and Mill Rate vs.
+Improvement Total and Mill Rate vs.
 Number of Properties were performed.
+From our initial analysis, outliers in municipalities had been found.
+Data transformation - calculating the average total assessment, was used to reduce the effect of outliers.
 
-Mill rate is mainly affected by assessment, so some scatter plots of mill rate vs.
-various assessment were created to see the correlation between each pair of the two factors.
-KW analysis is also performed to see the correlation between mill rate vs.
+Mill rate is mainly affected by assessment, so scatter plots of mill rate vs.
+total, land, and improvement assessment were created to see the correlation between each pair of the two factors.
+Kruskal Wallis analysis was also performed to see the correlation between mill rate vs.
 tax class and mill rate vs.

### 2020-03-19 src/2nd_draft.Rmd
assign
assign more

@@
 Assessment, Mill Rate vs.
-Land Total, Mill Rate vs.
-Improvement Total and Mill Rate vs.
-Number of Properties were performed.
+Land Total, and Mill Rate vs.
+Improvement Total were performed.
 From our initial analysis, outliers in municipalities had been found.
@@
 
+## TODO modify this part based on the comment in the following link (Leonie)
+## _______________________________
+https://github.com/STAT450-550/RealEstate/blob/450/doc/Peter_Han_Individual_edit.pdf
 ### Measure of goodness of fit and prediction power
@@
 We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.
+##_____________________________________________
 
@@
 assessment_transform %>%
+                   ggplot(aes(x=assessTotal,
+                              y=rate,group=AddressAssessorMunicipalityDesc,
+                              color=AddressAssessorMunicipalityDesc)) + 
+                   geom_point() + 
+                   ggtitle(sprintf("mill rate vs total assessment value")) + 
+                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+assessment_transform %>%
+                   ggplot(aes(x=landTotal,
+                              y=rate,group=AddressAssessorMunicipalityDesc,
+                              color=AddressAssessorMunicipalityDesc)) + 
+                   geom_point() + 
+                   ggtitle(sprintf("mill rate vs total land assessment value")) + 
+                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+assessment_transform %>%
+                   ggplot(aes(x=improvementTotal,
+                              y=rate,group=AddressAssessorMunicipalityDesc,
+                              color=AddressAssessorMunicipalityDesc)) + 
+                   geom_point() + 
+                   ggtitle(sprintf("mill rate vs total improvement assessment value")) + 
+                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+
+assessment_transform %>%
                    ggplot(aes(x=AvgTotalAssessmentValue,
@@
 ```
+
+## ___________________________
+## TODO KW result table, explanation of all plots (Yuting)
+## ___________________________
+
+##_____________________________
+## TODO modify this part using boxplot and tables, change desrciptions (Leonie)
+###Linear Model
+
+**Table 2: Result from Model Used**
+
+Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE 
+------|--------|--------|--------|--------
+      OLR full | 0.8874| 0.8707| 1.9843| 2.5902
+      OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237
+      Ridge|  0.8868| NA| 1.9896| 2.5675
+      LASSO| 0.8873| NA| 1.9855| 2.5280
+      Elastic Net| 0.8625| NA| 2.2366| 2.5486
+
+Above is the goodness of fit of five models.
+Except for the Elastic Net model which has higher MSE and PMSE, there are no significant differences between these models.
+We have not used cross-validation to reduce model overfitting.
+Although the results are relatively similar and have high R_squared across all models, we believe after using cross-validation there will be a clearer difference between the goodness of fit of each model.
+
+For a dtailed look of how the models were fitted and evaluated, please refer to the appendix.
 
@@
 ```
-
-###Linear Model
-
-**Table 2: Result from Model Used**
-
-Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE 
-------|--------|--------|--------|--------
-      OLR full | 0.8874| 0.8707| 1.9843| 2.5902
-      OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237
-      Ridge|  0.8868| NA| 1.9896| 2.5675
-      LASSO| 0.8873| NA| 1.9855| 2.5280
-      Elastic Net| 0.8625| NA| 2.2366| 2.5486
-
-Above is the goodness of fit of five models.
-Except for the Elastic Net model which has higher MSE and PMSE, there are no significant differences between these models.
-We have not used cross-validation to reduce model overfitting.
-Although the results are relatively similar and have high R_squared across all models, we believe after using cross-validation there will be a clearer difference between the goodness of fit of each model.
-
-For a dtailed look of how the models were fitted and evaluated, please refer to the appendix.
+##_______________________________________
 

### 2020-03-19 src/2nd_draft.Rmd
assigned more

@@
 ```
+##_______________________________________
 
@@
 ```
-##_______________________________________
 
+##_____________________________
+## TODO: delete pct change part, add prediction based on our chosen model (Peter)
 ## Conclusion
@@
 For our next report, we are going to use cross-validation to reduce the effect of overfitting.
+##_______________________________
 

### 2020-03-19 src/2nd_draft.Rmd
final assign
one more assign

@@
 assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))[,-1]
-#assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
-#assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
+assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
+assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
 ```
@@
 
+##_______________________________________
+##TODO: make prediction work for 2020 data (Yuetong)
+##(it's missing AvgTotalAssessmentValue, looks like we need to calculate them)
 ```{r fit, echo=FALSE}
@@
 
+\pagebreak
 ## References
@@
 
+##——————————————————————————
+##TODO: more styling (Yuting)
+##___________________________
 ## Appendix
+
+##__________________________
+## TODO: add one more section: (Yuetong)
+## Include the code relevant to summarizing and aggregating the data from the client, which produces in the dataframe used for analysis.
+##__________________________
 

### 2020-03-19 src/2nd_draft.Rmd
Data Discription

@@
 
-For exploratory data analysis and building predictive models, we used past property assessment data, which was provided by our client.
-It contained information about some commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019.
-We selected variables that we believe were relevant to the property tax prediction.
-These selected variables are:
+Our client provides us the past 5 years' property assessment data in BC.
+Since we are only interested in predicting mill rate for metro Vancouver and specific class code, we get a subset of properties that satisfy our interest:
 
-- Mill Rates ("rate": 2016-2019)
-- Assessment ("assessTotal": 2016-2019)
-- Tax Class Code ("TaxClassCode": 01, 05, 06)
-- Municipality ("AddressAssessorMunicipalityDesc"")
+- TaxClass Code in (01,05,06)
+- Municipality in Metro Vancouver
+
+Moreover, we select 5 features that could be relevant to mill rate, which are:
+
+- Tax Year
+- Municipality 
+- Tax Class
+- Assessment Type
+- Assessment Value
+
+There are 1801 missing value in mill rate. 1509 are imputed, and 292 are removed from the data frame. The imputation method is mentioned in appendix.
+
+Oo reduce the dimension of our data, we aggregate all properties in the same region, classcode, and year are aggregate into a group, becasue these properties have same mill rate,  which is our response variable.
+Here is the summary statistics for these groups:
+
+- Mill rate 
+- Total Assessment 
+- Total Land Assessment 
+- Total Improvement Assessment 
+- Total number of properties 
+- Taxclasscode
+- Municipality
 - Year
-- Land Total ("landTotal": 2016 - 2019)
-- Improvement Total ("improvementTotal": 2016 - 2019)
-- Number of properties ("propertyCount")
-
-We decided to calculate the sum of TotalAssessedValue, AverageTotalAssessedValue, LandAssessedValue, and ImprovementAssessedValue for each combination of municipality and the tax class to reduce the dimension of our data.
-The mill rate is unique for each combination of municipality and tax class, so we have added this information into our data for each municipality and its tax classes.
 
@@
 assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
-assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
+# assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
 ```
@@
 assessment_2020.agg <- assessment_2020%>%
+  mutate(AvgTotalAssessmentValue = assessTotal/propertyCount)%>%
   mutate(predict_rate = predict(linear_1,newdata = .))
@@
 
-## Final result
-municipality.list = c(
-  "Burnaby", 
-  "Coquitlam", 
-  "Delta", 
-  "Langley - City", 
-  "Langley - Township",
-  "Maple Ridge",
-  "Maple Ridge Rural", 
-  "North Vancouver - City",
-  "North Vancouver - Dist",
-  "Pitt Meadows", 
-  "Port Coquitlam", 
-  "Port Moody", 
-  "Richmond", 
-  "Surrey", 
-  "Vancouver", 
-  "White Rock", 
-  "West Vancouver", 
-  "Bowen Island", 
-  "Anmore", 
-  "Belcarra",
-  "Lions Bay",
-  "New Westminster")
-
-## Raw Data
-a <- assessment_2020.full%>%
-  filter(TaxClassCode %in% c("01","05","06"),
-         AddressAssessorMunicipalityDesc %in% municipality.list)
+# ## Final result
+# municipality.list = c(
+#   "Burnaby", 
+#   "Coquitlam", 
+#   "Delta", 
+#   "Langley - City", 
+#   "Langley - Township",
+#   "Maple Ridge",
+#   "Maple Ridge Rural", 
+#   "North Vancouver - City",
+#   "North Vancouver - Dist",
+#   "Pitt Meadows", 
+#   "Port Coquitlam", 
+#   "Port Moody", 
+#   "Richmond", 
+#   "Surrey", 
+#   "Vancouver", 
+#   "White Rock", 
+#   "West Vancouver", 
+#   "Bowen Island", 
+#   "Anmore", 
+#   "Belcarra",
+#   "Lions Bay",
+#   "New Westminster")
+# 
+# ## Raw Data
+# a <- assessment_2020.full%>%
+#   filter(TaxClassCode %in% c("01","05","06"),
+#          AddressAssessorMunicipalityDesc %in% municipality.list)
 
@@
 b <- assessment_2020.agg%>%
-  select(-c(Year, assessTotal : rate))%>%
+  select(-c(assessTotal : propertyCount, AvgTotalAssessmentValue))%>%
   mutate(TaxClassCode = as.factor(paste0("0",TaxClassCode)))
 
-assessment_2020.final<- merge(a, b, by =c("AddressAssessorMunicipalityDesc", "TaxClassCode"), all.x = TRUE)
+# assessment_2020.final<- merge(a, b, by =c("AddressAssessorMunicipalityDesc", "TaxClassCode"), all.x = TRUE)
 
-write.csv(assessment_2020.final, "../Data/Assessment/assessment_2020.final")
-write.csv(b, "../Data/assessment_2020.csv")
+# write.csv(assessment_2020.final, "../Data/Assessment/assessment_2020.final")
+write.csv(b, "../Data/assessment_2020_fit.csv")
 ```
@@
 
-###Here is some exceptions found:
+#### Here is some exceptions found:
 
@@
 - In Delta, properties in different neighbourhoods have slightly different mill rates. Since the variance is not significant, we take the mean as the overall mill rate in groups.
+
+- In New Westminister, 2019, Class 06, one property's mill rate is different from others. It is regarded as an outlier.
 

### 2020-03-19 src/2nd_draft.Rmd
Update 2nd_draft.Rmd
null model added.

@@
 N <- 50
-mspe1.n<-mspe1.el<-mspe1.la <- mspe1.f <- mspe1.ri <- mspe1.reduced <- rep(0, N)
-mspe.n<-mspe.el<-mspe.la <- mspe.f <- mspe.ri <- mspe.reduced <- rep(0, N)
+mspe1.n<-mspe1.la <- mspe1.f <- mspe1.ri <- mspe1.reduced <- rep(0, N)
+mspe.n<-mspe.la <- mspe.f <- mspe.ri <- mspe.reduced <- rep(0, N)
 
@@
   ii <- sample(ii)
-  pr1.n<-pr1.el<-pr1.la <- pr1.f <- pr1.ri <- pr1.reduced <- rep(0, n)
-  pr.n<-pr.el<-pr.la <- pr.f <- pr.ri <- pr.reduced <- rep(0, n)
+  pr1.n<-pr1.la <- pr1.f <- pr1.ri <- pr1.reduced <- rep(0, n)
+  pr.n<-pr.la <- pr.f <- pr.ri <- pr.reduced <- rep(0, n)
   for (j in 1:k) {
@@
                    data = x.data.frame[ii != j, ])
+    tmp.n <- lm(rate ~ 1, 
+                      data = x.data.frame[ii != j, ])
 
@@
     pr1.f[ii != j] <- predict(tmp.full, newdata = x.data.frame[ii != j, ])
+    pr1.n[ii != j] <-predict(tmp.n, newdata = x.data.frame[ii != j, ])
 
@@
     pr.f[ii == j] <- predict(tmp.full, newdata = x.data.frame[ii == j, ])
+    pr.n[ii == j] <-predict(tmp.n, newdata = x.data.frame[ii == j, ])
+
   }
@@
   mspe1.f[i]<-mean((y - pr1.f)^2)
+  mspe1.n[i]<-mean((y - pr1.n)^2)
 
@@
   mspe.f[i]<-mean((y - pr.f)^2)
+  mspe.n[i]<-mean((y - pr.n)^2)
 }
@@
         mspe1.f, 
+        mspe1.n,
         names = c("LASSO", 
@@
                   "Reduced", 
-                  "Full"), 
+                  "Full",
+                  "Null"), 
         col = c("steelblue","gray80",
-                # "pink", 
+                "pink", 
                 "tomato", "springgreen"), 
@@
         mspe.f, 
+        mspe.n,
         names = c("LASSO", "Ridge", 
-                  "Reduced", "Full"), 
-        col = c("steelblue","gray80", 
+                  "Reduced", 
+                  "Full",
+                  "Null"), 
+        col = c("steelblue","gray80", "pink",
                 "tomato", "springgreen"), 

### 2020-03-19 src/2nd_draft.Rmd
result section updated

@@
 
-Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE 
-------|--------|--------|--------|--------
-      OLR full | 0.8874| 0.8707| 1.9843| 2.5902
-      OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237
-      Ridge|  0.8868| NA| 1.9896| 2.5675
-      LASSO| 0.8873| NA| 1.9855| 2.5280
-      Elastic Net| 0.8625| NA| 2.2366| 2.5486
+<!-- Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE  -->
+<!-- ------|--------|--------|--------|-------- -->
+<!--       OLR full | 0.8874| 0.8707| 1.9843| 2.5902 -->
+<!--       OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237 -->
+<!--       Ridge|  0.8868| NA| 1.9896| 2.5675 -->
+<!--       LASSO| 0.8873| NA| 1.9855| 2.5280 -->
+<!--       Elastic Net| 0.8625| NA| 2.2366| 2.5486 -->
 
-Above is the goodness of fit of five models.
-Except for the Elastic Net model which has higher MSE and PMSE, there are no significant differences between these models.
-We have not used cross-validation to reduce model overfitting.
-Although the results are relatively similar and have high R_squared across all models, we believe after using cross-validation there will be a clearer difference between the goodness of fit of each model.
+Below is a comparison of the five models (Lasso, Ridge, OLR reduced, OLR full and OLR null) as for goodness of fit and predcition power.
+Goodness of fit is measured by MSE of the training data, whereas prediction power is measured by MSPE of the testing data.
+Generally, smaller MSEs and MSPEs indicate better fit and prediction power respectively.
+
+To examine the fit and prediciton power, a 50-run of 10-fold cross validation was performed.
+For each run, a 10-fold corss validation was used to train the five models and make predictions on training and testing data respectively.
+Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding mdoels.
+After the 50 runs, a vector of MSEs and a vectors of MSPEs for each model were therefore constructed successfully.
+
+As for goodness of fit, Lasso, Ridge and OLR reduced have a similar performance; MSEs are around 4.6, but OLR reduced perform slightly better and has a smaller spread (variance) of MSE.
+OLR full performs worse than the the three, around 5.4, and OLR null has the greatest MSE, around 35.
+
+Similarly, as for prediction power, Lasso, Ridge and OLR reduced perform roughly the same; MSPEs are close to 6. OLR full is worse with MSPE around 6.9. OLR Null performs the worst with MSPE over 35.
+
+Therefore, we conclude that from the results of a 50-run of 10-fold cross validation, Lasso, Ridge and OLR reduced have the best goodness of fit and prediction power.
+OLR null perfroms much worse than all the other four models.
 
@@
                   "Full",
-                  "Null"), 
+                  "Null"
+                  ), 
         col = c("steelblue","gray80",
@@
                   "Full",
-                  "Null"), 
-        col = c("steelblue","gray80", "pink",
+                  "Null"
+                  ), 
+        col = c("steelblue","gray80", 
+                "pink",
                 "tomato", "springgreen"), 

### 2020-03-19 src/2nd_draft.Rmd
method and result section updated. name of transformed model fixed.

@@
 
-In this study, we built a few linear models and evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.
+In this study, we built linear models and evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.
 The definition of goodness of fit and prediction power is given below.
@@
 
-### Linear model
+### Ordinary Linear models
 
-For our linear model, Year, TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties and tax (budget) are considered.
-We first explore the full linear model using all the available features.
-From our exploratory data analysis, we have found Vancouver, Richmond, Surrey and Burnaby are outliers.
-Further prediction model will be built based on these findings.
-We are also intesterd in how different tax class can affect our predictive model, so models based on different tax classes will also be considered in the future.
-For our reduced model, we build our linear predictor using only the significant variables from the full model (See table 4).
-The reduced model's performence in terms of goodness of fit will decrease with fewer features, however, the performance of prediction power might be improved.
+TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties are considered.
+The full linear model (ORL full) using all the available features was built first.
+Based on the results of EDA, we selected a list of significant variables and included them in another linear model (OLR transformed).
+To compare the effect of linear models with and without features, a null model (OLR null) with no features used, was also constructed.
 
-### Ridge, Lasso and Elastic Net
+### Ridge and Lasso
 
-Other than simple linear regression, we are also interested in the performance of Ridge Regression, Lasso, and Elastic Net.
-They are more advanced than simple linear regression as they reduce the variance of the model.
-In these three models, weights are assigned to features.
-Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively.
-Elastic net is a combination of Ridge and Lasso.
+Other than ordinary linear regressions, we are also interested in the performance of more advanced linear regressions like Ridge and Lasso.
+Ridge and Lasso have a different objective function to optimize; they take penalty in sum of absolute values and sum of squared absolute values of weights respectively.
+These models have an interesting characteristic, that is, a weight is assigned to each feature.
+This might be relevant to our study as mill rate can be affected by each feature to a different extent.
+Also, another advantage of these model is that MSPE is more stable as the variance of MSPE is reduced.
 
-We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.
+### Cross Validation
+To examine the goodness of fit and prediciton power, a 50-run of 10-fold cross validation would be performed in this study.
+For each run, a 10-fold corss validation was used to train the five models and make predictions on training and testing data respectively.
+Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding mdoels.
+
+After the 50 runs, a vector of MSEs and a vectors of MSPEs for each model were therefore constructed successfully.
+Based on these vectors, boxplots would be used to show the mean and spread of MSE and MSPE aross all models respectively.
+
 ##_____________________________________________
@@
 
-Below is a comparison of the five models (Lasso, Ridge, OLR reduced, OLR full and OLR null) as for goodness of fit and predcition power.
+Below is a comparison of the five models (Lasso, Ridge, OLR transformed, OLR full and OLR null) as for goodness of fit and predcition power.
 Goodness of fit is measured by MSE of the training data, whereas prediction power is measured by MSPE of the testing data.
@@
 
-To examine the fit and prediciton power, a 50-run of 10-fold cross validation was performed.
-For each run, a 10-fold corss validation was used to train the five models and make predictions on training and testing data respectively.
-Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding mdoels.
-After the 50 runs, a vector of MSEs and a vectors of MSPEs for each model were therefore constructed successfully.
-
-As for goodness of fit, Lasso, Ridge and OLR reduced have a similar performance; MSEs are around 4.6, but OLR reduced perform slightly better and has a smaller spread (variance) of MSE.
+As for goodness of fit, Lasso, Ridge and OLR reduced have a similar performance; MSEs are around 4.6, but OLR transformed perform slightly better and has a smaller spread (variance) of MSE.
 OLR full performs worse than the the three, around 5.4, and OLR null has the greatest MSE, around 35.
 
-Similarly, as for prediction power, Lasso, Ridge and OLR reduced perform roughly the same; MSPEs are close to 6. OLR full is worse with MSPE around 6.9. OLR Null performs the worst with MSPE over 35.
+Similarly, as for prediction power, Lasso, Ridge and OLR transformed perform roughly the same; MSPEs are close to 6.
+OLR full is worse with MSPE around 6.9.
+OLR Null performs the worst with MSPE over 35.
 
-Therefore, we conclude that from the results of a 50-run of 10-fold cross validation, Lasso, Ridge and OLR reduced have the best goodness of fit and prediction power.
+Therefore, we conclude that from the results of a 50-run of 10-fold cross validation, Lasso, Ridge and OLR transformed have the best goodness of fit and prediction power.
 OLR null perfroms much worse than all the other four models.
@@
                   "Ridge",
-                  "Reduced", 
+                  "Transformed", 
                   "Full",
@@
        mspe.n,
-        names = c("LASSO", "Ridge", 
-                  "Reduced", 
+        names = c("LASSO", 
+                  "Ridge", 
+                  "Transformed", 
                   "Full",

### 2020-03-20 src/2nd_draft.Rmd
eda updated

@@
 ---
-title: "2nd_draft"
+title: 'STAT 450 Project: Real Estate'
 author: "Xuechun Lu, Yuting Wen, Peter Han, Yuetong Liu"
@@
 output:
-  github_document
+  pdf_document
 ---
@@
 
-Oo reduce the dimension of our data, we aggregate all properties in the same region, classcode, and year are aggregate into a group, becasue these properties have same mill rate,  which is our response variable.
+To reduce the dimension of our data, we aggregate all properties in the same region, tax class code, and year are aggregate into a group, becasue these properties have same mill rate,  which is our response variable.
 Here is the summary statistics for these groups:
@@
 
-## TODO modify this part based on the comment in the following link (Leonie)
-## _______________________________
-https://github.com/STAT450-550/RealEstate/blob/450/doc/Peter_Han_Individual_edit.pdf
 ### Measure of goodness of fit and prediction power
@@
 ### Cross Validation
+
 To examine the goodness of fit and prediciton power, a 50-run of 10-fold cross validation would be performed in this study.
@@
 
-##_____________________________________________
-
 \pagebreak
@@
 ###Exploratory Analysis
+
+#### Continuous Variables
 
@@
 
+Since mill rate is mainly affected by total assessment, which is sum of total land assessment and total improvement assessment, scatter plots between mill rate and all kinds of assessment among different municipalities are created to show the relationship in Figure 1, Figure 2, Figure 3
+
 ```{r, echo=FALSE}
@@
 
-# log transformation and take avg
+# transformation: taking avg
 assessment_transform[,9] <- assessment_transform$assessTotal/assessment_transform$propertyCount
@@
 
-cor(assessment_transform$rate,assessment_transform$AvgTotalAssessmentValue)
+# rename column for better visulization
+assessment_eda <- assessment_transform
+names(assessment_eda)[2] <- paste("Municipality")
+names(assessment_eda)[3] <- paste("Tax Class")
+names(assessment_eda)[4] <- paste("Total Assessment")
+names(assessment_eda)[5] <- paste("Total Land Assessment")
+names(assessment_eda)[6] <- paste("Total Improvement Assessment")
+names(assessment_eda)[8] <- paste("Mill Rate")
+names(assessment_eda)[9] <- paste("Average Total Assessment")
 
-assessment_transform %>%
-                   ggplot(aes(x=assessTotal,
-                              y=rate,group=AddressAssessorMunicipalityDesc,
-                              color=AddressAssessorMunicipalityDesc)) + 
+assessment_eda %>%
+   ggplot(aes(x=`Total Assessment`,
+              y=`Mill Rate`,group=Municipality,
+              color=Municipality)) + 
    geom_point() + 
-                   ggtitle(sprintf("mill rate vs total assessment value")) + 
-                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+   ggtitle(sprintf("Figure 1: Mill Rate vs Total Assessment")) + 
+   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
+   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 
-assessment_transform %>%
-                   ggplot(aes(x=landTotal,
-                              y=rate,group=AddressAssessorMunicipalityDesc,
-                              color=AddressAssessorMunicipalityDesc)) + 
+assessment_eda %>%
+   ggplot(aes(x=`Total Land Assessment`,
+              y=`Mill Rate`,group=Municipality,
+              color=Municipality)) + 
    geom_point() + 
-                   ggtitle(sprintf("mill rate vs total land assessment value")) + 
-                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
-assessment_transform %>%
-                   ggplot(aes(x=improvementTotal,
-                              y=rate,group=AddressAssessorMunicipalityDesc,
-                              color=AddressAssessorMunicipalityDesc)) + 
+   ggtitle(sprintf("Figure 2: Mill Rate vs Total Land Assessment")) + 
+   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
+   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+
+assessment_eda %>%
+   ggplot(aes(x=`Total Improvement Assessment`,
+              y=`Mill Rate`,group=Municipality,
+              color=Municipality)) + 
    geom_point() + 
-                   ggtitle(sprintf("mill rate vs total improvement assessment value")) + 
-                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+   ggtitle(sprintf("Figure 3: Mill Rate vs Total Improvement Assessment")) + 
+   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
+   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+```
 
-assessment_transform %>%
-                   ggplot(aes(x=AvgTotalAssessmentValue,
-                              y=rate,group=AddressAssessorMunicipalityDesc,
-                              color=AddressAssessorMunicipalityDesc)) + 
+There is no clear trend between mill rate and any kind of assessment. Plots also have shown that most points are condensed on the left horizontal axis, since some municipalities have large assessment.
+
+To reduce the effect of large assessment of some municipalities, total assessment across municipalities are transformed by taking total assessment dividing by number of properties of each municipality and tax class.
+The transformed data is named **"Average Total Assessment"**.
+A scatter plot between mill rate and average total assessment is showed in Figure 4.
+
+```{r, echo=FALSE}
+assessment_eda %>%
+ ggplot(aes(x=`Average Total Assessment`,
+            y=`Mill Rate`,group=Municipality,
+            color=Municipality)) + 
  geom_point() + 
-                   ggtitle(sprintf("mill rate vs average total assessment value")) + 
-                   geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+ ggtitle(sprintf("Figure 4: Mill rate v.s. Average Total Assessment")) + 
+ geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") +
+theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+```
 
+Plot from Figure 4 has shown that mill rate tends to decrease as average total assessment increase. Also they have moderately strong linear correlation.
+
+#### Categorical Variables
+
+Here categorical variables were taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 5 and Figure 6.
+
+```{r, echo=FALSE}
 # ______________________________________________________________________________________________
 # CATEGORICAL variables
-for(i in colnames(assessment_transform)[2:3]){
-  print(ggplot(assessment_aggregate, aes(x=as.factor(!!as.name(i)),y=rate, fill=as.factor(!!as.name(i)))) + 
-  geom_boxplot()+ 
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
-}
+assessment_eda %>%
+     ggplot(aes(x=Municipality,
+                y=`Mill Rate`,fill=Municipality)) + 
+     geom_boxplot() + 
+     ggtitle(sprintf("Figure 5: Mill rate across Municipalities")) +
+     theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
 
+assessment_eda %>%
+     ggplot(aes(x=as.factor(`Tax Class`),
+                y=`Mill Rate`,fill=`Tax Class`)) + 
+     geom_boxplot() + 
+     ggtitle(sprintf("Figure 6: Mill rate across Tax Classes")) +
+     theme(legend.position = "none") +
+     labs(x="Tax Class") +
+     scale_x_discrete(labels=c("1", "5", "6"))
+```
+
+Figure 5 has shown that most municipalities have different mean and variance in terms of mill rates.
+
+Figure 6 also supports that there is unequal mean and variance across tax classes.
+
+A statistical test called Kruskal-Wallis Test is also performed to test their distribution.
+It is used to decide if population distributions are identical, and the corresponding p-value which is smaller than 0.05 can indicate that the data has nonidentical distributions.
+
+The results of Kruskal-Wallis Test is shown in Table 1.
+
+**Table 1: Kruskal-Wallis Test of Mill Rate across Municipality and Tax Class**
+
+Distribution  | p-value
+------|--------
+Mill Rate across Municipality | 0.00675
+Mill Rate across Tax Class | < 2.2e-16
+
+The p-values in Table 1 supports that there is nonidentical distributions of mill rates across municipalities and tax classes.
+
+```{r, echo=FALSE, results='hide'}
 # kruskal.test
@@
 
-## ___________________________
-## TODO KW result table, explanation of all plots (Yuting)
-## ___________________________
-
-##_____________________________
-## TODO modify this part using boxplot and tables, change desrciptions (Leonie)
+\pagebreak
 ###Linear Model
-
-**Table 2: Result from Model Used**
-
-<!-- Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE  -->
-<!-- ------|--------|--------|--------|-------- -->
-<!--       OLR full | 0.8874| 0.8707| 1.9843| 2.5902 -->
-<!--       OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237 -->
-<!--       Ridge|  0.8868| NA| 1.9896| 2.5675 -->
-<!--       LASSO| 0.8873| NA| 1.9855| 2.5280 -->
-<!--       Elastic Net| 0.8625| NA| 2.2366| 2.5486 -->
 
@@
 
-For a dtailed look of how the models were fitted and evaluated, please refer to the appendix.
+For a dtailed look of how the models were fitted and evaluated, please refer to the **Appendix**.
 
@@
 ```
-##_______________________________________
 
-##_______________________________________
-##TODO: make prediction work for 2020 data (Yuetong)
-##(it's missing AvgTotalAssessmentValue, looks like we need to calculate them)
 ```{r fit, echo=FALSE}
@@
 
-##——————————————————————————
-##TODO: more styling (Yuting)
-##___________________________
 ## Appendix
-
-##__________________________
-## TODO: add one more section: (Yuetong) 
-## Include the code relevant to summarizing and aggregating the data from the client, which produces in the dataframe used for analysis.
-##__________________________
 

### 2020-03-20 src/2nd_draft.Rmd
formatting and eda updated

@@
 
-## R Markdown
-libraries
-```{r, message=F, warning=F}
+```{r, message=F, warning=F, echo=FALSE}
 library(tidyverse)
@@
 library(dummies)
+set.seed(123)
 ```
@@
 assessment_2020 <- read.csv(here("data","assessment_2020.csv"))[,-1]
-# assessment_2020.full <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
 ```
@@
 
-**Table 1: Kruskal-Wallis Test of Mill Rate across Municipality and Tax Class**
+        Table 1: Kruskal-Wallis Test of Mill Rate across Municipality and Tax Class
 
@@
 \pagebreak
+
 ###Linear Model
@@
 
-As for goodness of fit, Lasso, Ridge and OLR reduced have a similar performance; MSEs are around 4.6, but OLR transformed perform slightly better and has a smaller spread (variance) of MSE.
-OLR full performs worse than the the three, around 5.4, and OLR null has the greatest MSE, around 35.
+The distributions of goodness of fit and prediction power across models are displyed in Figure 7 and Figure 8, respectively, the distribution of null model is removed since it has larger values compared to all other models.
 
-Similarly, as for prediction power, Lasso, Ridge and OLR transformed perform roughly the same; MSPEs are close to 6.
-OLR full is worse with MSPE around 6.9.
-OLR Null performs the worst with MSPE over 35.
-
-Therefore, we conclude that from the results of a 50-run of 10-fold cross validation, Lasso, Ridge and OLR transformed have the best goodness of fit and prediction power.
-OLR null perfroms much worse than all the other four models.
-
-For a dtailed look of how the models were fitted and evaluated, please refer to the **Appendix**.
+The distributions of MSPE across models are also displayed in Table 2.
 
@@
 ## model fitting
-set.seed(123)
 lambdas <- 10^seq(2, -3, by = -.1)
@@
                         nfolds = 10, alpha = 1, family = "gaussian")
-    tmp.reduced <- lm(rate ~ factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+AvgTotalAssessmentValue, 
+    tmp.reduced <- lm(rate ~ factor(TaxClassCode)+
+                        factor(AddressAssessorMunicipalityDesc)+
+                        AvgTotalAssessmentValue, 
                       data = x.data.frame[ii != j, ])
-    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, 
+    tmp.full <- lm(rate ~ factor(AddressAssessorMunicipalityDesc)+
+                     factor(TaxClassCode)+
+                     assessTotal+
+                     landTotal+
+                     improvementTotal+
+                     propertyCount, 
                    data = x.data.frame[ii != j, ])
@@
         mspe1.f, 
-        mspe1.n,
         names = c("LASSO", 
@@
                   "Transformed", 
-                  "Full",
-                  "Null"
+                  "Full"
                   ), 
@@
                 "pink", 
-                "tomato", "springgreen"), 
+                "tomato"), 
         cex.axis = 1, 
@@
         cex.main = 2,
-        main = "Goodness of Fit",
+        main = " Figure 7: Goodness of Fit",
         ylab = expression(hat(MSE)))
@@
         mspe.f,
-       mspe.n,
         names = c("LASSO", 
@@
                 "pink",
-                "tomato", "springgreen"), 
+                "tomato"), 
         cex.axis = 1, 
@@
         cex.main = 2,
-        main = "Prediction Power",
+        main = "Figure 8: Prediction Power",
         ylab = expression(hat(MSPE)))
 ```
+
+                Table 2: Prediction Power of All Models
+
+Model | PMSE:Min | PMSE:Mean | PMSE:Max
+-----|-----|-----|-----|
+Full Model | 6.50 | 6.97 | 7.59 |
+Transformed Model | 5.60 | 5.92 | 6.51 |
+Null Model | 36.67 | 36.87 | 37.26 |
+The Lasso for Full Model | 5.58 | 5.91 | 6.40 |
+Ridge Regression for Full Model | 5.53 | 5.88 | 6.39 |
+
+As for goodness of fit, Lasso, Ridge and OLR reduced have a similar performance; MSEs are around 4.6, but OLR transformed perform slightly better and has a smaller spread (variance) of MSE.
+OLR full performs worse than the the three, around 5.4, and OLR null has the greatest MSE, around 35.
+
+Similarly, as for prediction power, Lasso, Ridge and OLR transformed perform roughly the same; MSPEs are close to 6.
+OLR full is worse with MSPE around 6.9.
+OLR Null performs the worst with MSPE over 35.
+
+Therefore, we conclude that from the results of a 50-run of 10-fold cross validation, Lasso, Ridge and OLR transformed have the best goodness of fit and prediction power.
+OLR null perfroms much worse than all the other four models.
+
+For a detailed look of how the models were fitted and evaluated, please refer to the **Appendix**.
 
@@
 resid<-assessment_2020.agg$predict_rate - assessment_2020.agg$rate
-sqrt(sum(resid^2)/nrow(assessment_2020.agg))
-
-# ## Final result
-# municipality.list = c(
-#   "Burnaby", 
-#   "Coquitlam", 
-#   "Delta", 
-#   "Langley - City", 
-#   "Langley - Township",
-#   "Maple Ridge",
-#   "Maple Ridge Rural", 
-#   "North Vancouver - City",
-#   "North Vancouver - Dist",
-#   "Pitt Meadows", 
-#   "Port Coquitlam", 
-#   "Port Moody", 
-#   "Richmond", 
-#   "Surrey", 
-#   "Vancouver", 
-#   "White Rock", 
-#   "West Vancouver", 
-#   "Bowen Island", 
-#   "Anmore", 
-#   "Belcarra",
-#   "Lions Bay",
-#   "New Westminster")
-# 
-# ## Raw Data
-# a <- assessment_2020.full%>%
-#   filter(TaxClassCode %in% c("01","05","06"),
-#          AddressAssessorMunicipalityDesc %in% municipality.list)
+#sqrt(sum(resid^2)/nrow(assessment_2020.agg))
 
@@
 
-# assessment_2020.final<- merge(a, b, by =c("AddressAssessorMunicipalityDesc", "TaxClassCode"), all.x = TRUE)
-
-# write.csv(assessment_2020.final, "../Data/Assessment/assessment_2020.final")
 write.csv(b, "../Data/assessment_2020_fit.csv")
@@
 \pagebreak
-## References
-
-Links to source of data:
-
-- Schedule 706 (https://www2.gov.bc.ca/gov/content/governments/local-governments/facts-framework/statistics/statistics)
-
-Code repository:
-
-- Data Cleaning (https://github.com/STAT450-550/RealEstate/blob/450/src/Data_Cleaning.Rmd)
-
-- Exploratory Data Analysis and Model Fitting (https://github.com/STAT450-550/RealEstate/blob/450/src/EDA%26mode_fitting.Rmd)
 
@@
 
-**Table 3: Data with missing mill rate**
+                    Table 3: Data with missing mill rate
 
@@
 2019 | Maple Ridge Rural | 05 | 38
+
+\pagebreak
+
+## References
+
+Code repository:
+
+- Data Cleaning (https://github.com/STAT450-550/RealEstate/blob/450/src/Data_Cleaning.Rmd)
+
+- Exploratory Data Analysis and Model Fitting (https://github.com/STAT450-550/RealEstate/blob/450/src/EDA%26mode_fitting.Rmd)

### 2020-03-20 src/2nd_draft.Rmd
grammar change, fully revised, added conclusion

@@
 Our client provides us the past 5 years' property assessment data in BC.
-Since we are only interested in predicting mill rate for metro Vancouver and specific class code, we get a subset of properties that satisfy our interest:
+Since the only interest is in predicting mill rate for metro Vancouver and specific class code, a subset of properties that satisfy our interest have been selected:
 
@@
 
-Moreover, we select 5 features that could be relevant to mill rate, which are:
+Moreover, 5 features have been selected that could be relevant to the mill rate:
 
@@
 
-To reduce the dimension of our data, we aggregate all properties in the same region, tax class code, and year are aggregate into a group, becasue these properties have same mill rate,  which is our response variable.
+To reduce the dimension of our data, all properties in the same region, tax class code, and year are aggregated into a group because these properties have the same mill rate,  which is our response variable.
 Here is the summary statistics for these groups:
 
-- Mill rate 
-- Total Assessment 
-- Total Land Assessment 
-- Total Improvement Assessment 
-- Total number of properties 
-- Taxclasscode
-- Municipality
+- Mill rate ("rate": 2016-2019)
+- Total Assessment ("assessTotal": 2016-2019)
+- Total Land Assessment ("landTotal": 2016 - 2019)
+- Total Improvement Assessment ("improvementTotal": 2016 - 2019)
+- Total number of properties ("propertyCount")
+- Taxclasscode ("TaxClassCode": 01, 05, 06)
+- Municipality ("AddressAssessorMunicipalityDesc")
 - Year
@@
 
-Before any prediction on the future mill rate of Metro Vancouver’s real estate market was made, exploratory data analysis was performed to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs.
+Before any prediction on the future mill rate of Metro Vancouver’s real estate market was made, exploratory data analysis was performed to explore and visualize the main characteristics of our dataset.
+Correlation analyses between Mill Rate vs.
 Assessment, Mill Rate vs.
@@
 
-Mill rate is mainly affected by assessment, so scatter plots of mill rate vs.
+Mill rate was mainly affected by assessment, so scatter plots of mill rate vs.
 total, land, and improvement assessment were created to see the correlation between each pair of the two factors.
@@
 
-In this study, we built linear models and evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.
-The definition of goodness of fit and prediction power is given below.
+In this study,  linear models were built and the performance of each model was evaluated by the goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.
+The definition of the goodness of fit and prediction power is given below.
 
  - **Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data.
-R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.
+R-squared and adjusted R-squared are the most well-known measures of goodness of fit and higher R-squared and adjusted R squares indicate better goodness of fit.
 
- - **Prediction power** measures how well models can predict the future values.
-We use mean squared prediction power to compare the prediction performance across all of our fitted linear predictive models.
-To do that, we will divide the data set into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.
+ - **Prediction power** measures how well models can predict future values.
+Mean squared prediction power was used to compare the prediction performance across all of our fitted linear predictive models.
+To do that, the data set was divided into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.
 
@@
 
-TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties are considered.
-The full linear model (ORL full) using all the available features was built first.
+The full linear model (ORL full) was built first.
+TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties were considered in this model.
 Based on the results of EDA, we selected a list of significant variables and included them in another linear model (OLR transformed).
@@
 Other than ordinary linear regressions, we are also interested in the performance of more advanced linear regressions like Ridge and Lasso.
-Ridge and Lasso have a different objective function to optimize; they take penalty in sum of absolute values and sum of squared absolute values of weights respectively.
+Ridge and Lasso have a different objective function to optimize; they take a penalty in the sum of absolute values and sum of squared absolute values of weights respectively.
 These models have an interesting characteristic, that is, a weight is assigned to each feature.
-This might be relevant to our study as mill rate can be affected by each feature to a different extent.
-Also, another advantage of these model is that MSPE is more stable as the variance of MSPE is reduced.
+This might be relevant to our study as the mill rate can be affected by each feature to a different extent.
+Also, another advantage of these models is that MSPE is more stable as the variance of MSPE is reduced.
 
@@
 
-To examine the goodness of fit and prediciton power, a 50-run of 10-fold cross validation would be performed in this study.
-For each run, a 10-fold corss validation was used to train the five models and make predictions on training and testing data respectively.
-Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding mdoels.
+To examine the goodness of fit and prediction power, a 50-run of 10-fold cross-validation was performed in this study.
+For each run, 10-fold cross-validation was used to train the five models and make predictions on training and testing data respectively.
+Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding models.
 
-After the 50 runs, a vector of MSEs and a vectors of MSPEs for each model were therefore constructed successfully.
-Based on these vectors, boxplots would be used to show the mean and spread of MSE and MSPE aross all models respectively.
+After the 50 runs, a vector of MSEs and a vector of MSPEs for each model were therefore constructed successfully.
+Based on these vectors, side-by-side boxplots were used to show the mean and the spread of MSE and MSPE across all models respectively.
 
@@
 
-Since mill rate is mainly affected by total assessment, which is sum of total land assessment and total improvement assessment, scatter plots between mill rate and all kinds of assessment among different municipalities are created to show the relationship in Figure 1, Figure 2, Figure 3
+Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and all kinds of assessment among different municipalities are created to show the relationship in Figure 1, Figure 2, Figure 3
 
@@
 
-There is no clear trend between mill rate and any kind of assessment. Plots also have shown that most points are condensed on the left horizontal axis, since some municipalities have large assessment.
+There is no clear trend between the mill rate and any kind of assessment.
+Plots have shown that most points are condensed on the left horizontal axis since some municipalities have larger assessment values than others.
 
-To reduce the effect of large assessment of some municipalities, total assessment across municipalities are transformed by taking total assessment dividing by number of properties of each municipality and tax class.
+To reduce the effect of large assessment in some municipalities, total assessment across all municipalities is transformed by taking total assessment dividing by the number of properties of each municipality and tax class.
 The transformed data is named **"Average Total Assessment"**.
-A scatter plot between mill rate and average total assessment is showed in Figure 4.
+A scatter plot between mill rates and average total assessments is shown in Figure 4.
 
@@
 
-Plot from Figure 4 has shown that mill rate tends to decrease as average total assessment increase. Also they have moderately strong linear correlation.
+The plot from Figure 4 has shown that the mill rate tends to decrease as the average total assessment increase. Also, they have a moderately strong linear correlation.
 
@@
 
-Figure 5 has shown that most municipalities have different mean and variance in terms of mill rates.
+Figure 5 shows that most municipalities have different mean and variance in terms of mill rates.
 
@@
 
-A statistical test called Kruskal-Wallis Test is also performed to test their distribution.
-It is used to decide if population distributions are identical, and the corresponding p-value which is smaller than 0.05 can indicate that the data has nonidentical distributions.
+A statistical test called the Kruskal-Wallis Test was performed to test their distribution.
+It was used to decide if population distributions were identical, and the corresponding p-value which is smaller than 0.05 indicated that the data have nonidentical distributions.
 
@@
 
-The p-values in Table 1 supports that there is nonidentical distributions of mill rates across municipalities and tax classes.
+The p-values in Table 1 support that there is nonidentical distributions of mill rates across municipalities and tax classes.
 
@@
 
-Below is a comparison of the five models (Lasso, Ridge, OLR transformed, OLR full and OLR null) as for goodness of fit and predcition power.
-Goodness of fit is measured by MSE of the training data, whereas prediction power is measured by MSPE of the testing data.
+Below is a comparison of the five models (Lasso, Ridge, OLR transformed, OLR full and OLR null) using the goodness of fit and prediction power.
+The goodness of fit was measured by MSE of the training data, whereas prediction power was measured by MSPE of the testing data.
 Generally, smaller MSEs and MSPEs indicate better fit and prediction power respectively.
 
-The distributions of goodness of fit and prediction power across models are displyed in Figure 7 and Figure 8, respectively, the distribution of null model is removed since it has larger values compared to all other models.
+The distributions of the goodness of fit and prediction power across models are displayed in Figure 7 and Figure 8, respectively, the distribution of null model is removed since it has larger values compared to all other models.
 
@@
         cex.lab = 1, 
-        cex.main = 2,
+        cex.main = 1,
         main = " Figure 7: Goodness of Fit",
@@
         cex.lab = 1, 
-        cex.main = 2,
+        cex.main = 1,
         main = "Figure 8: Prediction Power",
@@
 
-As for goodness of fit, Lasso, Ridge and OLR reduced have a similar performance; MSEs are around 4.6, but OLR transformed perform slightly better and has a smaller spread (variance) of MSE.
-OLR full performs worse than the the three, around 5.4, and OLR null has the greatest MSE, around 35.
+As for goodness of fit, Lasso, Ridge, and OLR transformed have similar performance; MSEs across these models were around 4.6, but OLR transformed performed slightly better and had a smaller spread (variance) of MSE.
+OLR full performed worse than the other three, around 5.4, and OLR null had the greatest MSE, around 35.
 
-Similarly, as for prediction power, Lasso, Ridge and OLR transformed perform roughly the same; MSPEs are close to 6.
-OLR full is worse with MSPE around 6.9.
-OLR Null performs the worst with MSPE over 35.
+Similarly, as for prediction power, Lasso, Ridge, and OLR transformed performed roughly the same; MSPEs across these models were close to 6.
+OLR full was worse with MSPE around 6.9.
+OLR Null performed the worst with MSPE over 35.
 
-Therefore, we conclude that from the results of a 50-run of 10-fold cross validation, Lasso, Ridge and OLR transformed have the best goodness of fit and prediction power.
-OLR null perfroms much worse than all the other four models.
+Therefore, we conclude that from the results of a 50-run of 10-fold cross-validation, Lasso, Ridge, and OLR transformed had the best goodness of fit and prediction power.
+OLR null perfromed much worse than all the other four models.
 
-For a detailed look of how the models were fitted and evaluated, please refer to the **Appendix**.
+For a detailed look at how the models were fitted and evaluated, please refer to the **Appendix**.
 
@@
 
-##_____________________________
-## TODO: delete pct change part, add prediction based on our chosen model (Peter)
 ## Conclusion
 
-From our exploratory data analysis, we can see that assessment total, land total, tax code, year and municipalities are all significant in our model.
-The client suggested that we can transform numerical factors such as assessment total into percentage change.
-This data transformation does not perform as well as we expected, which only yields an R^2 of around 0.2 across all of our fitted models.
-We believe the method that the client has suggested might leave out some important information about the housing market in each municipality.
-We have also found that when comparing the correlation between mill rate against features in our models, Vancouver is an outlier.
-We would like to further investigate Vancouver as a special case.
-We are also interested in the effect of different tax classes in predicting mill rate.
-Further analysis and prediction will be performed based on our hypothesis.
+From the exploratory data analysis, we have found that Total Assessments contain outliers in major municipalities such as Vancouver and Burnaby.
+In order to reduce the effect of outliers, the Total Assessments were transformed by taking their average over the number of properties in each municipality.
+The correlation analyses had shown that the transformed assessment total is the only continuous variable that has a relatively strong correlation with the mill rate.
+The Kruskal-Wallis had suggested that mill rates in each municipality and tax class are significantly different.
+From these results, we have decided to use Average Assessment Total, TaxClassCode and Municipality to fit our reduced model.
 
-All fitted linear model were able to make accurate prediction based on means squared error, but there is still a very high chance that our model is overfitted.
-For our next report, we are going to use cross-validation to reduce the effect of overfitting.
+Transformed OLR, Ridge Regression, and LASSO were able to make good predictions based on mean squared error and mean squared prediction error from cross-validation.
+Since the client prefers a simpler model, we choose the transformed model to make our 2020 prediction.
 ##_______________________________

### 2020-03-21 src/2nd_draft.Rmd
reference added.

@@
 The full linear model (ORL full) was built first.
-TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties were considered in this model.
+TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total and Number of properties were considered in this model.
 Based on the results of EDA, we selected a list of significant variables and included them in another linear model (OLR transformed).
@@
 Also, another advantage of these models is that MSPE is more stable as the variance of MSPE is reduced.
+For more detials about the two models, please refer to the **Appendix**.
 
@@
 To examine the goodness of fit and prediction power, a 50-run of 10-fold cross-validation was performed in this study.
-For each run, 10-fold cross-validation was used to train the five models and make predictions on training and testing data respectively.
+For each run, a 10-fold cross-validation was used to train the five models and make predictions on training and testing data respectively.
 Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding models.
@@
 Based on these vectors, side-by-side boxplots were used to show the mean and the spread of MSE and MSPE across all models respectively.
+Refer to the **Appendix** for algorithms.
 
@@
 - Exploratory Data Analysis and Model Fitting (https://github.com/STAT450-550/RealEstate/blob/450/src/EDA%26mode_fitting.Rmd)
+
+Lasso and Ridge:
+https://web.stanford.edu/class/stats202/content/lec14-cond.pdf
+
+Cross Validation:
+https://github.com/msalibian/STAT406/tree/master/Lecture2

### 2020-03-23 src/2nd_draft.Rmd
typo corrected
Format List

@@
 
-Our goal is to predict mill rates in 2020 in the following 22 municipalities:
-
-- Burnaby
-- Coquitlam
-- Delta
-- Langley - City
-- Langley - Township
-- Maple Ridge
-- Maple Ridge Rural
-- New Westminster
-- North Vancouver - City
-- North Vancouver - Dist
-- Pitt Meadows
-- Port Coquitlam
-- Port Moody
-- Richmond
-- Surrey
-- Vancouver
-- White Rock
-- West Vancouver
-- Bowen Island
-- Anmore
-- Belcarra
-- Lions Bay.
+Our goal is to predict mill rates in 2020 in Metro Vancouver.
 
@@
 - TaxClass Code in (01,05,06)
-- Municipality in Metro Vancouver
+- Municipality in (Burnaby, Coquitlam, Delta, Langley - City, Langley - Township, Maple Ridge, Maple Ridge Rural, New Westminster, North Vancouver - City, North Vancouver - Dist, Pitt Meadows, Port Coquitlam, Port Moody, Richmond, Surrey, Vancouver, White Rock, West Vancouver,  Bowen Island, Anmore, Belcarra, Lions Bay)
 
@@
 
-There are 1801 missing value in mill rate. 1509 are imputed, and 292 are removed from the data frame. The imputation method is mentioned in appendix.
+There are 1801 missing value in mill rate. 1509 are imputed, and 292 are removed from the data frame. The imputation method is mentioned in **Appendix**.
 
@@
 
-- Mill rate ("rate": 2016-2019)
-- Total Assessment ("assessTotal": 2016-2019)
-- Total Land Assessment ("landTotal": 2016 - 2019)
-- Total Improvement Assessment ("improvementTotal": 2016 - 2019)
-- Total number of properties ("propertyCount")
-- Taxclasscode ("TaxClassCode": 01, 05, 06)
-- Municipality ("AddressAssessorMunicipalityDesc")
+- Mill rate (rate)
+- Total Assessment (assessTotal)
+- Total Land Assessment (landTotal)
+- Total Improvement Assessment (improvementTotal)
+- Total number of properties (propertyCount)
+- Taxclasscode (TaxClassCode)
+- Municipality (AddressAssessorMunicipalityDesc)
 - Year
@@
 
-###Exploratory Analysis
+### Exploratory Analysis
 
@@
 
-###Linear Model
+#### Linear Model
 

### 2020-03-23 doc/Peter_Han_Individual_edit.Rmd

---
title: "Peter Han Individual Report Edit"
author: "Mallory Flynn"
date: \today
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overall Suggestions and Impressions

- The report is concise and mostly well organized, with clear breakdown of report sections using subtitles.  I would suggest another revision with the intent goal of rearranging ideas into their appropriate subsections (some examples of this are elaborated below).

- You clearly understand the client's goals. This report presents an appropriate first attempt at analyzing the relationship between variables of interest to the client.

- Watch use of tense in each section, and try to stay consistent.  Also try to reword sentences so that you don't need to refer to "us" or "we".  For example, "We decided to calculate the sum..." could become "The sum was calculated..."

- Revise again for mistakes in spelling and grammar, typos, as well as proper spacing after a period.

- Code section needs major revision to make it easier to follow.  Include code comments, avoid multiple renaming of variables, and block libraries and seed-setting into the start of the code, or at least the start of a section.

See the following sections for a more detailed edit of the sections of the report.

# Summary

-  Revise again for spelling and grammar correction, paying careful attention to words that should be plural.

- Begin the summary with a concise description and motivation of the problem.

- Summary clearly identifies the client's main goals, but does not even address the linear models that were used for this, or their results.  This was a large component of the report and should be summarized here also.

# Intro

- The introduction jumps right into the particulars of what the client wants and the exact data analysis used.  The intro should instead mainly be motivating the problem and providing an overview and background of the topic (for example, explaining property tax generally and why one would care about being able to predict this).  Why is this significant, and why do we care?  Then it should ease into your particular (broader) objectives for this report.

- Your intro finishes with an overview of your analysis methods. Instead, a more broad description of what's to come should finish your introduction, which goes over what's to come in the following body paragraphs of the report.

- Also be careful not to switch between which tense you are using.  The intro starts in the present tense ("we did this", "we saw this") and ends in the future tense ("we will do this"), but all analysis included has already been finished.  When revising the intro, try to rewrite using the present tense consistently throughout the intro.

# Statistical Analysis
## Data Description

- Again, be careful with remaining consistent in one tense, especially in one sentence.  For example "we used past property assessment data, which is provided by our client" uses two different tenses.  Stick to past tense for describing the data analysis which was carried out, and present tense to report the findings.

- Maybe consider putting variable names in quotation (for example, "TotalAssessedValue").  To add to this point, you have made it clear which variables you used by listing them above, but it might be made clearer by also providing the column name in this list, since you are going to refer to it by this in the report.  For example, "- Assessment Values ("TotalAssessedValue") for 2016-2019" as the bullet point for assessment.

## Methods

- It is also possible to label figures in your Rmd report by adding to the code at the beginning of the chunk, and then referring to specific figures just by using ```\ref{figurelabel}``` in your usual body sentences.  This would help the readability of this section of the report, especially the *Exploratory Data Analysis* section, where you could refer the reader to specific figures since there are quite a few included.

- You mention some results you saw for exploratory data analysis.  Since you have also included a *Results* section that is broken down into parts, I would not include results in the methods section.

- Are you evaluating goodness of fit solely on $R^2$ values?  Is there other ways you could do this?  Is there a downside or limitation to only looking at $R^2$ values when comparing models?  You might also mention that $R^2$ is computed on training data.

- For *Prediction power*, I think a little more could be said about how you chose training and test data (is this done by choosing random observations?), if you did several rounds of cross validation, etc.  It is relatively easy to understand for the client, and helps either support your methods or highlight limitations.

- For the *Linear Model* section, you mention Vancouver, Richmond, Surrey, and Burnaby were outliers.  How was this determined?  Maybe expand on this and move to the results section.

- "The reduced model’s performance in terms of goodness of fit will decrease with fewer features, however, the performance of prediction power might be improved.": This sentence might be modified a little and better suited to the paragraph above, "Measure of goodness of fit and prediction power", where it would complement with the explanation here.

- What was the motivation for attempting Ridge, Lasso and Elastic Net? These methods are typically motivated by certain features of the data.  I would defend the consideration of these frameworks by providing further clarification as to why you might want to consider these methods.

- While doing the above, you could explain Ridge, Lasso and Elastic Net in a little more "client friendly" way (that is, discussing that they use a penalty may be correct but not particularly meaningful to the client unless it is clear what this means and why one would want to try this).  Maybe compare to the ordinary linear model, if the client has a solid understanding of that.

- The *Neural Network* paragraph is very clear.  Since this is not a part of this report's analysis or findings, you could consider including this information in the conclusion as a future direction of this study.

# Results
## Exploratory Analysis

- Plots should have a caption, where Figure number is specified (this can be done automatically, making it easy to refer to figure numbers within the body of the report, and the numbers update automatically if new figures are inserted or deleted - see description above)

- "Dollars ($)" might be more informative as an axis title than "amount"

- For Vancouver, Harry mentioned the "school tax" which began just recently, resulting in a different calculation for amount of tax being paid by properties worth over $2 million (this is all tax class code 1 properties in our dataset).  Was this incorporated to the appropriate years to calculate total tax income collected in Vancouver for the relevant years?

## Correlation Analysis

- Plots a very condensed on the left horizontal axis.  Instead of plotting raw totals assessment dollar amounts, consider rescaling.  Since different municipalities have drastically different numbers of properties, this may be one reason amounts are so much higher for some areas.  Maybe rescale by number of properties and see how that plots instead?

- Also, consider transforming your data, for example by taking logarithms.  So much clustering could be spread out by an appropriate transformation, making the correlations clearer.  (Could try Box-Cox to see which transformation might be appropriate)

- You visually identified potential outliers - is there another way to determine if they are outliers?

## ANOVA Analysis

- I like the tables you included - this makes results clear and easy to compare.  However, explain what "MSE" and "PMSE" stand for in the body paragraphs or in a table caption - this may be a common abbreviation to you but it is not likely in the everyday lingo of the client.  Also, your client might wonder why there are NA's in the table for some of your models.  Maybe a brief mention of this

- Do you expect municipality to be significant?  Some plots I have seen from your previous analysis showed tax rates are all very similar, following similar trends for the municipalities you considered.  Including that plot might help explain why this came up as insignificant.   It might be worth considering that this may not be the case once the model is expanded to greater BC.

- Also, "R_squared" should be $R^2$ or "R-squared" in both body paragraphs and table column headers.

- MSE and PMSE are just quoted as values.  How does the client interpret whether these values are good or bad?

# Conclusion

- "The client suggested that we can transform numerical factors such as assessment total into percentage change. This data transformation does not perform as well as we expected, which only yields an Rˆ2 of around 0.2 across all of our fitted models. We believe the method that the client has suggested might leave out some important information about the housing market in each municipality. " This is new information, that is maybe better to include briefly in the methods and results sections rather than the conclusions.

- If you are going to also consider a neural net as a future piece of this analysis, include this here as another future direction.

# References and Appendix

- I like that you included the link to the GitHub repo.

- Use a page break to put references and appendix each on a new page of their own.

- Consider including references to support ridge regression, elastic net, and LASSO.  There are lots of sources you could point the client to that explain these methods clearly, if he were to want to do additional reading.

- Make sure to adjust your YAML to specify that you don't want code to run out of the page margins. I think Gaby suggested a fix for this on Slack.

- If the client would like to follow through your code as you directed him to in the body of the report, my impression is that he will need significantly more guidance in the form of comments throughout, as to what you are doing and why.

- Is this the code for the rate change of mill rate?  Be more clear on which model you are fitting.

- Include the code relevant to summarizing and aggregating the data from the client, which produces in the dataframe used for analysis.

- Also for readability, consider putting all library's used and setting seed at the beginning of the code appendix, rather than throughout the code.  At the very least, they should be at the beginning of each section of code used for various models.

- Using comments, maybe highlight what code produced figures you included in the report.

- Following through the code as in the appendix, it looks like you are training your lasso model on the whole data set, not a training set.  Is this the case?  How might this affect robustness to new data?

- Change relevant variables to the appropriate type (for example, factors) before analysis, rather than specifying ```factor(variable)``` every time it is used.

- Since you are only creating one training set and one test set, this may as well be done once at the beginning of the code, rather than repeatedly throughout with each model.  That will eliminate the repetitive naming of objects 'x' and 'y' which are constantly being rewritten from one part to the next.

- Since the reduced model is a nested version of the full model, you could try a stepwise calculation of AIC values to see if this supports the reduced model you chose.

- Missing data and imputation should be discussed be discussed in the *Data Description* section, rather than at the end of the code appendix.  Include a commentary on how and why you imputed values the way you did.

### 2020-03-23 src/2nd_draft.Rmd
more modification

@@
 Exploratory analysis has shown there are strong relationships between mill rate and tax class and mill rate and municipalities; it has also shown there is a fairly strong correlation between mill rate and average assessment per property in different municipalities.
-Ordinary linear model, reduced ordinary linear model, Ridge Regression and LASSO are used to predict the mill rate.
+*Ordinary linear model*, *reduced ordinary linear model*, *Ridge Regression* and *the LASSO* are used to predict the mill rate.
 A full assessment of the performance - prediction power and goodness of fit of these models, is shown below.
@@
 Our client provides us the past 5 years' property assessment data in BC.
-Since the only interest is in predicting mill rate for metro Vancouver and specific class code, a subset of properties that satisfy our interest have been selected:
+Since the only interest is in predicting mill rate for metro Vancouver and specific tax class code, a subset of properties that satisfy our interest have been selected:
 
-- TaxClass Code in (01,05,06)
+- Tax Class in (01,05,06)
 - Municipality in (Burnaby, Coquitlam, Delta, Langley - City, Langley - Township, Maple Ridge, Maple Ridge Rural, New Westminster, North Vancouver - City, North Vancouver - Dist, Pitt Meadows, Port Coquitlam, Port Moody, Richmond, Surrey, Vancouver, White Rock, West Vancouver,  Bowen Island, Anmore, Belcarra, Lions Bay)
@@
 - Total number of properties (propertyCount)
-- Taxclasscode (TaxClassCode)
+- Tax Class Code (TaxClassCode)
 - Municipality (AddressAssessorMunicipalityDesc)
@@
 
-The full linear model (ORL full) was built first.
-TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total and Number of properties were considered in this model.
-Based on the results of EDA, we selected a list of significant variables and included them in another linear model (OLR transformed).
-To compare the effect of linear models with and without features, a null model (OLR null) with no features used, was also constructed.
+The full linear model (*ORL full*) was built first.
+TaxClassCode, Municipalities, assessTotal, landTotal, improvementTotal and propertyCount were considered in this model.
+Based on the results of EDA, we selected a list of significant variables and included them in another linear model (*OLR transformed*).
+To compare the effect of linear models with and without features, a null model (*OLR null*) with no features used, was also constructed.
 
@@
 
-Other than ordinary linear regressions, we are also interested in the performance of more advanced linear regressions like Ridge and Lasso.
-Ridge and Lasso have a different objective function to optimize; they take a penalty in the sum of absolute values and sum of squared absolute values of weights respectively.
+Other than ordinary linear regressions, we are also interested in the performance of more advanced linear regressions like *Ridge* and *Lasso*.
+*Ridge* and *Lasso* have a different objective function to optimize; they take a penalty in the sum of absolute values and sum of squared absolute values of weights respectively.
 These models have an interesting characteristic, that is, a weight is assigned to each feature.
@@
 Also, another advantage of these models is that MSPE is more stable as the variance of MSPE is reduced.
-For more detials about the two models, please refer to the **Appendix**.
+For more detials about the two models, please refer to the **References**.
 
@@
 Based on these vectors, side-by-side boxplots were used to show the mean and the spread of MSE and MSPE across all models respectively.
-Refer to the **Appendix** for algorithms.
+For more details about cross-validation, please refer to the **References**.
 
@@
 
-#### Continuous Variables
+#### - Continuous Variables
 
@@
 
-Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and all kinds of assessment among different municipalities are created to show the relationship in Figure 1, Figure 2, Figure 3
+Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and all kinds of assessment among different municipalities are created to show the relationship in Figure 1, Figure 2, Figure 3.
+Each color in the figures belongs to three tax classes of one municipality.
 
@@
 
-#### Categorical Variables
+#### - Categorical Variables
 
-Here categorical variables were taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 5 and Figure 6.
+Here categorical variables are taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 5 and Figure 6.
 
@@
      theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
+```
 
+```{r, echo=FALSE}
 assessment_eda %>%
@@
 
-Below is a comparison of the five models (Lasso, Ridge, OLR transformed, OLR full and OLR null) using the goodness of fit and prediction power.
+Below is a comparison of the five models (*Lasso*, *Ridge*, *OLR transformed*, *OLR full* and *OLR null*) using the goodness of fit and prediction power.
 The goodness of fit was measured by MSE of the training data, whereas prediction power was measured by MSPE of the testing data.
@@
 
-As for goodness of fit, Lasso, Ridge, and OLR transformed have similar performance; MSEs across these models were around 4.6, but OLR transformed performed slightly better and had a smaller spread (variance) of MSE.
-OLR full performed worse than the other three, around 5.4, and OLR null had the greatest MSE, around 35.
+As for goodness of fit, *Lasso*, *Ridge*, and *OLR transformed* have similar performance; MSEs across these models were around 4.6, but *OLR transformed* performed slightly better and had a smaller spread (variance) of MSE.
+*OLR full* performed worse than the other three, around 5.4, and *OLR null* had the greatest MSE, around 35.
 
-Similarly, as for prediction power, Lasso, Ridge, and OLR transformed performed roughly the same; MSPEs across these models were close to 6.
-OLR full was worse with MSPE around 6.9.
-OLR Null performed the worst with MSPE over 35.
+Similarly, as for prediction power, *Lasso*, *Ridge*, and *OLR transformed* performed roughly the same; MSPEs across these models were close to 6.
+*OLR full* was worse with MSPE around 6.9.
+*OLR Null* performed the worst with MSPE over 35.
 
-Therefore, we conclude that from the results of a 50-run of 10-fold cross-validation, Lasso, Ridge, and OLR transformed had the best goodness of fit and prediction power.
-OLR null perfromed much worse than all the other four models.
+Therefore, we conclude that from the results of a 50-run of 10-fold cross-validation, *Lasso*, *Ridge*, and *OLR transformed* had the best goodness of fit and prediction power.
+*OLR null* perfromed much worse than all the other four models.
 
@@
 
-Transformed OLR, Ridge Regression, and LASSO were able to make good predictions based on mean squared error and mean squared prediction error from cross-validation.
-Since the client prefers a simpler model, we choose the transformed model to make our 2020 prediction.
-##_______________________________
+*Transformed OLR*, *Ridge Regression*, and *the LASSO* were able to make good predictions based on mean squared error and mean squared prediction error from cross-validation.
+Since the client prefers a simpler model, we choose the *transformed model* to make our 2020 prediction.
 

### 2020-03-23 src/2nd_draft.Rmd
MSE and MSPE added. Reference fixed.

@@
  - **Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data.
-R-squared and adjusted R-squared are the most well-known measures of goodness of fit and higher R-squared and adjusted R squares indicate better goodness of fit.
+MSE (Mean Squared Error) is a measure of goodness of fit and a smaller MSE indicates better goodness of fit.
 
@@
 Mean squared prediction power was used to compare the prediction performance across all of our fitted linear predictive models.
-To do that, the data set was divided into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.
+Prior to examine the prediction power, the data set was divided into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.
+MSPE (Mean Squared Predction Error) on testing set is used to measure the prediction power in this study.
+
+Refer to **Appendix** for forumlas of MSE and MSPE.
 
@@
 Also, another advantage of these models is that MSPE is more stable as the variance of MSPE is reduced.
-For more detials about the two models, please refer to the **References**.
+For more detials about the two models, please refer to **Appendix**.
 
@@
 Based on these vectors, side-by-side boxplots were used to show the mean and the spread of MSE and MSPE across all models respectively.
-For more details about cross-validation, please refer to the **References**.
+For more details about cross-validation, please refer to **Appendix**.
 
@@
 
-For a detailed look at how the models were fitted and evaluated, please refer to the **Appendix**.
-
 ```{r fit, echo=FALSE}
@@
 
+Measure of Goodness of Fit and Prediction Power:
+https://channabasavagola.github.io/2018-01-09-metrics/
+
 Lasso and Ridge:

### 2020-03-24 src/draft.Rmd

@@
 ---
-title: "EDA"
-author: "Xuechun Lu, Yuting Wen, Peter Han"
+title: "draft"
+author: "Xuechun Lu"
 date: "25/02/2020"
-output:
-  github_document
+output: html_document
 ---
@@
 ## R Markdown
-libraries
 ```{r}
 library(tidyverse)
-library(ggplot2)
 library(dplyr)
@@
 library(readxl)
-
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-  numPlots = length(plots)
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
- if (numPlots==1) {
-    print(plots[[1]])
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-
 ```
 
-1. data processing
-```{r, warning=F, message=F}
-# all municipalities
-municipality.list = c(
-  "Burnaby", 
-  "Coquitlam", 
-  "Delta", 
-  "Langley - City", 
-  "Langley - Township",
-  "Maple Ridge",
-  "Maple Ridge Rural", 
-  "North Vancouver - City",
-  "North Vancouver - Dist",
-  "Pitt Meadows", 
-  "Port Coquitlam", 
-  "Port Moody", 
-  "Richmond", 
-  "Surrey", 
-  "Vancouver", 
-  "White Rock", 
-  "West Vancouver", 
-  "Bowen Island", 
-  "Anmore", 
-  "Belcarra",
-  "Lions Bay")
-
-# budget and assessment data import and process
+1. read the data and aggregate them
+```{r}
 tax_pct<-read_csv(here("data", "tax_pct.csv"))
@@
 tax_pct<-tax_pct[,-c(1)]
-#head(tax_pct)
+
 ```
@@
 
-2019 tax pct is missing, we imputed 2019 tax by the average of previous tax pct
-```{r}
-tax_pct[,4]<-(tax_pct$pct_2017+tax_pct$pct_2018)/2
-#tax_pct
-names(tax_pct)[4]<-paste("pct_2019")
-#head(tax_pct)
-```
-
-aggregate tax and assessment
 ```{r}
@@
 assessment2019<-assessment_pct  %>%  filter(Year=="2019")
-tax2019<-tax_pct[,c(1,4)]
-assessment2019<-assessment2019 %>% left_join(tax2019, by = c("Municipalities"))
-names(assessment2019)[9]<-paste("tax")
 
-pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
-write.csv(pct_final, here("data","data_final.csv"))
+data_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
+write.csv(data_final, here("data","data_final.csv"))
+```
 
-# all pct data aggregated
-#head(pct_final)
+2. plots
+```{r}
 
 ```
-
-2019 tax is missing, we imputed 2019 tax using previous values
-```{r}
-tax_final <- read.csv(here("data","tax_final.csv"))
-#head(tax_final)
-
-# tax_final is the budget
-tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))
-
-# tax_modified contains all the budget values from 2016 to 2019
-#head(tax_modified)
-```
-
-global variables
-```{r}
-# all pct aggregated
-pct_final
-# all budget values
-tax_modified
-```
-
-2. val_plots
-```{r}
-# mill rate vesus year for each municipality and tax class
-assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
-dat_rate <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, rate)
-#head(dat_rate)
-
-dat_rate_1 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
-#head(dat_rate_1)
-rate_class1 <- dat_rate_1 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 1") + theme(legend.position = "none")
-#rate_class1
-
-dat_rate_5 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
-#head(dat_rate_5)
-rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 5") + theme(legend.position = "none")
-#rate_class5
-
-dat_rate_6 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6") 
-#head(dat_rate_6)
-rate_class6 <- dat_rate_6 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("mill rate for taxClass 6") + theme(legend.position = "none")
-#rate_class6
-
-#multiplot(rate_class1, rate_class5, rate_class6)
-
-# plot assessed value vesus year for each municipality and tax class
-dat_assess <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, assessTotal)
-#head(dat_assess)
-
-dat_assess_1 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
-#head(dat_assess_1)
-assess_class1 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 1") + theme(legend.position = "none")
-#assess_class1
-
-dat_assess_5 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
-#head(dat_assess_5)
-assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 5") + theme(legend.position = "none")
-#assess_class5
-
-dat_assess_6 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6") 
-#head(dat_assess_6)
-assess_class6 <- dat_assess_1 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_line() + ggtitle("assessTotal for taxClass 6") + theme(legend.position = "none")
-#assess_class6
-
-#multiplot(assess_class1, assess_class5, assess_class6)
-
-# plot budgets vesus year
-dat_tax_long <- gather(tax_modified, Year, TaxAmount, X2016:X2019, factor_key=TRUE)
-#dat_tax_long
-budget_plot <- dat_tax_long %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_line() +ggtitle("Tax v.s Year")
-#budget_plot
-
-multiplot(rate_class1,assess_class1,budget_plot)
-multiplot(rate_class5,assess_class5,budget_plot)
-multiplot(rate_class6,assess_class6,budget_plot)
-```
-
-3. pct_plots
-pct change of mill rates for each tax class
-```{r}
-pct_rate1 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "01") %>%
-  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass1 pct rate v.s year") + theme(legend.position = "none")
-
-pct_rate5 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "05") %>%
-  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass5 pct rate v.s year") + theme(legend.position = "none")
-
-pct_rate6 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "06") %>%
-  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass6 pct rate v.s year") + theme(legend.position = "none")
-
-pct_assess1 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "01") %>%
-  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass1 pct assessment v.s year") + theme(legend.position = "none")
-
-pct_assess5 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "05") %>%
-  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass5 pct assessment v.s year") + theme(legend.position = "none")
-
-pct_assess6 <- pct_final %>% 
-  filter(Municipalities %in% municipality.list) %>%
-  filter(TaxClassCode == "06") %>%
-  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
-    geom_line() + ggtitle("taxClass6 pct assessment v.s year") + theme(legend.position = "none")
-
-multiplot(pct_rate1,pct_assess1)
-multiplot(pct_rate5,pct_assess5)
-multiplot(pct_rate6,pct_assess6)
-```
-pct change of mill rate for each municipality
-```{r}
-# for each municipality
-tax.list <- c("01","05","06")
-for (i in 1:21){
-  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=rate_pct, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line() + ggtitle(municipality.list[i])
-  print(plot)
-}
-```
-
-plot pct change of assessment
-```{r}
-# total trend of assessment 
-for (i in 1:21){
-  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line() + ggtitle(municipality.list[i])
-  print(plot)
-}
-```
-
-plot pct change of tax(budget)
-```{r}
-# total trend for budget 
-for (i in 1:21){
-  plot<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode)) +
-    geom_line() + ggtitle(municipality.list[i])
-  print(plot)
-}
-```
-
-```{r}
-# plot rate, assessment, and budget for each municipality.
-
-for (i in 1:21){
-  test1<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=rate_pct, group=TaxClassCode, color=TaxClassCode)) +
-   geom_line()
-
-  test2<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
-   geom_line()
-
-  test3<-pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
-  filter(TaxClassCode %in% tax.list) %>% 
-  ggplot(aes(x=Year, y=tax)) +
-   geom_line() + theme(legend.position = "none")
-
-  multiplot(test1,test2,test3)
-
-}
-```
-
-4. box plots of mill rates, grouped by region
-```{r}
-rate_boxes <- ggplot(dat_rate, aes(x=AddressAssessorMunicipalityDesc,y=rate,fill=TaxClassCode)) + geom_boxplot() + facet_wrap(~TaxClassCode) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
-rate_boxes 
-```
-
-5. Test correlations between past mill rates and assessed values and government budget.
-```{r}
-#tax_final
-#assessment_aggregate
-
-#colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
-#head(tax_modified)
-#dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)
-
-#tax.all <- full_join(assessment_aggregate, tax_modified, by = 'Municipalities') 
-#tax.all
-#(tax.all <- tax.all[,-5])
-# correlation by Municipalities?
-#mill.budget.cor <- tax.all  %>% summarize(cor_2017 = cor(pct_2017, X2017), cor_2018 = cor(pct_2018, X2018), cor_2019 = cor(pct_2019, X2019))
-#mill.budget.cor
-
-years <- c(2017, 2018, 2019)
-city.cor <- data.frame(Municipality = municipality.list)
-corr.list <- vector(length = 21)
-corr.list
-for(city in municipality.list){
-  corr <- assessment_aggregate %>% filter(AddressAssessorMunicipalityDesc == city) %>% summarize(cor = cor(assessTotal, rate))
-  #print(corr)
-  corr.list <- cbind(corr.list, corr)
-}
-corr.list
-city.cor <- cbind(city.cor,corr.list)
-#city.cor
-```
-
-6. Compute autocorrelations of past mill rates to test the independence of mill rates. 
-```{r}
-# all white noise
-for (i in 1:21) {
-  print(municipality.list[i])
-  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "1") %>% select(rate)
-  if (municipality.list[i] != "Maple Ridge Rural") {
-    as.ts(dat_tmp)
-    acf(dat_tmp)
-  }
-}
-for (i in 1:21) {
-  print(municipality.list[i])
-  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "5") %>% select(rate)
-  if (municipality.list[i] != "Maple Ridge Rural" & municipality.list[i] != "White Rock" & municipality.list[i] != "Anmore" & municipality.list[i] != "Belcarra" & municipality.list[i] != "Lions Bay") {
-    as.ts(dat_tmp)
-    acf(dat_tmp)
-  }
-}
-for (i in 1:21) {
-  print(municipality.list[i])
-  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "6") %>% select(rate)
-  if (municipality.list[i] != "Maple Ridge Rural") {
-    as.ts(dat_tmp)
-    acf(dat_tmp)
-  }
-}
-```
-
-model fitting
-```{r}
-linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
-summary(linear1)
-
-linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
-summary(linear2)
-```

### 2020-03-24 src/2nd_draft.Rmd
added references

@@
 
-Refer to **Appendix** for forumlas of MSE and MSPE.
+Refer to **Reference** for forumlas of MSE and MSPE.
 
@@
 Also, another advantage of these models is that MSPE is more stable as the variance of MSPE is reduced.
-For more detials about the two models, please refer to **Appendix**.
+For more detials about the two models, please refer to **Reference**.
 
@@
 Based on these vectors, side-by-side boxplots were used to show the mean and the spread of MSE and MSPE across all models respectively.
-For more details about cross-validation, please refer to **Appendix**.
+For more details about cross-validation, please refer to **Reference**.
 
@@
 The p-values in Table 1 support that there is nonidentical distributions of mill rates across municipalities and tax classes.
+Refer to **Reference** for more information on the Kruskal Wallis Test.
 
@@
 https://github.com/msalibian/STAT406/tree/master/Lecture2
+
+Kruskal Wallis:
+http://www.biostathandbook.com/kruskalwallis.html

### 2020-03-26 src/eda - s550/eda.R
Add s550 plots and eda code. -gd

@@ -1,0 +1,221 @@
+# Gian Carlo EDA
+
+# preamble ####
+library(tidyverse)
+library(readr)
+library(readxl)
+library(ggplot2)
+ggplot2::theme_set(theme_classic())
+library(viridis)
+
+# data wrangling ####
+
+# data import
+columnnames <- readxl::read_xlsx("colnames.xlsx") %>% 
+  colnames()
+
+real.estate_full <- readr::read_csv("2016 - 2020 Raw.csv",
+                                    na = c("", "NA", "NULL", "NULL_1"),
+                                    col_names = columnnames,
+                                    col_types = "ciccccccccccccccdddddiicccddddc")
+# data wrangling
+
+# assessments by PIC and year
+assessments <- real.estate_full %>% 
+  dplyr::select(PIC, Year, # relevant variables
+                AssessedValueAmt, AssetTypeDesc) %>% 
+  dplyr::rename(year = Year,
+                assessment = AssessedValueAmt,
+                assessment.type = AssetTypeDesc) %>% 
+  dplyr::group_by(PIC, year, assessment.type) %>% 
+  dplyr::summarise(assessment = sum(assessment)) %>% 
+  dplyr::ungroup() %>% 
+  tidyr::spread(assessment.type, assessment) %>% 
+  dplyr::rename(improvement.assessment = Improvement,
+                land.assessment = Land) %>% 
+  dplyr::select(PIC, year, improvement.assessment, land.assessment) %>% 
+  dplyr::mutate(total.assessment = improvement.assessment + land.assessment)
+
+
+re <- real.estate_full %>% 
+  dplyr::select(PIC, Year, AddressAssessorMunicipalityDesc, # relevant variables
+                TaxClassCode, TaxOwingAmountTotalCalculated, TaxClassTaxRate) %>% 
+  dplyr::rename(year = Year,
+                municipality = AddressAssessorMunicipalityDesc, # human-readable names
+                tax.class = TaxClassCode,
+                tax = TaxOwingAmountTotalCalculated,
+                mill.rate = TaxClassTaxRate) %>% 
+  dplyr::filter(tax.class %in% c("01", "05", "06")) %>%  # relevant values for tax class
+  dplyr::distinct() %>% 
[…]

### 2020-03-26 src/2nd_draft.Rmd
formatted
EDA refactor

@@
 
+Refer to **Reference** for more information on the Kruskal Wallis Test.
+
 ### Measure of goodness of fit and prediction power
@@
 
-Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and all kinds of assessment among different municipalities are created to show the relationship in Figure 1, Figure 2, Figure 3.
-Each color in the figures belongs to three tax classes of one municipality.
+Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and total assessment among different municipalities is created to show the relationship in Figure 1.
+Each color in the figure belongs to three tax classes of one municipality.
+
+Scatter plots between mill rate and total land assessment, mill rate and total improvement assessment are similar to Figure 1. For more details about them, please refer to **Appendix**.
 
@@
    theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
-
-assessment_eda %>%
-   ggplot(aes(x=`Total Land Assessment`,
-              y=`Mill Rate`,group=Municipality,
-              color=Municipality)) + 
-   geom_point() + 
-   ggtitle(sprintf("Figure 2: Mill Rate vs Total Land Assessment")) + 
-   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
-   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
-
-assessment_eda %>%
-   ggplot(aes(x=`Total Improvement Assessment`,
-              y=`Mill Rate`,group=Municipality,
-              color=Municipality)) + 
-   geom_point() + 
-   ggtitle(sprintf("Figure 3: Mill Rate vs Total Improvement Assessment")) + 
-   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
-   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
 ```
 
-There is no clear trend between the mill rate and any kind of assessment.
-Plots have shown that most points are condensed on the left horizontal axis since some municipalities have larger assessment values than others.
+There is no clear trend between the mill rate and total assessment.
+Plot has shown that most points are condensed on the left horizontal axis since some municipalities have larger assessment values than others.
 
@@
 The transformed data is named **"Average Total Assessment"**.
-A scatter plot between mill rates and average total assessments is shown in Figure 4.
+A scatter plot between mill rates and average total assessments is shown in Figure 2.
 
@@
  geom_point() + 
- ggtitle(sprintf("Figure 4: Mill rate v.s. Average Total Assessment")) + 
+ ggtitle(sprintf("Figure 2: Mill rate v.s. Average Total Assessment")) + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") +
@@
 
-The plot from Figure 4 has shown that the mill rate tends to decrease as the average total assessment increase. Also, they have a moderately strong linear correlation.
+The plot from Figure 2 has shown that the mill rate tends to decrease as the average total assessment increase. Also, they have a moderately strong linear correlation.
 
@@
 
-Here categorical variables are taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 5 and Figure 6.
+Here categorical variables are taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 3 and Figure 4.
 
@@
      geom_boxplot() + 
-     ggtitle(sprintf("Figure 5: Mill rate across Municipalities")) +
+     ggtitle(sprintf("Figure 3: Mill rate across Municipalities")) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
@@
      geom_boxplot() + 
-     ggtitle(sprintf("Figure 6: Mill rate across Tax Classes")) +
+     ggtitle(sprintf("Figure 4: Mill rate across Tax Classes")) +
      theme(legend.position = "none") +
@@
 
-Figure 5 shows that most municipalities have different mean and variance in terms of mill rates.
+Figure 3 shows that most municipalities have different mean and variance in terms of mill rates.
 
-Figure 6 also supports that there is unequal mean and variance across tax classes.
+Figure 4 also supports that there is unequal mean and variance across tax classes.
 
@@
 The p-values in Table 1 support that there is nonidentical distributions of mill rates across municipalities and tax classes.
-Refer to **Reference** for more information on the Kruskal Wallis Test.
 
@@
 
-The distributions of the goodness of fit and prediction power across models are displayed in Figure 7 and Figure 8, respectively, the distribution of null model is removed since it has larger values compared to all other models.
+The distributions of the goodness of fit and prediction power across models are displayed in Figure 5 and Figure 6, respectively, the distribution of null model is removed since it has larger values compared to all other models.
 
@@
         cex.main = 1,
-        main = " Figure 7: Goodness of Fit",
+        main = " Figure 5: Goodness of Fit",
         ylab = expression(hat(MSE)))
@@
         cex.main = 1,
-        main = "Figure 8: Prediction Power",
+        main = "Figure 6: Prediction Power",
         ylab = expression(hat(MSPE)))
@@
 ## Appendix
+
+### Scatter plots:
+
+#### - mill rate v.s. total land assessment
+
+```{r, echo=FALSE}
+assessment_eda %>%
+   ggplot(aes(x=`Total Land Assessment`,
+              y=`Mill Rate`,group=Municipality,
+              color=Municipality)) + 
+   geom_point() + 
+   ggtitle(sprintf("Figure 7: Mill Rate vs Total Land Assessment")) + 
+   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
+   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+```
+
+#### - mill rate v.s. total improvement assessment
+
+```{r, echo=FALSE}
+assessment_eda %>%
+   ggplot(aes(x=`Total Improvement Assessment`,
+              y=`Mill Rate`,group=Municipality,
+              color=Municipality)) + 
+   geom_point() + 
+   ggtitle(sprintf("Figure 8: Mill Rate vs Total Improvement Assessment")) + 
+   geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
+   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+```
 

### 2020-03-30 src/2nd_draft.Rmd
1. replaced MSE with MSPE. 2. introduction of Lasso and Ridge updated

@@
  - **Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data.
-MSE (Mean Squared Error) is a measure of goodness of fit and a smaller MSE indicates better goodness of fit.
+MSPE (Mean Squared Prediction Error) on training sets is a measure of goodness of fit and a smaller MSPE indicates better goodness of fit.
 
@@
 Prior to examine the prediction power, the data set was divided into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.
-MSPE (Mean Squared Predction Error) on testing set is used to measure the prediction power in this study.
+MSPE (Mean Squared Predction Error) on testing sets is used to measure the prediction power in this study.
 
@@
 *Ridge* and *Lasso* have a different objective function to optimize; they take a penalty in the sum of absolute values and sum of squared absolute values of weights respectively.
-These models have an interesting characteristic, that is, a weight is assigned to each feature.
-This might be relevant to our study as the mill rate can be affected by each feature to a different extent.
-Also, another advantage of these models is that MSPE is more stable as the variance of MSPE is reduced.
-For more detials about the two models, please refer to **Reference**.
+A reason to consider Ridge Regression is that it helps deal with multicollinearity of the explanatory variables.
+This might be relevant to our study as some features used in this study are correlated.
+Lasso Regression is also included in this study because it does variable selections automatically by imposing a constraint on model parameters that causes some regression coefficients to shrink to zero.
+
+Also, another advantage of these models is that MSPE is more stable as the variance of MSPE is reduced. For more detials about the two models, please refer to **Reference**.
 
@@
         main = " Figure 5: Goodness of Fit",
-        ylab = expression(hat(MSE)))
+        ylab = expression(hat(MSPE)))
 
@@
 
-As for goodness of fit, *Lasso*, *Ridge*, and *OLR transformed* have similar performance; MSEs across these models were around 4.6, but *OLR transformed* performed slightly better and had a smaller spread (variance) of MSE.
-*OLR full* performed worse than the other three, around 5.4, and *OLR null* had the greatest MSE, around 35.
+As for goodness of fit, *Lasso*, *Ridge*, and *OLR transformed* have similar performance; MSPEs across these models were around 4.6, but *OLR transformed* performed slightly better and had a smaller spread (variance) of MSPE.
+*OLR full* performed worse than the other three, around 5.4, and *OLR null* had the greatest MSPE, around 35.
 

### 2020-03-30 src/2nd_draft.Rmd
linear regression assumptions added.

@@
 Refer to **Reference** for more information on the Kruskal Wallis Test.
+
+### Assumption of Linear Regression
+Models used in this study are mainly linear regressions.
+Before model construction, it is important to test the two assumptions of linear regression, i.e.
+normality of data and equal variance across groups for each categorical variables.
+If the two assumptions are satisfied, then the test results can be trusted.
+
+A histogram of Mill Rate and Normal Q-Q Plot were used to test the normality assumption.
+If the data is normally distributed, the histogram would be symmetric with peak in the middle and decreasing frequency as approaching the two tails and the data in the Normal Q-Q Plot would mostly fall on a 45 degree line.
+
+Fligner-Killeen Test was used to test the homogeneity of variances.
+The test does not rely on the assumption of normality and thus it is more robust to non-normal data.
+We applied the test to each categorical variable to test if there is a equal variance aross all sub-groups in each variable.
+If p-value is less than 5\% for one categorical variable, then at 5\% significane level, there is a significant difference of variance aross groups for that variable.
 
@@
 ```
+#### Assumptions of Linear Regression
+
+1. Normality Assumption
+
+```{r,echo=F, warning=F}
+hist(assessment_transform$rate, main = "Figure 5: Histogram of rate", xlab="Mill Rate")
+qqnorm(assessment_transform$rate, main = "Figure 6: Normal Q-Q Plot of rate", ylab = "Mill Rate")
+
+```
+Figure 5 exhibits the distribution of mill rates and it is observed that rate is not normal and it has heavy tails on both ends, especially on the left hand side.
+
+Figure 6 shows that most of data do not fall on the 45 degree line and curve away from the line at each end in opposite directions due to the “heavy tails” at the each end of the histogram.
+This, again, verifies that rate is not normally distributed.
+
+2. Equal variance
+
+```{r, echo=F, results='hide'}
+attach(assessment_transform)
+fligner.test(rate~Year) # p-value = 0.004907
+fligner.test(rate~TaxClassCode) # p-value < 2.2e-16
+fligner.test(rate~AddressAssessorMunicipalityDesc) # p-value = 0.1299
+```
+
+        Table 2: Fligner Killeen Test of Mill Rate across Municipality, Tax Class and Year
+
+Distribution  | p-value
+------|--------
+Mill Rate across Municipality | 0.1299
+Mill Rate across Tax Class | < 2.2e-16
+Mill Rate across Year | 0.004907
+
+Table 2 shows a non-significant difference between the group variances of Municipality, but a significant difference of variances in Year and Tax Class since p-values for Tax Class and Year are less than 5\%.
+Therefore, the equal variance assumption is not satisfied for Tax Class and Year.
+This finding is also consistent with Figure 4 where it is clear that variance (spread) is not identical across all three tax classes.
+
+Hence, assumptions of linear regression are not met in this study.
+This means the test results, for example, t-statisitcs and p-values, are biased and thus cannot be trusted.
+Since this study only focuses on prediction accuracy, the failure of linear regression assumptions might not matter much.
 
@@
 
-The distributions of MSPE across models are also displayed in Table 2.
+The distributions of MSPE across models are also displayed in Table 3.
 
@@
         cex.main = 1,
-        main = " Figure 5: Goodness of Fit",
+        main = " Figure 7: Goodness of Fit",
         ylab = expression(hat(MSPE)))
@@
         cex.main = 1,
-        main = "Figure 6: Prediction Power",
+        main = "Figure 8: Prediction Power",
         ylab = expression(hat(MSPE)))
@@
 
-                    Table 2: Prediction Power of All Models
+                    Table 3: Prediction Power of All Models
 

### 2020-03-30 src/2nd_draft.Rmd
reverted and tiny change

--- 2020-03-30 src/2nd_draft.Rmd
@@
 
-#### Linear Model
+#### Linear Models
 

### 2020-03-31 src/2nd_draft.Rmd
more modified

@@
 Data cleaning is performed to aggregate our data into summary statistics.
-Exploratory analysis has shown there are strong relationships between mill rate and tax class and mill rate and municipalities; it has also shown there is a fairly strong correlation between mill rate and average assessment per property in different municipalities.
-*Ordinary linear model*, *reduced ordinary linear model*, *Ridge Regression* and *the LASSO* are used to predict the mill rate.
+Exploratory analysis shows that there are strong relationships between mill rate and tax class and mill rate and municipalities; it also shows there is a fairly strong correlation between mill rate and average assessment per property in different municipalities.
+*Ordinary linear model*, *reduced ordinary linear model*, *Ridge Regression* and *LASSO* are used to predict the mill rate.
 A full assessment of the performance - prediction power and goodness of fit of these models, is shown below.
@@
 We also seek to identify which explanatory variables are the most important in determining mill rates.
-Every year, the assessment value of each property is released at the beginning of the year, however, the mill rate is still unknown until Spring.
+Every year, the assessment value of each property is released at the beginning of the year; however, the mill rate is still unknown until Spring.
 Prediction of mill rate is a focus of interest because it gives an approximate property tax to pay for property owners.
@@
 
-Correlation between the mill rate and each explanatory variable will be used to pick the essential variables in our model.
-Then, a variety of linear models will be fitted using our selected variable.
+Correlations between the mill rate and each explanatory variable are used to pick the essential variables in our model.
+Then, a variety of linear models are fitted using our selected variables.
 The best model is selected based on its prediction power and goodness of fit.
@@
 
-Our client provides us the past 5 years' property assessment data in BC.
-Since the only interest is in predicting mill rate for metro Vancouver and specific tax class code, a subset of properties that satisfy our interest have been selected:
+Our client provides us with the past 5 years' property assessment data in BC.
+Since the only interest is in predicting mill rate for metro Vancouver and specific tax classes, a subset of properties that satisfy our interest have been selected:
 
@@
 Before any prediction on the future mill rate of Metro Vancouver’s real estate market was made, exploratory data analysis was performed to explore and visualize the main characteristics of our dataset.
-Correlation analyses between Mill Rate vs.
-Assessment, Mill Rate vs.
-Land Total, and Mill Rate vs.
-Improvement Total were performed.
-From our initial analysis, outliers in municipalities had been found.
+Correlation analyses between Mill Rate and Total Assessment, Mill Rate and Total Land Assessment, and Mill Rate and Total Improvement Assessment were performed.
+From our initial analysis, outliers in municipalities were found.
 Data transformation - calculating the average total assessment, was used to reduce the effect of outliers.
@@
 total, land, and improvement assessment were created to see the correlation between each pair of the two factors.
-Kruskal Wallis analysis was also performed to see the correlation between mill rate vs.
-tax class and mill rate vs.
-municipality.
+Kruskal Wallis analysis was also performed to see the correlation between mill rate and tax class and mill rate and municipality.
 
@@
 
-In this study,  linear models were built and the performance of each model was evaluated by the goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.
-The definition of the goodness of fit and prediction power is given below.
+In this study, linear models were built and the performance of each model was evaluated by the goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values, respectively.
+The definitions of the goodness of fit and prediction power are given below.
+
+Prior to examine the goodness of fit and prediction power, the dataset was divided into training sets - used to build our models, and test sets - used to evaluate the prediction power of our models.
 
@@
  - **Prediction power** measures how well models can predict future values.
-Mean squared prediction power was used to compare the prediction performance across all of our fitted linear predictive models.
-Prior to examine the prediction power, the data set was divided into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.
-MSPE (Mean Squared Predction Error) on testing sets is used to measure the prediction power in this study.
+Mean squared prediction power is used to compare the prediction performance across all of our fitted linear predictive models.
+MSPE (Mean Squared Predction Error) on test sets is used to measure the prediction power in this study and a smaller MSPE indicates better prediction power.
 
-Refer to **Reference** for forumlas of MSE and MSPE.
+Refer to **Reference** for forumlas of MSPE.
 
@@
 TaxClassCode, Municipalities, assessTotal, landTotal, improvementTotal and propertyCount were considered in this model.
-Based on the results of EDA, we selected a list of significant variables and included them in another linear model (*OLR transformed*).
+Based on the results of EDA, a list of significant variables were selected and included in another linear model (*OLR transformed*).
 To compare the effect of linear models with and without features, a null model (*OLR null*) with no features used, was also constructed.
@@
 Other than ordinary linear regressions, we are also interested in the performance of more advanced linear regressions like *Ridge* and *Lasso*.
-*Ridge* and *Lasso* have a different objective function to optimize; they take a penalty in the sum of absolute values and sum of squared absolute values of weights respectively.
+*Ridge* and *Lasso* have different objective functions to optimize; they take penalties in the sum of absolute values and sum of squared absolute values of weights respectively.
 A reason to consider Ridge Regression is that it helps deal with multicollinearity of the explanatory variables.
-This might be relevant to our study as some features used in this study are correlated.
-Lasso Regression is also included in this study because it does variable selections automatically by imposing a constraint on model parameters that causes some regression coefficients to shrink to zero.
+This might be relevant to our study as some features used in this study are correlated (assessTotal, landTotal and improvementTotal).
+Lasso Regression is also included in this study because it does variable selections automatically by imposing a constraint on model parameters that possibably causes some regression coefficients to shrink to zero.
 
-Also, another advantage of these models is that MSPE is more stable as the variance of MSPE is reduced. For more detials about the two models, please refer to **Reference**.
+Also, another advantage of these models is that they help reduce the variance of MSPE, so that MSPE is more stable. For more detials about the two models, please refer to **Reference**.
 
@@
 To examine the goodness of fit and prediction power, a 50-run of 10-fold cross-validation was performed in this study.
-For each run, a 10-fold cross-validation was used to train the five models and make predictions on training and testing data respectively.
-Then, the MSEs calculated from the training data and MSPEs calculated from the testing data were stored in vectors of corresponding models.
+For each run, a 10-fold cross-validation was used to train the five models and make predictions on training and test sets respectively.
+Then, the MSPEs calculated from the training sets and MSPEs calculated from the test sets were stored in vectors of corresponding models.
 
-After the 50 runs, a vector of MSEs and a vector of MSPEs for each model were therefore constructed successfully.
-Based on these vectors, side-by-side boxplots were used to show the mean and the spread of MSE and MSPE across all models respectively.
+After the 50 runs, a vector of MSPEs on the training sets and a vector of MSPEs on the test sets for each model were therefore constructed successfully.
+Based on these vectors, side-by-side boxplots were used to show the mean and the spread of MSPEs on the training sets and test sets across all models respectively.
+
 For more details about cross-validation, please refer to **Reference**.
@@
 
-There is no clear trend between the mill rate and total assessment.
-Plot has shown that most points are condensed on the left horizontal axis since some municipalities have larger assessment values than others.
+There is no clear trend between the mill rate and total assessment from Figure 1.
+The plot has also shown that most points are condensed on the left horizontal axis since some municipalities have larger assessment values than others.
 
@@
 
-Here categorical variables are taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 3 and Figure 4.
+Here categorical variables are taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 3 and Figure 4, respectively.
 
@@
 
-Figure 3 shows that most municipalities have different mean and variance in terms of mill rates.
+Figure 3 has shown that most municipalities have different mean and variance in terms of mill rates.
 
-Figure 4 also supports that there is unequal mean and variance across tax classes.
+Figure 4 has also supported that there is unequal mean and variance across tax classes.
 
@@
 
-The p-values in Table 1 support that there is nonidentical distributions of mill rates across municipalities and tax classes.
+The p-values in Table 1 have supported that there is nonidentical distributions of mill rates across municipalities and tax classes.
 
@@
 
-Below is a comparison of the five models (*Lasso*, *Ridge*, *OLR transformed*, *OLR full* and *OLR null*) using the goodness of fit and prediction power.
-The goodness of fit was measured by MSE of the training data, whereas prediction power was measured by MSPE of the testing data.
-Generally, smaller MSEs and MSPEs indicate better fit and prediction power respectively.
+Below is a comparison of the five models (*OLR full*, *OLR transformed*, *Lasso*, *Ridge* and *OLR null*) using the goodness of fit and prediction power.
+The goodness of fit was measured by MSPE on the training sets, whereas prediction power was measured by MSPE on the test sets.
+Generally, smaller MSPEs on the training sets and test sets indicate better fit and prediction power respectively.
 
@@
 
-The distributions of MSPE across models are also displayed in Table 3.
+The distributions of MSPE on the training sets and test sets across all models are also displayed in Table 3 and Table 4 respectively.
 
@@
         ylab = expression(hat(MSPE)))
+```
 
+                          Table 2: Goodness of fit of All Models
+
+Model | PMSE:Min | PMSE:Mean | PMSE:Max
+-----|-----|-----|-----|
+Full Model | 5.23 | 5.43 | 5.52 |
+Transformed Model | 4.45 | 4.59 | 4.64 |
+Null Model | 36.33 | 36.52 | 36.64 |
+The Lasso for Full Model | 4.52 | 4.63 | 4.73 |
+Ridge Regression for Full Model | 4.50 | 4.63 | 4.72 |
+
+```{r, warning=FALSE, echo=FALSE}
 boxplot(mspe.la, 
@@
 
+\pagebreak
+
 ## Conclusion
@@
 From the exploratory data analysis, we have found that Total Assessments contain outliers in major municipalities such as Vancouver and Burnaby.
-In order to reduce the effect of outliers, the Total Assessments were transformed by taking their average over the number of properties in each municipality.
+In order to reduce the effect of outliers, the Total Assessment was transformed by taking its average over the number of properties in each municipality.
 The correlation analyses had shown that the transformed assessment total is the only continuous variable that has a relatively strong correlation with the mill rate.
 The Kruskal-Wallis had suggested that mill rates in each municipality and tax class are significantly different.
-From these results, we have decided to use Average Assessment Total, TaxClassCode and Municipality to fit our reduced model.
+From these results, we have decided to use Average Assessment Total, TaxClassCode and Municipality to fit our transformed model.
 
-*Transformed OLR*, *Ridge Regression*, and *the LASSO* were able to make good predictions based on mean squared error and mean squared prediction error from cross-validation.
+*Transformed OLR*, *Ridge Regression*, and *LASSO* were able to make good predictions based on mean squared prediction error on the training sets and test sets from cross-validation.
 Since the client prefers a simpler model, we choose the *transformed model* to make our 2020 prediction.
@@
 
-                    Table 3: Data with missing mill rate
+                    Table 4: Data with missing mill rate
 

### 2020-03-31 src/2nd_draft.Rmd
fix typo

@@
 This might be relevant to our study as some features used in this study are correlated (assessTotal, landTotal and improvementTotal).
-Lasso Regression is also included in this study because it does variable selections automatically by imposing a constraint on model parameters that possibably causes some regression coefficients to shrink to zero.
+Lasso Regression is also included in this study because it does variable selections automatically by imposing a constraint on model parameters that may possibably causes some regression coefficients to shrink to zero.
 

### 2020-03-31 src/2nd_draft.Rmd
typo

@@
 Mean squared prediction power is used to compare the prediction performance across all of our fitted linear predictive models.
-MSPE (Mean Squared Predction Error) on test sets is used to measure the prediction power in this study and a smaller MSPE indicates better prediction power.
+MSPE (Mean Squared Prediction Error) on test sets is used to measure the prediction power in this study and a smaller MSPE indicates better prediction power.
 
@@
 
-Also, another advantage of these models is that they help reduce the variance of MSPE, so that MSPE is more stable. For more detials about the two models, please refer to **Reference**.
+Also, another advantage of these models is that they help reduce the variance of MSPE, so that MSPE is more stable. For more details about the two models, please refer to **Reference**.
 
@@
 
-The p-values in Table 1 have supported that there is nonidentical distributions of mill rates across municipalities and tax classes.
+The p-values in Table 1 have supported that there are nonidentical distributions of mill rates across municipalities and tax classes.
 

### 2020-03-31 src/2nd_draft.Rmd
finalized

@@
 title: 'STAT 450 Project: Real Estate'
-author: "Xuechun Lu, Yuting Wen, Peter Han, Yuetong Liu"
-date: "15/03/2020"
+author: "Peter Han, Xuechun Lu, Yuetong Liu, Yuting Wen"
+date: "30/03/2020"
 output:
@@
 Null Model | 36.33 | 36.52 | 36.64 |
-The Lasso for Full Model | 4.52 | 4.63 | 4.73 |
+Lasso for Full Model | 4.52 | 4.63 | 4.73 |
 Ridge Regression for Full Model | 4.50 | 4.63 | 4.72 |
@@
 Null Model | 36.67 | 36.87 | 37.26 |
-The Lasso for Full Model | 5.58 | 5.91 | 6.40 |
+Lasso for Full Model | 5.58 | 5.91 | 6.40 |
 Ridge Regression for Full Model | 5.53 | 5.88 | 6.39 |

### 2020-03-31 src/2nd_draft.Rmd

@@
 
+For predicted mill rate in 2020, please refer to **Appendix** Table 4.
+
 \pagebreak
@@
 
+### 2020 Mill Rate prediction
+
+```{r, echo=FALSE, warning=FALSE}
+mill_rate_pre_2020 <- read.csv(here("data","assessment_2020_fit.csv"))[,-c(1,2,5)]
+```
+
+                Table 4: Predicted Mill Rate in 2020 based on transformed model (%)
+
+Municipality | Residential (%) | Light industry (%) | Business (%)                      
+------|------|------|------
+Anmore | 0.98 | NA | 9.91
+Belcarra | 1.19 | NA | 12.31
+Bowen Island | 1.63 | 14.69 | 11.15
+Burnaby | -2.76 | 12.80 | 10.88
+Coquitlam | 1.72 | 17.70 | 12.84
+Delta | 5.50 | 12.98 | 13.65
+Langley - City | 1.73 | 15.12 | 12.25
+Langley - Township | 3.77 | 12.46 | 12.49
+Lions Bay | 2.19 | NA | 11.66
+Maple Ridge | 5.11 | 16.88 | 13.60
+New Westminster | 4.65 | 18.04 | 15.78
+North Vancouver - City | -2.17 | 15.51 | 11.90
+North Vancouver - Dist | 1.29 | 13.68 | 11.82
+Pitt Meadows | 6.06 | 17.00 | 13.78
+Port Coquitlam | 4.91 | 17.58 | 14.50
+Port Moody | 4.16 | 17.66 | 13.44
+Richmond | 1.14 | 10.42 | 10.29
+Surrey | 0.82 | 12.47 | 10.53
+Vancouver | 2.69 | 12.38 | 8.79
+West Vancouver | 4.92 | 17.05 | 12.62
+White Rock | 1.88 | 16.33 | 12.17
+
 ### Missing Value:
@@
 
-                    Table 4: Data with missing mill rate
+                          Table 5: Data with missing mill rate
 

### 2020-03-31 doc/450/RealEstate_Xuechun_Lu.Rmd
moved to 450 folder

--- 2020-03-03 doc/RealEstate_Xuechun_Lu.Rmd
@@
 
-## 2. Measure of goodness of fit and prediction power
+## 2. Measure of Goodness of Fit and Prediction Power
 
@@
 
-assessment_aggregate <- read.csv(here("RealEstate/data","assessment_aggregate.csv"))
+assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
 
-tax_final <- read.csv(here("RealEstate/data","tax_final.csv"))
+tax_final <- read.csv(here("data","tax_final.csv"))
 #head(tax_final)

### 2020-04-02 src/a_rough_clean.R
Add files via upload
some prelim cleaning.. use with care and read comments!

@@ -1,0 +1,99 @@
+# Mallory ###
+# some cleaning.. with some simplifications we probably don't want
+library(tidyverse)
+library(readr)
+library(readxl)
+library(ggplot2)
+library(viridis)
+library(here)
+library(rnn)
+rm(list=ls())
+
+### DATA CLEANING 
+# import assessment and sales data for all years
+assess_all <- read.csv(here("data/Assessment/2016 - 2020 Raw.csv"),
+                       header = F)
+col_names <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
+colnames(assess_all) <- colnames(col_names)
+
+# select relevant features 
+assess <- assess_all %>%
+  select(PIC, Year, AssessorAreaCode:AddressAssessorMunicipalityDesc, 
+         TaxClassCode:TaxClassTaxRate)  %>% 
+  filter(TaxClassCode %in% c("01","05","06"))
+
+# change types
+assess <- assess %>% 
+  mutate_at(c(6, 10, 11), as.character) %>% 
+  mutate_at(c(6, 10, 11), as.numeric)
+
+# check for NA's (thanks to STAT450 for this code)
+for (feature in colnames(assess)){
+  variable = eval(parse(text = paste0("assess$",feature)))
+  null.count = sum(is.na(variable))
+  print(paste(feature, null.count, sep = ":"))
+} 
+
+# check for constant mill rate across municipalities
+test <- assess %>% 
+  group_by(Year, AddressAssessorMunicipalityDesc, TaxClassCode) %>% 
+  filter(!is.na(TaxClassTaxRate)) %>% 
+  summarise(test = var(TaxClassTaxRate)) %>%
+  filter(test !=0)
+
+### for now just took mean of mill rates..
+assess.rate <- assess %>%
+  group_by(Year, AddressAssessorMunicipalityDesc, TaxClassCode) %>% 
+  filter(!is.na(TaxClassTaxRate)) %>% 
+  mutate(TaxClassTaxRate = mean(TaxClassTaxRate))
+
+# check that this reduced variance in mill rates to 0 for all municipalities
+test1 <- assess.rate %>% 
+  group_by(Year, AddressAssessorMunicipalityDesc, TaxClassCode) %>%
+  summarise(test = var(TaxClassTaxRate)) %>%
+  filter(test !=0)
+
+# create dataset listing mill rates per municipality, year, and tax class
+assess.agg <- assess.rate%>%
+  group_by(Year, AddressAssessorMunicipalityDesc,TaxClassCode)%>%
[…]

### 2020-04-05 src/eda - s550/eda.R
Add test and train data set. -gd

@@ -186,11 +186,15 @@ ggsave("RealEstate/src/eda - s550/plots/4. scatter average year.png")
     dplyr::filter(!is.na(total.assessment), !is.na(mill.rate), year == 2020) %>% 
     dplyr::group_by(PIC, tax.class) %>% 
     dplyr::summarise(total.assessment = mean(total.assessment), mill.rate = mean(mill.rate)) %>% 
-    ggplot(aes(x = tax.class, y = log(mill.rate))) +
-    geom_violin() +
+    ggplot(aes(x = tax.class, y = log(mill.rate), fill = tax.class)) +
+    geom_violin(alpha = 0.5, width = 1) +
+    geom_boxplot(alpha = 0.75, width = 0.1) +
     labs(x = "tax class",
          y = "log mill rate",
-         title = "Mill rates accross tax classes")
+         title = "Mill rates accross tax classes") +
+    #scale_fill_viridis_d(begin=0, end=1) +
+    scale_fill_manual(values = c("#3E4A89FF", "#26828EFF","#B4DE2CFF" )) +
+    theme(legend.position = "none")
   ggsave("RealEstate/src/eda - s550/plots/7. violin mill rates.pdf")
   ggsave("RealEstate/src/eda - s550/plots/7. violin mill rates.png")
   
@@ -207,6 +211,22 @@ ggsave("RealEstate/src/eda - s550/plots/4. scatter average year.png")
          title = "Mill rates accross tax classes")
   ggsave("RealEstate/src/eda - s550/plots/8. boxplot mill rates.pdf")
   ggsave("RealEstate/src/eda - s550/plots/8. boxplot mill rates.png")
+  
+# export data set
+set.seed(2303)
+train <- re %>% 
+  dplyr::select(PIC) %>% 
+  dplyr::distinct() %>% 
+  dplyr::sample_frac(0.75) %>% 
+  dplyr::pull()
+
+re.out <- re %>% 
+  dplyr::mutate(test.train = ifelse(PIC %in% train, "train", "test"))
+
+readr::write_delim(re.out, "test_train_data.txt", delim = ",")
+# read for test
+dataset <- readr::read_delim("test_train_data.txt", delim = ",",
+                             col_types = "ciccdddddc")
 
 ## S450 data ####
 re.450 <- read_csv("RealEstate/data/assessment_aggregate.csv")

### 2020-04-07 src/rf.Rmd
rf

---
title: "random forest"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(randomForest))
suppressPackageStartupMessages(library(tidyverse))
```

```{r}
dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
dat <- dat %>% 
  mutate(next.assess = lead(total.assessment, order_by = year)) %>%
  arrange(year)
train1 <- dat %>%  filter(test.train == "train" & tax.class == "01") 
test1 <- dat %>%  filter(test.train == "test" & tax.class == "01")
train5 <- dat %>%  filter(test.train == "train" & tax.class == "05")
test5 <- dat %>%  filter(test.train == "test" & tax.class == "05")
train6 <- dat %>%  filter(test.train == "train" & tax.class == "06")
test6 <- dat %>%  filter(test.train == "test" & tax.class == "06")
```

```{r}
rf1.mill <- randomForest(
  mill.rate ~ tax + municipality + total.assessment + year,
  data=na.omit(train1)
)
```

### 2020-04-07 src/rf.Rmd
changed ordering

@@
 dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
-dat <- dat %>% 
+dat <- dat %>% group_by(PIC) %>% 
   mutate(next.assess = lead(total.assessment, order_by = year)) %>%
-  arrange(year)
+  arrange(PIC)
+filter(dat, PIC == "CA-BC-003-6500000")
 train1 <- dat %>%  filter(test.train == "train" & tax.class == "01") 

### 2020-04-07 src/rf.Rmd
added prop counts

--- 2020-04-07 src/rf.Rmd
@@
 dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
+dat$municipality <- as.factor(dat$municipality)
 dat <- dat %>% 
+  group_by(PIC) %>% 
   mutate(next.assess = lead(total.assessment, order_by = year)) %>%
-  arrange(year)
+  arrange(PIC) %>% 
+  group_by(municipality, year) %>% 
+  mutate(n.prop = n()) %>% 
+  arrange(desc(n.prop)) %>% 
+  distinct(municipality, .keep_all = T)
+
 train1 <- dat %>%  filter(test.train == "train" & tax.class == "01") 

### 2020-04-08 doc/450/RealEstate_final_report.Rmd
final report RMD, pdf, supported file added

--- 2020-03-31 src/2nd_draft.Rmd
@@
 author: "Peter Han, Xuechun Lu, Yuetong Liu, Yuting Wen"
-date: "30/03/2020"
+date: "07/04/2020"
 output: 
-  pdf_document
+  rmarkdown::pdf_document:
+    fig_caption: yes        
+    includes:  
+      in_header: my_header.tex
 ---
@@
 library(dummies)
+library(scales)
 set.seed(123)
@@
 *Ordinary linear model*, *reduced ordinary linear model*, *Ridge Regression* and *LASSO* are used to predict the mill rate.
-A full assessment of the performance - prediction power and goodness of fit of these models, is shown below.
+A full assessment of the performance --- prediction power and goodness of fit of these models, is shown below.
 
@@
 
-Our client provides us with the past 5 years' property assessment data in BC.
+Our client provided us with the past 5 years' property assessment data in BC.
 Since the only interest is in predicting mill rate for metro Vancouver and specific tax classes, a subset of properties that satisfy our interest have been selected:
@@
 To reduce the dimension of our data, all properties in the same region, tax class code, and year are aggregated into a group because these properties have the same mill rate,  which is our response variable.
-Here is the summary statistics for these groups:
+Here are the summary statistics for these groups:
 
@@
 - Total Improvement Assessment (improvementTotal)
-- Total number of properties (propertyCount)
+- Total Number of Properties (propertyCount)
 - Tax Class Code (TaxClassCode)
 - Municipality (AddressAssessorMunicipalityDesc)
-- Year
 
@@
 From our initial analysis, outliers in municipalities were found.
-Data transformation - calculating the average total assessment, was used to reduce the effect of outliers.
+Data transformation --- calculating the average total assessment, was used to reduce the effect of outliers.
 
@@
 
-The full linear model (*ORL full*) was built first.
+The full linear model (*OLR full*) was built first.
 TaxClassCode, Municipalities, assessTotal, landTotal, improvementTotal and propertyCount were considered in this model.
-Based on the results of EDA, a list of significant variables were selected and included in another linear model (*OLR transformed*).
+Based on the results of EDA, a list of significant variables were selected and included in another linear model (*OLR transformed*), as shown in **Table 2**.
 To compare the effect of linear models with and without features, a null model (*OLR null*) with no features used, was also constructed.
+
+Refer to **Appendix** for variables selected in both *OLR full* and *OLR transformed* models.
 
@@
 This might be relevant to our study as some features used in this study are correlated (assessTotal, landTotal and improvementTotal).
-Lasso Regression is also included in this study because it does variable selections automatically by imposing a constraint on model parameters that may possibably causes some regression coefficients to shrink to zero.
+Lasso Regression is also included in this study because it does variable selections automatically by imposing a constraint on model parameters that may possibly cause some regression coefficients to shrink to zero.
 
@@
 
-Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and total assessment among different municipalities is created to show the relationship in Figure 1.
+Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and total assessment among different municipalities is created to show the relationship in **Figure 1**.
 Each color in the figure belongs to three tax classes of one municipality.
 
-Scatter plots between mill rate and total land assessment, mill rate and total improvement assessment are similar to Figure 1. For more details about them, please refer to **Appendix**.
+Scatter plots between mill rate and total land assessment, mill rate and total improvement assessment are similar to **Figure 1**. For more details about them, please refer to **Appendix**.
 
@@
 names(assessment_eda)[9] <- paste("Average Total Assessment")
+```
 
+```{r, echo=FALSE, fig.cap="Mill Rate vs Total Assessment"}
 assessment_eda %>%
@@
    geom_point() + 
-   ggtitle(sprintf("Figure 1: Mill Rate vs Total Assessment")) + 
+  xlab("Total Assessment (in Billion)") +
    geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
-   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+  scale_x_continuous(labels = unit_format(unit = "Billion", scale = 1e-9))
 ```
 
-There is no clear trend between the mill rate and total assessment from Figure 1.
+There is no clear trend between the mill rate and total assessment from **Figure 1**.
 The plot has also shown that most points are condensed on the left horizontal axis since some municipalities have larger assessment values than others.
@@
 The transformed data is named **"Average Total Assessment"**.
-A scatter plot between mill rates and average total assessments is shown in Figure 2.
+A scatter plot between mill rates and average total assessments is shown in **Figure 2**.
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, fig.cap="Mill rate v.s. Average Total Assessment"}
 assessment_eda %>%
@@
  geom_point() + 
- ggtitle(sprintf("Figure 2: Mill rate v.s. Average Total Assessment")) + 
+  xlab("Average Total Assessment (in Million)") +
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") +
-theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+  scale_x_continuous(labels = unit_format(unit = "Million", scale = 1e-6))
 ```
 
-The plot from Figure 2 has shown that the mill rate tends to decrease as the average total assessment increase. Also, they have a moderately strong linear correlation.
+The plot from **Figure 2** has shown that the mill rate tends to decrease as the average total assessment increase. Also, they have a moderately strong linear correlation.
 
@@
 
-Here categorical variables are taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in Figure 3 and Figure 4, respectively.
+Here categorical variables are taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in **Figure 3** and **Figure 4**, respectively.
 
-```{r, echo=FALSE}
+```{r, echo=FALSE,fig.cap="Mill rate across Municipalities"}
 # ______________________________________________________________________________________________
@@
      geom_boxplot() + 
-     ggtitle(sprintf("Figure 3: Mill rate across Municipalities")) +
-     theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
+     theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none", axis.title.x=element_blank())
 ```
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, fig.cap="Mill rate across Tax Classes", fig.width=6,fig.height=3}
 assessment_eda %>%
@@
      geom_boxplot() + 
-     ggtitle(sprintf("Figure 4: Mill rate across Tax Classes")) +
-     theme(legend.position = "none") +
-     labs(x="Tax Class") +
-     scale_x_discrete(labels=c("1", "5", "6"))
+     theme(legend.position = "none", axis.title.x=element_blank()) +
+     scale_x_discrete(labels=c("Residential (1)", "Light industry (5)", "Business and other (6)"))
 ```
 
-Figure 3 has shown that most municipalities have different mean and variance in terms of mill rates.
+**Figure 3** has shown that most municipalities have different mean and variance in terms of mill rates.
 
-Figure 4 has also supported that there is unequal mean and variance across tax classes.
+**Figure 4** has also supported that there is unequal mean and variance across tax classes.
 
@@
 
-The results of Kruskal-Wallis Test is shown in Table 1.
+The results of Kruskal-Wallis Test is shown in **Table 1**.
 
-        Table 1: Kruskal-Wallis Test of Mill Rate across Municipality and Tax Class
+| Distribution  | p-value |
+| :------------ | :-----------: | 
+| Mill Rate across Municipality | 0.00675 |
+| Mill Rate across Tax Class | < 2.2e-16 |
 
-Distribution  | p-value
-------|--------
-Mill Rate across Municipality | 0.00675
-Mill Rate across Tax Class | < 2.2e-16
+Table: Kruskal-Wallis Test of Mill Rate across Municipality and Tax Class
 
-The p-values in Table 1 have supported that there are nonidentical distributions of mill rates across municipalities and tax classes.
+The p-values in **Table 1** have supported that there are nonidentical distributions of mill rates across municipalities and tax classes.
 
@@
 ```
+
+In conclusion, Average Total Assessment, Tax Class Code and Municipality are selected to fit another linear model, named **OLR transformed**. Refer to **Appendix** for more detail.
 
@@
 
-The distributions of the goodness of fit and prediction power across models are displayed in Figure 5 and Figure 6, respectively, the distribution of null model is removed since it has larger values compared to all other models.
+The distributions of the goodness of fit and prediction power across models are displayed in **Figure 5** and **Figure 6**, respectively, the distribution of null model is removed since it has larger values compared to all other models.
 
-The distributions of MSPE on the training sets and test sets across all models are also displayed in Table 3 and Table 4 respectively.
+The distributions of MSPE on the training sets and test sets across all models are also displayed in **Table 2** and **Table 3** respectively.
 
@@
 }
+```
 
+```{r, warning=FALSE, echo=FALSE, fig.cap="Goodness of Fit"}
 boxplot(mspe1.la, 
@@
         cex.lab = 1, 
-        cex.main = 1,
-        main = " Figure 7: Goodness of Fit",
         ylab = expression(hat(MSPE)))
@@
 
-                          Table 2: Goodness of fit of All Models
+| Model | PMSE:Min | PMSE:Mean | PMSE:Max |
+| :-----: | :-----: | :-----: | :-----: |
+| Full Model | 5.23 | 5.43 | 5.52 |
+| Transformed Model | 4.45 | 4.59 | 4.64 |
+| Null Model | 36.33 | 36.52 | 36.64 |
+| Lasso for Full Model | 4.52 | 4.63 | 4.73 |
+| Ridge Regression for Full Model | 4.50 | 4.63 | 4.72 |
 
-Model | PMSE:Min | PMSE:Mean | PMSE:Max
------|-----|-----|-----|
-Full Model | 5.23 | 5.43 | 5.52 |
-Transformed Model | 4.45 | 4.59 | 4.64 |
-Null Model | 36.33 | 36.52 | 36.64 |
-Lasso for Full Model | 4.52 | 4.63 | 4.73 |
-Ridge Regression for Full Model | 4.50 | 4.63 | 4.72 |
+Table: Goodness of fit of All Models
 
-```{r, warning=FALSE, echo=FALSE}
+```{r, warning=FALSE, echo=FALSE, fig.cap="Prediction Power"}
 boxplot(mspe.la, 
@@
         cex.lab = 1, 
-        cex.main = 1,
-        main = "Figure 8: Prediction Power",
         ylab = expression(hat(MSPE)))
@@
 
-                            Table 3: Prediction Power of All Models
+| Model | PMSE:Min | PMSE:Mean | PMSE:Max |
+| :-----: | :-----: | :-----: | :-----: |
+| Full Model | 6.50 | 6.97 | 7.59 |
+| Transformed Model | 5.60 | 5.92 | 6.51 |
+| Null Model | 36.67 | 36.87 | 37.26 |
+| Lasso for Full Model | 5.58 | 5.91 | 6.40 |
+| Ridge Regression for Full Model | 5.53 | 5.88 | 6.39 |
 
-Model | PMSE:Min | PMSE:Mean | PMSE:Max
------|-----|-----|-----|
-Full Model | 6.50 | 6.97 | 7.59 |
-Transformed Model | 5.60 | 5.92 | 6.51 |
-Null Model | 36.67 | 36.87 | 37.26 |
-Lasso for Full Model | 5.58 | 5.91 | 6.40 |
-Ridge Regression for Full Model | 5.53 | 5.88 | 6.39 |
+Table: Prediction Power of All Models
 
@@
 
-write.csv(b, "../Data/assessment_2020_fit.csv")
+write.csv(b, "../../Data/assessment_2020_fit.csv")
 ```
@@
 
+For predicted mill rate in 2020, please refer to **Appendix** **Table 6**.
+
 \pagebreak
@@
 ## Appendix
+
+### Variables selected in different models
+
+| Parameter  | Type | 
+| :------------ | :-----------: | 
+| Total Assessment     | Quantitative   | 
+| Total Land Assessment     | Quantitative   | 
+| Total Improvement Assessment     | Quantitative   | 
+| Total Number of Properties     | Quantitative   | 
+| Tax Class Code     | Categorical  | 
+| Municipality       | Categorical |
+
+Table: Selected Explanatory Variables of OLR full
+
+*Ridge Regression* and *Lasso* models were fitted based on *OLR full* model.
+
+| Parameter  | Type | 
+| :------------ | :-----------: | 
+| Average Total Assessment     | Quantitative   | 
+| Tax Class Code     | Categorical  | 
+| Municipality       | Categorical |
+
+Table: Selected Explanatory Variables of OLR transformed
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, fig.cap="Mill Rate vs Total Land Assessment",fig.width=6,fig.height=4}
 assessment_eda %>%
@@
    geom_point() + 
-   ggtitle(sprintf("Figure 7: Mill Rate vs Total Land Assessment")) + 
+  xlab("Total Land Assessment (in Billion)") +
    geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
-   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+  scale_x_continuous(labels = unit_format(unit = "B", scale = 1e-9))
 ```
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, fig.cap="Mill Rate vs Total Improvement Assessment",fig.width=6,fig.height=4}
 assessment_eda %>%
@@
    geom_point() + 
-   ggtitle(sprintf("Figure 8: Mill Rate vs Total Improvement Assessment")) + 
+  xlab("Total Improvement Assessment (in Billion)") +
    geom_smooth(aes(group = 1), size = 0.5, method = "glm", se = FALSE, colour = "black") +
-   theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
+  scale_x_continuous(labels = unit_format(unit = "Billion", scale = 1e-9))
 ```
+
+### 2020 Mill Rate prediction
+
+```{r, echo=FALSE, warning=FALSE}
+mill_rate_pre_2020 <- read.csv(here("data","assessment_2020_fit.csv"))[,-c(1,2,5)]
+```
+
+| Municipality | Residential (%) | Light industry (%) | Business (%) |                    
+| :------: | :------: | :------: | :------: |
+| Anmore | 0.98 | NA | 9.91 |
+| Belcarra | 1.19 | NA | 12.31 |
+| Bowen Island | 1.63 | 14.69 | 11.15 |
+| Burnaby | -2.76 | 12.80 | 10.88 |
+| Coquitlam | 1.72 | 17.70 | 12.84 |
+| Delta | 5.50 | 12.98 | 13.65 |
+| Langley - City | 1.73 | 15.12 | 12.25 |
+| Langley - Township | 3.77 | 12.46 | 12.49 |
+| Lions Bay | 2.19 | NA | 11.66 |
+| Maple Ridge | 5.11 | 16.88 | 13.60 |
+| New Westminster | 4.65 | 18.04 | 15.78 |
+| North Vancouver - City | -2.17 | 15.51 | 11.90 |
+| North Vancouver - Dist | 1.29 | 13.68 | 11.82 |
+| Pitt Meadows | 6.06 | 17.00 | 13.78 |
+| Port Coquitlam | 4.91 | 17.58 | 14.50 |
+| Port Moody | 4.16 | 17.66 | 13.44 |
+| Richmond | 1.14 | 10.42 | 10.29 |
+| Surrey | 0.82 | 12.47 | 10.53 |
+| Vancouver | 2.69 | 12.38 | 8.79 |
+| West Vancouver | 4.92 | 17.05 | 12.62 |
+| White Rock | 1.88 | 16.33 | 12.17 |
+
+Table: Predicted Mill Rate in 2020 based on transformed model
 
@@
 
-                    Table 4: Data with missing mill rate
+| Year  | Region | Class | Number of Properties |
+| :------: | :--------: | :-------: | :----------: |
+| 2016 | Belcarra | 06 | 9 |
+| 2016 | Lions Bay | 01 | 40 | 
+| 2016 | Lions Bay | 06 | 25 |
+| 2016 | Maple Ridge Rural | 05 | 36 |
+| 2017 | Belcarra | 06 | 9 |
+| 2017 | Lions Bay | 01 | 39 |
+| 2017 | Lions Bay | 06 | 24 |
+| 2017 | Maple Ridge Rural | 05 | 36 |
+| 2018 | Maple Ridge Rural | 05 | 36 |
+| 2019 | Maple Ridge Rural | 05 | 38 |
 
-Year  | Region | Class |Number of Properties
-------|--------|-------|----------
-2016 | Belcarra | 06 | 9 
-2016 | Lions Bay | 01 | 40 
-2016 | Lions Bay | 06 | 25 
-2016 | Maple Ridge Rural | 05 | 36 
-2017 | Belcarra | 06 | 9 
-2017 | Lions Bay | 01 | 39 
-2017 | Lions Bay | 06 | 24 
-2017 | Maple Ridge Rural | 05 | 36 
-2018 | Maple Ridge Rural | 05 | 36 
-2019 | Maple Ridge Rural | 05 | 38
+Table: Data with missing mill rate
 

### 2020-04-08 src/eda - s550/eda.R
Update eda code. -gd

@@ -7,6 +7,7 @@ library(readxl)
 library(ggplot2)
 ggplot2::theme_set(theme_classic())
 library(viridis)
+library(GGally)
 
 # data wrangling ####
 
@@ -49,6 +50,18 @@ re <- real.estate_full %>%
   dplyr::distinct() %>% 
   dplyr::left_join(assessments, by = c("PIC" = "PIC", "year" = "year")) # add assessment
 
+# properties by municipality 2020
+re %>% 
+  dplyr::select(municipality, PIC) %>% 
+  #dplyr::filter(year == 2020) %>% 
+  dplyr::group_by(municipality) %>% 
+  dplyr::summarise(n.properties = n()) %>% 
+  dplyr::arrange(desc(n.properties)) %>% 
+  dplyr::mutate(perc.properties = n.properties / sum(n.properties),
+                acum.properties = cumsum(perc.properties)) %>% 
+  View()
+  
+  
 
 # data viz ####
 # tax classes dictionary
@@ -68,10 +81,11 @@ re %>%
   ggplot(aes(x = log(total.assessment), y = log(mill.rate), color = factor(year))) +
   geom_point() +
   facet_wrap(.~tax.class, labeller = tax.classes) +
-  labs(x = "log assessment",
-       y = "log mill rate",
+  labs(x = "mean log assessment value",
+       y = "mean log mill rate",
        color = "year") +
-  scale_color_viridis_d()
+  scale_color_viridis_d() +
+  theme(text = element_text(size = 18))
 ggsave("RealEstate/src/eda - s550/plots/1. scatter with year.pdf")
 ggsave("RealEstate/src/eda - s550/plots/1. scatter with year.png")
 
@@ -86,9 +100,9 @@ re %>%
   ggplot(aes(x = log(total.assessment), y = log(mill.rate))) +
   geom_point(color = viridis(20)[3]) +
   facet_wrap(.~tax.class, labeller = tax.classes) +
-  labs(x = "log assessment",
-       y = "log mill rate",
-       title = "Assessment vs mill rates by municipality for 2020")
+  labs(x = "mean log assessment value",
+       y = "mean log mill rate") +
+  theme(text = element_text(size = 18))
 ggsave("RealEstate/src/eda - s550/plots/2. scatter 2020 by municipality.pdf")
 ggsave("RealEstate/src/eda - s550/plots/2. scatter 2020 by municipality.png")
 
[…]

### 2020-04-08 src/rf.Rmd
extend rf

--- 2020-04-07 src/rf.Rmd
@@
 dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
+dat$municipality <- as.factor(dat$municipality)
+
+factors_tbl = dat %>% 
+  group_by(municipality) %>% 
+  count(name="mun_count", sort = TRUE) %>% 
+  ungroup() %>% 
+  mutate(perc = mun_count/sum(mun_count),
+         cum_perc = cumsum(perc)) %>% 
+  arrange(desc(mun_count)) %>% 
+  mutate(rank = row_number(),
+         municipality = fct_reorder(municipality, rank)) %>% 
+  mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
+  select(municipality, col_municipality)
+
 dat <- dat %>% 
-  mutate(next.assess = lead(total.assessment, order_by = year)) %>%
-  arrange(year)
-train1 <- dat %>%  filter(test.train == "train" & tax.class == "01") 
-test1 <- dat %>%  filter(test.train == "test" & tax.class == "01")
-train5 <- dat %>%  filter(test.train == "train" & tax.class == "05")
-test5 <- dat %>%  filter(test.train == "test" & tax.class == "05")
-train6 <- dat %>%  filter(test.train == "train" & tax.class == "06")
-test6 <- dat %>%  filter(test.train == "test" & tax.class == "06")
+  left_join(factors_tbl) %>% 
+  select(-c(municipality)) %>% 
+  rename(municipality = col_municipality) %>% 
+  group_by(PIC) %>% 
+  mutate(next.assess = lead(total.assessment, order_by = year),
+         past.mill = lag(mill.rate, order_by = year)) %>%
+  arrange(PIC) %>% 
+  group_by(municipality, year) %>% 
+  mutate(n.prop = n()) %>% 
+  arrange(desc(n.prop)) %>% 
+  distinct(municipality, .keep_all = T)
+
+train <- dat %>%  filter(test.train == "train") 
+test <- dat %>%  filter(test.train == "test")
 ```
@@
 ```{r}
-rf1.mill <- randomForest(
-  mill.rate ~ tax + municipality + total.assessment + year,
-  data=na.omit(train1)
+rf.mill <- randomForest(
+  mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit,
+  data=train
 )
 ```
+
+```{r}
+predict(rf.mill, test)
+```

### 2020-04-08 src/geocoding/geocoder.py

@@ -1,0 +1,49 @@
+import os
+from geopy import geocoders
+from osgeo import ogr, osr
+
+def geocode(address):
+    g = geocoders.Nominatim(user_agent="myGeocoder")
+    place, (lat, lng) = g.geocode(address)
+    return place, lat, lng
+
+def parse_file(filepath, output_shape):
+    # create the shapefile
+    drv = ogr.GetDriverByName("ESRI Shapefile")
+    if os.path.exists(output_shape):
+        drv.DeleteDataSource(output_shape)
+    ds = drv.CreateDataSource(output_shape)
+    # spatial reference
+    sr = osr.SpatialReference()
+    sr.ImportFromProj4('+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs')
+    lyr = ds.CreateLayer(output_shape, sr, ogr.wkbPoint)
+    # fields
+    featDefn = lyr.GetLayerDefn()
+    fld_id = ogr.FieldDefn('id', ogr.OFTInteger)
+    fld_address = ogr.FieldDefn('ADDRESS', ogr.OFTString)
+    fld_address.SetWidth(255)
+    lyr.CreateField(fld_id)
+    lyr.CreateField(fld_address)
+    # read text addresses file
+    i = 0
+    f = open(filepath, 'r')
+    for address in f:
+        try:
+            address = address.replace("\"", "")
+            print("Trying address " + address)
+            place, lat, lng = geocode(address)
+            point = ogr.Geometry(ogr.wkbPoint)
+            point.SetPoint(0, lng, lat)
+            feat = ogr.Feature(lyr.GetLayerDefn())
+            feat.SetGeometry(point)
+            feat.SetField('id', i)
+            feat.SetField('ADDRESS', address)
+            lyr.CreateFeature(feat)
+            feat.Destroy()
+            i = i + 1
+            print("Done address " + address)
+        except:
+            print('Error, skipping address...' + address)
+
+parse_file(os.path.join("..", "..", "data","Assessment", "addresses.csv"),
+           os.path.join("..", "..", "data", "Assessment", "addresses_points.shp"))

### 2020-04-08 src/geocoding/get_addresses.py

@@ -1,0 +1,13 @@
+from geopy.extra.rate_limiter import RateLimiter
+from geopy.geocoders import Nominatim
+import os
+import pandas as pd
+import numpy as np
+
+column_names = pd.read_excel(os.path.join("..", "..", "data", "Assessment", "2020 Only.xlsx")).columns
+column_names = np.array(column_names)
+data = pd.read_csv(os.path.join("..", "..", "data", "Assessment", "2016 - 2020 Raw.csv"), header=None,
+                   names=column_names, usecols=['AddressFullAddress'])
+data.drop_duplicates(inplace=True)
+
+data.to_csv(os.path.join("..", "..", "data", "Assessment", "addresses.csv"), index=False, header=False)

### 2020-04-08 src/geocoding/shp_to_csv.py

@@ -1,0 +1,15 @@
+import os
+import geopandas as gpd
+import pandas as pd
+
+shapes = gpd.read_file(os.path.join("..", "..", "data", "Assessment", "addresses_points.shp"), rows=100)
+
+new_data = []
+for pt in shapes.values:
+    address = pt[1].replace("\"", "")
+    latitude = pt[2].y
+    longitude = pt[2].x
+    new_data.append([address, latitude, longitude])
+
+df = pd.DataFrame(new_data)
+df.to_csv(os.path.join("..", "..", "data", "Assessment", "addresses_points.csv"))

### 2020-04-09 src/shiny_app/global.R
shiny app in progress

@@ -1,0 +1,21 @@
+library(shiny)
+library(readr)
+library(ggplot2)
+library(dplyr)
+library(DT)
+library(shinyjs)
+library(tidyr)
+library(shinyjs)
+
+#PIC using for testing
+#CA-BC-022-04490000
+
+
+# for bookmarking button
+enableBookmarking("url")
+
+# read data
+dat <- readr::read_delim("./../../data/test_train_data.txt", 
+                         delim = ",", col_types = "ciccdddddf")
+
+

### 2020-04-09 src/shiny_app/helpers.R

@@ -1,0 +1,40 @@
+# This file will have helpers for the model and the loading of data for models
+
+# function to modify data for random forest (if needed)
+rfData <- function(data) {
+  
+  rfdat <- data
+  
+  rfdat$municipality <- as.factor(rfdat$municipality)
+  
+  factors_tbl = rfdat %>% 
+    group_by(municipality) %>% 
+    count(name="mun_count", sort = TRUE) %>% 
+    ungroup() %>% 
+    mutate(perc = mun_count/sum(mun_count),
+           cum_perc = cumsum(perc)) %>% 
+    arrange(desc(mun_count)) %>% 
+    mutate(rank = row_number(),
+           municipality = fct_reorder(municipality, rank)) %>% 
+    mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
+    select(municipality, col_municipality)
+  
+  
+  
+  rfdat <- rfdat %>% 
+    left_join(factors_tbl) %>% 
+    select(-c(municipality)) %>% 
+    rename(municipality = col_municipality) %>% 
+    group_by(PIC) %>% 
+    mutate(next.assess = lead(total.assessment, order_by = year),
+           past.mill = lag(mill.rate, order_by = year)) %>%
+    arrange(PIC) %>% 
+    group_by(municipality, year) %>% 
+    mutate(n.prop = n()) %>% 
+    arrange(desc(n.prop)) %>% 
+    distinct(municipality, .keep_all = T) 
+  
+  rfdat
+}
+
+

### 2020-04-09 src/shiny_app/server.R

@@ -1,0 +1,249 @@
+# Mallory - STAT 550 2020###
+# This is the server file of the shiny app
+
+# in case modified data needs to be accessed
+source("helpers.R")
+
+server <- function(input, output, session) {
+  
+  filtered <- reactive({
+    # SINCE PLOTS FOR MILL RATE AND ASSESSMENT VALUES DIFFER, WE NEED TO DISTINGUISH THIS HERE
+    # IF A PIC IS AVAILABLE (IN CASE OF MILL RATE - PLOT FOR MUNCIPALITY.  IN CASE OF ASSESSMENTS
+    # - PLOT THE PROPERTY(OR NOTHING))
+    
+    # Update when following inputs are changed
+    input$updateButton
+    
+    newdata <- dat
+    d <- NULL
+    print(dim(dat))
+    
+    # Filter data based on the user inputs
+    isolate({
+      #if (input$typeInput != 'Select'){
+        print("The type is not Select, it is:")
+        
+          print(input$typeInput)
+          
+          # If using PIC:
+          if(input$picInput && input$identInput!="")
+            {
+            print(paste("PIC input is", input$picInput, "and is:", sep = " "))
+            print(input$identInput)
+            
+            d <- newdata  %>% 
+              filter(PIC == input$identInput)
+          }
+          
+          
+          # If not using PIC, filter by municipality and tax class:
+          if(!input$picInput){
+            print(paste("PIC input is:", input$picInput, sep = " "))
+            d <- newdata %>% 
+              filter(tax.class == input$taxclassInput,
+                     municipality == input$municipalityInput)
+          }
+      
+      })
+    
+    # return filtered data
+    if(dim(d)[1]==0){
+      d <- NULL
+    }
+    d
+    print(dim(d))
+    print(head(d))
+    })
+  
+   ########## PLOTTING TAB ###################
+   # Add plots of either mill rate or assessment value over time to plot tab
+  munic <- reactive(input$municipalityInput)
+  
+  millrateplot <- reactive({
+    input$updateButton
+    
+    data <- filtered()
+    print("plotting mill rate plot")
+    print(head(data))
+    
+          
+    isolate({
+    if(is.null(data)){
+      p <- paste("No corresponding data.")
+    }
+
+    # plot mill rates over time for municipality chosen for mill rate predicitons
+    if (input$typeInput == 'Mill Rate'){
+      p <- ggplot(data, aes(x = year, y = mill.rate)) +
+        geom_line(color="#FF3333") +
+        geom_point(color="#FF3333") +
[…]

### 2020-04-09 src/shiny_app/ui.R

@@ -1,0 +1,102 @@
+# Mallory - STAT 550 2020###
+# This is the user interface version of the shiny app
+
+library(shinythemes)
+
+ui <- fluidPage(theme = shinytheme("paper"),
+                # header
+                div(id = "headerSection",
+                    h2("BC Mill Rate & Assessment Value Predictions"),
+                    
+                    span(
+                      style = "font-size: 1em",
+                      # authors
+                      span("Created by "),
+                      a("Gian Carlo Diluvi, Vittorio Romaniello, Sophia Li & Mallory Flynn"),
+                      HTML("&bull;"),
+                      # date
+                      span("April 2020"),
+                      HTML("&bull;"),
+                      # Shiny app code link
+                      span("Code"),
+                      a("on GitHub", href = "insert_link"),
+                    )
+                ),
+                
+                
+                # all content goes here, and is hidden initially until the page fully loads
+                sidebarLayout(
+                  sidebarPanel(
+                                    
+                    # Only show the following for assessment predictions:
+                    # Use PIC?
+                    checkboxInput("picInput", "Use PIC?", value = FALSE),
+                    selectInput("typeInput", "Estimate Type",
+                                c("Select", "Assessment Value", "Mill Rate"),
+                                selected = "Select"),
+                                     
+                    # If using PIC:
+                    conditionalPanel("input.picInput",
+                                     textInput("identInput", "PIC:", placeholder = NULL)),
+                                     
+                    # If PIC is not available:
+                    conditionalPanel("!input.picInput",
+                                     # selectInput("typeInput", "Estimate Type",
+                                     #             c("Select", "Assessment Value", "Mill Rate"),
+                                     #             selected = "Select"),
+                                     
+                                     # for municipality
+                                     selectInput("municipalityInput", "Municipality:", 
+                                                 c("-",sort(unique(dat$municipality))),
+                                                 selected = "-"),
+                                                     
+                                     # for Tax Class code
+                                     selectInput("taxclassInput", "Tax Class Code:",
+                                                 c("-",sort(unique(dat$tax.class))),
+                                                 selected = "-"),
+                                     
+                                     #conditional input for estimate type
+                                     conditionalPanel("input.typeInput == 'Assessment Value'",
+                                                      numericInput("assessmentInput",
+                                                                   "Current Assessment Value:",
+                                                                   value = 500000, min = 0, 
+                                                                   max = 6000000000))
+                                     ),
+                                    
+                    
[…]

### 2020-04-09 src/shiny_app/server.R
deleting to upload update
pushing from Rstudio not working.. this will do for now

@@ -1,249 +1,0 @@
-# Mallory - STAT 550 2020###
-# This is the server file of the shiny app
-
-# in case modified data needs to be accessed
-source("helpers.R")
-
-server <- function(input, output, session) {
-  
-  filtered <- reactive({
-    # SINCE PLOTS FOR MILL RATE AND ASSESSMENT VALUES DIFFER, WE NEED TO DISTINGUISH THIS HERE
-    # IF A PIC IS AVAILABLE (IN CASE OF MILL RATE - PLOT FOR MUNCIPALITY.  IN CASE OF ASSESSMENTS
-    # - PLOT THE PROPERTY(OR NOTHING))
-    
-    # Update when following inputs are changed
-    input$updateButton
-    
-    newdata <- dat
-    d <- NULL
-    print(dim(dat))
-    
-    # Filter data based on the user inputs
-    isolate({
-      #if (input$typeInput != 'Select'){
-        print("The type is not Select, it is:")
-        
-          print(input$typeInput)
-          
-          # If using PIC:
-          if(input$picInput && input$identInput!="")
-            {
-            print(paste("PIC input is", input$picInput, "and is:", sep = " "))
-            print(input$identInput)
-            
-            d <- newdata  %>% 
-              filter(PIC == input$identInput)
-          }
-          
-          
-          # If not using PIC, filter by municipality and tax class:
-          if(!input$picInput){
-            print(paste("PIC input is:", input$picInput, sep = " "))
-            d <- newdata %>% 
-              filter(tax.class == input$taxclassInput,
-                     municipality == input$municipalityInput)
-          }
-      
-      })
-    
-    # return filtered data
-    if(dim(d)[1]==0){
-      d <- NULL
-    }
-    d
-    print(dim(d))
-    print(head(d))
-    })
-  
-   ########## PLOTTING TAB ###################
-   # Add plots of either mill rate or assessment value over time to plot tab
-  munic <- reactive(input$municipalityInput)
-  
-  millrateplot <- reactive({
-    input$updateButton
-    
-    data <- filtered()
-    print("plotting mill rate plot")
-    print(head(data))
-    
-          
-    isolate({
-    if(is.null(data)){
-      p <- paste("No corresponding data.")
-    }
-
-    # plot mill rates over time for municipality chosen for mill rate predicitons
-    if (input$typeInput == 'Mill Rate'){
-      p <- ggplot(data, aes(x = year, y = mill.rate)) +
-        geom_line(color="#FF3333") +
-        geom_point(color="#FF3333") +
[…]

### 2020-04-09 src/shiny_app/ui.R
deleting to upload update..
pushing from rstudio not working - this will do for now

@@ -1,102 +1,0 @@
-# Mallory - STAT 550 2020###
-# This is the user interface version of the shiny app
-
-library(shinythemes)
-
-ui <- fluidPage(theme = shinytheme("paper"),
-                # header
-                div(id = "headerSection",
-                    h2("BC Mill Rate & Assessment Value Predictions"),
-                    
-                    span(
-                      style = "font-size: 1em",
-                      # authors
-                      span("Created by "),
-                      a("Gian Carlo Diluvi, Vittorio Romaniello, Sophia Li & Mallory Flynn"),
-                      HTML("&bull;"),
-                      # date
-                      span("April 2020"),
-                      HTML("&bull;"),
-                      # Shiny app code link
-                      span("Code"),
-                      a("on GitHub", href = "insert_link"),
-                    )
-                ),
-                
-                
-                # all content goes here, and is hidden initially until the page fully loads
-                sidebarLayout(
-                  sidebarPanel(
-                                    
-                    # Only show the following for assessment predictions:
-                    # Use PIC?
-                    checkboxInput("picInput", "Use PIC?", value = FALSE),
-                    selectInput("typeInput", "Estimate Type",
-                                c("Select", "Assessment Value", "Mill Rate"),
-                                selected = "Select"),
-                                     
-                    # If using PIC:
-                    conditionalPanel("input.picInput",
-                                     textInput("identInput", "PIC:", placeholder = NULL)),
-                                     
-                    # If PIC is not available:
-                    conditionalPanel("!input.picInput",
-                                     # selectInput("typeInput", "Estimate Type",
-                                     #             c("Select", "Assessment Value", "Mill Rate"),
-                                     #             selected = "Select"),
-                                     
-                                     # for municipality
-                                     selectInput("municipalityInput", "Municipality:", 
-                                                 c("-",sort(unique(dat$municipality))),
-                                                 selected = "-"),
-                                                     
-                                     # for Tax Class code
-                                     selectInput("taxclassInput", "Tax Class Code:",
-                                                 c("-",sort(unique(dat$tax.class))),
-                                                 selected = "-"),
-                                     
-                                     #conditional input for estimate type
-                                     conditionalPanel("input.typeInput == 'Assessment Value'",
-                                                      numericInput("assessmentInput",
-                                                                   "Current Assessment Value:",
-                                                                   value = 500000, min = 0, 
-                                                                   max = 6000000000))
-                                     ),
-                                    
-                    
[…]

### 2020-04-09 src/shiny_app/server.R
changed theme and commented debugging lines

@@ -1,0 +1,246 @@
+# Mallory - STAT 550 2020###
+# This is the server file of the shiny app
+
+# in case modified data needs to be accessed
+source("helpers.R")
+
+server <- function(input, output, session) {
+  
+  filtered <- reactive({
+    
+    # Update when following inputs are changed
+    input$updateButton
+    
+    newdata <- dat
+    d <- NULL
+    #print(dim(dat))
+    
+    # Filter data based on the user inputs
+    isolate({
+      #if (input$typeInput != 'Select'){
+        #print("The type is not Select, it is:")
+          #print(input$typeInput)
+          
+          # If using PIC:
+          if(input$picInput && input$identInput!="")
+            {
+            #print(paste("PIC input is", input$picInput, "and is:", sep = " "))
+            #print(input$identInput)
+            
+            d <- newdata  %>% 
+              filter(PIC == input$identInput)
+          }
+          
+          
+          # If not using PIC, filter by municipality and tax class:
+          if(!input$picInput){
+            #print(paste("PIC input is:", input$picInput, sep = " "))
+            d <- newdata %>% 
+              filter(tax.class == input$taxclassInput,
+                     municipality == input$municipalityInput)
+          }
+      
+      })
+    
+    # return filtered data
+    if(dim(d)[1]==0){
+      d <- NULL
+    }
+    d
+    #print(dim(d))
+    #print(head(d))
+    })
+  
+   ########## PLOTTING TAB ###################
+   # Add plots of either mill rate or assessment value over time to plot tab
+  munic <- reactive(input$municipalityInput)
+  
+  millrateplot <- reactive({
+    input$updateButton
+    
+    data <- filtered()
+    #print("plotting mill rate plot")
+    #print(head(data))
+    
+          
+    isolate({
+    if(is.null(data)){
+      p <- paste("No corresponding data.")
+    }
+
+    # plot mill rates over time for municipality chosen for mill rate predicitons
+    if (input$typeInput == 'Mill Rate'){
+      p <- ggplot(data, aes(x = year, y = mill.rate)) +
+        geom_line(color="#FF3333") +
+        geom_point(color="#FF3333") +
+        theme_minimal() +
+        xlab("Year") +
+        ylab("Mill Rate") +
+        ggtitle("Municipal Mill Rate Over Time")
+    }
+    })
+
+    return(p)
+
+  })
+
+  assessplot <- reactive({
+    input$updateButton
+    
+    data <- filtered()
+    #print("plotting assessplot")
[…]

### 2020-04-09 src/shiny_app/ui.R

@@ -1,0 +1,104 @@
+# Mallory - STAT 550 2020###
+# This is the user interface version of the shiny app
+
+library(shinythemes)
+
+ui <- fluidPage(theme = shinytheme("journal"), #maybe lumen theme?
+
+                # header
+                div(id = "headerSection",
+                    h2("BC Mill Rate & Assessment Value Predictions"),
+                    
+                    span(
+                      style = "font-size: 1em",
+                      # authors
+                      span("Created by "),
+                      a("Gian Carlo Diluvi, Vittorio Romaniello, Sophia Li & Mallory Flynn"),
+                      HTML("&bull;"),
+                      # date
+                      span("April 2020"),
+                      HTML("&bull;"),
+                      # Shiny app code link
+                      span("Code"),
+                      a("on GitHub", href = "insert_link"),
+                    )
+                ),
+                br(),
+                br(),
+                
+                # all content goes here, and is hidden initially until the page fully loads
+                sidebarLayout(
+                  sidebarPanel(
+                                    
+                    # Only show the following for assessment predictions:
+                    # Use PIC?
+                    checkboxInput("picInput", "Use PIC?", value = FALSE),
+                    selectInput("typeInput", "Estimate Type",
+                                c("Select", "Assessment Value", "Mill Rate"),
+                                selected = "Select"),
+                                     
+                    # If using PIC:
+                    conditionalPanel("input.picInput",
+                                     textInput("identInput", "PIC:", placeholder = NULL)),
+                                     
+                    # If PIC is not available:
+                    conditionalPanel("!input.picInput",
+                                     # selectInput("typeInput", "Estimate Type",
+                                     #             c("Select", "Assessment Value", "Mill Rate"),
+                                     #             selected = "Select"),
+                                     
+                                     # for municipality
+                                     selectInput("municipalityInput", "Municipality:", 
+                                                 c("-",sort(unique(dat$municipality))),
+                                                 selected = "-"),
+                                                     
+                                     # for Tax Class code
+                                     selectInput("taxclassInput", "Tax Class Code:",
+                                                 c("-",sort(unique(dat$tax.class))),
+                                                 selected = "-"),
+                                     
+                                     #conditional input for estimate type
+                                     conditionalPanel("input.typeInput == 'Assessment Value'",
+                                                      numericInput("assessmentInput",
+                                                                   "Current Assessment Value:",
+                                                                   value = 500000, min = 0, 
+                                                                   max = 6000000000))
+                                     ),
[…]

### 2020-04-09 src/shiny_app/ui.R
added UBC hyperlink and trying to fix altus group img issue

@@ -13,7 +13,8 @@ ui <- fluidPage(theme = shinytheme("journal"), #maybe lumen theme?
                       style = "font-size: 1em",
                       # authors
                       span("Created by "),
-                      a("Gian Carlo Diluvi, Vittorio Romaniello, Sophia Li & Mallory Flynn"),
+                      a("Gian Carlo Diluvi, Vittorio Romaniello, Sophia Li & Mallory Flynn",
+                        href = "https://www.stat.ubc.ca"),
                       HTML("&bull;"),
                       # date
                       span("April 2020"),
@@ -72,14 +73,14 @@ ui <- fluidPage(theme = shinytheme("journal"), #maybe lumen theme?
                     actionButton("updateButton", "Update"),
                       
                     
-                    # source of data as a footer
+                    # source of data as a footer - Altus Group image not loading
                     br(),
                     br(),
                     p("Generated using data from ",
                       a("the Altus Group Ltd.",
                         href = "https://www.altusgroup.com",
                         target = "_blank")),
-                    a(img(src = "AltusGroup.svg", alt = "Altus Group"),
+                    a(img(src = "altusgroupimg.png", alt = "Altus Group"),
                       href = "https://www.altusgroup.com",
                       target = "_blank"),
                     bookmarkButton()

### 2020-04-09 src/rf.Rmd

@@
 ---
-title: "random forest"
+title: "random forest and lme"
 output: github_document
@@
 suppressPackageStartupMessages(library(tidyverse))
+suppressPackageStartupMessages(library(nlme))
+suppressPackageStartupMessages(library(ROCR))
 ```
@@
 
+# Random Forest
 ```{r}
+set.seed(0)
+
 rf.mill <- randomForest(
-  mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit,
+  mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit, mtry = 4,
+  data=train, ntree=500
+)
+#Evaluate variable importance
+importance(rf.mill)
+varImpPlot(rf.mill)
+
+rf.as <- randomForest(
+  next.assess ~  tax.class + municipality + total.assessment + mill.rate,na.action = na.omit,mtry = 4,
   data=train
 )
+
+importance(rf.as)
+varImpPlot(rf.as)
+
+yhat.bag <- predict(rf.mill,newdata=test)
+plot(yhat.bag, test$mill.rate, xlab="Predicted Mill Rate Using Test Set", ylab="Actual Mill Rate")
+abline(0,1)
+
+yhat.bag1 <- predict(rf.as,newdata=test)
+plot(yhat.bag1, test$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value")
+abline(0,1)
 ```
 
+# Linear Mixed Effect Model
 ```{r}
-predict(rf.mill, test)
+# both random slope and random intercept
+# different rate of change of assessment value and mill rate as well as initial assessment value and mill rate for each municipality 
+lme.mill <- lme( mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit, data=train, random = ~ 1 + total.assessment + past.mill | municipality, method = "ML")
+
+summary(lme.mill)
+
+lme.as <- lme(next.assess ~  tax.class + municipality + total.assessment + mill.rate ,na.action = na.omit, data=train, random = ~ 1 + total.assessment + mill.rate | municipality, method = "ML")
+
+summary(lme.as)
 ```

### 2020-04-09 src/rf.Rmd

@@
 suppressPackageStartupMessages(library(tidyverse))
-suppressPackageStartupMessages(library(nlme))
+suppressPackageStartupMessages(library(lme4))
 suppressPackageStartupMessages(library(ROCR))
+suppressPackageStartupMessages(library(predictmeans))
 ```
@@
 # different rate of change of assessment value and mill rate as well as initial assessment value and mill rate for each municipality
-lme.mill <- lme( mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit, data=train, random = ~ 1 + total.assessment + past.mill | municipality, method = "ML")
+
+lme.mill <- lmer(mill.rate ~ 1+ (1+year|municipality) + as.factor(year) + tax.class  + total.assessment + past.mill, data = train)
 
@@
 
-lme.as <- lme(next.assess ~  tax.class + municipality + total.assessment + mill.rate ,na.action = na.omit, data=train, random = ~ 1 + total.assessment + mill.rate | municipality, method = "ML")
+lme.as <- lmer(next.assess ~ 1+ (1+year|municipality) + as.factor(year) + tax.class + total.assessment + mill.rate, data = train)
 
@@
 ```
+
+```{r}
+# check assumptions of lme
+
+# Homogeneity of Variance
+residplot(lme.mill)
+residplot(lme.as)
+```

### 2020-04-10 src/shiny_app/server.R
updated style of shiny app and fixed broken img

@@ -15,9 +15,9 @@ server <- function(input, output, session) {
     d <- NULL
     #print(dim(dat))
     
+    #if(input$csvfile == NULL){
     # Filter data based on the user inputs
     isolate({
-      #if (input$typeInput != 'Select'){
         #print("The type is not Select, it is:")
           #print(input$typeInput)
           

### 2020-04-10 src/shiny_app/ui.R

@@ -2,8 +2,9 @@
 # This is the user interface version of the shiny app
 
 library(shinythemes)
+library(png)
 
-ui <- fluidPage(theme = shinytheme("journal"), #maybe lumen theme?
+ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?
 
                 # header
                 div(id = "headerSection",
@@ -30,7 +31,8 @@ ui <- fluidPage(theme = shinytheme("journal"), #maybe lumen theme?
                 # all content goes here, and is hidden initially until the page fully loads
                 sidebarLayout(
                   sidebarPanel(
-                                    
+                    # tabsetPanel(
+                    #   tabPanel("User Inputs",                
                     # Only show the following for assessment predictions:
                     # Use PIC?
                     checkboxInput("picInput", "Use PIC?", value = FALSE),
@@ -71,6 +73,24 @@ ui <- fluidPage(theme = shinytheme("journal"), #maybe lumen theme?
                     # button to update the data
                     shiny::hr(),
                     actionButton("updateButton", "Update"),
+                    # ),
+                    # 
+                    # tabPanel("Upload data",
+                    #          br(),
+                    #          # Input: Select a file ----
+                    #          fileInput("csvfile", "Upload Data from .csv file",
+                    #                    multiple = TRUE,
+                    #                    accept = c("text/csv",
+                    #                               "text/comma-separated-values,text/plain",
+                    #                               ".csv")),
+                    #          
+                    #          # Horizontal line ----
+                    #          tags$hr(),
+                    #          
+                    # )
+                    # ),
+                    
+                    
                       
                     
                     # source of data as a footer - Altus Group image not loading
@@ -80,7 +100,8 @@ ui <- fluidPage(theme = shinytheme("journal"), #maybe lumen theme?
                       a("the Altus Group Ltd.",
                         href = "https://www.altusgroup.com",
                         target = "_blank")),
-                    a(img(src = "altusgroupimg.png", alt = "Altus Group"),
+                    a(img(src = "altusgroupimg.png", alt = "Altus Group",
+                          height = 63, width = 150),
                       href = "https://www.altusgroup.com",
                       target = "_blank"),
                     br(),

### 2020-04-10 src/shiny_app/global.R
fixed title bug with assessment predictions

@@ -3,11 +3,10 @@ library(readr)
 library(ggplot2)
 library(dplyr)
 library(DT)
-library(shinyjs)
 library(tidyr)
 library(shinyjs)
 
-#PIC using for testing
+#PIC used for some testing
 #CA-BC-022-04490000
 
 

### 2020-04-10 src/shiny_app/server.R

@@ -176,8 +176,7 @@ server <- function(input, output, session) {
        }
         
      if(input$typeInput == 'Assessment Value'){
-       return(paste("prediction for", input$identInput,
-                    "is...", sep = " "))  # RETURN PREDICTION
+       return(paste("Predicted next assessment value is...", sep = " "))  # RETURN PREDICTION
        }
         
      if(input$typeInput == 'Select'){

### 2020-04-10 src/shiny_app/server.R
cleaned up github repo

@@ -1,6 +1,11 @@
 # Mallory - STAT 550 2020###
 # This is the server file of the shiny app
 
+# fix mill rate title when PIC is used
+# fix main title when PIC is checked but empty 
+
+
+
 # in case modified data needs to be accessed
 source("helpers.R")
 

### 2020-04-10 src/shiny_app/ui.R

@@ -22,7 +22,8 @@ ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?
                       HTML("&bull;"),
                       # Shiny app code link
                       span("Code"),
-                      a("on GitHub", href = "insert_link"),
+                      a("on GitHub", 
+                        href = "https://github.com/STAT450-550/RealEstate/tree/master/src/shiny_app"),
                     )
                 ),
                 br(),
@@ -77,14 +78,13 @@ ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?
                     # 
                     # tabPanel("Upload data",
                     #          br(),
-                    #          # Input: Select a file ----
+                    #          # Input: Select a file 
                     #          fileInput("csvfile", "Upload Data from .csv file",
                     #                    multiple = TRUE,
                     #                    accept = c("text/csv",
                     #                               "text/comma-separated-values,text/plain",
                     #                               ".csv")),
                     #          
-                    #          # Horizontal line ----
                     #          tags$hr(),
                     #          
                     # )

### 2020-04-10 src/modeling - 550/a_rough_clean.R
no longer needed

@@ -1,99 +1,0 @@
-# Mallory ###
-# some cleaning.. with some simplifications we probably don't want
-library(tidyverse)
-library(readr)
-library(readxl)
-library(ggplot2)
-library(viridis)
-library(here)
-library(rnn)
-rm(list=ls())
-
-### DATA CLEANING 
-# import assessment and sales data for all years
-assess_all <- read.csv(here("data/Assessment/2016 - 2020 Raw.csv"),
-                       header = F)
-col_names <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
-colnames(assess_all) <- colnames(col_names)
-
-# select relevant features 
-assess <- assess_all %>%
-  select(PIC, Year, AssessorAreaCode:AddressAssessorMunicipalityDesc, 
-         TaxClassCode:TaxClassTaxRate)  %>% 
-  filter(TaxClassCode %in% c("01","05","06"))
-
-# change types
-assess <- assess %>% 
-  mutate_at(c(6, 10, 11), as.character) %>% 
-  mutate_at(c(6, 10, 11), as.numeric)
-
-# check for NA's (thanks to STAT450 for this code)
-for (feature in colnames(assess)){
-  variable = eval(parse(text = paste0("assess$",feature)))
-  null.count = sum(is.na(variable))
-  print(paste(feature, null.count, sep = ":"))
-} 
-
-# check for constant mill rate across municipalities
-test <- assess %>% 
-  group_by(Year, AddressAssessorMunicipalityDesc, TaxClassCode) %>% 
-  filter(!is.na(TaxClassTaxRate)) %>% 
-  summarise(test = var(TaxClassTaxRate)) %>%
-  filter(test !=0)
-
-### for now just took mean of mill rates..
-assess.rate <- assess %>%
-  group_by(Year, AddressAssessorMunicipalityDesc, TaxClassCode) %>% 
-  filter(!is.na(TaxClassTaxRate)) %>% 
-  mutate(TaxClassTaxRate = mean(TaxClassTaxRate))
-
-# check that this reduced variance in mill rates to 0 for all municipalities
-test1 <- assess.rate %>% 
-  group_by(Year, AddressAssessorMunicipalityDesc, TaxClassCode) %>%
-  summarise(test = var(TaxClassTaxRate)) %>%
-  filter(test !=0)
-
-# create dataset listing mill rates per municipality, year, and tax class
-assess.agg <- assess.rate%>%
-  group_by(Year, AddressAssessorMunicipalityDesc,TaxClassCode)%>%
[…]

### 2020-04-10 src/modeling - 550/rf_mill.Rmd
modified mill model rf

---
title: "Random forest mill rates"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(randomForest))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(ROCR))
suppressPackageStartupMessages(library(predictmeans))
```

```{r}
dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
dat$municipality <- as.factor(dat$municipality)

factors_tbl = dat %>% 
  group_by(municipality) %>% 
  count(name="mun_count", sort = TRUE) %>% 
  ungroup() %>% 
  mutate(perc = mun_count/sum(mun_count),
         cum_perc = cumsum(perc)) %>% 
  arrange(desc(mun_count)) %>% 
  mutate(rank = row_number(),
         municipality = fct_reorder(municipality, rank)) %>% 
  mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
  select(municipality, col_municipality)

avg_assessment_by_municipality = dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  filter(test.train == "train") %>% 
  group_by(municipality, tax.class, year) %>% 
  mutate(avg_assessment = mean(total.assessment, na.rm=TRUE)) %>% 
  select(municipality, year, tax.class, avg_assessment) %>% 
  distinct(municipality, .keep_all = T)

dat_1 <- dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  left_join(avg_assessment_by_municipality) %>% 
  group_by(PIC) %>% 
  mutate(next.assess = lead(avg_assessment, order_by = year),
         past.mill = lag(mill.rate, order_by = year)) %>%
  arrange(PIC) %>% 
  group_by(municipality, year) %>% 
  mutate(n.prop = n()) %>% 
  arrange(desc(n.prop)) %>% 
  distinct(municipality, .keep_all = T) %>% 
  select(-c(tax, improvement.assessment, land.assessment, total.assessment))
# 
# notwant <- c("CA-BC-003-6500000", "CA-BC-223-00100030000", "CA-BC-227-0000020000", "CA-BC-234-01001006", "CA-BC-301-0000001", "CA-BC-315-00001000","CA-BC-316-010012122009", "CA-BC-336-00001000", "CA-BC-338-1000004884000", "CA-BC-556-00001000", "CA-BC-727-03025100", "CA-BC-757-05141000", "CA-BC-759-000147000", "CA-BC-760-001203000", "CA-BC-765-00002001", "CA-BC-771-00003000")
train <- dat_1 %>%  filter(test.train == "train")
test <- dat_1 %>%  filter(test.train == "test")
```

# Random Forest
```{r}
set.seed(0)

rf.mill <- randomForest(
  mill.rate ~  tax.class + municipality + avg_assessment + past.mill, na.action = na.omit, mtry = 4,
  data=train, ntree=500
)

save(rf.mill, file = "../src/shiny_app/rf.mill.rda")
#Evaluate variable importance
importance(rf.mill)
varImpPlot(rf.mill)

# rf.as <- randomForest(
#   next.assess ~ tax.class + municipality + total.assessment + mill.rate,na.action = na.omit,mtry = 4,
#   data=train
# )
# 
# importance(rf.as)
# varImpPlot(rf.as)
# 
# yhat.bag <- predict(rf.mill,newdata=test)
# plot(yhat.bag, test$mill.rate, xlab="Predicted Mill Rate Using Test Set", ylab="Actual Mill Rate")
# abline(0,1)
# 
# yhat.bag1 <- predict(rf.as,newdata=test)
# plot(yhat.bag1, test$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value")
# abline(0,1)
```

### 2020-04-10 src/modeling - 550/rf.Rmd

--- 2020-04-09 src/rf.Rmd
@@
 dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
+train <- readr::read_delim("train_data.txt", delim = ",", col_types = "ciccddddd")
+test <- readr::read_delim("test_data.txt", delim = ",", col_types = "ciccddddd")
 dat$municipality <- as.factor(dat$municipality)
@@
 
-train <- dat %>%  filter(test.train == "train") 
+notwant <- c("CA-BC-003-6500000", "CA-BC-223-00100030000", "CA-BC-227-0000020000", "CA-BC-234-01001006", "CA-BC-301-0000001", "CA-BC-315-00001000","CA-BC-316-010012122009", "CA-BC-336-00001000", "CA-BC-338-1000004884000", "CA-BC-556-00001000", "CA-BC-727-03025100", "CA-BC-757-05141000", "CA-BC-759-000147000", "CA-BC-760-001203000", "CA-BC-765-00002001", "CA-BC-771-00003000")
+train <- dat %>%  filter(test.train == "train" & !PIC %in% notwant)
 test <- dat %>%  filter(test.train == "test")
@@
 )
+
+save(rf.mill, file = "~/Desktop/git_docs/RealEstate/src/shiny_app/rf.mill.rda")
 #Evaluate variable importance
@@
 ```
+
+# Linear Regression
+```{r}
+lm.as <- lm(
+  next.assess ~ PIC + tax.class + municipality + total.assessment + mill.rate, na.action = na.omit, data=train)
+
+predict(lm.as, test)
+```

### 2020-04-10 src/modeling - 550/rf_mill.Rmd
modified rf file

@@
 
-dat_1 <- dat %>% 
+dat_mill <- dat %>% 
   left_join(factors_tbl) %>% 
@@
   select(-c(tax, improvement.assessment, land.assessment, total.assessment))
+
+dat_as <- dat %>% 
+  left_join(factors_tbl) %>% 
+  select(-c(municipality)) %>% 
+  rename(municipality = col_municipality) %>% 
+  group_by(PIC) %>% 
+  arrange(year) %>% 
+  mutate(next.assess = lead(total.assessment, order_by = year),
+         past.mill = lag(mill.rate, order_by = year))
+
 # 
 # notwant <- c("CA-BC-003-6500000", "CA-BC-223-00100030000", "CA-BC-227-0000020000", "CA-BC-234-01001006", "CA-BC-301-0000001", "CA-BC-315-00001000","CA-BC-316-010012122009", "CA-BC-336-00001000", "CA-BC-338-1000004884000", "CA-BC-556-00001000", "CA-BC-727-03025100", "CA-BC-757-05141000", "CA-BC-759-000147000", "CA-BC-760-001203000", "CA-BC-765-00002001", "CA-BC-771-00003000")
-train <- dat_1 %>%  filter(test.train == "train")
-test <- dat_1 %>%  filter(test.train == "test")
+
+train_mill <- dat_mill %>%  filter(test.train == "train")
+test_mill <- dat_mill %>%  filter(test.train == "test")
+train_as <- dat_as %>%  filter(test.train == "train")
+test_as <- dat_as %>%  filter(test.train == "test")
 ```
@@
   mill.rate ~  tax.class + municipality + avg_assessment + past.mill, na.action = na.omit, mtry = 4,
-  data=train, ntree=500
+  data=train_mill, ntree=500
 )
 
-save(rf.mill, file = "../src/shiny_app/rf.mill.rda")
+save(rf.mill, file = "rf.mill.rda")
+
 #Evaluate variable importance
@@
 
-# rf.as <- randomForest(
-#   next.assess ~ tax.class + municipality + total.assessment + mill.rate,na.action = na.omit,mtry = 4,
-#   data=train
-# )
-# 
-# importance(rf.as)
-# varImpPlot(rf.as)
-# 
-# yhat.bag <- predict(rf.mill,newdata=test)
-# plot(yhat.bag, test$mill.rate, xlab="Predicted Mill Rate Using Test Set", ylab="Actual Mill Rate")
-# abline(0,1)
-# 
-# yhat.bag1 <- predict(rf.as,newdata=test)
-# plot(yhat.bag1, test$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value")
-# abline(0,1)
+rf.as <- randomForest(
+  next.assess ~ tax.class + municipality + total.assessment + mill.rate, na.action = na.omit,mtry = 4,
+  data=train_as
+)
+
+save(rf.as, file = "rf.as.rda")
+importance(rf.as)
+varImpPlot(rf.as)
+
+yhat.bag <- predict(rf.mill,newdata=test_mill)
+plot(yhat.bag, test_mill$mill.rate, xlab="Predicted Mill Rate Using Test Set", ylab="Actual Mill Rate")
+abline(0,1)
+
+yhat.bag1 <- predict(rf.as,newdata=test_as)
+plot(yhat.bag1, test_as$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value")
+abline(0,1)
 ```

### 2020-04-11 src/modeling - 550/rf_mill.Rmd
deleted comment

@@
   mutate(next.assess = lead(total.assessment, order_by = year),
-         past.mill = lag(mill.rate, order_by = year))
-
-# 
-# notwant <- c("CA-BC-003-6500000", "CA-BC-223-00100030000", "CA-BC-227-0000020000", "CA-BC-234-01001006", "CA-BC-301-0000001", "CA-BC-315-00001000","CA-BC-316-010012122009", "CA-BC-336-00001000", "CA-BC-338-1000004884000", "CA-BC-556-00001000", "CA-BC-727-03025100", "CA-BC-757-05141000", "CA-BC-759-000147000", "CA-BC-760-001203000", "CA-BC-765-00002001", "CA-BC-771-00003000")
+         past.mill = lag(mill.rate, order_by = year)) %>% 
+  na.omit()
 

### 2020-04-11 src/modeling - 550/rf_mill.Rmd
filtered top 5 prop for each muni for rf.as

@@
          past.mill = lag(mill.rate, order_by = year)) %>% 
-  na.omit()
+  group_by(municipality) %>% 
+  top_n(5, wt = total.assessment) %>% 
+  arrange(municipality)
 

### 2020-04-11 src/modeling - 550/rf_mill.Rmd
top 25

--- 2020-04-11 src/modeling - 550/rf_mill.Rmd
@@
          past.mill = lag(mill.rate, order_by = year)) %>% 
-  na.omit()
+  group_by(municipality) %>% 
+  top_n(25, wt = total.assessment) %>% 
+  arrange(municipality)
 

### 2020-04-11 src/modelling - 550/rf.Rmd
shiny app update

--- 2020-04-10 src/modeling - 550/rf.Rmd
@@
 dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
-train <- readr::read_delim("train_data.txt", delim = ",", col_types = "ciccddddd")
-test <- readr::read_delim("test_data.txt", delim = ",", col_types = "ciccddddd")
 dat$municipality <- as.factor(dat$municipality)

### 2020-04-11 src/shiny_app/helpers.R

@@ -7,7 +7,7 @@ rfData <- function(data) {
   
   rfdat$municipality <- as.factor(rfdat$municipality)
   
-  factors_tbl = rfdat %>% 
+  factors_tbl = dat %>% 
     group_by(municipality) %>% 
     count(name="mun_count", sort = TRUE) %>% 
     ungroup() %>% 

### 2020-04-11 src/shiny_app/server.R

@@ -1,7 +1,6 @@
 # Mallory - STAT 550 2020###
 # This is the server file of the shiny app
 
-# fix mill rate title when PIC is used
 # fix main title when PIC is checked but empty 
 
 
@@ -9,6 +8,11 @@
 # in case modified data needs to be accessed
 source("helpers.R")
 
+# load models - RF for mill rate predictions and ??? for assessment value predictions
+load("RegularizedLogisticRegression.rda")
+
+
+# server:
 server <- function(input, output, session) {
   
   filtered <- reactive({
@@ -146,6 +150,80 @@ server <- function(input, output, session) {
   estimates <- reactive({
     input$updateButton
     
+    data <- filtered()
+    
+    isolate({
+      if(is.null(data)){
+        return(NULL)
+      }
+      
+     # If using PIC: 
+      if(input$picInput){
+        if(input$identInput != ""){
+          
+          # If doing Mill Rate prediciton, use random forest
+          if(input$typeInput == 'Mill Rate'){
+            
+            # extract latest mill rate
+            past20 <- filtered() %>% 
+              filter(year == 2020)
+            
+            past19 <- filtered() %>% 
+              filter(year == 2019)
+            
+            # if 2020 column is nonempty of mill rates for this community:
+            if(sum(!is.na(past20$mill.rate>0))){
+              meanmillrate <- mean(past20$mill.rate)
+            }
+            
+            # otherwise use 2019 mill rate data
+            else{
+              meanmillrate <- mean(past19$mill.rate)
+            }
+            
+            # put data together in the way rfmill expects as input call it inputdata
+            # columns include tax.class, municipality, total.assessment, past.mill
+            meanassess <- mean(past20$total.assessment)
+            
+            pred.data <- cbind(filtered()$tax.class[1], 
+                               filtered()$municipality[1],
+                               meanassess,
+                               meanmillrate)
+            pred.data <- as.data.frame(pred.data)
+            pred.data$municipality <- as.factor(pred.data$municipality)
+                               
+            colnames(pred.data) <- c('tax.class', 'municipality', 'total.assessment',
+                                  'past.mill')
+                               
+            # predict next mill rate using random forest
+            predict(rf.mill, newdata = pred.data)
+          }
+          
+          # If doing Assessment Value prediction, use something else...
+          if(input$typeInput == 'Assessment Value'){
+            #...
+          }
[…]

### 2020-04-11 src/shiny_app/server.R
shiny app update and renaming folder

@@ -2,7 +2,8 @@
 # This is the server file of the shiny app
 
 # fix main title when PIC is checked but empty 
-
+# load rdas for each so that estimates can be made
+# fix select input to choose only the top 52 categories and other
 
 
 # in case modified data needs to be accessed

### 2020-04-11 src/eda&modelling - 450/EDA&mode_fitting.Rmd

--- 2020-03-01 src/EDA&mode_fitting.Rmd
@@
 assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and assess - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+cor(assessment_aggregate$assessTotal, assessment_aggregate$rate)
 
-tax_class <- assessment_aggregate %>% ggplot(aes(x=predTaxIncome,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+cor(assessment_aggregate$tax, assessment_aggregate$rate)
 
-assess_class
+land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and landTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+cor(assessment_aggregate$landTotal, assessment_aggregate$rate)
 
-tax_class
+improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and improvementTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+cor(assessment_aggregate$improvementTotal, assessment_aggregate$rate)
+
+propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and propertyCount - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
+cor(assessment_aggregate$propertyCount, assessment_aggregate$rate)
+
+library(plotly)
+ggplotly(assess_class)
+ggplotly(tax_class)
+ggplotly(land_class)
+ggplotly(improvementTotal_class)
+ggplotly(propertyCount_class)
 
 ## Correlation for categorical variables
-TaxClassCode <- aov(rate ~ TaxClassCode, data = assessment_aggregate)
+TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
 summary(TaxClassCode)
 
-Year <- aov(rate_pct ~ Year, data = pct_final)
+Year <- aov(rate_pct ~ factor(Year), data = pct_final)
 summary(Year)
 
-Municipalities <- aov(rate ~ AddressAssessorMunicipalityDesc, data = assessment_aggregate)
+Municipalities <- aov(rate ~ factor(AddressAssessorMunicipalityDesc), data = assessment_aggregate)
 summary(Municipalities)
@@
 
+Summary on GOF
+```{r}
+
+GOF_table <- rbind(
+      c("OLR full", 0.8874, 0.8707, 1.9843, 2.5902),
+      c("OLR reduced", 0.8874, 0.8721, 1.9845, 2.5237),
+      c("Ridge", 0.8868, 0.8707, 1.9896, 2.5675),
+      c("LASSO", 0.8873, 0.8707, 1.9855, 2.5280),
+      c("Elastic Net", 0.8625, 0.8707, 2.2366, 2.5486))
+colnames(GOF_table) <- c("Model", "Mutiple R-Squared","Adjusted R_Squared", "MSE", "PMSE")
+
+```
+
 ## neural network

### 2020-04-11 src/shiny_app/global.R
updated shiny app with predict error

@@ -14,9 +14,13 @@ library(shinyjs)
 enableBookmarking("url")
 
 # read data
-# dat <- readr::read_delim("./../../data/test_train_data.txt", 
-#                          delim = ",", col_types = "ciccdddddf")
-
 dat <- readr::read_delim("test_train_data.txt", 
                          delim = ",", col_types = "ciccdddddf")
 
+# creates dataset with only the top 52 municipalities
+counts <- dat %>% 
+  count(municipality, sort = TRUE)
+
+datshort <- dat %>% 
+  filter(municipality %in% counts$municipality[1:52])
+  

### 2020-04-11 src/shiny_app/server.R

@@ -10,7 +10,8 @@
 source("helpers.R")
 
 # load models - RF for mill rate predictions and ??? for assessment value predictions
-load("RegularizedLogisticRegression.rda")
+load("rf.mill.rda")
+load("rf.as.rda")
 
 
 # server:
@@ -25,17 +26,13 @@ server <- function(input, output, session) {
     d <- NULL
     #print(dim(dat))
     
-    #if(input$csvfile == NULL){
     # Filter data based on the user inputs
     isolate({
-        #print("The type is not Select, it is:")
-          #print(input$typeInput)
           
           # If using PIC:
           if(input$picInput && input$identInput!="")
             {
-            #print(paste("PIC input is", input$picInput, "and is:", sep = " "))
-            #print(input$identInput)
+          
             
             d <- newdata  %>% 
               filter(PIC == input$identInput)
@@ -44,7 +41,7 @@ server <- function(input, output, session) {
           
           # If not using PIC, filter by municipality and tax class:
           if(!input$picInput){
-            #print(paste("PIC input is:", input$picInput, sep = " "))
+            
             d <- newdata %>% 
               filter(tax.class == input$taxclassInput,
                      municipality == input$municipalityInput)
@@ -57,20 +54,18 @@ server <- function(input, output, session) {
       d <- NULL
     }
     d
-    #print(dim(d))
-    #print(head(d))
+    
     })
   
    ########## PLOTTING TAB ###################
    # Add plots of either mill rate or assessment value over time to plot tab
   munic <- reactive(input$municipalityInput)
   
+  # create mill rate plot that reacts to inputs
   millrateplot <- reactive({
     input$updateButton
     
     data <- filtered()
-    #print("plotting mill rate plot")
-    #print(head(data))
     
           
     isolate({
@@ -94,12 +89,11 @@ server <- function(input, output, session) {
 
   })
 
+  # create assessment plots that react to user inputs
   assessplot <- reactive({
     input$updateButton
     
     data <- filtered()
-    #print("plotting assessplot")
-    #print(head(data))
 
     isolate({
       if(is.null(data)){
@@ -107,7 +101,6 @@ server <- function(input, output, session) {
         }
       
     # plot assessment values over time 
-      #print("trying to plot")
       if(input$typeInput == 'Assessment Value'){
           #print("ggplotting assessment values")
           p <- ggplot(data, aes(x = year, y = total.assessment)) +
[…]

### 2020-04-11 src/shiny_app/ui.R

@@ -47,13 +47,11 @@ ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?
                                      
                     # If PIC is not available:
                     conditionalPanel("!input.picInput",
-                                     # selectInput("typeInput", "Estimate Type",
-                                     #             c("Select", "Assessment Value", "Mill Rate"),
-                                     #             selected = "Select"),
                                      
                                      # for municipality
                                      selectInput("municipalityInput", "Municipality:", 
-                                                 c("-",sort(unique(dat$municipality))),
+                                                 c("-",sort(unique(datshort$municipality)),
+                                                   "Other"),
                                                  selected = "-"),
                                                      
                                      # for Tax Class code
@@ -74,22 +72,6 @@ ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?
                     # button to update the data
                     shiny::hr(),
                     actionButton("updateButton", "Update"),
-                    # ),
-                    # 
-                    # tabPanel("Upload data",
-                    #          br(),
-                    #          # Input: Select a file 
-                    #          fileInput("csvfile", "Upload Data from .csv file",
-                    #                    multiple = TRUE,
-                    #                    accept = c("text/csv",
-                    #                               "text/comma-separated-values,text/plain",
-                    #                               ".csv")),
-                    #          
-                    #          tags$hr(),
-                    #          
-                    # )
-                    # ),
-                    
                     
                       
                     

### 2020-04-11 src/modelling - 550/rf_mill.Rmd
changes tax class to factor

--- 2020-04-11 src/modeling - 550/rf_mill.Rmd
@@
 dat$municipality <- as.factor(dat$municipality)
+dat$tax.class <- as.factor(dat$tax.class)
 
@@
 yhat.bag1 <- predict(rf.as,newdata=test_as)
-plot(yhat.bag1, test_as$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value")
+plot(yhat.bag1, test_as$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value", ylim=c(0, 6e8))
 abline(0,1)

### 2020-04-12 src/shiny_app/global.R
prediction functionality and lots of bug fixes to shiny

@@ -7,7 +7,7 @@ library(tidyr)
 library(shinyjs)
 
 #PIC used for some testing
-#CA-BC-022-04490000
+# CA-BC-200-001019632060000
 
 
 # for bookmarking button
@@ -23,4 +23,11 @@ counts <- dat %>%
 
 datshort <- dat %>% 
   filter(municipality %in% counts$municipality[1:52])
+
+
+# load dataset used for rf.mill
+rfdatload <- readRDS("rf_data.rds")
+
+# create dataset used for rf.as
+asdatload <- readRDS("as_data.rds")
   

### 2020-04-12 src/shiny_app/helpers.R

@@ -1,11 +1,14 @@
 # This file will have helpers for the model and the loading of data for models
 
 # function to modify data for random forest (if needed)
+
+# used to create mill rate data for random forest
+# no longer needed for shiny app; included for future use if needed
 rfData <- function(data) {
   
-  rfdat <- data
-  
-  rfdat$municipality <- as.factor(rfdat$municipality)
+  dat <- data
+  dat$municipality <- as.factor(dat$municipality)
+  dat$tax.class <- as.factor(dat$tax.class)
   
   factors_tbl = dat %>% 
     group_by(municipality) %>% 
@@ -20,21 +23,78 @@ rfData <- function(data) {
     select(municipality, col_municipality)
   
   
-  
-  rfdat <- rfdat %>% 
+  avg_assessment_by_municipality = dat %>% 
     left_join(factors_tbl) %>% 
     select(-c(municipality)) %>% 
     rename(municipality = col_municipality) %>% 
+    filter(test.train == "train") %>% 
+    group_by(municipality, tax.class, year) %>% 
+    mutate(avg_assessment = mean(total.assessment, na.rm=TRUE)) %>% 
+    select(municipality, year, tax.class, avg_assessment) %>% 
+    distinct(municipality, .keep_all = T)
+  
+  dat_mill <- dat %>% 
+    left_join(factors_tbl) %>% 
+    select(-c(municipality)) %>% 
+    rename(municipality = col_municipality) %>% 
+    left_join(avg_assessment_by_municipality) %>% 
     group_by(PIC) %>% 
-    mutate(next.assess = lead(total.assessment, order_by = year),
+    mutate(next.assess = lead(avg_assessment, order_by = year),
            past.mill = lag(mill.rate, order_by = year)) %>%
     arrange(PIC) %>% 
     group_by(municipality, year) %>% 
     mutate(n.prop = n()) %>% 
     arrange(desc(n.prop)) %>% 
-    distinct(municipality, .keep_all = T) 
+    distinct(municipality, .keep_all = T) %>% 
+    select(-c(tax, improvement.assessment, land.assessment, total.assessment))
   
-  rfdat
+  dat_mill
+}
+
+# used to create assessment value data for random forest
+# no longer needed for shiny app; included for future use if needed
+asData <- function(data) {
+  
+  dat <- data
+  dat$municipality <- as.factor(dat$municipality)
[…]

### 2020-04-12 src/shiny_app/server.R

@@ -22,18 +22,14 @@ server <- function(input, output, session) {
     # Update when following inputs are changed
     input$updateButton
     
-    newdata <- dat
+    newdata <- datshort
     d <- NULL
     #print(dim(dat))
     
     # Filter data based on the user inputs
     isolate({
-          
           # If using PIC:
-          if(input$picInput && input$identInput!="")
-            {
-          
-            
+          if(input$picInput && input$identInput!=""){
             d <- newdata  %>% 
               filter(PIC == input$identInput)
           }
@@ -59,7 +55,6 @@ server <- function(input, output, session) {
   
    ########## PLOTTING TAB ###################
    # Add plots of either mill rate or assessment value over time to plot tab
-  munic <- reactive(input$municipalityInput)
   
   # create mill rate plot that reacts to inputs
   millrateplot <- reactive({
@@ -70,7 +65,7 @@ server <- function(input, output, session) {
           
     isolate({
     if(is.null(data)){
-      p <- paste("No corresponding data.")
+      p <- paste("No corresponding data to plot.")
     }
 
     # plot mill rates over time for municipality chosen for mill rate predicitons
@@ -85,7 +80,7 @@ server <- function(input, output, session) {
     }
     })
 
-    return(p)
+    p
 
   })
 
@@ -97,23 +92,28 @@ server <- function(input, output, session) {
 
     isolate({
       if(is.null(data)){
-        p <- paste("No corresponding data.")
+        p <- paste("No corresponding data to plot.")
         }
       
     # plot assessment values over time 
-      if(input$typeInput == 'Assessment Value'){
+      if(input$picInput && input$identInput!=""){
+        if(input$typeInput == 'Assessment Value') {
           #print("ggplotting assessment values")
           p <- ggplot(data, aes(x = year, y = total.assessment)) +
-            geom_line(color="#56B4E9") +
-            geom_point(color="#56B4E9") +
+            geom_line(color = "#56B4E9") +
+            geom_point(color = "#56B4E9") +
             theme_minimal() +
             xlab("Year") +
             ylab("Assessment Value") +
             ggtitle("Assessment Values Over Time")
+        }
+      }
+      else{
+        p <- paste("No corresponding data to plot.")
       }
       })
     
-    return(p)
+    p
     
   })
   
@@ -121,18 +121,13 @@ server <- function(input, output, session) {
    output$coolplot <- renderPlot({
      if (input$typeInput != 'Select'){
        if(input$typeInput == 'Assessment Value'){
[…]

### 2020-04-12 src/shiny_app/global.R
almost done..

@@ -26,8 +26,8 @@ datshort <- dat %>%
 
 
 # load dataset used for rf.mill
-rfdatload <- readRDS("rf_data.rds")
+rfdat <- readRDS("rf_data.rds")
 
 # create dataset used for rf.as
-asdatload <- readRDS("as_data.rds")
+asdat <- readRDS("as_data.rds")
   

### 2020-04-12 src/shiny_app/server.R

@@ -5,11 +5,12 @@
 # load rdas for each so that estimates can be made
 # fix select input to choose only the top 52 categories and other
 
+library(randomForest)
 
 # in case modified data needs to be accessed
 source("helpers.R")
 
-# load models - RF for mill rate predictions and ??? for assessment value predictions
+# load models - RF for mill rate predictions and for assessment value predictions
 load("rf.mill.rda")
 load("rf.as.rda")
 
@@ -140,18 +141,17 @@ server <- function(input, output, session) {
   estimates <- reactive({
     input$updateButton
     
-    data <- filtered()
-    
     isolate({
-      if(is.null(data)){
+      if(is.null(filtered())){
         pred <- paste("No data.")
       }
 
       else{
         # If using PIC: 
+        
         # If doing Mill Rate prediciton:
         if(input$typeInput == 'Mill Rate'){
-          print("doing mill rate prediction")
+          #print("doing mill rate prediction")
           
           # extract latest mill rate
           past20 <- filtered() %>% 
@@ -257,7 +257,7 @@ server <- function(input, output, session) {
           pred <- round(predict(rf.as, newdata = pred.data),2)
           
           if(last.assess==0 || is.na(pred)){
-            pred <- paste("No previous assessment value found.")
+            pred <- paste("Missing required data.")
           }
         }
       }
@@ -356,6 +356,10 @@ server <- function(input, output, session) {
           if(input$typeInput == 'Select'){
             return(paste("Select prediction type."))
           }
+        }
+        
+        else{
+          return(paste(""))
         }
       }
       

### 2020-04-12 src/shiny_app/global.R
last edits to shiny app

@@ -5,6 +5,7 @@ library(dplyr)
 library(DT)
 library(tidyr)
 library(shinyjs)
+library(randomForest)
 
 #PIC used for some testing
 # CA-BC-200-001019632060000

### 2020-04-12 src/shiny_app/helpers.R

@@ -97,4 +97,4 @@ asData <- function(data) {
   dat_as
 }
 
-
+saveRDS(asData(dat),"as_data.rds")

### 2020-04-12 src/shiny_app/server.R

@@ -5,7 +5,6 @@
 # load rdas for each so that estimates can be made
 # fix select input to choose only the top 52 categories and other
 
-library(randomForest)
 
 # in case modified data needs to be accessed
 source("helpers.R")
@@ -308,8 +307,18 @@ server <- function(input, output, session) {
          }
       
        if(input$typeInput == 'Assessment Value'){
+         md <- asdat %>% 
+           filter(municipality == input$municipalityInput)
+         
+         minm <- min(na.omit(md$total.assessment))
+         print(minm)
+
+         maxm <- max(na.omit(md$total.assessment))
+         print(maxm)
+         
          return(paste("Predicted next assessment value - \n", 
-                      estimates(), sep = " "))  # RETURN PREDICTION
+                      estimates(), "\n Valid prediction range for this municipality is",
+                      minm, "-", maxm, sep = " "))  # RETURN PREDICTION
          }
         
        if(input$typeInput == 'Select'){

### 2020-04-12 src/shiny_app/ui.R

@@ -33,7 +33,8 @@ ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?
                 sidebarLayout(
                   sidebarPanel(
                     # tabsetPanel(
-                    #   tabPanel("User Inputs",                
+                    #   tabPanel("User Inputs",        
+                    
                     # Only show the following for assessment predictions:
                     # Use PIC?
                     checkboxInput("picInput", "Use PIC?", value = FALSE),
@@ -50,8 +51,7 @@ ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?
                                      
                                      # for municipality
                                      selectInput("municipalityInput", "Municipality:", 
-                                                 c("-",sort(unique(datshort$municipality)),
-                                                   "Other"),
+                                                 c("-",sort(unique(datshort$municipality))),
                                                  selected = "-"),
                                                      
                                      # for Tax Class code
@@ -63,8 +63,9 @@ ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?
                                      conditionalPanel("input.typeInput == 'Assessment Value'",
                                                       numericInput("assessmentInput",
                                                                    "Current Assessment Value:",
-                                                                   value = 500000, min = 0, 
-                                                                   max = 6000000000))
+                                                                   value = 70000000, 
+                                                                   min = 4241700, 
+                                                                   max = 10000000000))
                                      ),
                                     
                     

### 2020-04-12 src/modelling - 550/rf_mill.Rmd
last edits to shiny app
change mean to median in mill rate RF

@@
   group_by(municipality, tax.class, year) %>% 
-  mutate(avg_assessment = mean(total.assessment, na.rm=TRUE)) %>% 
+  mutate(avg_assessment = median(total.assessment, na.rm=TRUE)) %>% 
   select(municipality, year, tax.class, avg_assessment) %>% 
@@
 
-save(rf.as, file = "rf.as.rda")
+save(rf.as, file = "rf.as1.rda")
 importance(rf.as)

### 2020-04-12 src/modelling - 550/rf_mill.Rmd
changed to med

@@
 
-avg_assessment_by_municipality = dat %>% 
+med_assessment_by_municipality = dat %>% 
   left_join(factors_tbl) %>% 
@@
   group_by(municipality, tax.class, year) %>% 
-  mutate(avg_assessment = median(total.assessment, na.rm=TRUE)) %>% 
-  select(municipality, year, tax.class, avg_assessment) %>% 
+  mutate(med_assessment = median(total.assessment, na.rm=TRUE)) %>% 
+  select(municipality, year, tax.class, med_assessment) %>% 
   distinct(municipality, .keep_all = T)
@@
   rename(municipality = col_municipality) %>% 
-  left_join(avg_assessment_by_municipality) %>% 
+  left_join(med_assessment_by_municipality) %>% 
   group_by(PIC) %>% 
-  mutate(next.assess = lead(avg_assessment, order_by = year),
+  mutate(next.assess = lead(med_assessment, order_by = year),
          past.mill = lag(mill.rate, order_by = year)) %>%
@@
 rf.mill <- randomForest(
-  mill.rate ~  tax.class + municipality + avg_assessment + past.mill, na.action = na.omit, mtry = 4,
+  mill.rate ~  tax.class + municipality + med_assessment + past.mill, na.action = na.omit, mtry = 4,
   data=train_mill, ntree=500
@@
 yhat.bag1 <- predict(rf.as,newdata=test_as)
-plot(yhat.bag1, test_as$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value", ylim=c(0, 6e8))
+plot(yhat.bag1, test_as$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value", ylim=c(0, 10e8))
 abline(0,1)

### 2020-04-12 src/shiny_app/server.R
correction to range in empty case

@@ -307,14 +307,17 @@ server <- function(input, output, session) {
          }
       
        if(input$typeInput == 'Assessment Value'){
-         md <- asdat %>% 
-           filter(municipality == input$municipalityInput)
          
-         minm <- min(na.omit(md$total.assessment))
-         print(minm)
-
-         maxm <- max(na.omit(md$total.assessment))
-         print(maxm)
+         if(!is.null(filtered())){
+           md <- asdat %>%
+             filter(municipality == input$municipalityInput)
+           
+           minm <- min(na.omit(md$total.assessment))
+           print(minm)
+           
+           maxm <- max(na.omit(md$total.assessment))
+           print(maxm)
+         }
          
          return(paste("Predicted next assessment value - \n", 
                       estimates(), "\n Valid prediction range for this municipality is",

### 2020-04-13 src/shiny_app/helpers.R
fixing ranges in shiny app

@@ -97,4 +97,4 @@ asData <- function(data) {
   dat_as
 }
 
-saveRDS(asData(dat),"as_data.rds")
+#saveRDS(asData(dat),"as_data.rds")

### 2020-04-13 src/shiny_app/server.R

@@ -311,18 +311,18 @@ server <- function(input, output, session) {
          if(!is.null(filtered())){
            md <- asdat %>%
              filter(municipality == input$municipalityInput)
-           
+
            minm <- min(na.omit(md$total.assessment))
            print(minm)
-           
+
            maxm <- max(na.omit(md$total.assessment))
            print(maxm)
          }
-         
-         return(paste("Predicted next assessment value - \n", 
+
+         return(paste("Predicted next assessment value - \n",
                       estimates(), "\n Valid prediction range for this municipality is",
                       minm, "-", maxm, sep = " "))  # RETURN PREDICTION
-         }
+       }
         
        if(input$typeInput == 'Select'){
          return("Enter prediction type.")

### 2020-04-13 src/shiny_app/server.R
fixed error statement when user inputs aren't complete.

@@ -318,6 +318,11 @@ server <- function(input, output, session) {
            maxm <- max(na.omit(md$total.assessment))
            print(maxm)
          }
+         
+         if(input$municipalityInput == '-' || 
+            input$taxclassInput == '-'){
+           return(paste("Complete user inputs."))
+         }
 
          return(paste("Predicted next assessment value - \n",
                       estimates(), "\n Valid prediction range for this municipality is",

### 2020-04-13 doc/450/RealEstate_final_report.Rmd
grammar fix according to Estalla's comments

@@
 author: "Peter Han, Xuechun Lu, Yuetong Liu, Yuting Wen"
-date: "07/04/2020"
+date: "12/04/2020"
 output: 
@@
 
-The main objective of our project is to accurately predict future mill rate (property tax) in metro Vancouver for the following 3 property tax classes: Tax class 1: Residential, Tax class 5: Light industry and Tax class 6: Business and other.
-Data cleaning, exploratory data analysis are used in this project to analyze the relationship between mill rate and other factors.
+The main objective of our project is to accurately predict future mill rate (property tax) in metro Vancouver for the following 3 property tax classes: 1) Tax class 1 - Residential; 2) Tax class 5 - Light industry; 3) Tax class 6 - Business and other.
+Data cleaning and exploratory data analysis are used to analyze the relationship between mill rate and other factors.
 Data cleaning is performed to aggregate our data into summary statistics.
 Exploratory analysis shows that there are strong relationships between mill rate and tax class and mill rate and municipalities; it also shows there is a fairly strong correlation between mill rate and average assessment per property in different municipalities.
-*Ordinary linear model*, *reduced ordinary linear model*, *Ridge Regression* and *LASSO* are used to predict the mill rate.
-A full assessment of the performance --- prediction power and goodness of fit of these models, is shown below.
+*Ordinary linear model*, *reduced ordinary linear model*, *Ridge Regression*, and *LASSO* are used to predict the mill rate.
 
@@
 Our goal is to predict mill rates in 2020 in Metro Vancouver.
-
 We also seek to identify which explanatory variables are the most important in determining mill rates.
@@
 Our client provided us with the past 5 years' property assessment data in BC.
-Since the only interest is in predicting mill rate for metro Vancouver and specific tax classes, a subset of properties that satisfy our interest have been selected:
+Since the only interest is in predicting mill rate for metro Vancouver and specific tax classes, a subset of properties that satisfy our interest are selected:
 
@@
 
-Moreover, 5 features have been selected that could be relevant to the mill rate:
+Moreover, 5 features that are selected can be relevant to the mill rate:
 
@@
 
-To reduce the dimension of our data, all properties in the same region, tax class code, and year are aggregated into a group because these properties have the same mill rate,  which is our response variable.
+All properties in the same region, tax class code, and year are aggregated into a group to reduce the dimension of our data; these properties have the same mill rate, which is our response variable.
 Here are the summary statistics for these groups:
@@
 Before any prediction on the future mill rate of Metro Vancouver’s real estate market was made, exploratory data analysis was performed to explore and visualize the main characteristics of our dataset.
-Correlation analyses between Mill Rate and Total Assessment, Mill Rate and Total Land Assessment, and Mill Rate and Total Improvement Assessment were performed.
-From our initial analysis, outliers in municipalities were found.
-Data transformation --- calculating the average total assessment, was used to reduce the effect of outliers.
+Correlation analyses between Mill Rate and Total Assessment, Mill Rate and Total Land Assessment, and Mill Rate and Total Improvement Assessment were performed respectively.
+From our initial analysis, outliers in municipalities were found, and the effects of these outliers were reduced by calculating the average total assessment.
 
@@
 
-In this study, linear models were built and the performance of each model was evaluated by the goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values, respectively.
+In this study, linear models were built and the performance of each model was evaluated by the goodness of fit and prediction power, that is, how well the model would explain the data and how well it could predict future values, respectively.
 The definitions of the goodness of fit and prediction power are given below.
@@
 
-Also, another advantage of these models is that they help reduce the variance of MSPE, so that MSPE is more stable. For more details about the two models, please refer to **Reference**.
+Another advantage of these models is that they help reduce the variance of MSPE, so that MSPE is more stable. For more details about the two models, please refer to **Reference**.
 
@@
 
-After the 50 runs, a vector of MSPEs on the training sets and a vector of MSPEs on the test sets for each model were therefore constructed successfully.
+After the 50 runs, a vector of MSPEs on the training sets and a vector of MSPEs on the test sets for each model were constructed successfully.
 Based on these vectors, side-by-side boxplots were used to show the mean and the spread of MSPEs on the training sets and test sets across all models respectively.
@@
 
-Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and total assessment among different municipalities is created to show the relationship in **Figure 1**.
+Since the mill rate is mainly affected by the total assessment, which is the sum of total land assessment and total improvement assessment, scatter plots between mill rate and total assessment among different municipalities have been created created to show the relationship in **Figure 1**.
 Each color in the figure belongs to three tax classes of one municipality.
 
-Scatter plots between mill rate and total land assessment, mill rate and total improvement assessment are similar to **Figure 1**. For more details about them, please refer to **Appendix**.
+Scatter plots between mill rate and total land assessment, and mill rate and total improvement assessment are similar to **Figure 1**. For more details about them, please refer to **Appendix**.
 
@@
 There is no clear trend between the mill rate and total assessment from **Figure 1**.
-The plot has also shown that most points are condensed on the left horizontal axis since some municipalities have larger assessment values than others.
+The plot shows that most points are condensed on the left horizontal axis since some municipalities have larger assessment values than others.
 
 To reduce the effect of large assessment in some municipalities, total assessment across all municipalities is transformed by taking total assessment dividing by the number of properties of each municipality and tax class.
-The transformed data is named **"Average Total Assessment"**.
+The transformed data set is named **"Average Total Assessment"**.
 A scatter plot between mill rates and average total assessments is shown in **Figure 2**.
@@
 
-The plot from **Figure 2** has shown that the mill rate tends to decrease as the average total assessment increase. Also, they have a moderately strong linear correlation.
+The plot from **Figure 2** shows that the mill rate tends to decrease as the average total assessment increase. Also, they have a moderately strong linear correlation.
 
@@
 
-Here categorical variables are taken into account, boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in **Figure 3** and **Figure 4**, respectively.
+Here categorical variables are taken into account. Boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in **Figure 3** and **Figure 4**, respectively.
 
@@
 
-**Figure 3** has shown that most municipalities have different mean and variance in terms of mill rates.
+**Figure 3** shows that most municipalities have different mean and variance in terms of mill rates.
 
-**Figure 4** has also supported that there is unequal mean and variance across tax classes.
+**Figure 4** also shows that there is unequal mean and variance across tax classes.
 
-A statistical test called the Kruskal-Wallis Test was performed to test their distribution.
+A statistical test called the Kruskal-Wallis Test has been performed to test their distribution.
 It was used to decide if population distributions were identical, and the corresponding p-value which is smaller than 0.05 indicated that the data have nonidentical distributions.
 
-The results of Kruskal-Wallis Test is shown in **Table 1**.
+**Table 1** shows the results of Kruskal-Wallis Test.
 
@@
 
-The p-values in **Table 1** have supported that there are nonidentical distributions of mill rates across municipalities and tax classes.
+The p-values in **Table 1** support that there are nonidentical distributions of mill rates across municipalities and tax classes.
 
@@
 
-Below is a comparison of the five models (*OLR full*, *OLR transformed*, *Lasso*, *Ridge* and *OLR null*) using the goodness of fit and prediction power.
-The goodness of fit was measured by MSPE on the training sets, whereas prediction power was measured by MSPE on the test sets.
+Five models, *OLR full*, *OLR transformed*, *Lasso*, *Ridge*, and *OLR null* are compared using the goodness of fit and prediction power.
+The goodness of fit is measured by MSPE on the training sets, whereas prediction power is measured by MSPE on the test sets.
 Generally, smaller MSPEs on the training sets and test sets indicate better fit and prediction power respectively.
 
-The distributions of the goodness of fit and prediction power across models are displayed in **Figure 5** and **Figure 6**, respectively, the distribution of null model is removed since it has larger values compared to all other models.
+The distributions of the goodness of fit and prediction power across models are displayed in **Figure 5** and **Figure 6**, respectively; the distribution of null model is removed since it has larger values compared to all other models.
 
@@
 
-As for goodness of fit, *Lasso*, *Ridge*, and *OLR transformed* have similar performance; MSPEs across these models were around 4.6, but *OLR transformed* performed slightly better and had a smaller spread (variance) of MSPE.
-*OLR full* performed worse than the other three, around 5.4, and *OLR null* had the greatest MSPE, around 35.
+As for goodness of fit, *Lasso*, *Ridge*, and *OLR transformed* have similar performance; MSPEs across these models are around 4.6, but *OLR transformed* performs slightly better and has a smaller spread (variance) of MSPE.
+*OLR full* performs worse than the other three, around 5.4, and *OLR null* has the greatest MSPE, around 35.
 
-Similarly, as for prediction power, *Lasso*, *Ridge*, and *OLR transformed* performed roughly the same; MSPEs across these models were close to 6.
-*OLR full* was worse with MSPE around 6.9.
-*OLR Null* performed the worst with MSPE over 35.
+Similarly, as for prediction power, *Lasso*, *Ridge*, and *OLR transformed* performs roughly the same; MSPEs across these models are close to 6.
+*OLR full* is worse with MSPE around 6.9.
+*OLR Null* performs the worst with MSPE over 35.
 
-Therefore, we conclude that from the results of a 50-run of 10-fold cross-validation, *Lasso*, *Ridge*, and *OLR transformed* had the best goodness of fit and prediction power.
-*OLR null* perfromed much worse than all the other four models.
+Therefore, we conclude that from the results of a 50-run of 10-fold cross-validation, *Lasso*, *Ridge*, and *OLR transformed* have the best goodness of fit and prediction power.
+*OLR null* performs much worse than all the other four models.
 
@@
 From the exploratory data analysis, we have found that Total Assessments contain outliers in major municipalities such as Vancouver and Burnaby.
-In order to reduce the effect of outliers, the Total Assessment was transformed by taking its average over the number of properties in each municipality.
-The correlation analyses had shown that the transformed assessment total is the only continuous variable that has a relatively strong correlation with the mill rate.
-The Kruskal-Wallis had suggested that mill rates in each municipality and tax class are significantly different.
+In order to reduce the effect of outliers, the Total Assessment has been transformed by taking its average over the number of properties in each municipality.
+The correlation analyses have shown that the transformed assessment total is the only continuous variable that has a relatively strong correlation with the mill rate.
+The Kruskal-Wallis has suggested that mill rates in each municipality and tax class are significantly different.
 From these results, we have decided to use Average Assessment Total, TaxClassCode and Municipality to fit our transformed model.
@@
 *Transformed OLR*, *Ridge Regression*, and *LASSO* were able to make good predictions based on mean squared prediction error on the training sets and test sets from cross-validation.
-Since the client prefers a simpler model, we choose the *transformed model* to make our 2020 prediction.
+Since the client prefers a simpler model, we have chosen the *transformed model* to make our 2020 prediction.
 

### 2020-04-13 src/modelling - 550/rf.Rmd
changed doc to reflect changing avg to median and factors in mill rate and assessment data, and redid outputs used in final latex report to relfect this

--- 2020-04-10 src/modeling - 550/rf.Rmd
@@
 
-lme.mill <- lmer(mill.rate ~ 1+ (1+year|municipality) + as.factor(year) + tax.class  + total.assessment + past.mill, data = train)
+lme.mill <- lmer(mill.rate ~ 1+ (1+year|municipality) + as.factor(year) + tax.class  + avg_assessment + past.mill, data = train_mill)
 
@@
 
-lme.as <- lmer(next.assess ~ 1+ (1+year|municipality) + as.factor(year) + tax.class + total.assessment + mill.rate, data = train)
+lme.as <- lmer(next.assess ~ 1+ (1+year|municipality) + as.factor(year) + tax.class + total.assessment + mill.rate, data = train_as)
 

### 2020-04-14 doc/450/RealEstate_final_report.Rmd
tiny fix

@@
 
-Here categorical variables are taken into account. Boxplots of mill rate across municipalities and tax classes are plotted to display the distributions in **Figure 3** and **Figure 4**, respectively.
+Here categorical variables are taken into account.
+Boxplots of mill rate across municipalities and tax classes have been plotted to display the distributions in **Figure 3** and **Figure 4**, respectively.
 

### 2020-04-14 doc/550/s550 no fig repor.tex
Final tweaking for submission. -gd

\documentclass{article}
\usepackage[utf8]{inputenc}

\input{latex-macros/report-template}
\input{latex-macros/typesetting-macros}
\input{latex-macros/basic-math-macros} 
\input{latex-macros/code-environments}

\bibliography{RealEstate.bib}

\begin{document}

\printtitle

\section*{Summary}
Property taxes are the single greatest operating expense for property owners in British Columbia, and mainly depend on two factors: property assessment---published each year in January---and municipal mill rates---published in April. Accurately projecting annual mill rates between January and April and future years' assessment values between May and December would allow businesses and individuals to budget for these expenses. For this purpose, multiple competing models are fitted based on a data set containing assessment values, mill rates, and other property information from 2016 to 2020 of over 200,000 properties. For predicting mill rates early in the year, a random forest regression model is found to be the most accurate. Likewise, another random forest regression model is also the most accurate for predicting future years' assessment values. Both models are incorporated in an interactive Shiny app that provides a straightforward user interface with a property-specific tax assessment.

\section{Introduction}
For property owners in British Columbia, property taxes are the single greatest operating expense. Mill rates are determined by individual municipalities and depend on a number of factors, including assessment values and municipal budgets.  While property assessments are published each year in January, municipal mill rates are not determined until April of the same calendar year, making it difficult for property owners to budget for this large expense.  Thus, accurately projecting the annual mill rates between the months of January and April and the next year's assessment values between May and December is a value-add to real estate consulting services. Furthermore, it also assists property owners in financial planning and fund allocation.  The objective of this written report is to build accurate statistical predictive models for both future municipal mill rates and property-specific assessment values. This will be done for tax class codes 1 (residential), 5 (light industrial), and 6 (commercial). Guided by the exploratory data analysis in Section \ref{section:eda}, several modeling techniques are presented in Section \ref{section:model}. Mill rate and assessment predictions using these models are available through a straightforward user interface via a Shiny app, which is discussed in the same section. Finally, Section \ref{section:conclusion} includes some concluding remarks.

\section{Statistical analysis} \label{section:analysis}

\subsection{Data} \label{sec:data}

\citet{s450-report} use the same dataset and offer a thorough description of the data in their analysis. In this report, however, all B.C. municipalities are included in analysis. The variables are summarised in Table \ref{table:variables}.  Missing mill rate values were imputed by using the average mill rate given by the data for the corresponding year and tax class code.

An important feature of the data is that spatial relationships exist within municipalities for assessment values and between municipalities for mill rates.  In other words, it is not sensible to assume independence in neither assessment values nor mill rates between different properties because geographical location (and thus proximity) affects these variables. This significantly influences possible model choices as the independence assumption is required for several standard modeling techniques. Furthermore, the data set lacks appropriate spatial information that would allow for geographical variables to be taken into account.

\subsection{Exploratory data analysis} \label{section:eda}

The exploratory data analysis (EDA) further investigates aspects of the EDA done by \citet{s450-report}. Both assessment values and mill rates have right-skewed distributions, and so they were log-transformed for visualization purposes. A visual inspection of Figure \ref{fig:scatter_year} shows that mill rates differ drastically from tax class to tax class, with industrial properties having the highest mill rates. However, neither mill rates nor assessment values seem to vary much over time, which suggests that current values of these variables will be a good indicator of their (short-term) future values. Furthermore, there does not appear to be a clear relationship between assessment values and mill rates when tax class is taken into account---except in the residential tax class, where a slightly negative relationship is present between the logarithms of average municipal assessment values and municipal mill rates. This remains true when only taking into account values from 2020, as can be seen in Figure \ref{fig:scatter_2020}.

The fact that mill rates remain relatively constant throughout the years is further verified in Figure \ref{fig:evolution_millrate}, where industrial properties are again seen to have the largest mill rates on average. However, Figure \ref{fig:evolution_assessment} suggests that assessment values do not follow this pattern. Municipality plays a role in a property's assessments, something taken into account in the modeling process.

Finally, Figure \ref{fig:violin} shows that the distributions of both mill rates and assessment values are indeed right-skewed (the values shown are in logarithmic scale), and further confirms the stark difference in mill rates between the three tax classes. Notably, Figure \ref{fig:violin_assessment} suggests that this difference is not present in assessment values, which indicates that tax class may not be that relevant for predicting future years' assessment values.

\FloatBarrier

\subsection{Modeling} \label{section:model}

This section describes the models used to predict mill rate and assessment values. Each of the following models were trained on a subset of 75\% of the data and tested on the remaining 25\%.

\subsubsection{Random forest regression}

% Pick one of the two
The random forest (RF) algorithm is an ensemble learning method for classification and regression that operates by constructing numerous decision trees and choosing the majority class output (classification) or the average prediction (regression) given by the individual decision trees. Decision trees are trained on different samples (with replacement) of the data, and only a random subset of the covariates is used to define the split at each node, based on a recursive partitioning strategy. RF is preferred over a single decision tree as it overcomes the problem of over-fitting.

%%%
A RF model was fitted to predict mill rate using tax class, municipality, past years' mill rate, and current years' total assessment as independent variables. Another RF model was similarly constructed using next year's assessment value as the dependent variable, and municipality, current years' mill rate, tax class, and current years' total assessment as covariates. The number of trees used for both models was 500, and all 4 predictors were evaluated at each node for splitting. The percent of total variability explained is 96.38 \% for the mill rates model and 97.48 \% for the assessment values model.

The importance of model covariates is indicated by a higher value of increase node purity (mean decrease accuracy). Municipality was deemed the most important variable in predicting mill rate (see Figure \ref{fig:4a}), followed by tax class and past year's mill rate. However, as shown in Figure  \ref{fig:4b}, current years' assessment value is the most important variable in predicting next years' assessment value, followed by municipality and mill rate.

% Don't think it is necessary to mention bootstrapping, this is done automatically in RF.
Figure \ref{fig:5} shows the evaluation of the model's performance on a test dataset. The RF model tends to underestimate mill rates, probably due to the greater difficulty of the task. However, RF shows good performance in predicting next year's assessment values.

\subsubsection{Mixed effects models and generalized estimating equation}
Since the data set contains repeated measurements for both municipalities and properties over the years, it is reasonable to consider a linear mixed effect model (LME) for predicting mill rates and assessment values. Both random slope and random intercept for each municipality are needed based on Figure \ref{fig:evolution_assessment}, which depicts different initial values and rates of change for both assessment values and municipal mill rates.  Based on the fixed effects shown in Figure \ref{fig:6b}, the direction of change in assessment values fluctuates over time. For example, assessment values increased from 2016 to 2017 but decreased from 2018 to 2020. The median municipal assessment value is positively associated with current mill rate, as is past year's mill rate (Figure \ref{fig:6a}). A negative relationship between mill rate and assessment values is observed in Figure \ref{fig:6b}, indicating that lower mill rates are associated with higher assessment values.

Unfortunately, the diagnostics plots verifying model assumptions are subpar for both LME models, as can be seen in Figure \ref{fig:7}. Residuals are not normally distributed, variance is heterogeneous, and random intercept is non-normal.  Thus it does not appear that an analysis using this type of model is appropriate for this data.

\subsubsection{Shiny app}

A straightforward user interface for making predictions using the RF models for mill rate and assessment values was built by way of a shiny app, which can be found in the GitHub repository \citep{gitrepo} or accessed on the web at {\color{blue} \underline{\url{https://malloryjflynn.shinyapps.io/STAT550_Real_Estate/}}}.

The shiny app calculates predictions of the next years' mill rate or assessment value under the `Estimate' tab of the main panel.  This can be done using either a Property Identifier Code (PIC) or user inputs for municipality, tax class code, and previous year's mill rate (mill rate predictions) or current assessment value (assessment value predictions).  To provide the user with a broader view of the trend in mill rate for the municipality selected, a plot of the mill rate for that municipality over the years (2016 - 2020) is visible under the `Plot' tab of the main panel.  When predicting assessment value using a PIC, a plot of the assessment value for that property over time will be displayed alternatively.

\section{Conclusions} \label{section:conclusion}

In this report we built models to predict mill rate and assessment value for properties in British Columbia's municipalities. Random forest was found to be the best fitting model, although it tended to overestimate mill rate predictions. In addition to this, we incorporated the predictive models in a shiny app that allows a user to perform predictions given a set of inputs and to visualize trends in mill rate and assessment value.

The main limitation of our analysis lies in the fact that the models assume independence of the data. However, as was argued in Section \ref{sec:data}, mill rates and assessment values are highly dependent on spatial information. An attempt to obtain the properties' spatial information was made by transforming the addresses into spatial coordinates using an open source geocoding system provided by the British Columbia Government. However, the open source geocoding system did not recognize some of the addresses in the dataset and provided incorrect coordinates for a large number of properties.  Therefore it is recommended that (1) addresses be formatted according to the standards required by the B.C. government, or (2) a private geocoding system that may be able to find addresses using the current format in the data is used. Provided spatial information is available, several model alternatives (e.g. factor models) could be tested to improve prediction accuracy, which take spatial information into account.

%%%% appendix
\printbibliography

\clearpage
\appendix
\section*{Appendix}

\section{EDA code}

\begin{lstlisting}[style=R, caption={Code used for the exploratory data analysis.}, captionpos=b]
# preamble ####
library(tidyverse)
library(readr)
library(readxl)
library(ggplot2)
ggplot2::theme_set(theme_classic())
library(viridis)

# data wrangling ####

# data import
columnnames <- readxl::read_xlsx("colnames.xlsx") %>% 
  colnames()

real.estate_full <- readr::read_csv("2016 - 2020 Raw.csv",
                                    na = c("", "NA", "NULL", "NULL_1"),
                                    col_names = columnnames,
                                    col_types = "ciccccccccccccccdddddiicccddddc")
# data wrangling

# assessments by PIC and year
assessments <- real.estate_full %>% 
  dplyr::select(PIC, Year, # relevant variables
                AssessedValueAmt, AssetTypeDesc) %>% 
  dplyr::rename(year = Year,
                assessment = AssessedValueAmt,
                assessment.type = AssetTypeDesc) %>% 
  dplyr::group_by(PIC, year, assessment.type) %>% 
  dplyr::summarise(assessment = sum(assessment)) %>% 
  dplyr::ungroup() %>% 
  tidyr::spread(assessment.type, assessment) %>% 
  dplyr::rename(improvement.assessment = Improvement,
                land.assessment = Land) %>% 
  dplyr::select(PIC, year, improvement.assessment, land.assessment) %>% 
  dplyr::mutate(total.assessment = improvement.assessment + land.assessment)

re <- real.estate_full %>% 
  dplyr::select(PIC, Year, AddressAssessorMunicipalityDesc, # relevant variables
                TaxClassCode, TaxOwingAmountTotalCalculated, TaxClassTaxRate) %>% 
  dplyr::rename(year = Year,
                municipality = AddressAssessorMunicipalityDesc, # human-readable names
                tax.class = TaxClassCode,
                tax = TaxOwingAmountTotalCalculated,
                mill.rate = TaxClassTaxRate) %>% 
  dplyr::filter(tax.class %in% c("01", "05", "06")) %>%  # relevant values for tax class
  dplyr::distinct() %>% 
  dplyr::left_join(assessments, by = c("PIC" = "PIC", "year" = "year")) # add assessment

# data viz ####
# tax classes dictionary
tax.classes <- as_labeller(c(
  `01` = "01 - Residential",
  `05` = "05 - Industrial",
  `06` = "06 - Commercial"
))

# facet scatter plots with year
re %>% 
  dplyr::select(-PIC) %>% 
  dplyr::filter(!is.na(total.assessment), !is.na(mill.rate)) %>% 
  dplyr::group_by(year, municipality, tax.class) %>% 
  dplyr::summarise(total.assessment = mean(total.assessment), mill.rate = mean(mill.rate)) %>% 
  ggplot(aes(x = log(total.assessment), y = log(mill.rate), color = factor(year))) +
  geom_point() +
  facet_wrap(.~tax.class, labeller = tax.classes) +
  labs(x = "log assessment value",
       y = "log mill rate",
       color = "year") +
  scale_color_viridis_d() +
  theme(text = element_text(size = 18))
ggsave("RealEstate/src/eda - s550/plots/1. scatter with year.pdf")
ggsave("RealEstate/src/eda - s550/plots/1. scatter with year.png")

# facet scatter plots for 2020 by municipality
re %>% 
  dplyr::filter(year == 2020) %>% 
  dplyr::select(-PIC, -year) %>% 
  dplyr::filter(!is.na(total.assessment), !is.na(mill.rate)) %>% 
  dplyr::group_by(municipality, tax.class) %>% 
  dplyr::summarise(total.assessment = mean(total.assessment), mill.rate = mean(mill.rate)) %>% 
  ggplot(aes(x = log(total.assessment), y = log(mill.rate))) +
  geom_point(color = viridis(20)[3]) +
  facet_wrap(.~tax.class, labeller = tax.classes) +
  labs(x = "log assessment value",
       y = "log mill rate") +
  theme(text = element_text(size = 18))
ggsave("RealEstate/src/eda - s550/plots/2. scatter 2020 by municipality.pdf")
ggsave("RealEstate/src/eda - s550/plots/2. scatter 2020 by municipality.png")

# facet line trends sample of 10
  re %>% 
    dplyr::select(-PIC) %>% 
    dplyr::filter(!is.na(total.assessment), !is.na(mill.rate)) %>% 
    dplyr::group_by(year, tax.class) %>% 
    dplyr::sample_n(size = 10) %>% 
    dplyr::ungroup() %>% 
    dplyr::group_by(year, tax.class, municipality) %>% 
    dplyr::summarise(total.assessment = mean(total.assessment), mill.rate = mean(mill.rate)) %>% 
    ggplot(aes(x = year, y = log(mill.rate), group = municipality)) +
    geom_line(color = viridis(20)[3]) +
    facet_wrap(.~tax.class, labeller = tax.classes) + 
    #scale_color_viridis_d() +
    #theme(legend.position = "none") +
    labs(x = "year",
         y = "log mill rate") +
    theme(text = element_text(size = 18),
          axis.text.x = element_text(angle = 45, hjust = 1))
  ggsave("RealEstate/src/eda - s550/plots/6. mill rate evolution sample.pdf")
  ggsave("RealEstate/src/eda - s550/plots/6. mill rate evolution sample.png")

    # facet line trends sample of 10 assessment
  re %>% 
    dplyr::select(-PIC) %>% 
    dplyr::filter(!is.na(total.assessment), !is.na(mill.rate)) %>% 
    dplyr::group_by(year, tax.class) %>% 
    dplyr::sample_n(size = 10) %>% 
    dplyr::ungroup() %>% 
    dplyr::group_by(year, tax.class, municipality) %>% 
    dplyr::summarise(total.assessment = mean(total.assessment), mill.rate = mean(mill.rate)) %>% 
    ggplot(aes(x = year, y = log(total.assessment), group = municipality)) +
    geom_line(color = viridis(20)[3]) +
    facet_wrap(.~tax.class, labeller = tax.classes) + 
    #scale_color_viridis_d() +
    #theme(legend.position = "none") +
    labs(x = "year",
         y = "log assessment value") +
    theme(text = element_text(size = 18),
          axis.text.x = element_text(angle = 45, hjust = 1))
  ggsave("RealEstate/src/eda - s550/plots/6.1 assessment evolution sample.pdf")
  ggsave("RealEstate/src/eda - s550/plots/6.1 assessment evolution sample.png")

  # violin plots of mill rates accross tax classes, for 2020
  re %>%  
    dplyr::filter(!is.na(total.assessment), !is.na(mill.rate), year == 2020) %>% 
    dplyr::group_by(PIC, tax.class) %>% 
    dplyr::summarise(total.assessment = mean(total.assessment), mill.rate = mean(mill.rate)) %>% 
    ggplot(aes(x = tax.class, y = log(mill.rate), fill = tax.class)) +
    geom_violin(alpha = 0.5, width = 1) +
    geom_boxplot(alpha = 0.75, width = 0.1) +
    labs(x = "tax class",
         y = "log mill rate") +
    #scale_fill_viridis_d(begin=0, end=1) +
    scale_fill_manual(values = c("#3E4A89FF", "#26828EFF","#B4DE2CFF" )) +
    theme(legend.position = "none") +
    scale_x_discrete(labels = c("01 - Residential", "05 - Industrial", "06 - Commercial")) +
    theme(text = element_text(size = 18))
  ggsave("RealEstate/src/eda - s550/plots/7. violin mill rates.pdf")
  ggsave("RealEstate/src/eda - s550/plots/7. violin mill rates.png")

  # violin plots of assessment values accross tax classes, for 2020
  re %>%  
    dplyr::filter(!is.na(total.assessment), !is.na(mill.rate), year == 2020) %>% 
    dplyr::group_by(PIC, tax.class) %>% 
    dplyr::summarise(total.assessment = mean(total.assessment), mill.rate = mean(mill.rate)) %>% 
    ggplot(aes(x = tax.class, y = log(total.assessment), fill = tax.class)) +
    geom_violin(alpha = 0.5, width = 1) +
    geom_boxplot(alpha = 0.75, width = 0.1) +
    labs(x = "tax class",
         y = "log assessment value") +
    #scale_fill_viridis_d(begin=0, end=1) +
    scale_fill_manual(values = c("#3E4A89FF", "#26828EFF","#B4DE2CFF" )) +
    theme(legend.position = "none") +
    scale_x_discrete(labels = c("01 - Residential", "05 - Industrial", "06 - Commercial")) +
    theme(text = element_text(size = 18))
  ggsave("RealEstate/src/eda - s550/plots/7.1 violin assessment values.pdf")
  ggsave("RealEstate/src/eda - s550/plots/7.1 violin assessment values.png")
\end{lstlisting}

\section{Modelling code}

\begin{lstlisting}[style=R, caption={Code used for the modeling.}, captionpos=b]
# preamble

suppressPackageStartupMessages(library(randomForest))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(ROCR))
suppressPackageStartupMessages(library(predictmeans))

###########################
# data prep
dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
dat$municipality <- as.factor(dat$municipality)
dat$tax.class <- as.factor(dat$tax.class)

factors_tbl = dat %>% 
  group_by(municipality) %>% 
  count(name="mun_count", sort = TRUE) %>% 
  ungroup() %>% 
  mutate(perc = mun_count/sum(mun_count),
         cum_perc = cumsum(perc)) %>% 
  arrange(desc(mun_count)) %>% 
  mutate(rank = row_number(),
         municipality = fct_reorder(municipality, rank)) %>% 
  mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
  select(municipality, col_municipality)

med_assessment_by_municipality = dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  filter(test.train == "train") %>% 
  group_by(municipality, tax.class, year) %>% 
  mutate(med_assessment = median(total.assessment, na.rm=TRUE)) %>% 
  select(municipality, year, tax.class, med_assessment) %>% 
  distinct(municipality, .keep_all = T)

dat_mill <- dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  left_join(med_assessment_by_municipality) %>% 
  group_by(PIC) %>% 
  mutate(next.assess = lead(med_assessment, order_by = year),
         past.mill = lag(mill.rate, order_by = year)) %>%
  arrange(PIC) %>% 
  group_by(municipality, year) %>% 
  mutate(n.prop = n()) %>% 
  arrange(desc(n.prop)) %>% 
  distinct(municipality, .keep_all = T) %>% 
  select(-c(tax, improvement.assessment, land.assessment, total.assessment))

dat_as <- dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  group_by(PIC) %>% 
  arrange(year) %>% 
  mutate(next.assess = lead(total.assessment, order_by = year),
         past.mill = lag(mill.rate, order_by = year)) %>% 
  group_by(municipality) %>% 
  top_n(25, wt = total.assessment) %>% 
  arrange(municipality)

train_mill <- dat_mill %>%  filter(test.train == "train")
test_mill <- dat_mill %>%  filter(test.train == "test")
train_as <- dat_as %>%  filter(test.train == "train")
test_as <- dat_as %>%  filter(test.train == "test")

####################################
# Random Forest
set.seed(0)

rf.mill <- randomForest(
  mill.rate ~  tax.class + municipality + med_assessment + past.mill, na.action = na.omit, mtry = 4,
  data=train_mill, ntree=500
)

save(rf.mill, file = "rf.mill.rda")

#Evaluate variable importance
importance(rf.mill)
varImpPlot(rf.mill)

rf.as <- randomForest(
  next.assess ~ tax.class + municipality + total.assessment + mill.rate, na.action = na.omit,mtry = 4,
  data=train_as
)

save(rf.as, file = "rf.as1.rda")
importance(rf.as)
varImpPlot(rf.as)

yhat.bag <- predict(rf.mill,newdata=test_mill)
plot(yhat.bag, test_mill$mill.rate, xlab="Predicted Mill Rate Using Test Set", ylab="Actual Mill Rate")
abline(0,1)

yhat.bag1 <- predict(rf.as,newdata=test_as)
plot(yhat.bag1, test_as$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value", ylim=c(0, 10e8))
abline(0,1)

################################
# Linear Mixed Effects Model

# both random slope and random intercept
# different rate of change of assessment value and mill rate as well as initial assessment value and mill rate for each municipality

lme.mill <- lmer(mill.rate ~ 1+ (1+year|municipality) + as.factor(year) + tax.class  + avg_assessment + past.mill, data = train_mill)

summary(lme.mill)

lme.as <- lmer(next.assess ~ 1+ (1+year|municipality) + as.factor(year) + tax.class + total.assessment + mill.rate, data = train_as)

summary(lme.as)

# check assumptions of lme

# Homogeneity of Variance
residplot(lme.mill)
residplot(lme.as)
\end{lstlisting}

\section{Shiny app code}

\begin{lstlisting}[style=R, caption={Code used for the Shiny app.}, captionpos=b]

#global.R file

library(shiny)
library(readr)
library(ggplot2)
library(dplyr)
library(DT)
library(tidyr)
library(shinyjs)
library(randomForest)

#PIC used for some testing
# CA-BC-200-001019632060000

# for bookmarking button
enableBookmarking("url")

# read data
dat <- readr::read_delim("test_train_data.txt", 
                         delim = ",", col_types = "ciccdddddf")

# creates dataset with only the top 52 municipalities
counts <- dat %>% 
  count(municipality, sort = TRUE)

datshort <- dat %>% 
  filter(municipality %in% counts$municipality[1:52])

# load dataset used for rf.mill
rfdat <- readRDS("rf_data.rds")

# create dataset used for rf.as
asdat <- readRDS("as_data.rds")

##################################
#################################
# ui.R file

# Mallory - STAT 550 2020###
# This is the user interface version of the shiny app

library(shinythemes)
library(png)

ui <- fluidPage(theme = shinytheme("cerulean"), #maybe journal theme?

                # header
                div(id = "headerSection",
                    h2("BC Mill Rate & Assessment Value Predictions"),

                    span(
                      style = "font-size: 1em",
                      # authors
                      span("Created by "),
                      a("Gian Carlo Diluvi, Vittorio Romaniello, Sophia Li & Mallory Flynn",
                        href = "https://www.stat.ubc.ca"),
                      HTML("&bull;"),
                      # date
                      span("April 2020"),
                      HTML("&bull;"),
                      # Shiny app code link
                      span("Code"),
                      a("on GitHub", 
                        href = "https://github.com/STAT450-550/RealEstate/tree/master/src/shiny_app"),
                    )
                ),
                br(),
                br(),

                # all content goes here, and is hidden initially until the page fully loads
                sidebarLayout(
                  sidebarPanel(
                    # tabsetPanel(
                    #   tabPanel("User Inputs",

                    # Only show the following for assessment predictions:
                    # Use PIC?
                    checkboxInput("picInput", "Use PIC?", value = FALSE),
                    selectInput("typeInput", "Estimate Type",
                                c("Select", "Assessment Value", "Mill Rate"),
                                selected = "Select"),

                    # If using PIC:
                    conditionalPanel("input.picInput",
                                     textInput("identInput", "PIC:", placeholder = NULL)),

                    # If PIC is not available:
                    conditionalPanel("!input.picInput",

                                     # for municipality
                                     selectInput("municipalityInput", "Municipality:", 
                                                 c("-",sort(unique(datshort$municipality))),
                                                 selected = "-"),

                                     # for Tax Class code
                                     selectInput("taxclassInput", "Tax Class Code:",
                                                 c("-",sort(unique(dat$tax.class))),
                                                 selected = "-"),

                                     #conditional input for estimate type
                                     conditionalPanel("input.typeInput == 'Assessment Value'",
                                                      numericInput("assessmentInput",
                                                                   "Current Assessment Value:",
                                                                   value = 70000000, 
                                                                   min = 4241700, 
                                                                   max = 10000000000))
                                     ),

                    # button to update the data
                    shiny::hr(),
                    actionButton("updateButton", "Update"),

                    # source of data as a footer - Altus Group image not loading
                    br(),
                    br(),
                    p("Generated using data from ",
                      a("the Altus Group Ltd.",
                        href = "https://www.altusgroup.com",
                        target = "_blank")),
                    a(img(src = "altusgroupimg.png", alt = "Altus Group",
                          height = 63, width = 150),
                      href = "https://www.altusgroup.com",
                      target = "_blank"),
                    br(),
                    br(),
                    br(),
                    br(),
                    bookmarkButton()
                    ),

                  # main panel with Estimate tab and plot tab with mill rates
                  # or assessment values over time
                  mainPanel(h4(textOutput("resultsText")),
                            tabsetPanel(
                              tabPanel("Estimate", 
                                       br(),
                                       verbatimTextOutput("results")),
                               tabPanel("Plot", 
                                        br(),
                                        plotOutput("coolplot"))
                              )
                  )
                )
                )

########################################
########################################
# server.R file

# Mallory - STAT 550 2020###
# This is the server file of the shiny app

# fix main title when PIC is checked but empty 
# load rdas for each so that estimates can be made
# fix select input to choose only the top 52 categories and other

# in case modified data needs to be accessed
source("helpers.R")

# load models - RF for mill rate predictions and for assessment value predictions
load("rf.mill.rda")
load("rf.as.rda")

# server:
server <- function(input, output, session) {

  filtered <- reactive({

    # Update when following inputs are changed
    input$updateButton

    newdata <- datshort
    d <- NULL
    #print(dim(dat))

    # Filter data based on the user inputs
    isolate({
          # If using PIC:
          if(input$picInput && input$identInput!=""){
            d <- newdata  %>% 
              filter(PIC == input$identInput)
          }

          # If not using PIC, filter by municipality and tax class:
          if(!input$picInput){

            d <- newdata %>% 
              filter(tax.class == input$taxclassInput,
                     municipality == input$municipalityInput)
          }

      })

    # return filtered data
    if(dim(d)[1]==0){
      d <- NULL
    }
    d

    })

   ########## PLOTTING TAB ###################
   # Add plots of either mill rate or assessment value over time to plot tab

  # create mill rate plot that reacts to inputs
  millrateplot <- reactive({
    input$updateButton

    data <- filtered()

    isolate({
    if(is.null(data)){
      p <- paste("No corresponding data to plot.")
    }

    # plot mill rates over time for municipality chosen for mill rate predicitons
    if (input$typeInput == 'Mill Rate'){
      p <- ggplot(data, aes(x = year, y = mill.rate)) +
        geom_line(color="#FF3333") +
        geom_point(color="#FF3333") +
        theme_minimal() +
        xlab("Year") +
        ylab("Mill Rate") +
        ggtitle("Municipal Mill Rate Over Time")
    }
    })

    p

  })

  # create assessment plots that react to user inputs
  assessplot <- reactive({
    input$updateButton

    data <- filtered()

    isolate({
      if(is.null(data)){
        p <- paste("No corresponding data to plot.")
        }

    # plot assessment values over time 
      if(input$picInput && input$identInput!=""){
        if(input$typeInput == 'Assessment Value') {
          #print("ggplotting assessment values")
          p <- ggplot(data, aes(x = year, y = total.assessment)) +
            geom_line(color = "#56B4E9") +
            geom_point(color = "#56B4E9") +
            theme_minimal() +
            xlab("Year") +
            ylab("Assessment Value") +
            ggtitle("Assessment Values Over Time")
        }
      }
      else{
        p <- paste("No corresponding data to plot.")
      }
      })

    p

  })

  # output one of the above plots onto UI
   output$coolplot <- renderPlot({
     if (input$typeInput != 'Select'){
       if(input$typeInput == 'Assessment Value'){
         assessplot()
       }

       else{
         millrateplot()
       }
       }

     else{
       return()
     }
     })

  ###### ESTIMATE TAB ####################
  # give predictions given user inputs for mill rate or assessment value
  estimates <- reactive({
    input$updateButton

    isolate({
      if(is.null(filtered())){
        pred <- paste("No data.")
      }

      else{
        # If using PIC:

        # If doing Mill Rate prediciton:
        if(input$typeInput == 'Mill Rate'){
          #print("doing mill rate prediction")

          # extract latest mill rate
          past20 <- filtered() %>% 
            filter(year == 2020)

          # if 2020 column is empty, it will break by condition on mean mill rate=0
          meanmillrate <- mean(na.omit(past20$mill.rate))
          print(meanmillrate)

          # put data together in the way rfmill expects as input call it inputdata
          # columns include tax.class, municipality, total.assessment, past.mill
          meanassess <- mean(na.omit(past20$total.assessment))
          print(meanassess)

          pred.data <- cbind(filtered()$tax.class[1], 
                             filtered()$municipality[1],
                             meanassess,
                             meanmillrate)
          pred.data <- as.data.frame(pred.data, stringsAsFactors = FALSE)
          colnames(pred.data) <- c('tax.class', 'municipality',
                                   'avg_assessment','past.mill')

          pred.data$past.mill <- as.numeric(pred.data$past.mill)
          pred.data$avg_assessment <- as.numeric(pred.data$avg_assessment)
          pred.data$municipality <- factor(pred.data$municipality,
                                           levels = levels(rfdat$municipality))
          pred.data$tax.class <- factor(pred.data$tax.class, 
                                        levels = levels(rfdat$tax.class))
          print(pred.data)

          # predict next mill rate using random forest
          pred <- round(predict(rf.mill, newdata = pred.data), 2)

          if(meanmillrate==0 || is.na(pred)){
            pred <- paste("No previous mill rate found in data.")
          }

          pred

          }

        # If doing Assessment Value prediction:
        if(input$typeInput == 'Assessment Value'){

          print("doing assessment value prediction")

          # extract latest assessment value
          past20 <- filtered() %>% 
            filter(year == 2020)

          # if using PIC and 2020 column is NA for this property's
          # assessment value, it will break 
          if(input$picInput){
            print("using PIC")
            if(input$identInput != ""){
              if(!is.na(past20$total.assessment)){
                last.assess <- past20$total.assessment
                }
              else{
                last.assess <- 0
              }
              }
            print(last.assess)
            }

          # if not using PIC, assessment value must come from user input
          else{
            print("not using PIC")
            if(input$assessmentInput != ""){
              last.assess <- input$assessmentInput
              }
            else{
              last.assess <- 0
              }
            print(last.assess)
            }

          # put data together in the way rfmill expects as input call it inputdata
          # columns include tax.class, municipality, total.assessment, and mill.rate
          print(head(filtered()))
          pred.data <- cbind(filtered()$tax.class[1], 
                             filtered()$municipality[1],
                             last.assess,
                             past20$mill.rate[1])

          pred.data <- as.data.frame(pred.data, stringsAsFactors = FALSE)
          colnames(pred.data) <- c('tax.class', 'municipality',
                                   'total.assessment', 'mill.rate')

          pred.data$mill.rate <- as.numeric(pred.data$mill.rate)
          pred.data$total.assessment <- as.numeric(pred.data$total.assessment)
          pred.data$municipality <- factor(pred.data$municipality,
                                           levels = levels(asdat$municipality))
          pred.data$tax.class <- factor(pred.data$tax.class,
                                        levels = levels(asdat$tax.class))
          print(pred.data)

          # predict next assessment value using random forest
          pred <- round(predict(rf.as, newdata = pred.data),2)

          if(last.assess==0 || is.na(pred)){
            pred <- paste("Missing required data.")
          }
        }
      }
      })

    pred
    print(pred)

    })

   # create estimates as text for output
   estimatestext <- reactive({
     input$updateButton

     # If using PIC:
     if(input$picInput){
       if(input$identInput!=""){
         if(input$typeInput == 'Mill Rate'){   #need to be changed to extract values
           return(paste("Mill rate prediction for class", 
                         filtered()$tax.class[1],
                         "in", filtered()$municipality[1], "- \n", 
                         estimates(), sep = " "))  #  RETURN PREDICTION
           }

         if(input$typeInput == 'Assessment Value'){
           return(paste("Predicted next assessment value of property \n", input$identInput,
                        "-", estimates(), sep = " "))  # RETURN PREDICTION
           }

         if(input$typeInput == 'Select'){
           return("Enter prediction type.")
         }
         }

       else{
         return("Please enter PIC.")
       }
       }

     # If not using PIC:
     else{
       if(input$typeInput == 'Mill Rate'){
         return(paste("Mill rate prediction for class", input$taxclassInput,
                      "in", input$municipalityInput, "- \n", 
                      estimates(), sep = " ")) # RETURN PREDICTION
         }

       if(input$typeInput == 'Assessment Value'){

         if(!is.null(filtered())){
           md <- asdat %>%
             filter(municipality == input$municipalityInput)

           minm <- min(na.omit(md$total.assessment))
           print(minm)

           maxm <- max(na.omit(md$total.assessment))
           print(maxm)
         }

         if(input$municipalityInput == '-' || 
            input$taxclassInput == '-'){
           return(paste("Complete user inputs."))
         }

         return(paste("Predicted next assessment value - \n",
                      estimates(), "\n Valid prediction range for this municipality is",
                      minm, "-", maxm, sep = " "))  # RETURN PREDICTION
       }

       if(input$typeInput == 'Select'){
         return("Enter prediction type.")
       }
       }
     })

   # output the estimates text in the main panel
   output$results <- renderText({
     estimatestext()
     })

  # Titles text for main panel title - describes prediction type or PIC 
  # number if applicable
  titles <- reactive({
    input$updateButton

    data <- filtered()

    if(!input$picInput && input$typeInput == 'Select' && input$municipalityInput == '-' &&
       input$taxclassInput == '-'){
      return(paste(""))
    }

    if(is.null(data)){
      return(paste("Could not find matching data."))
    }

    else{
      if(input$picInput){
        if(input$identInput!=""){

          if(input$typeInput == 'Assessment Value'){
            return(paste("Assessment value for", input$identInput, sep = " ")) #ADD PREDICTION HERE
          }

          if(input$typeInput == 'Mill Rate'){
            return(paste("Class", filtered()$tax.class[1], 
                         "mill rate for", filtered()$municipality[1], sep = " ")) #ADD PREDICTION HERE
          }

          if(input$typeInput == 'Select'){
            return(paste("Select prediction type."))
          }
        }

        else{
          return(paste(""))
        }
      }

      else{
          if(input$typeInput == 'Mill Rate'){
            return(paste("Class", input$taxclassInput, 
                         "mill rate for", 
                         input$municipalityInput, sep = " ")) 
          }

          else{
            if(input$typeInput == 'Select'){
              return(paste("Select prediction type."))
            }
            else{
              return(paste("Assessment Value prediction for class", 
                           input$taxclassInput, "property \n in",
                           input$municipalityInput, sep = " "))
            }
          }
        }
      }
  })

  output$resultsText <- renderText({     
    titles()
  })
}

######################################
#####################################
# helpers.R file
# This file will have helpers for the model and the loading of data for models

# function to modify data for random forest (if needed)

# used to create mill rate data for random forest
# no longer needed for shiny app; included for future use if needed
rfData <- function(data) {

  dat <- data
  dat$municipality <- as.factor(dat$municipality)
  dat$tax.class <- as.factor(dat$tax.class)

  factors_tbl = dat %>% 
    group_by(municipality) %>% 
    count(name="mun_count", sort = TRUE) %>% 
    ungroup() %>% 
    mutate(perc = mun_count/sum(mun_count),
           cum_perc = cumsum(perc)) %>% 
    arrange(desc(mun_count)) %>% 
    mutate(rank = row_number(),
           municipality = fct_reorder(municipality, rank)) %>% 
    mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
    select(municipality, col_municipality)

  avg_assessment_by_municipality = dat %>% 
    left_join(factors_tbl) %>% 
    select(-c(municipality)) %>% 
    rename(municipality = col_municipality) %>% 
    filter(test.train == "train") %>% 
    group_by(municipality, tax.class, year) %>% 
    mutate(avg_assessment = mean(total.assessment, na.rm=TRUE)) %>% 
    select(municipality, year, tax.class, avg_assessment) %>% 
    distinct(municipality, .keep_all = T)

  dat_mill <- dat %>% 
    left_join(factors_tbl) %>% 
    select(-c(municipality)) %>% 
    rename(municipality = col_municipality) %>% 
    left_join(avg_assessment_by_municipality) %>% 
    group_by(PIC) %>% 
    mutate(next.assess = lead(avg_assessment, order_by = year),
           past.mill = lag(mill.rate, order_by = year)) %>%
    arrange(PIC) %>% 
    group_by(municipality, year) %>% 
    mutate(n.prop = n()) %>% 
    arrange(desc(n.prop)) %>% 
    distinct(municipality, .keep_all = T) %>% 
    select(-c(tax, improvement.assessment, land.assessment, total.assessment))

  dat_mill
}

# used to create assessment value data for random forest
# no longer needed for shiny app; included for future use if needed
asData <- function(data) {

  dat <- data
  dat$municipality <- as.factor(dat$municipality)
  dat$tax.class <- as.factor(dat$tax.class)

  factors_tbl = dat %>% 
    group_by(municipality) %>% 
    count(name="mun_count", sort = TRUE) %>% 
    ungroup() %>% 
    mutate(perc = mun_count/sum(mun_count),
           cum_perc = cumsum(perc)) %>% 
    arrange(desc(mun_count)) %>% 
    mutate(rank = row_number(),
           municipality = fct_reorder(municipality, rank)) %>% 
    mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
    select(municipality, col_municipality)

  avg_assessment_by_municipality = dat %>% 
    left_join(factors_tbl) %>% 
    select(-c(municipality)) %>% 
    rename(municipality = col_municipality) %>% 
    filter(test.train == "train") %>% 
    group_by(municipality, tax.class, year) %>% 
    mutate(avg_assessment = mean(total.assessment, na.rm=TRUE)) %>% 
    select(municipality, year, tax.class, avg_assessment) %>% 
    distinct(municipality, .keep_all = T)

  dat_as <- dat %>% 
    left_join(factors_tbl) %>% 
    select(-c(municipality)) %>% 
    rename(municipality = col_municipality) %>% 
    group_by(PIC) %>% 
    arrange(year) %>% 
    mutate(next.assess = lead(total.assessment, order_by = year),
           past.mill = lag(mill.rate, order_by = year)) %>% 
    group_by(municipality) %>% 
    top_n(25, wt = total.assessment) %>% 
    arrange(municipality)

  dat_as
}

\end{lstlisting}

\end{document}

### 2020-04-14 doc/550/s550 project proposal.tex
Final tweaking for submission. -gd

\documentclass{article}
\usepackage[utf8]{inputenc}

\input{latex-macros/report-template}
\input{latex-macros/typesetting-macros}
\input{latex-macros/basic-math-macros} 
\input{latex-macros/code-environments}

\begin{document}

\printtitle

\section*{Summary}

Property taxes are the single greatest operating expense for property owners in British Columbia, and depend on two factors: property assessment---published each year in January---and municipal mill rates---published in April. Accurately projecting annual mill rates between January and April and future years assessment values between May and December would allow businesses and individuals to budget for these expenses, a value-add to the client's consulting services. This project will build statistical models to accurately predict mill rates and property assessment values. For this purpose, an expanded exploratory data analysis will be conducted to inform relationships between relevant property-related variables. It will also scrutinize trends uncovered by the STAT 450 analysis. Multiple statistical models---including stepwise mixed effects models, random forest regression, and simultaneous equations model---will then be explored. After the most suitable model is chosen, a Shiny app will be created to provide the client with a straightforward user interface which incorporates model predictions into a property-specific tax assessment.

\section{Objectives}

This project will address two research questions, building on the analysis of STAT 450 students. First, it will determine how assessment values in a given year, as well as other relevant property-related variables, affect mill rates before they are released in April. Second, it will explore how a given year's assessment values and mill rates affect the next year's assessment values.

In order to address these research questions, our analysis will build predictive models that accurately forecast both mill rates and assessment values for municipalities and properties in British Columbia. Mill rates for three tax classes will be predicted - residential (class 1), light industrial (class 5), and commercial (class 6).  An exploratory data analysis (EDA) will also be conducted, which will further investigate aspects of the \href{https://github.com/STAT450-550/RealEstate/blob/450/src/2nd_draft.pdf}{\color{blue} \underline{EDA}} completed by STAT 450 students.

\section{Data}\label{data}

A thorough description of the data can be found on the STAT 450 students'  \href{https://github.com/STAT450-550/RealEstate/blob/450/src/2nd_draft.pdf}{\color{blue} \underline{report}}. The data set used in this report is the same, with the caveat that information from all municipalities---not only Metro Vancouver---is also considered. The variables are summarised in Table \ref{table:variables}.

\begin{table}[H]
\centering
\rowcolors{2}{gray!20}{white}
\begin{tabular}{l|ll}
\rowcolor{gray!40}
\textbf{Variable}                                                               & \textbf{Type}                                                                                                              & \textbf{Included as}                 \\ \hline
Mill rate                                                              & Continuous                                                                                                        & Dependent variable          \\
Total assessment                                                       & Continuous                                                                                                        & Independent variable        \\
Total land assessment                                                  & Continuous                                                                                                        & Independent variable        \\
\begin{tabular}[c]{@{}l@{}}Total improvement\\ assessment\end{tabular} & Continuous                                                                                                        & Independent variable        \\
Tax class code                                                         & \begin{tabular}[c]{@{}l@{}}Categorical. \\ One of 1 (residential), \\ 5 (industrial), 6 (commercial)\end{tabular} & Independent variable        \\
Municipality                                                           & Categorical                                                                                                       & Independent variable        \\
Year & \begin{tabular}[c]{@{}l@{}}Discrete. \\ Values of 2016, 2017, 2018, and 2019\end{tabular}  & \begin{tabular}[c]{@{}l@{}} Independent variable. \\ Random slope/intercept\end{tabular}
\end{tabular}
\caption{Brief description of the variables included in the data set.}
\label{table:variables}
\end{table}

\section{Analysis outline}
The analysis will be divided in two parts: (1) exploratory data analysis (EDA) and (2) modeling and evaluation.

\subsection{Exploratory data analysis}

Our EDA will be similar to the one conducted by STAT 450 students. However, we will try to gain more insight from visualisations, meeting the client's request to understand relationships between the variables. More specifically, we plan on observing how mill rate changes not only within a municipality but also within tax classes. Figure 6 in the STAT 450 final report shows a difference in mill rate across tax classes, and so we deem necessary to further explore the impact of this variable at the level of municipalities. Furthermore, we have observed a negative relationship between mill rate and average total assessment found in Figure 4 of the STAT 450 report. Such an inverse relationship makes intuitive sense because mill rate and average total assessment for each municipalities should balance each other out in reality.

In addition to informing our analysis through visualisations, we plan on exploring whether transformations help improve understanding of the data and finally, model performance. In particular, a log transformation may help better analyze information related to property assessment, given the high variance and right-skewness present in the data.

%Lastly, since a high correlation is present between total, land, and improvement assessment values, we will use Principal Component Analysis to create independent features that summarise the information in the three variables. (Not anymore.)

%Plot mill rate for each municipality for each year to determine if a random slope/intercept is needed.

\subsection{Modeling}

%The modeling part is divided into statistical techniques and evaluation methods.

%\subsubsection{Statistical techniques}

To answer the research questions we will explore a variety of models, which we outline below.

\paragraph{Mixed effects models and Generalized Estimating Equation}

The data contains repeated measurements taken over years for each subject.  Based on the EDA, linear mixed models (LME) or nonlinear mixed effect models (NLME) will be fit to the data, with random effects on municipality and/or individual houses.  For mill rates, municipalities will be treated as subjects, while for assessment values, individual properties will be treated as subjects.

The NLME and LME models will be compared with General Estimating Equation (GEE) using diagnostics tools. GEE has poor interpretability but good prediction capabilities since it does not have any distributional assumptions on the data. NLME and LME have better interpretability, but assumptions will need to be verified.

\paragraph{Random forest regression}
Given the clients' interest in prediction accuracy, we plan to exploit the predictive power offered by random forest in this setting. Since the response variable is continuous, we propose random forest regression.

Random forest is a boosted decision tree-based algorithm that is preferred over a single regression tree as it helps prevent overfitting and reduces the variance of predictions. Both problems are likely to appear in our situation, as the data contains a limited number of data points per property.

\paragraph{Simultaneous equations model}
The mixed effects model and random forest regression can only predict one outcome at the time.  To answer the research questions, we will need to build two models under this methodology - one to predict mill rate, and one to predict the next year's assessment values. Using two models, where one takes as input the output of the other, increases prediction variance and can generally lead to highly biased predictions as the error accumulates over the models.

We address this problem proposing a simultaneous equations model. This model, popular in the economics literature, models a set of linear relationships where the response variable appears as explanatory variable for the other response variable (e.g. modeling supply and demand in economics). Such approach allows to answer both research questions with one model.

%\paragraph{Neural networks}
%The models discussed above aim at predicting mill rate and use its relationship with tax rate to predict next year's property assessment. While this approach is valid, it does not take advantage of all the information present in the dataset provided by the client. In fact, computing mill rate requires aggregating property information, therefore reducing the number of data points available for our analysis.

%A different approach is to directly predict a property assessment value for next year for each of the properties in the dataset, hence disregarding the relationship with mill rate. Given the amount of data available for this task, we will explore using a neural network to model next year's property assessment value.

%Since neural networks can learn hidden features in the data, we believe that a neural network with high generalization ability will intrinsically learn the relationship between property value and mill rate.

\bigskip
Initial covariates will be chosen based on previous work by STAT 450 students and client suggestion.  Feature selection will be performed using stepwise model selection based on Akaike information criterion (AIC), and multicollinearity will be assessed using the variance inflation factor (VIF). In order to evaluate each model, the data will be randomly split into two data sets: a training set and a test set. Each model will be fitted using the training data. Each model will then be evaluated by its prediction accuracy on observations in the test set. Because the client is interested in high accuracy predictions, the model with the most precise test set predictions will be chosen as the final model. This process will be repeated for both research questions.

\section{Shiny app}

After the modelling process is over and the final models are selected, they will be implemented in a Shiny app. Once the government publishes the assessments in January, the client will be able to estimate that year's tax payment based on previous years' information and the current years' assessment. After the mill rates are released in April, the client will then be able to estimate next year's assessment. The app will allow the client to export the information into Excel, which will make further analyses easier to carry out.

% \begin{itemize}
%     \item For the EDA part, be specific about which variables/relations you want to illustrate, explore, etc. Specify which model assumptions need to be tested, why and how
%     \item For the modelling part, specify which statistical tool/method (not R function) you are planning to use to address each scientific question
% \end{itemize}

% Note: the points above are not mandatory and you should try not to overlap a lot with S450. For example, if S450 has performed an exhausting EDA, you don't need to include that in your report, you can reference to it.

%\section{Code}

%You can copy paste your \textsf{R} code below.

%\begin{lstlisting}[style=R, caption={You can caption you code}, captionpos=b]
%    # some comment
%    rnorm(200)
%    
%    ggplot(aes(x, y)) +
%        geom_line() 
%        
%    xbar <- mean(x)
%    s <- sd(x)
%\end{lstlisting}

%You can also include generic code like this:

%\begin{lstlisting}[style=regular, caption={You can caption you code}, captionpos=b]
%    Algorithm 1
%        1. For i=1, ...
%            If(something)
%            DO
%            Else
%            DO NOT
%        End
%\end{lstlisting}

\end{document}

### 2020-04-14 doc/550/s550 report.tex
Final tweaking for submission. -gd

--- 2020-04-14 doc/550/s550 no fig repor.tex
@@
 
-\bibliography{RealEstate.bib}
+\bibliography{../../ref/RealEstate.bib}
 
@@
 The variables are summarised in Table \ref{table:variables}.
-Missing mill rate values were imputed by using the average mill rate given by the data for the corresponding year and tax class code.
+Missing mill rate values were imputed by using the average municipality mill rate for the corresponding year and tax class code.
+This is sensible because mill rates are constant across municipalities, and so a more complex imputation technique is not necessary.
 
@@
 Furthermore, the data set lacks appropriate spatial information that would allow for geographical variables to be taken into account.
+
+\begin{table}[ht]
+\centering
+\rowcolors{2}{gray!20}{white}
+\begin{tabular}{ll}
+\rowcolor{gray!40}
+\textbf{Variable}                & \textbf{Type}               \\ \hline
+Mill rate                                                              & Continuous         \\
+Total assessment                                                       & Continuous         \\
+Total land assessment                                                  & Continuous         \\
+Total improvement assessment & Continuous         \\
+Tax class code                                                         & \begin{tabular}[c]{@{}l@{}}Categorical. \\ One of 1 (residential), 5 (industrial), 6 (commercial)\end{tabular}       \\
+Municipality                                                           & Categorical        \\
+Year & \begin{tabular}[c]{@{}l@{}}Discrete. \\ Values of 2016, 2017, 2018, and 2019\end{tabular} 
+\end{tabular}
+\caption{Brief description of the variables included in the data set.}
+\label{table:variables}
+\end{table}
 
@@
 
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/1. scatter with year.pdf}
+  \caption{Scatterplot with year in different colors.}
+  \label{fig:scatter_year}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/2. scatter 2020 by municipality.pdf}
+  \caption{Scatterplot only for 2020.}
+  \label{fig:scatter_2020}
+\end{subfigure}
+\caption{Average log assessment values against log mill rates by municipality across the three different tax classes of interest.}
+\label{fig:scatter}
+\end{figure}
+
+%\begin{figure}
+%    \centering
+%    \includegraphics[scale=0.55]{figures/1. scatter with year.pdf}
+%    \caption{Log assessment values against log mill rates across the three different tax classes of interest, with year in different colors.}
+%    \label{fig:scatter_year}
+%\end{figure}
+
 The fact that mill rates remain relatively constant throughout the years is further verified in Figure \ref{fig:evolution_millrate}, where industrial properties are again seen to have the largest mill rates on average.
@@
 Notably, Figure \ref{fig:violin_assessment} suggests that this difference is not present in assessment values, which indicates that tax class may not be that relevant for predicting future years' assessment values.
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/6. mill rate evolution sample.pdf}
+  \caption{Evolution of (log) mill rates.}
+  \label{fig:evolution_millrate}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/6.1 assessment evolution sample.pdf}
+  \caption{Evolution of (log) assessment values.}
+  \label{fig:evolution_assessment}
+\end{subfigure}
+\caption{Evolution of target variables over the years for some randomly-selected properties.}
+\label{fig:evolution}
+\end{figure}
+
+%\begin{figure}[ht]
+%    \centering
+%    \includegraphics[scale=0.45]{figures/6. mill rate evolution sample.pdf}
+%    \caption{Evolution of (log) mill rates over the years for some randomly-selected properties.}
+%    \label{fig:rates_by_year}
+%\end{figure}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/7. violin mill rates.pdf}
+  \caption{Distribution of (log) mill rates.}
+  \label{fig:violin_millrate}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/7.1 violin assessment values.pdf}
+  \caption{Distribution of (log) assessment values.}
+  \label{fig:violin_assessment}
+\end{subfigure}
+\caption{Violin and boxplots of (log) mill rates and (log) assessment values of individual properties across the three tax classes of interest, taking into account only the year 2020.}
+\label{fig:violin}
+\end{figure}
+
+%\begin{figure}[ht]
+%    \centering
+%    \includegraphics[scale=0.45]{figures/7. violin mill rates.pdf}
+%    \caption{Violin and boxplots of (log) mill rates of individual properties across the three tax classes of interest, taking into account only the year 2020.}
+%    \label{fig:violin}
+%\end{figure}
 
@@
 RF is preferred over a single decision tree as it overcomes the problem of over-fitting.
+The following models were constructed using the 'randomForest' package in R \citep{rfpack}.
 
@@
 
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/nodepur.png}
+  \caption{Variable importance plot for mill rates.}
+  \label{fig:4a}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/asnodepur.png}
+  \caption{Variable importance plot for assessment values.}
+  \label{fig:4b}
+\end{subfigure}
+\caption{Variable importance plot based on increase in node purity for mill rate and total assessment value predictions.}
+\label{fig:4}
+\end{figure}
+
 % Don't think it is necessary to mention bootstrapping, this is done automatically in RF.
@@
 However, RF shows good performance in predicting next year's assessment values.
+
+% Bagging (aggregate bootstrap) was implemented by the random forest algorithm and performance is shown in Figure \ref{fig:5}.
+Bagging uses bootstrapping---sampling the data with replacement---and aggregation---combining several model predictions to obtain the best prediction.
+The accuracy of predicting mill rates is less optimal for higher mill rates.
+However, overall, the prediction power of the random forest models are desirable, particularly for predicting next year's assessment value.
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/bagmi.png}
+  \caption{Performance for mill rates prediction.}
+  \label{fig:5a}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/bagas.png}
+  \caption{Performance for assessment value prediction.}
+  \label{fig:5b}
+\end{subfigure}
+\caption{Models' performance on test set.}
+\label{fig:5}
+\end{figure}
+\FloatBarrier
 
@@
 
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/lmemill.png}
+  \caption{Summary output for predicting mill rates.}
+  \label{fig:6a}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/lmeas.png}
+  \caption{Summary output for predicting assessment values.}
+  \label{fig:6b}
+\end{subfigure}
+\caption{Models summary output.}
+\label{fig:6}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/diagmill.png}
+  \caption{Diagnostics plots for LME model for mill rate.}
+  \label{fig:7a}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/diagas.png}
+  \caption{Diagnostics plots for LME model for assessment values.}
+  \label{fig:7b}
+\end{subfigure}
+\caption{Diagnostics plots.}
+\label{fig:7}
+\end{figure}
+
 Unfortunately, the diagnostics plots verifying model assumptions are subpar for both LME models, as can be seen in Figure \ref{fig:7}.
@@
 However, as was argued in Section \ref{sec:data}, mill rates and assessment values are highly dependent on spatial information.
-An attempt to obtain the properties' spatial information was made by transforming the addresses into spatial coordinates using an open source geocoding system provided by the British Columbia Government.
+An attempt to obtain the properties' spatial information was made by transforming the addresses into spatial coordinates using an open source geocoding system provided by the British Columbia Government (see the documentation of \citet{geopy}).
 However, the open source geocoding system did not recognize some of the addresses in the dataset and provided incorrect coordinates for a large number of properties.
@@
 
-%%%% appendix
+%%%% appendix and bibliography
+\newpage
 \printbibliography
